### 2014-05-23 arch.tex
Initial draft.

\section{Monitoring Architecture Model}
\label{sec:arch}

% Run-time monitoring overview
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/monitoring_architecture.pdf}
    \vspace{-0.2in}
    \caption{Overview of run-time monitoring architecture.}
    \label{fig:arch.overview} 
    \vspace{-0.2in}
  \end{center}
\end{figure}

Figure~\ref{fig:arch.overview} shows an overview of the run-time monitoring
model that is assumed in this paper. 
% The \emph{main task} is a computation task that performs the original function
% of the real-time system and is run on the \emph{main core}. 
The \emph{main program} is a computation task that performs the original
function of the system and is run on the \emph{main core}.
% The main task has a
% worst-case execution time (WCET) which can be upper-bounded using a variety
% of existing techniques \cite{wcetsurvey-tecs08}. This WCET is used for timing
% analysis in the design of the real-time system. As long as the execution
% time of the main task does not exceed this bound, then all timing guarantees
% from the analysis hold.
On certain events
during the main program, such as the execution of certain types of
instructions, the \emph{monitor} performs a series of \emph{monitoring
operations}. The monitor operates in parallel to the main core. These events
are referred to as
\emph{monitoring events}. Depending on the type of a monitoring event, different
monitoring operations are executed. Monitoring events are buffered
in a FIFO
structure to decouple the running of the main core and the monitor. If the FIFO
is full, then the main core is forced to stall on a monitoring event until a
FIFO entry becomes available. For programmable monitors, these stalls are a
major source of overhead because the monitor may take several cycles to process
a single event from the main core. We refer to these stalls and other
overheads,
such as contention for shared resources, as
\emph{monitoring overheads}. If the monitor detects an %inconsistent or
undesired behavior in the monitoring events, then an error is detected. 
% We do
% not focus here on the question on how to handle such an error. However, there are
% several options on how to handle an error such as raising an exception,
% notifying the user, or switching to a simpler, more trusted main task
% \cite{sha-simplex-sw01}.

There are many possible monitoring schemes that can be implemented on this
type of fine-grained parallel monitoring architecture such as
array bound checks \cite{devietti-hardbound-asplos08}, memory protection \cite{mmp02}, 
information flow tracking \cite{suh-dift-asplos04, greathouse-testudo-micro08}, 
soft error detection \cite{argus-micro07}, data-race detection \cite{cord-hpca06}, etc.
One example is an
uninitialized memory check (UMC) where 
monitoring is used to detect when software attempts to read from a memory location that
was not previously initialized. This can be done by forwarding each load and store
instruction from the main core to the monitor.
%the memory address
%of each store and load instruction on the main core to the monitor as
%well as whether the instruction was a store or load. 
For every memory location,
the monitor keeps one bit of metadata. On a store to a
memory location, the monitor marks the corresponding metadata bit to
indicate that the memory location has been initialized. On a load,
the monitor checks that the corresponding metadata bit has
been previously marked as initialized.

There are multiple options for implementing programmable parallel monitors. For example, the
Log-Based Architecture \cite{chen08-lba} uses processor cores in a multi-core
system as monitors. The FlexCore architecture \cite{deng-flexcore-micro10}
instead uses an FPGA-fabric to implement the monitor. The approach we describe
in this paper applies to any of these parallel monitors. However, for experiments,
we model an FPGA-based monitor similar to FlexCore. 
%Specifically, this
%architecture uses an on-chip FPGA fabric to
%allow high-performance and reconfigurable run-time monitoring. 
The on-chip FPGA fabric is
used to implement the ``Monitor'' block in Figure~\ref{fig:arch.overview}
while the FIFO from the main core and metadata cache are implemented as ASICs.

### 2014-05-23 drop.tex
Initial draft.

\section{Enabling Coverage/Overhead Trade-off}
\label{sec:drop}

Figure~\ref{dummy} shows the overheads of array bounds check with filtering
based on initialized metadata (see Section~\ref{sec:evaluation} for details on
the experimental methodology). The overheads range from XX\% to XX\%. In some
contexts, this overhead may still be considered to be too high. In this
section, we discuss how we can selectively perform monitoring in order to
match a specified overhead target.

The high-level idea is to drop monitoring operations in order to remain within
the specified overhead budget. Given an overhead budget, the system estimates
the run-time overhead caused by monitoring (Section~\ref{sec:drop.slack}). When
this overhead exceeds the budget, then the system drops monitoring operations.
In order to prevent false positives, we must be able to identify flows that
have been dropped. This can be done by extending the dataflow engine to keep
track of an additional ``drop'' flag (Section~\ref{sec:drop.false_positives}).

\subsection{Measuring Overheads}
\label{sec:drop.slack}

In order to attempt to match an overhead budget, we need a way to estimate the
actual overheads incurred due to monitoring. More precisely, what we care about
is the difference between the actual overheads and the overhead budget.
Slack is essentially a measure of this overhead error. We can measure slack by
doing the following.

% Slack
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/slack.pdf}
    \vspace{-0.3in}
    \caption{Slack and its effect on monitoring over time.}
    \label{fig:drop.slack}
    \vspace{-0.2in}
  \end{center}
\end{figure}

% Slack tracking module
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/stm.pdf}
    \vspace{-0.3in}
    \caption{Slack tracking and drop decision hardware.}
    \label{fig:drop.stm}
    \vspace{-0.1in}
  \end{center}
\end{figure}

In order to decide whether a monitoring event should be dropped or not, we need
a way to keep track of overheads. In this section, we present a specific method
for measuring execution time overheads. However, we note that other methods
could be used in order to target budgets on IPC, power, energy, etc.  For
example, one way to measure execution time overheads would be to insert
checkpoints in the program. By recording the execution time of these
checkpoints without monitoring, the overheads incurred while running with
monitoring could be determined. Although this would give an accurate
measurement of overheads, it would require modifying the program binary. In
addition, the overhead budget for monitoring would only be updated at
checkpoints.

Instead, we develop a model to estimate the execution time overhead budget in a
gradual manner and without the need to modify the main program. Specifically,
our model of execution time budget is that a percentage of the main program's
executed cycles is the number of cycles of overhead that we allow for
monitoring.  We define \emph{slack} as the number of cycles of monitoring
overhead that can be incurred while staying within the budget target.  For
example, if no monitoring overheads occur during 1000 cycles of the main
program's execution and the designer sets a 20\% overhead target, then the
slack that is built up during this period is 200 cycles. This slack can be used
to perform monitoring without exceeding the overhead budget.
Figure~\ref{fig:drop.slack} shows an example of how slack can change over time.
As monitoring overheads occur, the slack is decremented. If the slack falls
below the overhead expected for handling a monitoring event, then the event is
dropped. In addition to this accumulated slack, a small amount of initial slack
can be given in order for monitoring to be performed at the start of a program.

Figure~\ref{fig:drop.stm} shows a hardware slack tracking module (STM) for
keeping track of slack. The STM uses a counter that increments on every $k$-th
cycle of the main core. This $k$ can be calculated by taking the reciprocal of
the target budget. For example, if the target budget is 20\%, then the counter
increments on every 5th clock cycle.  The value of this counter is the
accumulated slack.  Whenever the main core is stalled due to the monitor, slack
is decremented.

In addition to keeping track of overheads, a policy must be developed to
determine when monitoring event should be dropped. Complex policies involving
multiple conditions and safety factors could be used to finely tune
the trade off between monitoring and performance overhead. Similarly, introducing
randomness into the policy could allow for better sampling in a cooperative
debugging application of monitoring. Here, we present one simple
dropping policy that is based on the accumulated slack.
When the monitor is initialized, the monitor will load
its needed slack in order to perform a monitoring operation into a register. When
there is a monitoring event in the FIFO to be processed, a hardware comparator
checks whether the accumulated slack is greater than or equal to the necessary
slack for full monitoring. If enough slack exists, the monitor is signaled to
perform the monitoring operation. Otherwise, the monitoring event is dropped.

\subsection{Preventing False Positives}
\label{sec:drop.false_positives}

Dropping a monitoring operation implies that some functionality of the monitor
has been lost. This may cause either false negatives, where an error that
occurs in the main program's execution is not detected, or false positives,
where the monitor incorrectly believes an error has occurred. For example, a
false positive can occur for UMC if a store monitoring event is dropped. This
causes the store memory location to not be marked as initialized. A subsequent
load for the memory location will incorrectly cause an error to be raised.  We
accept false negatives as the loss in coverage that we trade off in order to
limit overheads.  However, we must safely drop monitoring events in such a way
as to avoid false positives so that the system does not incorrectly raise an
error.

Dropping monitoring events can create false negatives and false positives.
False negatives are tolerated. We handle false positives by invalidating
metadata.

% Flag propagation rules for BC
\begin{table}[tb]
  \begin{center}
    \begin{small}
    \input{tabs/propagation_rules}
    \end{small}
    \caption{Flag propagation rules for array bounds check. X indicates that no
    propagation is done while 0 or 1 indicates that the destination flag is set
    to that value.}
    \label{tab:drop.propagation_rules}
    \vspace{-0.2in}
  \end{center}
\end{table}

Since metadata derived from dropped metadata is also invalid. We propagate the
drop flag in a similar manner to the initialization flags using the dataflow
engine. Since certain metadata is invalid, we can filter out useless monitoring
done on this invalid metadata. In the typical case, the propagation rule is
that if any of the source operands is marked as dropped, then the destination
operand is marked as dropped. This is slightly different that the
initialization flags which propagate an uninitialized state only if both
sources are uninitialized. Table~\ref{tab:drop.propagation_rules} shows the
propagation rules that we use for array bounds check.  However, we note that
the propagation rules can be customized depending on the monitor by
reconfiguring the Filter Decision Table.

\subsection{Dropping Decision}

% Varying slack's impact on coverage for UMC
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/dataflow_graph.pdf}
    \vspace{-0.3in}
    \caption{Example dataflow graph for metadata.}
    \label{fig:drop.dataflow_graph}
    \vspace{-0.1in}
  \end{center}
\end{figure}

One of the issues with waiting until the overhead budget is reached (i.e.,
slack is zero) to start dropping monitoring events is that this can result in
wasted work. For example, consider the dataflow graph shown in
Figure~\ref{fig:drop.dataflow_graph}. If event {\tt E} is dropped, then this
means that monitoring for events {\tt C} and {\tt D} was wasted.

One method to eliminate this wasted work is to only make dropping decisions at
the root of these dataflow graphs. That is, we will decide to either monitor or
not monitor an entire metadata flow. We refer to this dropping decision policy
as \emph{source dropping} and we refer to the previous policy of waiting until
the overhead budget is met as \emph{last-minute dropping}. Source dropping will
result in no wasted work. However, because of the coarser-grained decision, it
may be more difficult to closely match overheads. Source dropping is expected
to work well when there are a large number of independent dataflows.

Another possible method is to make the dropping decisions somewhere in between
\emph{source dropping} and \emph{last-minute dropping}. From
Figure~\ref{fig:drop.dataflow_graph}, what we would like to do is make a
decision at event {\tt C} whether to perform monitoring on that branch of the
dataflow. If we choose to perform monitoring, then we want to perform all
monitoring on that flow. Instead, if we choose to drop event {\tt C}, then the
entire flow is dropped with no wasted work. Similarly, dropping event {\tt F}
drops the bottom flow with no wasted work and event {\tt A} can be dropped to
drop the entire shown dataflow with no wasted work. If the program can be
analyzed or profiled to identify these instructions which, if dropped, lead to
no wasted work, then this information can be used at run-time to minimize the
amount of wasted work. We refer to this policy as \emph{min-point dropping}.
Min-point dropping is a coarser-grained decision that last-minute dropping, but
finer-grained that source dropping. Thus, its ability to match overheads is
expected to sit between these other two policies. Similarly, there is still a
chance of wasted work due to strange splitting and merging dataflow graph
shapes but typically we expect much less wasted work than last-minute
dropping. Thus, we expect the coverage achieved by min-point dropping for a
certain overhead to fall between source and last-minute dropping. These three
dropping policies create a trade-off space between coverage achieved and
ability to match an overhead target.

\subsection{Multiple-Run Coverage}
Coverage can be considered for a single or average run instance. It can also be
measured over multiple users/runs in order to capture total bug coverage of a
program.

The system presented so far operates in a deterministic manner and so each run
will perform the same monitoring checks. However, we can introduce randomness
in order to achieve different dropping patterns over runs and achieve higher
multiple-run coverage.

\subsection{Multi-Core Implementation}

This can be extended to mulit-core by using multiple dataflow engines and a
similar coherency protocol as that used for data caches.

### 2014-05-23 evaluation.tex
Initial draft.

\section{Evaluation}
\label{sec:evaluation}

\subsection{Experimental Setup}

\subsection{Filtering of Monitoring Events}

Effect of using initialized flag to filter out monitoring events and lower overheads.

\subsection{Coverage with Adjustable Overheads}
Coverage and overhead budget accuracy for source, last-minute, and min-point dropping.

\subsection{Randomness for Multi-Run Coverage}

\subsection{Custom Hardware Monitors}

\subsection{Main Core Scaling}

\subsection{Area and Power Overheads}

### 2014-05-23 extensions.tex
Initial draft.

\section{Application to Monitoring Schemes}
\label{sec:extensions}

This section describes how three example monitoring schemes map to the proposed
architecture.

\subsection{Array Bounds Check}

\subsection{Uninitialized Memory Check}

\subsection{Dynamic Information Flow Tracking}

\subsection{Applicability to Other Schemes}

### 2014-05-23 filter.tex
Initial draft.

\section{Reducing Monitoring Overheads}
\label{sec:filter}

The instruction-grained monitoring techniques we consider forward all
instructions of certain types (e.g., load instructions) to the monitoring core.
However, not all of these forwarded instructions have relevant monitoring to be
done. For example, for array bounds check, all load and store instructions are
forwarded. However, only loads and stores associated with array pointers have
relevant checking to be done. Thus, we could greatly reduce the amount of
monitoring done by only forwarding instructions that correspond to data that
correspond to array pointers and have been marked with base and bound metadata.
More generally, we can reduce monitoring overheads by filtering out certain
events that correspond to empty (i.e., uninitialized) metadata.

% Overview of inserting dataflow engine
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/dataflow_overview.pdf}
    \vspace{-0.2in}
    \caption{The dataflow engine is inserted between the main core and the monitoring core.}
    \label{fig:filter.overview} 
    \vspace{-0.1in}
  \end{center}
\end{figure}

In this section, we describe how a DIFT-like hardware engine can be used to track
whether metadata has been initialized. We use this information to filter out
monitoring operations for uninitialized metadata. This engine sits between the
main core and the monitoring core (see Figure~\ref{fig:filter.overview}).

\subsection{Initialized Metadata}

% Detailed architecture of dataflow engine
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/dataflow_architecture.pdf}
    \vspace{-0.3in}
    \caption{Hardware architecture of dataflow engine.}
    \label{fig:filter.architecture} 
    \vspace{-0.1in}
  \end{center}
\end{figure*}

Figure~\ref{fig:filter.architecture} shows a block diagram of the hardware for keeping track of
initialized metadata. When metadata is initialized, the monitor sends a signal
to the DIFT engine to mark a corresponding initialization flag. Similarly, when
the monitor propagates metadata, it also sends a signal to initialize the
metadata that has been written to. When a new event arrives at the DIFT engine,
it will read in up to two initialization flags. A table is used to configure
which initialization flags should be read in depending on the instruction type of
the event and information from the monitoring event. A pair of simple ALUs is
used to allow calculating memory addresses based on information from the
monitoring event. Typically, the initialization flags read correspond to the
source operands of the monitored instruction.

Given this set of initialization flags, a Filter Decision Table decides, based
on instruction type, whether an event can be filtered out. This table is
configured based on the specific semantics of a monitoring technique.
Typically, if all initialization flags are marked as uninitialized, then the
event can be filtered out. For example, for array bounds check, if the source
register's metadata is marked as uninitialized, then this means that the
register does not correspond to an array pointer. Thus, we can filter out
performing monitoring for this event.

\subsection{Propagating Initialization Flags}

If the source operands are marked as uninitialized, then typically we can
filter out this event. However, in these cases, the monitor also typically
resets the destination operand's metadata as cleared or uninitialized.
Thus, when we filter out one of these events, we need to also propagate the
uninitialized flag to the destination. The Filter Decision Table includes this
information about whether the destination operand metadata should be marked as
uninitialized.

### 2014-05-23 introduction.tex
Initial draft.

\section{Introduction}
\label{sec:intro}

% Parallel run-time useful
Run-time monitoring techniques have been shown to be useful for improving the
reliability, security, and debugging capabilities of computer systems. For
example, Hardbound is a hardware-assisted technique to detect out-of-bound
memory accesses, which can cause undesired behavior or create a security
vulnerability if uncaught \cite{hardbound-asplos08}.  Intel has
recently announced plans to support buffer overflow protection similar to
Hardbound in future architectures \cite{intel-mpx}. Similarly, run-time
monitoring can enable many other new security, reliability, and debugging
capabilities such as fine-grained memory protection \cite{mondrian-asplos02},
information flow tracking \cite{dift-asplos04, testudo-micro08}, hardware error
detection \cite{argus-micro07}, data-race detection \cite{radish-isca12,
cord-hpca06}, etc.

% Programmable hardware monitors
These run-time monitors introduce performance, power, and energy overheads.
Implementing run-time monitoring in software using binary instrumentation or
other similar methods introduces high overheads. For example,
dynamic information flow tracking (DIFT) implemented in software suffers a 3.6x
slowdown \cite{lift-micro06}. Implementing monitors in hardware greatly decreases
the performance impact by performing monitoring in parallel to a program's
execution. A dedicated hardware implementation of DIFT reduces performance
overheads to just 1.1\% \cite{dift-asplos04}. However, fixed hardware loses
the programmability of software, and cannot be updated or changed.

% Recent studies proposed programmable parallel hardware as a way to achieve both
% the performance advantages of hardware and the flexibility of software
% \cite{lba-isca08, flexcore-micro10, harmoni-dsn12}.  While more
% efficient than software, these programmable hardware monitors can still show
% significant performance overheads.  For example,
% Figure~\ref{fig:intro.full_mon} shows the execution time for three monitoring
% techniques normalized to the execution time with no monitoring for an
% FPGA-based parallel monitor. The program being monitored is run on an in-order
% embedded core running at 500 MHz while the FPGA-based monitor runs at 250 MHz.
% We can see that, depending on the monitoring scheme, overheads can easily be
% several tens of percent.  A faster or more complex (e.g., superscalar or
% out-of-order) core is expected to show even more significant overheads.

% % Run-time monitoring overview
% \begin{figure}
%   \begin{center}
%     \vspace{-0.2in}
%     \includegraphics[width=\columnwidth]{figs/full_mon.pdf}
%     \vspace{-0.5in}
%     \caption{Performance overheads of FPGA-based monitors.}
%     \vspace{-0.2in}
%     \label{fig:intro.full_mon} 
%   \end{center}
% \end{figure}

In this paper, we present an architecture that combines hardware-based dynamic
information flow tracking with a software-based monitor in order to reduce the
overheads of monitoring. By using the DIFT engine to filter out events that do
not correspond to valid metadata, we can greatly reduce the amount of
monitoring that must be handled by the monitor. For example, for an array
bounds check, we are able to filter out instruction that do not correspond to
pointer data. Since the monitor is implemented in software, the system allows a
large range of monitors to be implemented and even updated or changed at
run-time.

With this filtering, we see overheads of XX\% - XX\% on average for
uninitialized memory check, array bounds check, and multi-bit dynamic
information flow tracking. Although full monitoring coverage is always desired,
if these overheads are considered too high by the designer or user then
monitoring must be disabled.
% % Current schemes are all-or-nothing: suffer full performance impact for full
% % coverage.
% Traditionally, these overheads have been an all-or-nothing cost. If the
% designer finds that the overheads exceed their desired budget, then monitoring
% cannot be enabled.  In this paper, we present an architecture that enables a
% trade-off between the overheads and the coverage of a monitoring technique by
% dynamically adjusting the number of monitoring operations.  This system allows
% monitoring to still be performed even when the overhead budget is not
% sufficient for full monitoring.
% Use cases
Trading off coverage for
reduced overheads can still be very useful. For example, this adjustable
monitoring enables a level of protection even for systems where the full
monitoring overheads are too high.  This is especially true for energy- or
power-constrained systems or soft real-time systems where the monitoring
overhead should not exceed energy/power limits or real-time deadlines. 
Additionally, adjustable monitoring can be used for debugging purposes
especially in the context of sampling-based or cooperative debugging techniques
which expect low coverage per run but use a large number of runs and users to
collect debugging information~\cite{liblit-pldi05,
chilimbi-asplos04, greathouse-cgo11}.

Thus, we develop an architecture that enables a
trade-off between the overheads and the coverage of a monitoring technique by
dynamically adjusting the number of monitoring operations. By extending the
hardware DIFT engine to track 2-bits of information, we are able to use it for
filtering empty metadata as well as tracking dropped monitoring flows in order
to prevent false positives. We introduce three policies of where dropping
decisions are made and show the trade-off space achieved between accuracy of
overhead matching and coverage achieved.

% %Invalidation in order to prevent false positives.
% In order to adjust monitoring overheads, a system must be able to drop a
% portion of the monitoring operations. Unfortunately, however, simply skipping
% monitoring operations can lead to false positives (false alarms) due to stale
% metadata.  One possible solution is to rewrite individual monitors to create
% ``drop-enabled'' versions.  However, such monitor-specific customizations are
% difficult and time consuming.  Instead, after analyzing a range of monitoring
% techniques, we found that by adding a 1-bit ``invalid'' flag to the bookkeeping
% metadata managed by the monitor, we were able to create a general mechanism to
% handle false positives across a broad range of monitoring techniques.  The
% result is that we have designed a hardware module that sits between a
% processing core and a programmable monitor, and allows adjustable overheads.
% This hardware module can be applied to a range of monitoring techniques and
% requires no modifications to the monitor.

% Evaluation
In order to evaluate our approach, we applied it to 3 different monitoring
techniques. These monitoring techniques vary in what events they monitor, the
size and semantics of their metadata, and the operations performed on metadata.
Our experiments show that with 10\% overheads, we can still provide 49-90\% of
coverage on average depending on the monitoring technique. By increasing the
overhead budget, the coverage rate can be increased. Our architecture is able
to closely meet the specified budgets. For all but one benchmark, the overheads
seen were within 2\% of the target overhead. Our results show that X policy is
able to achieve very close overhead matching (within X\% on average) with an
average coverage of X\%. On the other hand, by using Y policy, the overhead
matching is not as high (within Y\% on average) but average coverage increases
to (Y\%).

% Section summary
This paper is organized as follows. Section~\ref{sec:monitoring} presents the
parallel run-time monitoring model that this paper assumes.
Section~\ref{sec:filter} presents our architecture that use hardware-based DIFT
in order to reduce monitoring overheads. In Section~\ref{sec:drop} we present
how the DIFT engine can be extended to enable adjustable overheads.
Section~\ref{sec:extensions} covers example implementations of three monitoring
schemes. Section~\ref{sec:evaluation} presents our evaluation methodology and
results. Finally, we discuss related work in Section~\ref{sec:related} and
conclude in Section~\ref{sec:conclusion}.

### 2014-05-23 monitoring.tex
Initial draft.

\section{Parallel Run-Time Monitoring}
\label{sec:monitoring}

% Run-time monitoring overview
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/monitoring_architecture.pdf}
    \vspace{-0.2in}
    \caption{Overview of run-time monitoring architecture.}
    \label{fig:arch.overview} 
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:arch.overview} shows an overview of the run-time monitoring
model that is assumed in this paper.  The \emph{main program} is a computation
task that performs the original function of the system and is run on the
\emph{main core}.  On certain events during the main program, such as the
execution of certain types of instructions, the \emph{monitor} performs a
series of \emph{monitoring operations}. The monitor operates in parallel to the
main core. These events are referred to as \emph{monitoring events}. Depending
on the type of a monitoring event, different monitoring operations are
executed. Monitoring events are buffered in a FIFO structure to decouple the
running of the main core and the monitor. If the FIFO is full, then the main
core is forced to stall on a monitoring event until a FIFO entry becomes
available. For programmable monitors, these stalls are a major source of
overhead because the monitor may take several cycles to process a single event
from the main core. We refer to these stalls and other overheads, such as
contention for shared resources, as \emph{monitoring overheads}. If the monitor
detects an inconsistent or undesired behavior in the monitoring events, then
an error is detected.

There are many possible monitoring schemes that can be implemented on this type
of fine-grained parallel monitoring architecture such as memory protection
\cite{mondrian-asplos02}, information flow tracking \cite{dift-asplos04,
testudo-micro08}, soft error detection \cite{argus-micro07}, data-race
detection \cite{cord-hpca06}, etc.  For example, an array bounds check (BC)
\cite{hardbound-asplos08} can be implemented where monitoring is used to detect
when software attempts to read or write to a memory location outside of an
array's bounds. This can be done by associating base and bound address metadata
to an array pointer when it is allocated. On loads or stores with the array
pointer, the monitor checks that the memory address is within the base and
bound. In addition, on certain ALU instructions, this base and bound metadata
is propagated to track the creation of new array pointers.
% One example is an uninitialized memory check (UMC) where monitoring is used to
% detect when software attempts to read from a memory location that was not
% previously initialized. This can be done by forwarding each load and store
% instruction from the main core to the monitor. For every memory location, the
% monitor keeps one bit of metadata. On a store to a memory location, the monitor
% marks the corresponding metadata bit to indicate that the memory location has
% been initialized. On a load, the monitor checks that the corresponding metadata
% bit has been previously marked as initialized.

There are multiple options for implementing programmable parallel monitors. For
example, the Log-Based Architecture \cite{lba-isca08} uses processor cores in a
multi-core system as monitors. The FlexCore architecture
\cite{flexcore-micro10} instead uses an FPGA-fabric to implement the
monitor. The approach we describe in this paper applies to any of these
parallel monitors. However, for experiments, we model an FPGA-based monitor
similar to FlexCore.  The on-chip FPGA fabric is used to implement the
``Monitor'' block in Figure~\ref{fig:arch.overview} while the FIFO from the
main core and metadata cache are implemented as ASICs.

### 2014-05-23 related.tex
Initial draft.

\section{Related Work}
\label{sec:related}

% Monitoring
Monitoring

% Filtering
Hardbound filters out non-pointer data \cite{hardbound-asplos08}. FlexiTaint filters out some monitoring
and performs it in hardware. FADE provides a more general filtering engine. It
filters out redundant updates and clean checks by reading metadata values. Our
dataflow engine does not require reading the actual metadata values which may
be large. We do not handle filtering redundant updates, but we do filter out
clean checks as well as clean (i.e., uninitialized) propagations. We are able
to do this because the dataflow engine can also propagate the uninitialized
information. In addition, our dataflow engine can be used to provide adjustable
overheads without false positives which is not a capability described by FADE.

% Adjustable overheads
Adjustable overhead/cooperative debugging: Testudo \cite{testudo-micro08}, QVM
\cite{qvm-oopsla08}, Greathouse et al. \cite{greathouse-cgo11}, Huang et al.
\cite{huang-sttt12}, Lo et al. \cite{lo-rtas14}

### 2014-05-23 title.tex
Initial draft.

\title{
Using Dynamic Dataflow Tracking to Reduce Overheads for \\ Parallel
Run-Time Monitoring
}

\ifanonymized{\author{Anonymized}}
{\author{Daniel Lo, Tao Chen, Mohamed Ismail, and G. Edward Suh}}

\date{}
\maketitle

\thispagestyle{empty}

\begin{abstract}

Recent studies have proposed various parallel run-time monitoring techniques to
improve the reliability, security, and debugging capabilities of
computer systems. These run-time monitors introduce overheads in performance
and energy to the system. Traditionally, these overheads have been an
all-or-nothing cost; monitoring cannot be used if the overheads are considered too large.
% where if monitoring is enabled, then the program will incur, in full, the
% associated overheads. 
%That is, if the overheads are considered too large, then monitoring cannot be
%enabled. 
In this paper, we present a run-time monitoring architecture that
enables a trade-off between the coverage and the overheads.
The system allows an overhead or coverage target to be specified,
and limits the amount of monitoring performed at run-time to meet the target. 
However, simply skipping monitoring operations can lead to false
positives, signaling errors when none have occurred. Our proposed architecture
prevents false positives using a method that invalidates certain metadata
when monitoring is not performed. Dedicated hardware that performs dynamic
information flow tracking (DIFT) is used to propagate this invalidation
information in order to prevent false positives. In addition, we can use this
DIFT hardware to also track the presence of valid metadata. This allows us to
filter out events in order to reduce monitoring overheads.
% Our experiments show that the system is able to
% closely meet the target overhead and achieve significant monitoring coverage
% for reasonable overhead targets.
In our experiments, we see that the system is able to come
within 2\% of the specified execution time overhead in almost all cases.  In addition, with a
performance overhead target of 10\%, 49-90\% of monitoring coverage is achieved
on average depending on the monitoring technique.

\end{abstract}

### 2014-05-23 filter.tex
Edits on Sections 1-3.

@@
-\section{Reducing Monitoring Overheads}
+\section{Filtering Monitoring Events}
 \label{sec:filter}
@@
 forwarded. However, only loads and stores associated with array pointers have
-relevant checking to be done. Thus, we could greatly reduce the amount of
-monitoring done by only forwarding instructions that correspond to data that
-correspond to array pointers and have been marked with base and bound metadata.
-More generally, we can reduce monitoring overheads by filtering out certain
-events that correspond to empty (i.e., uninitialized) metadata.
+relevant monitoring operations to be done. Thus, we can greatly reduce the amount of
+monitoring done by only forwarding instructions to the monitoring core that 
+correspond to array pointers. These accesses can be identified because their
+corresponding metadata has been set with base and bound information.
+More generally, we can reduce monitoring overheads by filtering out 
+events that correspond to clean or uninitialized metadata.
 
@@
 
-In this section, we describe how a DIFT-like hardware engine can be used to track
-whether metadata has been initialized. We use this information to filter out
-monitoring operations for uninitialized metadata. This engine sits between the
-main core and the monitoring core (see Figure~\ref{fig:filter.overview}).
+In this section, we describe how we can use a dataflow tracking engine to
+efficiently track clean metadata. This engine sits between the main core and
+the monitoring core (see Figure~\ref{fig:filter.overview}) in order to filter
+out monitoring operations for clean metadata.
 
-\subsection{Initialized Metadata}
+\subsection{Tracking Clean Metadata}
 
@@
     \vspace{-0.3in}
-    \caption{Hardware architecture of dataflow engine.}
+    \caption{Hardware architecture for dataflow engine.}
     \label{fig:filter.architecture} 
@@
 
-Figure~\ref{fig:filter.architecture} shows a block diagram of the hardware for keeping track of
-initialized metadata. When metadata is initialized, the monitor sends a signal
-to the DIFT engine to mark a corresponding initialization flag. Similarly, when
-the monitor propagates metadata, it also sends a signal to initialize the
-metadata that has been written to. When a new event arrives at the DIFT engine,
-it will read in up to two initialization flags. A table is used to configure
-which initialization flags should be read in depending on the instruction type of
-the event and information from the monitoring event. A pair of simple ALUs is
+Figure~\ref{fig:filter.architecture} shows a block diagram of the dataflow
+tracking engine for
+keeping track of clean metadata. Initially, all metadata is marked as clean.
+When the monitoring core writes a non-null metadata value, it also sends a
+signal to the dataflow engine to mark the metadata as dirty.
+
+When a monitoring event arrives at the dataflow engine,
+it reads in up to two clean flags. A table is used to configure
+which clean flags should be read in depending on the instruction type of
+the event and information from the monitoring event. A pair of simple ALUs are
 used to allow calculating memory addresses based on information from the
-monitoring event. Typically, the initialization flags read correspond to the
+monitoring event. Typically, the clean flags read correspond to the
 source operands of the monitored instruction.
 
-Given this set of initialization flags, a Filter Decision Table decides, based
+Given this set of clean flags, a Filter Decision Table decides, based
 on instruction type, whether an event can be filtered out. This table is
 configured based on the specific semantics of a monitoring technique.
-Typically, if all initialization flags are marked as uninitialized, then the
+Typically, if all clean flags are marked as clean, then the
 event can be filtered out. For example, for array bounds check, if the source
-register's metadata is marked as uninitialized, then this means that the
-register does not correspond to an array pointer. Thus, we can filter out
-performing monitoring for this event.
-
-\subsection{Propagating Initialization Flags}
-
-If the source operands are marked as uninitialized, then typically we can
-filter out this event. However, in these cases, the monitor also typically
-resets the destination operand's metadata as cleared or uninitialized.
-Thus, when we filter out one of these events, we need to also propagate the
-uninitialized flag to the destination. The Filter Decision Table includes this
-information about whether the destination operand metadata should be marked as
-uninitialized.
+register's metadata is marked as clean, then this means that the
+register does not correspond to an array pointer. Thus, we can filter out the
+monitoring operation for this event. 
+It is important that we also propagate the clean information to the destination
+of the instruction.  The Filter Decision Table also includes a bit indicating
+whether the destination operand metadata should be marked as uninitialized.

### 2014-05-23 introduction.tex
Edits on Sections 1-3.

@@
 Implementing run-time monitoring in software using binary instrumentation or
-other similar methods introduces high overheads. For example,
+other similar methods has been shown to introduce high overheads. For example,
 dynamic information flow tracking (DIFT) implemented in software suffers a 3.6x
@@
 
-In this paper, we present an architecture that combines hardware-based dynamic
-information flow tracking with a software-based monitor in order to reduce the
-overheads of monitoring. By using the DIFT engine to filter out events that do
-not correspond to valid metadata, we can greatly reduce the amount of
-monitoring that must be handled by the monitor. For example, for an array
-bounds check, we are able to filter out instruction that do not correspond to
-pointer data. Since the monitor is implemented in software, the system allows a
-large range of monitors to be implemented and even updated or changed at
-run-time.
+In this paper, we present an architecture that combines a hardware-based
+dataflow tracking engine with a software-based monitor in order to reduce the
+overheads of monitoring. By using the dataflow engine to track flows of
+initialized metadata, the system is able to filter out events that do not
+correspond to valid metadata, greatly reducing the amount of monitoring that
+must be handled in software. For example, for an array
+bounds check, the dataflow engine is able to filter out instruction that do not
+correspond to pointer data. Thus, we are able to have the flexibility of
+software-based monitors with greatly reduced overheads.
 
@@
 % Use cases
-Trading off coverage for
+Thus, we present an architecture that enables a trade-off between overheads and
+the monitoring coverage achieved. This is achieved by dynamically dropping
+monitoring operations when overheads are too high. We are able to prevent false
+positives by extending the dataflow engine to also track these dropped
+monitoring flows. We also present three policies on when dropping decisions are
+made that cover a trade-off space between ability to match a target overhead
+and the monitoring coverage achieved.
+
+Although full monitoring coverage is always desired, trading off coverage for
 reduced overheads can still be very useful. For example, this adjustable
@@
 which expect low coverage per run but use a large number of runs and users to
-collect debugging information~\cite{liblit-pldi05,
-chilimbi-asplos04, greathouse-cgo11}.
+collect debugging information~\cite{liblit-pldi05, chilimbi-asplos04,
+greathouse-cgo11}.
 
-Thus, we develop an architecture that enables a
-trade-off between the overheads and the coverage of a monitoring technique by
-dynamically adjusting the number of monitoring operations. By extending the
-hardware DIFT engine to track 2-bits of information, we are able to use it for
-filtering empty metadata as well as tracking dropped monitoring flows in order
-to prevent false positives. We introduce three policies of where dropping
-decisions are made and show the trade-off space achieved between accuracy of
-overhead matching and coverage achieved.
+The main contributions of this paper are as follows.
+\begin{itemize}
+  \item Using a dataflow tracking engine to filter out monitoring events in order to reduce overheads.
+  \item Extending the dataflow tracking engine to enable monitoring with adjustable overheads
+  \item Exploration of the design space and trade-offs in dropping policies for performing partial monitoring
+\end{itemize}
 

### 2014-05-23 monitoring.tex
Edits on Sections 1-3.

@@
 \emph{main core}.  On certain events during the main program, such as the
-execution of certain types of instructions, the \emph{monitor} performs a
-series of \emph{monitoring operations}. The monitor operates in parallel to the
+execution of certain types of instructions, the \emph{monitoring core} performs a
+series of \emph{monitoring operations}. The monitoring core operates in parallel to the
 main core. These events are referred to as \emph{monitoring events}. Depending
@@
 executed. Monitoring events are buffered in a FIFO structure to decouple the
-running of the main core and the monitor. If the FIFO is full, then the main
+running of the main core and the monitoring core. If the FIFO is full, then the main
 core is forced to stall on a monitoring event until a FIFO entry becomes
-available. For programmable monitors, these stalls are a major source of
-overhead because the monitor may take several cycles to process a single event
+available. These stalls are a major source of
+overhead because the monitoring core may take several cycles to process a single event
 from the main core. We refer to these stalls and other overheads, such as
-contention for shared resources, as \emph{monitoring overheads}. If the monitor
+contention for shared resources, as \emph{monitoring overheads}. If the monitoring core
 detects an inconsistent or undesired behavior in the monitoring events, then
@@
 to an array pointer when it is allocated. On loads or stores with the array
-pointer, the monitor checks that the memory address is within the base and
+pointer, the monitoring core checks that the memory address is within the base and
 bound. In addition, on certain ALU instructions, this base and bound metadata
@@
 
-There are multiple options for implementing programmable parallel monitors. For
-example, the Log-Based Architecture \cite{lba-isca08} uses processor cores in a
-multi-core system as monitors. The FlexCore architecture
-\cite{flexcore-micro10} instead uses an FPGA-fabric to implement the
-monitor. The approach we describe in this paper applies to any of these
-parallel monitors. However, for experiments, we model an FPGA-based monitor
-similar to FlexCore.  The on-chip FPGA fabric is used to implement the
-``Monitor'' block in Figure~\ref{fig:arch.overview} while the FIFO from the
-main core and metadata cache are implemented as ASICs.
+% There are multiple options for implementing programmable parallel monitors. For
+% example, the Log-Based Architecture \cite{lba-isca08} uses processor cores in a
+% multi-core system as monitors. The FlexCore architecture
+% \cite{flexcore-micro10} instead uses an FPGA-fabric to implement the
+% monitor. The approach we describe in this paper applies to any of these
+% parallel monitors. However, for experiments, we model an FPGA-based monitor
+% similar to FlexCore.  The on-chip FPGA fabric is used to implement the
+% ``Monitor'' block in Figure~\ref{fig:arch.overview} while the FIFO from the
+% main core and metadata cache are implemented as ASICs.

### 2014-05-23 title.tex
Edits on Sections 1-3.

@@
 when monitoring is not performed. Dedicated hardware that performs dynamic
-information flow tracking (DIFT) is used to propagate this invalidation
+dataflow tracking is used to propagate this invalidation
 information in order to prevent false positives. In addition, we can use this
-DIFT hardware to also track the presence of valid metadata. This allows us to
-filter out events in order to reduce monitoring overheads.
+dataflow hardware to also track the presence of initialized metadata. This
+allows us to filter out events on uninitialized metadata in order to reduce
+monitoring overheads.
 % Our experiments show that the system is able to
@@
 % for reasonable overhead targets.
-In our experiments, we see that the system is able to come
-within 2\% of the specified execution time overhead in almost all cases.  In addition, with a
+In our experiments, we see that the system is able to come within 2\% of the
+specified execution time overhead in almost all cases.  In addition, with a
 performance overhead target of 10\%, 49-90\% of monitoring coverage is achieved

### 2014-05-23 drop.tex
Updates to Sections 3 & 4.

@@
 section, we discuss how we can selectively perform monitoring in order to
-match a specified overhead target.
+further reduce overheads and match a specified overhead budget.
 
@@
 have been dropped. This can be done by extending the dataflow engine to keep
-track of an additional ``drop'' flag (Section~\ref{sec:drop.false_positives}).
+track of an additional \emph{invalid flag} (Section~\ref{sec:drop.false_positives}).
 
@@
 \label{sec:drop.slack}
-
-In order to attempt to match an overhead budget, we need a way to estimate the
-actual overheads incurred due to monitoring. More precisely, what we care about
-is the difference between the actual overheads and the overhead budget.
-Slack is essentially a measure of this overhead error. We can measure slack by
-doing the following.
 
@@
 for measuring execution time overheads. However, we note that other methods
-could be used in order to target budgets on IPC, power, energy, etc.  For
-example, one way to measure execution time overheads would be to insert
-checkpoints in the program. By recording the execution time of these
-checkpoints without monitoring, the overheads incurred while running with
-monitoring could be determined. Although this would give an accurate
-measurement of overheads, it would require modifying the program binary. In
-addition, the overhead budget for monitoring would only be updated at
-checkpoints.
+could be used in order to target budgets on IPC, power, energy, etc.  
+% For example, one way to measure execution time overheads would be to insert
+% checkpoints in the program. By recording the execution time of these
+% checkpoints without monitoring, the overheads incurred while running with
+% monitoring could be determined. Although this would give an accurate
+% measurement of overheads, it would require modifying the program binary. In
+% addition, the overhead budget for monitoring would only be updated at
+% checkpoints.
 
-Instead, we develop a model to estimate the execution time overhead budget in a
-gradual manner and without the need to modify the main program. Specifically,
-our model of execution time budget is that a percentage of the main program's
-executed cycles is the number of cycles of overhead that we allow for
-monitoring.  We define \emph{slack} as the number of cycles of monitoring
-overhead that can be incurred while staying within the budget target.  For
+% Instead, we develop a model to estimate the execution time overhead budget in a
+% gradual manner and without the need to modify the main program. Specifically,
+We specify the overhead budget as a percentage of the main program's execution
+cycles without monitoring. We define \emph{slack} as the number of cycles of
+monitoring overhead that can be incurred while staying within the budget
+target. Slack is essentially the difference between the actual overheads seen
+and the budget specified. Slack is generated as the main program runs.  For
 example, if no monitoring overheads occur during 1000 cycles of the main
@@
 Figure~\ref{fig:drop.slack} shows an example of how slack can change over time.
-As monitoring overheads occur, the slack is decremented. If the slack falls
-below the overhead expected for handling a monitoring event, then the event is
-dropped. In addition to this accumulated slack, a small amount of initial slack
+As monitoring overheads occur, the slack is decremented. 
+% If the slack falls below the overhead expected for handling a monitoring event,
+% then the event is dropped. 
+If the slack falls below zero (i.e., the overhead budget is exceeded), then
+monitoring events are dropped.
+In addition to this accumulated slack, a small amount of initial slack
 can be given in order for monitoring to be performed at the start of a program.
@@
 
-In addition to keeping track of overheads, a policy must be developed to
-determine when monitoring event should be dropped. Complex policies involving
-multiple conditions and safety factors could be used to finely tune
-the trade off between monitoring and performance overhead. Similarly, introducing
-randomness into the policy could allow for better sampling in a cooperative
-debugging application of monitoring. Here, we present one simple
-dropping policy that is based on the accumulated slack.
-When the monitor is initialized, the monitor will load
-its needed slack in order to perform a monitoring operation into a register. When
-there is a monitoring event in the FIFO to be processed, a hardware comparator
-checks whether the accumulated slack is greater than or equal to the necessary
-slack for full monitoring. If enough slack exists, the monitor is signaled to
-perform the monitoring operation. Otherwise, the monitoring event is dropped.
+% In addition to keeping track of overheads, a policy must be developed to
+% determine when monitoring event should be dropped. Complex policies involving
+% multiple conditions and safety factors could be used to finely tune
+% the trade off between monitoring and performance overhead. Similarly, introducing
+% randomness into the policy could allow for better sampling in a cooperative
+% debugging application of monitoring. Here, we present one simple
+% dropping policy that is based on the accumulated slack.
+% When the monitor is initialized, the monitor will load
+% its needed slack in order to perform a monitoring operation into a register. When
+% there is a monitoring event in the FIFO to be processed, a hardware comparator
+% checks whether the accumulated slack is greater than or equal to the necessary
+% slack for full monitoring. If enough slack exists, the monitor is signaled to
+% perform the monitoring operation. Otherwise, the monitoring event is dropped.
 
@@
 occurs in the main program's execution is not detected, or false positives,
-where the monitor incorrectly believes an error has occurred. For example, a
-false positive can occur for UMC if a store monitoring event is dropped. This
-causes the store memory location to not be marked as initialized. A subsequent
-load for the memory location will incorrectly cause an error to be raised.  We
+where the monitor incorrectly believes an error has occurred. 
+% For example, a
+% false positive can occur for UMC if a store monitoring event is dropped. This
+% causes the store memory location to not be marked as initialized. A subsequent
+% load for the memory location will incorrectly cause an error to be raised.  
+For example, a false positive can occur for array bounds check if an event is
+dropped that handles copying a pointer. A subsequent array access using the new
+pointer will incorrectly cause an error to be raised since the new pointer's
+base and bounds will not be correctly set.
+We
 accept false negatives as the loss in coverage that we trade off in order to
@@
 
-Dropping monitoring events can create false negatives and false positives.
-False negatives are tolerated. We handle false positives by invalidating
-metadata.
+In analyzing various monitoring schemes, we found that monitoring operations
+are primarily of two types: \emph{checks} and \emph{metadata updates}. Monitors
+\emph{check} certain properties to ensure correct main program execution and
+they \emph{update} metadata for bookkeeping. Skipping a check operation can
+only cause false negatives and will never cause a false positive. Therefore, we
+may simply skip a check operation. Skipping an update operation can cause false
+negatives but may also cause false positives. Essentially, when an update
+operation is skipped, we can no longer trust the corresponding metadata.  Thus,
+by associating an invalid flag with each metadata we can mark metadata that is
+not correctly updated and use this information to prevent false positives.
 
@@
 
+We extend the dataflow engine to keep track of invalidation flags in addition
+to the clean flags discussed in Section~\ref{sec:filter}.
 Since metadata derived from dropped metadata is also invalid. We propagate the
-drop flag in a similar manner to the initialization flags using the dataflow
-engine. Since certain metadata is invalid, we can filter out useless monitoring
-done on this invalid metadata. In the typical case, the propagation rule is
-that if any of the source operands is marked as dropped, then the destination
-operand is marked as dropped. This is slightly different that the
-initialization flags which propagate an uninitialized state only if both
-sources are uninitialized. Table~\ref{tab:drop.propagation_rules} shows the
+drop flag in a similar manner to the clean flags.  In the typical case, the
+propagation rule is
+that if any of the source operands is marked as invalid, then the destination
+operand is marked as dropped. This is slightly different than the
+policy for clean flags which propagates a clean flag only if both
+sources are marked as clean. Table~\ref{tab:drop.propagation_rules} shows the
 propagation rules that we use for array bounds check.  However, we note that
@@
 reconfiguring the Filter Decision Table.
+Similarly to filtering for clean flags, we can also filter out
+useless monitoring for invalid metadata.
+Thus, we see that by tracking two bits of information, we are able to use the
+dataflow engine to both filter out monitoring operations for clean metadata as
+well as selectively perform monitoring without false positives.
 
@@
     \vspace{-0.3in}
-    \caption{Example dataflow graph for metadata.}
+    \caption{Example dependence graph for metadata. Blue (dark) nodes represent
+    events where checks are performed.}
     \label{fig:drop.dataflow_graph}
@@
 slack is zero) to start dropping monitoring events is that this can result in
-wasted work. For example, consider the dataflow graph shown in
-Figure~\ref{fig:drop.dataflow_graph}. If event {\tt E} is dropped, then this
-means that monitoring for events {\tt C} and {\tt D} was wasted.
+wasted work. For example, consider the metadata dependence graph shown in
+Figure~\ref{fig:drop.dataflow_graph}. Here, an edge from node {\tt A} to node
+{\tt B} represents that if event {\tt A} is dropped, then due to its
+invalidated metadata, it will cause event {\tt B} to be also dropped. In the
+example, we see that if event {\tt E} is dropped, then this meant that
+monitoring for events {\tt C} and {\tt D} was wasted. The results of those
+monitoring operations were not used for any monitoring checks.
 
 One method to eliminate this wasted work is to only make dropping decisions at
-the root of these dataflow graphs. That is, we will decide to either monitor or
+the root of these metadata flow graphs. That is, we will decide to either monitor or
 not monitor an entire metadata flow. We refer to this dropping decision policy
 as \emph{source dropping} and we refer to the previous policy of waiting until
-the overhead budget is met as \emph{last-minute dropping}. Source dropping will
+the overhead budget is reached as \emph{last-minute dropping}. Source dropping will
 result in no wasted work. However, because of the coarser-grained decision, it
 may be more difficult to closely match overheads. Source dropping is expected
-to work well when there are a large number of independent dataflows.
+to work well when there are a large number of independent metadata flows.
 
@@
 decision at event {\tt C} whether to perform monitoring on that branch of the
-dataflow. If we choose to perform monitoring, then we want to perform all
+metadata flow. If we choose to perform monitoring, then we want to perform all
 monitoring on that flow. Instead, if we choose to drop event {\tt C}, then the
@@
 drops the bottom flow with no wasted work and event {\tt A} can be dropped to
-drop the entire shown dataflow with no wasted work. If the program can be
+drop the entire shown metadata flow with no wasted work. If the program can be
 analyzed or profiled to identify these instructions which, if dropped, lead to
@@
 amount of wasted work. We refer to this policy as \emph{min-point dropping}.
-Min-point dropping is a coarser-grained decision that last-minute dropping, but
+Min-point dropping is a coarser-grained decision than last-minute dropping, but
 finer-grained that source dropping. Thus, its ability to match overheads is
 expected to sit between these other two policies. Similarly, there is still a
-chance of wasted work due to strange splitting and merging dataflow graph
+chance of wasted work due to strange splitting and merging metadata flow graph
 shapes but typically we expect much less wasted work than last-minute
@@
 dropping policies create a trade-off space between coverage achieved and
-ability to match an overhead target.
+ability to match an overhead target which we evaluate in detail in
+Section~\ref{sec:evaluation}.
 
@@
 
+Due to the multiplicative effect of randomness, long dependence flows with
+multiple points where randomness is introduced leads to a small chance that the
+entire flow is monitored. This is a significant issue for last-minute dropping
+where all events could be dropped. Min-point dropping reduces this effect by
+reducing the points where dropping decisions are made while source-dropping
+eliminates this issue entirely.
+
 \subsection{Multi-Core Implementation}

### 2014-05-23 introduction.tex
Updates to intro and related work.

@@
 
-In this paper, we present an architecture that combines a hardware-based
-dataflow tracking engine with a software-based monitor in order to reduce the
-overheads of monitoring. By using the dataflow engine to track flows of
-initialized metadata, the system is able to filter out events that do not
-correspond to valid metadata, greatly reducing the amount of monitoring that
-must be handled in software. For example, for an array
-bounds check, the dataflow engine is able to filter out instruction that do not
-correspond to pointer data. Thus, we are able to have the flexibility of
-software-based monitors with greatly reduced overheads.
+Recent studies have shown that filtering out monitoring checks on clean
+metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}. For example, for
+an array bounds check, checking non-pointer accesses is unnecessary. In this
+paper, we present an architecture that combines a hardware-based dataflow
+tracking engine with a software-based monitor in order to filter out both
+propagation and checking of clean metadata. By keeping a separate set of flags
+to track clean metadata, we are able to handle the common case of clean
+metadata without the need to read full metadata. By propagating these flags, we
+are also able to handle propagation of clean metadata efficiently in hardware.
+% By using reduce the
+% overheads of monitoringthe dataflow engine
+% to track flows of
+% initialized metadata, the system is able to filter out events that do not
+% correspond to valid metadata, greatly reducing the amount of monitoring that
+% must be handled in software. 
+Thus, we are able to have the flexibility of software-based monitors with
+greatly reduced overheads.
 

### 2014-05-23 related.tex
Edits on Sections 1-3.
Updates to intro and related work.

@@
 % Monitoring
-Monitoring
+Many monitoring schemes have been developed for various security, reliability,
+debugging, and other capabilities.  As we described in
+Section~\ref{sec:extensions}, there exist monitoring schemes for uninitialized
+memory check \cite{mondrian-asplos02}, array bounds check
+\cite{hardbound-asplos08, clause-ase07}, and dynamic information flow
+tracking \cite{dift-asplos04, raksha-isca07, loki-osdi08}.
+Memtracker \cite{memtracker-hpca07} uses monitoring to check for various
+memory-related bugs and errors.  Control flow verification techniques perform
+monitoring to check that program execution follows expected control flows
+\cite{schuette-comp87, impres-dac06,
+kayaalp-isca12}.  Our monitoring architecture is designed to
+be generally applicable to any of these instruction-grained monitoring schemes.
+In particular, our proposed architecture may be especially useful for some of
+these more complex schemes which also have higher overheads.  Similarly,
+although many of these schemes showed low overheads in custom hardware, our
+architecture can be used to reduce the overheads when applying these schemes on
+a multi-core platform.
 
@@
 Hardbound filters out non-pointer data \cite{hardbound-asplos08}. FlexiTaint filters out some monitoring
-and performs it in hardware. FADE provides a more general filtering engine. It
+and performs it in hardware. FADE \cite{fade-hpca14} provides a more general filtering engine. It
 filters out redundant updates and clean checks by reading metadata values. Our
@@
 
-% Adjustable overheads
-Adjustable overhead/cooperative debugging: Testudo \cite{testudo-micro08}, QVM
-\cite{qvm-oopsla08}, Greathouse et al. \cite{greathouse-cgo11}, Huang et al.
-\cite{huang-sttt12}, Lo et al. \cite{lo-rtas14}
+% Sampling
+Statistical sampling is a method that has been applied to various debugging
+techniques in order to reduce their performance overheads. For example, Liblit
+et al. sampled program runs of thousands of users in order to find bugs
+\cite{liblit-pldi05}. This idea has also been applied to detecting memory
+leaks~\cite{chilimbi-asplos04}, dataflow analysis~\cite{greathouse-cgo11}, and
+other run-time monitoring techniques. The Testudo project
+\cite{testudo-micro08} specifically targets hardware-based run-time monitors.
+These sampling-based techniques all limit the debugging capabilities in order
+to provide lower performance overheads, similar to our architecture's goals.
+However, these approaches are not able to explicitly set an overhead budget, as
+our system is able to do. In addition, with the exception of Testudo, these
+methods have not considered parallel monitoring. Finally, we note that the
+concept of distributing the monitoring over many users and runs to increase
+coverage can also be enabled using our architecture.
+
+% Limited Monitoring
+In terms of controlling overheads, the Quality Virtual Machine (QVM) is a
+modification of the Java Virtual Machine (JVM) that supports run-time
+monitoring with controllable overheads \cite{qvm-oopsla08}. Similarly, Huang et
+al. created a framework for controlling the overheads of software-based
+monitoring \cite{huang-sttt12}. These projects are targeting a similar problem
+to the one we address in this paper. However, both projects are focused on
+software-based monitoring and thus their mechanisms modify the software
+monitoring framework in order to enable or disable monitoring. Instead, our
+architecture provides a general hardware mechanism to control overheads of
+parallel monitoring architectures. In addition, these works prevented
+false positives by enabling and disabling monitoring using a source-dropping
+policy. Lo et al. \cite{lo-rtas14} developed a hardware architecture to limit
+monitoring in hard real-time systems. Their architecture performs last-minute
+dropping in order to guarantee strict deadlines. Our system allows for a
+trade-off between choosing source-dropping, last-minute dropping, or min-point
+dropping depending on the needs of the designer or user which has not been
+explored by these previous works.

### 2014-05-27 introduction.tex
Changed title. Edited intro.

@@
 
-% Recent studies proposed programmable parallel hardware as a way to achieve both
-% the performance advantages of hardware and the flexibility of software
-% \cite{lba-isca08, flexcore-micro10, harmoni-dsn12}.  While more
-% efficient than software, these programmable hardware monitors can still show
-% significant performance overheads.  For example,
-% Figure~\ref{fig:intro.full_mon} shows the execution time for three monitoring
-% techniques normalized to the execution time with no monitoring for an
-% FPGA-based parallel monitor. The program being monitored is run on an in-order
-% embedded core running at 500 MHz while the FPGA-based monitor runs at 250 MHz.
-% We can see that, depending on the monitoring scheme, overheads can easily be
-% several tens of percent.  A faster or more complex (e.g., superscalar or
-% out-of-order) core is expected to show even more significant overheads.
-
-% % Run-time monitoring overview
-% \begin{figure}
-%   \begin{center}
-%     \vspace{-0.2in}
-%     \includegraphics[width=\columnwidth]{figs/full_mon.pdf}
-%     \vspace{-0.5in}
-%     \caption{Performance overheads of FPGA-based monitors.}
-%     \vspace{-0.2in}
-%     \label{fig:intro.full_mon} 
-%   \end{center}
-% \end{figure}
-
+% Filtering
 Recent studies have shown that filtering out monitoring checks on clean
@@
 are also able to handle propagation of clean metadata efficiently in hardware.
-% By using reduce the
-% overheads of monitoringthe dataflow engine
-% to track flows of
-% initialized metadata, the system is able to filter out events that do not
-% correspond to valid metadata, greatly reducing the amount of monitoring that
-% must be handled in software. 
 Thus, we are able to have the flexibility of software-based monitors with
@@
 
+% Adjustable overhead
 With this filtering, we see overheads of XX\% - XX\% on average for
 uninitialized memory check, array bounds check, and multi-bit dynamic
-information flow tracking. Although full monitoring coverage is always desired,
-if these overheads are considered too high by the designer or user then
-monitoring must be disabled.
-% % Current schemes are all-or-nothing: suffer full performance impact for full
-% % coverage.
-% Traditionally, these overheads have been an all-or-nothing cost. If the
-% designer finds that the overheads exceed their desired budget, then monitoring
-% cannot be enabled.  In this paper, we present an architecture that enables a
-% trade-off between the overheads and the coverage of a monitoring technique by
-% dynamically adjusting the number of monitoring operations.  This system allows
-% monitoring to still be performed even when the overhead budget is not
-% sufficient for full monitoring.
-% Use cases
+information flow tracking. 
+Traditionally, these overheads have been an all-or-nothing cost. If the
+designer finds that the overheads exceed their desired budget, then monitoring
+cannot be enabled.  
 Thus, we present an architecture that enables a trade-off between overheads and
 the monitoring coverage achieved. This is achieved by dynamically dropping
-monitoring operations when overheads are too high. We are able to prevent false
+monitoring operations when overheads are too high. We prevent false
 positives by extending the dataflow engine to also track these dropped
-monitoring flows. We also present three policies on when dropping decisions are
-made that cover a trade-off space between ability to match a target overhead
+monitoring flows. We investigate three different policies on when dropping decisions are
+made that cover a trade-off space between the ability to match a target overhead
 and the monitoring coverage achieved.
@@
 Although full monitoring coverage is always desired, trading off coverage for
-reduced overheads can still be very useful. For example, this adjustable
+reduced overheads can be very useful. For example, this adjustable
 monitoring enables a level of protection even for systems where the full
@@
 
-The main contributions of this paper are as follows.
+Thus, the main contributions of this paper are summarized as follows.
 \begin{itemize}
-  \item Using a dataflow tracking engine to filter out monitoring events in order to reduce overheads.
-  \item Extending the dataflow tracking engine to enable monitoring with adjustable overheads
-  \item Exploration of the design space and trade-offs in dropping policies for performing partial monitoring
+  \item We use a dataflow tracking engine to efficiently filter out monitoring events in order to reduce overheads.
+  \item We enable a trade-off between monitoring coverage and overheads by extending the dataflow tracking engine to keep track of dropped monitoring flows.
+  \item We explore the design space and trade-offs in dropping policies for performing partial monitoring.
 \end{itemize}
@@
 parallel run-time monitoring model that this paper assumes.
-Section~\ref{sec:filter} presents our architecture that use hardware-based DIFT
+Section~\ref{sec:filter} presents our architecture that uses dataflow-guided filtering
 in order to reduce monitoring overheads. In Section~\ref{sec:drop} we present
-how the DIFT engine can be extended to enable adjustable overheads.
+how the dataflow tracking engine can be extended to enable adjustable overheads.
 Section~\ref{sec:extensions} covers example implementations of three monitoring

### 2014-05-27 monitoring.tex
Edits to monitoring architecture.

@@
 main core. These events are referred to as \emph{monitoring events}. Depending
-on the type of a monitoring event, different monitoring operations are
+on the type of monitoring event, different monitoring operations are
 executed. Monitoring events are buffered in a FIFO structure to decouple the
@@
 detection \cite{cord-hpca06}, etc.  For example, an array bounds check (BC)
-\cite{hardbound-asplos08} can be implemented where monitoring is used to detect
+\cite{hardbound-asplos08} can be implemented in order to detect
 when software attempts to read or write to a memory location outside of an
-array's bounds. This can be done by associating base and bound address metadata
-to an array pointer when it is allocated. On loads or stores with the array
-pointer, the monitoring core checks that the memory address is within the base and
-bound. In addition, on certain ALU instructions, this base and bound metadata
-is propagated to track the creation of new array pointers.
+array's bounds. This can be done by associating metadata with array pointers that 
+indicates the array's base and bound addresses. On loads or stores with the
+array
+pointer, the monitoring core checks that the memory address accessed is within the base and
+bound addresses. In addition, this base and bound metadata
+is propagated on certain ALU instructions to track the creation of new array pointers.
 % One example is an uninitialized memory check (UMC) where monitoring is used to

### 2014-05-27 filter.tex
Updates to Sections 3 & 4.
Edits to filtering section.

@@
 
-The instruction-grained monitoring techniques we consider forward all
-instructions of certain types (e.g., load instructions) to the monitoring core.
+For the instruction-grained monitoring techniques we consider, instructions in
+the main program's execution are forwarded to the monitoring core based on
+their instruction type (e.g., load, store, ALU, etc.).
 However, not all of these forwarded instructions have relevant monitoring to be
@@
 correspond to array pointers. These accesses can be identified because their
-corresponding metadata has been set with base and bound information.
+corresponding metadata has been initialized with base and bound address information.
 More generally, we can reduce monitoring overheads by filtering out 
@@
 
-\subsection{Tracking Clean Metadata}
+\subsection{Dataflow Engine for Clean Metadata}
 
@@
 tracking engine for
-keeping track of clean metadata. Initially, all metadata is marked as clean.
+keeping track of clean metadata. The engine keeps a \emph{clean flag}
+corresponding to each piece of register and memory metadata. Initially, all
+metadata is marked as clean.
 When the monitoring core writes a non-null metadata value, it also sends a
-signal to the dataflow engine to mark the metadata as dirty.
+signal to the dataflow engine to mark the metadata as dirty (i.e., initialized).
 
 When a monitoring event arrives at the dataflow engine,
-it reads in up to two clean flags. A table is used to configure
-which clean flags should be read in depending on the instruction type of
-the event and information from the monitoring event. A pair of simple ALUs are
-used to allow calculating memory addresses based on information from the
-monitoring event. Typically, the clean flags read correspond to the
-source operands of the monitored instruction.
+it reads in up to two clean flags. These typically correspond to the source
+operands of the monitored instruction but can be configured to other targets. 
+This is done by configuring a pair of address generation units during the
+initialization of the monitoring core.
+For example, this can be used to keep clean flags at the word-level rather than
+the byte-level by shifting memory addresses.
 
-Given this set of clean flags, a Filter Decision Table decides, based
+Given this pair of clean flags, a filter decision table decides, based
 on instruction type, whether an event can be filtered out. This table is
@@
 monitoring operation for this event.
+
 It is important that we also propagate the clean information to the destination
-of the instruction.  The Filter Decision Table also includes a bit indicating
-whether the destination operand metadata should be marked as uninitialized.
+of the instruction in order to prevent the use of stale (possibly non-clean) metadata.
+The filter decision table also indicates
+whether the destination operand metadata should be marked as clean.

### 2014-05-27 drop.tex
Edits to dropping section.

@@
 Figure~\ref{dummy} shows the overheads of array bounds check with filtering
-based on initialized metadata (see Section~\ref{sec:evaluation} for details on
+based on clean metadata (see Section~\ref{sec:evaluation} for details on
 the experimental methodology). The overheads range from XX\% to XX\%. In some
-contexts, this overhead may still be considered to be too high. In this
+contexts, this overhead may still be considered to be too high, forcing
+monitoring to be disabled. Instead, it would be useful if partial monitoring
+can be done in order to provide partial protection with further reduced
+overheads. In this
 section, we discuss how we can selectively perform monitoring in order to
-further reduce overheads and match a specified overhead budget.
+further reduce overheads.
 
-The high-level idea is to drop monitoring operations in order to remain within
-the specified overhead budget. Given an overhead budget, the system estimates
-the run-time overhead caused by monitoring (Section~\ref{sec:drop.slack}). When
-this overhead exceeds the budget, then the system drops monitoring operations.
-In order to prevent false positives, we must be able to identify flows that
-have been dropped. This can be done by extending the dataflow engine to keep
-track of an additional \emph{invalid flag} (Section~\ref{sec:drop.false_positives}).
+% Overview of dropping architecture
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\columnwidth]{figs/drop_overview.pdf}
+    \vspace{-0.3in}
+    \caption{Block diagram of architecture for monitoring with adjustable overheads.}
+    \label{fig:drop.overview}
+    \vspace{-0.1in}
+  \end{center}
+\end{figure}
+
+The goal is to be able to perform as much monitoring as possible within a specified overhead budget.
+The high-level approach is to keep track of monitoring overheads at run-time
+and drop monitoring operations when the overhead exceed the specified budget.
+A block diagram of the architecture we will present in this section is shown in
+Figure~\ref{fig:drop.overview}.  In order to do limit the overheads of
+monitoring, we need to first be able to track the gap between the specified
+overhead budget and the monitoring overhead seen at run-time. We refer to this
+as slack and discuss how we measure it in Section~\ref{sec:drop.slack}. 
+Dropping monitoring operations can cause false positives which we prevent by
+extending the dataflow engine to track and filter out dropped monitoring flows (see
+Section~\ref{sec:drop.false_positives}).
+Given that we know how much slack we have for monitoring and can drop
+monitoring operations without false positives, we must decide when and which
+monitoring operations to drop.  There exists some interesting trade-offs in the
+policy for this dropping decision which we discuss in
+Section~\ref{sec:drop.policies}.
 
@@
     \label{fig:drop.slack}
-    \vspace{-0.2in}
+    \vspace{-0.1in}
   \end{center}
@@
     \includegraphics[width=\columnwidth]{figs/stm.pdf}
-    \vspace{-0.3in}
+    \vspace{-0.2in}
     \caption{Slack tracking and drop decision hardware.}
@@
 target. Slack is essentially the difference between the actual overheads seen
-and the budget specified. Slack is generated as the main program runs.  For
+and the budget specified. Slack is generated as the main program runs and consumed as monitoring overheads occur.  For
 example, if no monitoring overheads occur during 1000 cycles of the main
 program's execution and the designer sets a 20\% overhead target, then the
-slack that is built up during this period is 200 cycles. This slack can be used
-to perform monitoring without exceeding the overhead budget.
-Figure~\ref{fig:drop.slack} shows an example of how slack can change over time.
-As monitoring overheads occur, the slack is decremented. 
-% If the slack falls below the overhead expected for handling a monitoring event,
-% then the event is dropped. 
-If the slack falls below zero (i.e., the overhead budget is exceeded), then
-monitoring events are dropped.
+slack that is built up during this period is 200 cycles. If the main core is
+then stalled for 50 cycles due to monitoring, then the remaining slack is 150
+cycles. 
 In addition to this accumulated slack, a small amount of initial slack
 can be given in order for monitoring to be performed at the start of a program.
+Figure~\ref{fig:drop.slack} shows an example of how slack can change over time.
+If the slack falls below zero (i.e., the overhead budget is exceeded), then
+monitoring events should be dropped.
 
-Figure~\ref{fig:drop.stm} shows a hardware slack tracking module (STM) for
-keeping track of slack. The STM uses a counter that increments on every $k$-th
+Figure~\ref{fig:drop.stm} shows a hardware slack tracking module for
+keeping track of slack. The slack tracking module uses a counter that increments on every $k$-th
 cycle of the main core. This $k$ can be calculated by taking the reciprocal of
@@
     \end{small}
-    \caption{Flag propagation rules for array bounds check. X indicates that no
+    \caption{Flag propagation rules for array bounds check. A dash indicates that no
     propagation is done while 0 or 1 indicates that the destination flag is set
@@
 
-We extend the dataflow engine to keep track of invalidation flags in addition
-to the clean flags discussed in Section~\ref{sec:filter}.
-Since metadata derived from dropped metadata is also invalid. We propagate the
-drop flag in a similar manner to the clean flags.  In the typical case, the
-propagation rule is
+We can track these invalid flags by extending the dataflow engine presented in
+Section~\ref{sec:filter} to keep track of an additional bit for each flag. That
+is, we use the dataflow engine to track 2-bit flags in order to keep track of
+both clean/dirty information as well as valid/invalid information.
+Similar to filtering for clean flags, we can also filter out
+monitoring corresponding to invalid metadata which would be useless to perform.
+However, it is important that we propagate the
+invalidation flag in these circumstances so that the entire flow is marked as invalid. 
+In the typical case, the propagation rule for this is
 that if any of the source operands is marked as invalid, then the destination
-operand is marked as dropped. This is slightly different than the
+operand is marked as invalid. This is slightly different than the
 policy for clean flags which propagates a clean flag only if both
 sources are marked as clean. Table~\ref{tab:drop.propagation_rules} shows the
-propagation rules that we use for array bounds check.  However, we note that
+propagation rules that are used for array bounds check. However, we note that
 the propagation rules can be customized depending on the monitor by
-reconfiguring the Filter Decision Table.
-Similarly to filtering for clean flags, we can also filter out
-useless monitoring for invalid metadata.
-Thus, we see that by tracking two bits of information, we are able to use the
-dataflow engine to both filter out monitoring operations for clean metadata as
-well as selectively perform monitoring without false positives.
+reconfiguring the filter decision table.
 
 \subsection{Dropping Decision}
+\label{sec:drop.policies}
 
@@
     \vspace{-0.3in}
-    \caption{Example dependence graph for metadata. Blue (dark) nodes represent
+    \caption{Example dependence graph for metadata. Square nodes represent
     events where checks are performed.} 
@@
 
-One of the issues with waiting until the overhead budget is reached (i.e.,
-slack is zero) to start dropping monitoring events is that this can result in
+The simplest policy of when to drop monitoring events is to drop all monitoring events as
+long as slack is less than or equal to zero.  However, this can result in
 wasted work. For example, consider the metadata dependence graph shown in
@@
 {\tt B} represents that if event {\tt A} is dropped, then due to its
-invalidated metadata, it will cause event {\tt B} to be also dropped. In the
-example, we see that if event {\tt E} is dropped, then this meant that
-monitoring for events {\tt C} and {\tt D} was wasted. The results of those
-monitoring operations were not used for any monitoring checks.
+invalidated metadata, it will cause event {\tt B} to be also dropped. Square
+nodes indicate events where monitoring checks are performed. In the
+example, event {\tt E} is meant to perform a check operation but is dropped.
+In this situation, the
+monitoring operations that were done for events {\tt C} and {\tt D} were wasted
+since their results not used for any monitoring checks.
+
+\begin{figure}
+  \begin{center}
+  \subfloat[all-event dropping]{
+    \includegraphics[width=\columnwidth]{figs/all_drop.pdf}
+    \label{fig:drop.all_drop}
+  }
+  \vspace{-0.1in}
+  \\
+  \subfloat[source-event dropping]{
+    \includegraphics[width=\columnwidth]{figs/source_drop.pdf}
+    \label{fig:drop.source_drop}
+  }
+  \vspace{-0.1in}
+  \\
+  \subfloat[min-event dropping]{
+    \includegraphics[width=\columnwidth]{figs/min_drop.pdf}
+    \label{fig:drop.min_drop}
+  }
+  \end{center}
+  \vspace{-0.2in}
+  \caption{Comparison of dropping policies. Blue (dark) nodes indicate where
+  dropping decisions are made.}
+  \label{fig:drop.policies}
+\end{figure}
 
 One method to eliminate this wasted work is to only make dropping decisions at
-the root of these metadata flow graphs. That is, we will decide to either monitor or
+the root of these metadata flows. That is, we will decide to either monitor or
 not monitor an entire metadata flow. We refer to this dropping decision policy
-as \emph{source dropping} and we refer to the previous policy of waiting until
-the overhead budget is reached as \emph{last-minute dropping}. Source dropping will
+as \emph{source-event dropping} and we refer to the previous policy of waiting until
+the overhead budget is reached as \emph{all-event dropping}.
+Figures~\ref{fig:drop.all_drop} and \ref{fig:drop.source_drop} shows a
+comparison of where dropping decisions are made for these two policies. Source
+dropping will
 result in no wasted work. However, because of the coarser-grained decision, it
@@
 Another possible method is to make the dropping decisions somewhere in between
-\emph{source dropping} and \emph{last-minute dropping}. From
+source-event dropping and all-event dropping. Looking at the graph from 
 Figure~\ref{fig:drop.dataflow_graph}, what we would like to do is make a
@@
 no wasted work, then this information can be used at run-time to minimize the
-amount of wasted work. We refer to this policy as \emph{min-point dropping}.
-Min-point dropping is a coarser-grained decision than last-minute dropping, but
+amount of wasted work. We refer to this policy as \emph{min-event dropping}.
+Figure~\ref{fig:drop.min_drop} shows the points were dropping decisions are
+made for this policy.
+Min-event dropping is a coarser-grained decision than all-event dropping, but
 finer-grained that source dropping. Thus, its ability to match overheads is
@@
 chance of wasted work due to strange splitting and merging metadata flow graph
-shapes but typically we expect much less wasted work than last-minute
-dropping. Thus, we expect the coverage achieved by min-point dropping for a
-certain overhead to fall between source and last-minute dropping. These three
+shapes, but typically we expect much less wasted work than all-event 
+dropping. Thus, we expect the coverage achieved by min-event dropping for a
+certain overhead to fall between source and all-event dropping. These three
 dropping policies create a trade-off space between coverage achieved and
@@
 
-\subsection{Multiple-Run Coverage}
-Coverage can be considered for a single or average run instance. It can also be
-measured over multiple users/runs in order to capture total bug coverage of a
-program.
+\textbf{[We probably need more details of how min-event dropping is
+implemented. How is analysis done? How is the information conveyed to
+run-time?]}
 
-The system presented so far operates in a deterministic manner and so each run
-will perform the same monitoring checks. However, we can introduce randomness
-in order to achieve different dropping patterns over runs and achieve higher
-multiple-run coverage.
-
-Due to the multiplicative effect of randomness, long dependence flows with
-multiple points where randomness is introduced leads to a small chance that the
-entire flow is monitored. This is a significant issue for last-minute dropping
-where all events could be dropped. Min-point dropping reduces this effect by
-reducing the points where dropping decisions are made while source-dropping
-eliminates this issue entirely.
-
-\subsection{Multi-Core Implementation}
-
-This can be extended to mulit-core by using multiple dataflow engines and a
-similar coherency protocol as that used for data caches.
+% \subsection{Multiple-Run Coverage}
+% Coverage can be considered for a single or average run instance. It can also be
+% measured over multiple users/runs in order to capture total bug coverage of a
+% program.
+% 
+% The system presented so far operates in a deterministic manner and so each run
+% will perform the same monitoring checks. However, we can introduce randomness
+% in order to achieve different dropping patterns over runs and achieve higher
+% multiple-run coverage.
+% 
+% Due to the multiplicative effect of randomness, long dependence flows with
+% multiple points where randomness is introduced leads to a small chance that the
+% entire flow is monitored. This is a significant issue for last-minute dropping
+% where all events could be dropped. Min-point dropping reduces this effect by
+% reducing the points where dropping decisions are made while source-dropping
+% eliminates this issue entirely.
+% 
+% \subsection{Multi-Core Implementation}
+% 
+% This can be extended to mulit-core by using multiple dataflow engines and a
+% similar coherency protocol as that used for data caches.

### 2014-05-28 extensions.tex
Filled out extensions section.

@@
 
-This section describes how three example monitoring schemes map to the proposed
-architecture.
+% Monitoring operations
+\begin{table*}[tb]
+  \begin{center}
+    \begin{small}
+    \input{tabs/monitor}
+    \end{small}
+    \caption{Monitoring operations for BC, UMC, and DIFT.}
+    \label{tab:extensions.monitor}
+    \vspace{-0.2in}
+  \end{center}
+\end{table*}
+
+% Filtering operations for clean flags
+\begin{table*}[tb]
+  \begin{center}
+    \begin{small}
+    \input{tabs/filter}
+    \end{small}
+    \caption{Filtering conditions and operations for clean flags.}
+    \label{tab:extensions.filter}
+    \vspace{-0.2in}
+  \end{center}
+\end{table*}
+
+% Filtering operations for invalid flags
+\begin{table*}[tb]
+  \begin{center}
+    \begin{small}
+    \input{tabs/drop}
+    \end{small}
+    \caption{Filtering conditions and operations performed for a dropped or filtered event for invalid flags.}
+    \label{tab:extensions.drop}
+    \vspace{-0.2in}
+  \end{center}
+\end{table*}
+
+In this section, we show examples of how our architecture can be applied to
+three specific monitoring schemes: array bounds check (BC), uninitialized
+memory check (UMC), and dynamic information flow tracking (DIFT).
+Table~\ref{tab:extensions.monitor} shows a summary of the monitoring operations for these
+monitoring schemes. Table~\ref{tab:extensions.filter} summarizes when when
+monitoring events can be filtered out due to clean metadata and what metadata
+needs to be marked as clean on a filtered event.
+Table~\ref{tab:extensions.drop} shows the analogous information for the valid
+flags. In this case, the operation listed is also the one performed when a
+monitoring event is dropped due to insufficient slack. Note that for simplicity
+the policies for clean and invalid flags are shown and discussed separately
+here. However, the dataflow engine actually operates on pairs of these flags
+using a combined policy.
 
@@
 
+As described in Section~\ref{sec:monitoring}, array bounds check is a
+monitoring scheme that attempts to detect buffer overflows where memory
+accesses go beyond the boundaries of an array.
+Buffer overflows can cause unreliable program execution or create openings for
+malicious attackers. One way to detect buffer overflows is a base and bounds
+approach \cite{hardbound-asplos08}. In this approach, metadata is
+associated with an array pointer when it is allocated that includes the base
+(start) address and bound (end) address of the corresponding array. On ALU
+instructions, this base and bound metadata is copied to the destination
+registers.  On store instructions, the metadata for the source register is
+copied to the metadata for the store address. Similarly, on load instructions,
+the metadata value for the load address is copied to the destination register's
+metadata. In addition, on load and store instructions, the memory address is
+checked to be within the base and bounds metadata. If it is not, then an
+out-of-bounds error is detected. These operations are summarized in
+Table~\ref{tab:extensions.monitor}.
+
+In terms of filtering using clean flags (Table~\ref{tab:extensions.filter}),
+array allocations are never filtered since these will always involve non-clean
+metadata (i.e., setting new base and bounds). For ALU operations, if both
+source operands are clean, then the event can be filtered by setting the
+destination register as clean. For load and store, only a single source needs
+to be checked to be clean in order for the event to be filtered. Similarly, the
+destination register or memory location needs to be marked as clean if the
+event is filtered.
+
+The invalid flags for BC operate similarly to the clean flags
+(Table~\ref{tab:extensions.drop}). When ALU, load, or store events are dropped or filtered,
+the destination register or memory location is marked as invalid. In addition,
+we need to handle the possibility of dropping an array allocation. In this
+case, the array pointer is marked as invalid. In terms of filtering out invalid
+events, for an ALU event, if either of the source registers is invalid then the
+event can be filtered. For load and store instructions, if the single source
+operand is invalid, then the event can be filtered.
+
 \subsection{Uninitialized Memory Check}
+
+Uninitialized memory check is a monitoring scheme that checks that memory is
+written to before it is read from. This is done by forwarding all load and
+store instructions to the monitoring core. For every byte of memory, the
+monitor keeps one bit of metadata. On a store to a memory location, the monitor
+marks the corresponding metadata bit to indicate that the memory location has
+been initialized. On a load, the monitor checks that the corresponding metadata
+bit has been previously marked as initialized.
+
+Rather than using the clean flags to mirror the initialization metadata that
+UMC stores, we use the clean flags as a coarse-grained version of the UMC
+metadata in order to more efficiently filter out monitoring operations. We use
+a clean flag per word (four bytes) of memory. A clean flag is set to dirty only
+if all four of the corresponding memory bytes have been marked as initialized.
+On a load, if the clean flag is marked as dirty, then we can filter the event
+since we know that all possible corresponding location memory locations have
+been marked as initialized. If the flag is marked as clean, then that indicates
+that at least one of the corresponding memory bytes has not been initialized
+and the monitoring core must perform a further check. Store instructions are
+never filtered out.
+
+On a dropped store instruction, the destination memory location is marked as
+invalid. On load instructions, we can filter them out if the source memory
+location has been invalidated. In this case, we do not have to perform any
+further operation since on load instructions, UMC only performs a check
+operation.
 
@@
 
+Dynamic information flow tracking (DIFT) \cite{dift-asplos04} is a security
+monitoring scheme that seeks to detect when information from untrusted sources
+is used to affect the control flow.  In its simplest form, DIFT keeps a 1-bit
+metadata tag for each memory location and each register, indicating tainted or
+untainted. When data is read from an untrusted source, it is marked as tainted.
+As this data is used for other operations, the results of these operations are
+also marked as tainted. On an indirect jump, the taint bit of the register used
+is checked. If the register is found to be tainted, then an error is detected.
+We focus on a more heavy-weight approach that also allows some information of
+where flows originate from by using a multi-bit (32-bits in our implementation)
+metadata tag. This can be used to assist debugging in the case of an error. In
+this case, when data is marked from an untrusted source, the software specifies
+an identifier to set the metadata tag. This multi-bit tag is propagated on
+load, store, and ALU operations. If two different tags are used as sources to
+an ALU operation, then a new identifier is calculated based on the sources.
+These monitoring operations are summarized in
+Table~\ref{tab:extensions.monitor}.
+
+Table~\ref{tab:extensions.filter} shows how the clean flags are used for DIFT.
+The rules for ALU, load, and store instructions are same as for array bounds
+check. Similar to array allocations, for DIFT, software setting of metadata
+tags is never filtered out due to clean metadata. On indirect jump
+instructions, if the source register is clean, then the event can be filtered
+out as no error has occurred.
+
+Again, the invalid flags for DIFT operate similar to their operation for BC
+(Table~\ref{tab:extensions.drop}. In addition to operations for invalidating
+destination operands for ALU, load, and store instructions, the destination
+memory location is also invalidated when set tag operations dropped. Indirect
+jump events are filtered out if the source register is marked as invalid.
+However, no further metadata needs to be marked as invalid in this case.
+
 \subsection{Applicability to Other Schemes}
+
+In this section, we have discussed how our dataflow engine can be applied to
+three example monitoring schemes to enable reduced and adjustable overheads.
+However, our approach can be generally applied to many monitoring schemes. This
+is possible because the use of clean and invalidation flags is largely agnostic
+to the semantics of the specific monitoring scheme. That is, the dataflow
+engine must be configured based on the dataflow of the monitoring scheme, but
+it does not need detailed information of what the monitoring scheme does.
+
+We do note some limitations to the proposed approach. Previous work
+\cite{fade-hpca14} that has looked at filtering in order to reduce monitoring
+overheads has also filtered out redundant updates where the source metadata
+matches the destination metadata. Since we do not read the actual metadata, our
+architecture is not able to identify these redundant updates. However, by using
+only 2-bit flags, the bandwidth need of our dataflow engine is reduced.  
+In addition, we note that filtering use clean flags works well when a large
+portion of instructions are not relevant to the monitoring scheme (e.g., array
+bounds check only cares about array pointers). However, if most instruction
+have dirty metadata, then very little filtering will be possible.
+
+In terms of using invalidation to perform partial monitoring, this 
+approach works best for monitoring schemes where there are many pieces of
+independent metadata. Since marking a piece of metadata as invalid will cause
+future metadata with dependences to also be marked as invalid, schemes with
+large metadata dependency chains can quickly see large amounts of metadata
+become invalid.
+
+Another limitation is that the dataflow engine, as described, only marks one
+flag location on each monitoring event. Similarly, the dataflow engine can only
+read in two sets of flags in order to determine whether an event can be
+filtered or not. This was done because we found that for many monitoring
+schemes, the monitoring operations corresponded to program instructions, which
+read up to two operands and write to one destination.  For monitoring schemes
+which update multiple pieces of metadata on a monitoring event, it may be
+adequate to associate a clean and invalid flag with a single piece.  For
+example, a scheme may use a data structure that includes several pieces of
+related metadata. On a drop, one of these pieces could be marked as invalid in
+order to indicate that the entire structure is invalid.  Similarly, marking one
+specified piece as clean may indicate that the entire data structure's metadata
+is clean.  Alternatively, the read and write capabilities of the dataflow
+engine could be expanded to support these more complex monitoring schemes.

### 2014-05-28 related.tex
Filled out extensions section.

@@
 % Filtering
-Hardbound filters out non-pointer data \cite{hardbound-asplos08}. FlexiTaint filters out some monitoring
-and performs it in hardware. FADE \cite{fade-hpca14} provides a more general filtering engine. It
+FADE \cite{fade-hpca14} provides a more general filtering engine. It
 filters out redundant updates and clean checks by reading metadata values. Our
@@
 policy. Lo et al. \cite{lo-rtas14} developed a hardware architecture to limit
-monitoring in hard real-time systems. Their architecture performs last-minute
+monitoring in hard real-time systems. Their architecture performs all-event 
 dropping in order to guarantee strict deadlines. Our system allows for a
-trade-off between choosing source-dropping, last-minute dropping, or min-point
+trade-off between choosing source-event dropping, all-event dropping, or min-event
 dropping depending on the needs of the designer or user which has not been

### 2014-05-28 title.tex
Updates to intro and related work.
Changed title. Edited intro.
Update abstract.

--- 2014-05-23 title.tex
@@
 \title{
-Using Dynamic Dataflow Tracking to Reduce Overheads for \\ Parallel
-Run-Time Monitoring
+Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring
 }
@@
 Recent studies have proposed various parallel run-time monitoring techniques to
-improve the reliability, security, and debugging capabilities of
-computer systems. These run-time monitors introduce overheads in performance
+improve the reliability, security, and debugging capabilities of computer
+systems. These run-time monitors can introduce large overheads in performance
 and energy to the system.  Traditionally, these overheads have been an
-all-or-nothing cost; monitoring cannot be used if the overheads are considered too large.
-% where if monitoring is enabled, then the program will incur, in full, the
-% associated overheads. 
-%That is, if the overheads are considered too large, then monitoring cannot be
-%enabled. 
-In this paper, we present a run-time monitoring architecture that
-enables a trade-off between the coverage and the overheads.
-The system allows an overhead or coverage target to be specified,
-and limits the amount of monitoring performed at run-time to meet the target. 
-However, simply skipping monitoring operations can lead to false
-positives, signaling errors when none have occurred. Our proposed architecture
-prevents false positives using a method that invalidates certain metadata
-when monitoring is not performed. Dedicated hardware that performs dynamic
-information flow tracking (DIFT) is used to propagate this invalidation
-information in order to prevent false positives. In addition, we can use this
-DIFT hardware to also track the presence of valid metadata. This allows us to
-filter out events in order to reduce monitoring overheads.
+all-or-nothing cost; monitoring cannot be used if the overheads are considered
+too large.  In this paper, we show how a hardware-based dataflow tracking
+engine can be used to filter out unnecessary monitoring and to enable partial
+monitoring.  We are able to filter out unnecessary monitoring related to clean
+metadata in order to greatly reduce the overheads of monitoring. In addition,
+we present a hardware architecture can further reduce overheads by enabling a
+trade-off between the amount of monitoring performed and the overheads
+incurred. In order to prevent false positives due to this partial monitoring,
+we extend the dataflow engine to track dropped monitoring flows. We investigate
+some of the trade-offs involved with partial monitoring in terms of how closely
+an overhead target can be matched versus the coverage that can be achieved.
 % Our experiments show that the system is able to
@@
 % for reasonable overhead targets.
-In our experiments, we see that the system is able to come
-within 2\% of the specified execution time overhead in almost all cases.  In addition, with a
-performance overhead target of 10\%, 49-90\% of monitoring coverage is achieved
-on average depending on the monitoring technique.
+In our experiments, we see that filtering is able to decrease monitoring
+overheads from XX\%-XX\% on average, depending on the monitoring scheme, to
+XX\%-XX\%. Additionally, we are able to still achieve XX\% coverage with a 10\%
+overhead target.
 

### 2014-05-28 tabs/drop.tex
Tables for Extensions section.

% Operation of hardware modules

\begin{tabular}{|l|l|l|l|}
\hline

{\bf Monitor} & {\bf Event} & {\bf Filter Condition} & {\bf Dropping/Filter Operation} \\ \hline \hline

\multirow{4}{*}{BC}  
& Array allocation & Never & Invalidate array pointer \\ \cline{2-4}
& ALU   & At least one invalid source register & Invalidate destination register \\ \cline{2-4}
& Load  & Invalid source memory location & Invalidate destination register \\ \cline{2-4}
& Store & Invalid source register & Invalidate destination memory location \\
\hline\hline

\multirow{2}{*}{UMC}  
& Store & Never & Invalidate destination memory location \\ \cline{2-4}
& Load & Invalid source memory location & Do nothing \\ 
\hline\hline

\multirow{5}{*}{DIFT}  
& Software set tag & Never & Invalidate destination memory location \\ \cline{2-4}
& ALU   & At least one invalid source register & Invalidate destination register \\ \cline{2-4}
& Load  & Invalid source memory location & Invalidate destination register \\ \cline{2-4}
& Store & Invalid source register & Invalidate destination memory location \\ \cline{2-4}
& Indirect jump & Invalid source metadata & Do nothing \\ 
\hline

\end{tabular}

### 2014-05-28 tabs/filter.tex
Tables for Extensions section.

% Operations for handling clean flags

\begin{tabular}{|l|l|l|l|}
\hline

{\bf Monitor} & {\bf Event} & {\bf Filter Condition} & {\bf Filter Operation} \\ \hline \hline

\multirow{4}{*}{BC}  
& Array allocation & Never & - \\ \cline{2-4}
& ALU   & Both source registers are clean & Mark destination register as clean \\ \cline{2-4}
& Load  & Source memory location is clean & Mark destination register as clean \\ \cline{2-4}
& Store & Source register is clean & Mark destination memory location as clean \\
\hline\hline

\multirow{2}{*}{UMC}  
& Store & Never & - \\ \cline{2-4}
& Load & Source memory location is dirty & Do nothing \\ 
\hline\hline

\multirow{5}{*}{DIFT}  
& Software set tag & Never & - \\ \cline{2-4}
& ALU   & Both source registers are clean & Mark destination register as clean \\ \cline{2-4}
& Load  & Source memory location is clean & Mark destination register as clean \\ \cline{2-4}
& Store & Source register is clean & Mark destination memory location as clean \\ \cline{2-4}
& Indirect jump & Source register is clean & Do nothing \\ 
\hline

\end{tabular}

### 2014-05-28 tabs/monitor.tex
Tables for Extensions section.

% Summary of monitoring operations

\begin{tabular}{|l|l|l|}
\hline

{\bf Monitor} & {\bf Event} & {\bf Monitoring Operation} \\ \hline \hline

\multirow{4}{*}{BC}  
 & Array allocation & Set base and bound metadata \\ \cline{2-3}
 & ALU & Copy metadata from register to register \\  \cline{2-3}
 & Load & Copy metadata from memory to register. Check memory access address. \\ \cline{2-3}
 & Store & Copy metadata from register to memory. Check memory access address. \\ 
\hline\hline

\multirow{2}{*}{UMC}  
 & Store & Set metadata tag (data initialized) \\ \cline{2-3}
 & Load & Check if metadata tag is set \\ 
\hline\hline
% & Software clear tag & Clear metadata tag & Non-critical \\ \hline\hline

\multirow{5}{*}{DIFT}  
 & Software set tag & Set tag to a specified value based on software command \\ \cline{2-3}
 & ALU & Output tag is the OR of the input tags \\ \cline{2-3}
 & Load & Set register tag as tag of loaded data \\ \cline{2-3}
 & Store & Set memory tag as register tag \\ \cline{2-3}
% & Software clear tag & Clear tag & Update \\ \cline{2-4}
 & Indirect jump & Check tag \\ 
 \hline

% \multirow{5}{*}{BC}  
%  & MOV/LD/ST          & Propagate colors (Tag[dest] = Tag[src]) & Critical \\ \cline{2-4}
%  & ADD/SUB/NOT        & Update colors (Tag[dest] = f(Tag[src])) & Critical \\ \cline{2-4}
%  & OR/XOR             & Clear colors & Critical \\ \cline{2-4}
%  & LD/ST              & Compare pointer and memory colors & Check \\ \cline{2-4}
%  & Software set color & Set pointer colors & Critical \\ \hline\hline

% Soft Error Check & Instruction & Check instruction computation & Check \\ \hline\hline
% 
% \multirow{2}{*}{Bank of Observers}  
%  & System output & Update state estimation & Critical \\ \cline{2-4}
%  & New state estimate & Check that states of all observers matches & Check \\ \hline

\end{tabular}

### 2014-05-28 drop.tex
Filled out extensions section.
Various edits.

@@
 A block diagram of the architecture we will present in this section is shown in
-Figure~\ref{fig:drop.overview}.  In order to do limit the overheads of
+Figure~\ref{fig:drop.overview}.  In order to limit the overheads of
 monitoring, we need to first be able to track the gap between the specified
@@
 
-% Flag propagation rules for BC
-\begin{table}[tb]
-  \begin{center}
-    \begin{small}
-    \input{tabs/propagation_rules}
-    \end{small}
-    \caption{Flag propagation rules for array bounds check. A dash indicates that no
-    propagation is done while 0 or 1 indicates that the destination flag is set
-    to that value.}
-    \label{tab:drop.propagation_rules}
-    \vspace{-0.2in}
-  \end{center}
-\end{table}
+% % Flag propagation rules for BC
+% \begin{table}[tb]
+%   \begin{center}
+%     \begin{small}
+%     \input{tabs/propagation_rules}
+%     \end{small}
+%     \caption{Flag propagation rules for array bounds check. A dash indicates that no
+%     propagation is done while a 1 indicates that the destination flag is set.}
+%     \label{tab:drop.propagation_rules}
+%     \vspace{-0.2in}
+%   \end{center}
+% \end{table}
 
@@
 Similar to filtering for clean flags, we can also filter out
-monitoring corresponding to invalid metadata which would be useless to perform.
+monitoring operations corresponding to invalid metadata which would be useless to perform.
 However, it is important that we propagate the
 invalidation flag in these circumstances so that the entire flow is marked as invalid. 
-In the typical case, the propagation rule for this is
+In the typical case, the propagation rule used is
 that if any of the source operands is marked as invalid, then the destination
@@
 policy for clean flags which propagates a clean flag only if both
-sources are marked as clean. Table~\ref{tab:drop.propagation_rules} shows the
-propagation rules that are used for array bounds check. However, we note that
+sources are marked as clean. 
+% Table~\ref{tab:drop.propagation_rules} shows the
+% propagation rules that are used for array bounds check. 
+We note that
 the propagation rules can be customized depending on the monitor by
@@
 
-The simplest policy of when to drop monitoring events is to drop all monitoring events as
+The simplest policy for when to drop monitoring events is to drop all monitoring events as
 long as slack is less than or equal to zero.  However, this can result in
@@
 nodes indicate events where monitoring checks are performed. In the
-example, event {\tt E} is meant to perform a check operation but is dropped.
+example, suppose that event {\tt E} is meant to perform a check operation but is dropped.
 In this situation, the
 monitoring operations that were done for events {\tt C} and {\tt D} were wasted
-since their results not used for any monitoring checks.
+since their results were not used for any monitoring checks.
 
@@
 not monitor an entire metadata flow. We refer to this dropping decision policy
-as \emph{source-event dropping} and we refer to the previous policy of waiting until
-the overhead budget is reached as \emph{all-event dropping}.
+as \emph{source-event dropping} and we refer to the previous policy of 
+dropping any event as \emph{all-event dropping}.
 Figures~\ref{fig:drop.all_drop} and \ref{fig:drop.source_drop} shows a
@@
 
-\textbf{[We probably need more details of how min-event dropping is
-implemented. How is analysis done? How is the information conveyed to
-run-time?]}
-
 % \subsection{Multiple-Run Coverage}

### 2014-05-28 extensions.tex
Various edits.

@@
 location has been invalidated. In this case, we do not have to perform any
-further operation since on load instructions, UMC only performs a check
-operation.
+further operation since on load instructions since UMC only performs a check
+operation for loads.
 
@@
 is checked. If the register is found to be tainted, then an error is detected.
-We focus on a more heavy-weight approach that also allows some information of
+We focus on a more heavy-weight approach that also tracks some information of
 where flows originate from by using a multi-bit (32-bits in our implementation)
@@
 Again, the invalid flags for DIFT operate similar to their operation for BC
-(Table~\ref{tab:extensions.drop}. In addition to operations for invalidating
+(Table~\ref{tab:extensions.drop}). In addition to operations for invalidating
 destination operands for ALU, load, and store instructions, the destination
@@
 only 2-bit flags, the bandwidth need of our dataflow engine is reduced.  
-In addition, we note that filtering use clean flags works well when a large
+In addition, we note that filtering using clean flags works well when a large
 portion of instructions are not relevant to the monitoring scheme (e.g., array
@@
 which update multiple pieces of metadata on a monitoring event, it may be
-adequate to associate a clean and invalid flag with a single piece.  For
+adequate to associate a clean and invalid flag with a single piece of metadata.  For
 example, a scheme may use a data structure that includes several pieces of
@@
 is clean.  Alternatively, the read and write capabilities of the dataflow
-engine could be expanded to support these more complex monitoring schemes.
+engine could be expanded to support more complex monitoring schemes.

### 2014-05-28 introduction.tex
Various edits.

@@
   \item We use a dataflow tracking engine to efficiently filter out monitoring events in order to reduce overheads.
-  \item We enable a trade-off between monitoring coverage and overheads by extending the dataflow tracking engine to keep track of dropped monitoring flows.
-  \item We explore the design space and trade-offs in dropping policies for performing partial monitoring.
+  \item We present an architecture that enables adjustable monitoring, creating a trade-off between monitoring coverage and overheads.
+  \item We show how the dataflow tracking engine can be used to prevent false positives for adjustable monitoring.
+  \item We explore the design space and trade-offs in dropping policies for performing adjustable monitoring.
 \end{itemize}

### 2014-05-28 related.tex
Various edits.

@@
 % Filtering
-FADE \cite{fade-hpca14} provides a more general filtering engine. It
-filters out redundant updates and clean checks by reading metadata values. Our
-dataflow engine does not require reading the actual metadata values which may
-be large. We do not handle filtering redundant updates, but we do filter out
-clean checks as well as clean (i.e., uninitialized) propagations. We are able
-to do this because the dataflow engine can also propagate the uninitialized
-information. In addition, our dataflow engine can be used to provide adjustable
-overheads without false positives which is not a capability described by FADE.
+FADE \cite{fade-hpca14} presents a general filtering engine in order to reduce
+the overheads of monitoring. It filters out redundant updates and clean checks
+by reading metadata values. We do not handle filtering redundant updates, but
+we do filter out clean checks as well as clean (i.e., uninitialized)
+propagations.  We are able to do this because the dataflow engine can also
+propagate the clean information.  In addition, our dataflow engine only
+requires reading and writing 2-bit flags rather than the entire metadata.  As
+we have shown, our dataflow engine also enables adjustable overheads without
+false positives which is not a capability described by FADE.
 

### 2014-05-28 tabs/monitor.tex
Various edits.

@@
 \multirow{5}{*}{DIFT}  
- & Software set tag & Set tag to a specified value based on software command \\ \cline{2-3}
- & ALU & Output tag is the OR of the input tags \\ \cline{2-3}
+ & Software set tag & Set tag to a specified identifier value based on software command \\ \cline{2-3}
+ & ALU & If input tags are different, output tag is a new identifier. Otherwise, output tag is the input tag. \\ \cline{2-3}
  & Load & Set register tag as tag of loaded data \\ \cline{2-3}

### 2014-05-28 title.tex
revised abstract

@@
 improve the reliability, security, and debugging capabilities of computer
-systems. These run-time monitors can introduce large overheads in performance
-and energy to the system.  Traditionally, these overheads have been an
-all-or-nothing cost; monitoring cannot be used if the overheads are considered
-too large.  In this paper, we show how a hardware-based dataflow tracking
-engine can be used to filter out unnecessary monitoring and to enable partial
-monitoring.  We are able to filter out unnecessary monitoring related to clean
-metadata in order to greatly reduce the overheads of monitoring. In addition,
-we present a hardware architecture can further reduce overheads by enabling a
-trade-off between the amount of monitoring performed and the overheads
-incurred. In order to prevent false positives due to this partial monitoring,
-we extend the dataflow engine to track dropped monitoring flows. We investigate
-some of the trade-offs involved with partial monitoring in terms of how closely
-an overhead target can be matched versus the coverage that can be achieved.
+systems. However, these run-time monitors can introduce large performance and energy
+overhead, especially when performed on programmable cores.
+%Traditionally, these overheads have been an
+%all-or-nothing cost; monitoring cannot be used if the overheads are considered
+%too large.  
+In this paper, we introduce a hardware dataflow tracking engine that can be
+used to filter out unnecessary monitoring and enable partial monitoring.
+The dataflow engine can identify events with clean metadata so that they can
+be filtered out. To further reduce the overhead of monitoring, we propose
+to enable a trade-off between monitoring coverage and overhead by dropping certain
+monitoring operations. For the partial monitoring, the dataflow engine is used
+to track dropped monitoring flows so that false positives can be avoided.
+Given the architecture, we also investigate how the dropping decisions should be
+made for partial monitoring and show that there exists a trade-off between closely
+matching an overhead target and coverage that can be achieved.
 % Our experiments show that the system is able to
@@
 % for reasonable overhead targets.
-In our experiments, we see that filtering is able to decrease monitoring
+The experimental results show that filtering can significantly reduce monitoring
 overheads from XX\%-XX\% on average, depending on the monitoring scheme, to
-XX\%-XX\%. Additionally, we are able to still achieve XX\% coverage with a 10\%
-overhead target.
+XX\%-XX\%. Additionally, the partial monitoring is shown to still achieve XX\% coverage 
+only with a 10\% overhead target.
 

### 2014-05-29 arch.tex
Reorganized paper to unify discussion of dataflow engine.

@@
-\section{Monitoring Architecture Model}
+\section{Dataflow Engine for Filtering and Dropping}
 \label{sec:arch}
 
-% Run-time monitoring overview
+% % Overview of inserting dataflow engine between main and monitoring core
+% \begin{figure}
+%   \begin{center}
+%     \includegraphics[width=\linewidth]{figs/dataflow_overview.pdf}
+%     \vspace{-0.2in}
+%     \caption{Overview of architecture for reduced and adjustable monitoring
+%     overheads.}
+%     \label{fig:arch.overview} 
+%     \vspace{-0.1in}
+%   \end{center}
+% \end{figure}
+
+% Overview of full architecture
 \begin{figure}
   \begin{center}
-    \includegraphics[width=\columnwidth]{figs/monitoring_architecture.pdf}
-    \vspace{-0.2in}
-    \caption{Overview of run-time monitoring architecture.}
+    \includegraphics[width=\columnwidth]{figs/architecture_overview.pdf}
+    \vspace{-0.3in}
+    \caption{Block diagram of architecture for monitoring with reduced and adjustable overheads.}
     \label{fig:arch.overview}
-    \vspace{-0.2in}
+    \vspace{-0.1in}
   \end{center}
@@
 
-Figure~\ref{fig:arch.overview} shows an overview of the run-time monitoring
-model that is assumed in this paper. 
-% The \emph{main task} is a computation task that performs the original function
-% of the real-time system and is run on the \emph{main core}. 
-The \emph{main program} is a computation task that performs the original
-function of the system and is run on the \emph{main core}.
-% The main task has a
-% worst-case execution time (WCET) which can be upper-bounded using a variety
-% of existing techniques \cite{wcetsurvey-tecs08}. This WCET is used for timing
-% analysis in the design of the real-time system. As long as the execution
-% time of the main task does not exceed this bound, then all timing guarantees
-% from the analysis hold.
-On certain events
-during the main program, such as the execution of certain types of
-instructions, the \emph{monitor} performs a series of \emph{monitoring
-operations}. The monitor operates in parallel to the main core. These events
-are referred to as
-\emph{monitoring events}. Depending on the type of a monitoring event, different
-monitoring operations are executed. Monitoring events are buffered
-in a FIFO
-structure to decouple the running of the main core and the monitor. If the FIFO
-is full, then the main core is forced to stall on a monitoring event until a
-FIFO entry becomes available. For programmable monitors, these stalls are a
-major source of overhead because the monitor may take several cycles to process
-a single event from the main core. We refer to these stalls and other
-overheads,
-such as contention for shared resources, as
-\emph{monitoring overheads}. If the monitor detects an %inconsistent or
-undesired behavior in the monitoring events, then an error is detected. 
-% We do
-% not focus here on the question on how to handle such an error. However, there are
-% several options on how to handle an error such as raising an exception,
-% notifying the user, or switching to a simpler, more trusted main task
-% \cite{sha-simplex-sw01}.
+In this section, we will describe our hardware architecture for reduced
+overheads by filtering out null metadata and enabling partial monitoring.
+Figure~\ref{fig:arch.overview} shows a high-level block diagram of the
+architecture. The basic idea is to insert hardware between the main core and
+the monitoring core that will reduce the number of monitoring events that reach
+the monitoring core and thus reduce the overheads of monitoring.
 
-There are many possible monitoring schemes that can be implemented on this
-type of fine-grained parallel monitoring architecture such as
-array bound checks \cite{devietti-hardbound-asplos08}, memory protection \cite{mmp02}, 
-information flow tracking \cite{suh-dift-asplos04, greathouse-testudo-micro08}, 
-soft error detection \cite{argus-micro07}, data-race detection \cite{cord-hpca06}, etc.
-One example is an
-uninitialized memory check (UMC) where 
-monitoring is used to detect when software attempts to read from a memory location that
-was not previously initialized. This can be done by forwarding each load and store
-instruction from the main core to the monitor.
-%the memory address
-%of each store and load instruction on the main core to the monitor as
-%well as whether the instruction was a store or load. 
-For every memory location,
-the monitor keeps one bit of metadata. On a store to a
-memory location, the monitor marks the corresponding metadata bit to
-indicate that the memory location has been initialized. On a load,
-the monitor checks that the corresponding metadata bit has
-been previously marked as initialized.
+As in the baseline architecture from Section~\ref{sec:monitoring}, events from
+the main core are first buffered in a FIFO. 
+Events are first filtered out based on whether they correspond to null or
+invalid metadata. A hardware dataflow engine is used to track the flow of null
+and invalid metadata (Section~\ref{sec:arch.dataflow}) and decide whether
+specific events can be filtered out.
+Next, if the event is not filtered out, then we
+check whether the event should be dropped in order to stay within the specified
+overhead budget. Whenever an event is dropped, its corresponding metadata is
+marked and tracked as invalid by the dataflow engine. The
+dropping decision hardware is described in Section~\ref{sec:arch.dropping}.
+Finally, if the event is not filtered or dropped, then it is forwarded to the
+monitoring core to be processed.
 
-There are multiple options for implementing programmable parallel monitors. For example, the
-Log-Based Architecture \cite{chen08-lba} uses processor cores in a multi-core
-system as monitors. The FlexCore architecture \cite{deng-flexcore-micro10}
-instead uses an FPGA-fabric to implement the monitor. The approach we describe
-in this paper applies to any of these parallel monitors. However, for experiments,
-we model an FPGA-based monitor similar to FlexCore. 
-%Specifically, this
-%architecture uses an on-chip FPGA fabric to
-%allow high-performance and reconfigurable run-time monitoring. 
-The on-chip FPGA fabric is
-used to implement the ``Monitor'' block in Figure~\ref{fig:arch.overview}
-while the FIFO from the main core and metadata cache are implemented as ASICs.
+\subsection{Filtering Using the Dataflow Engine}
+\label{sec:arch.dataflow}
+
+% Detailed architecture of dataflow engine
+\begin{figure*}
+  \begin{center}
+    \includegraphics[width=\linewidth]{figs/dataflow_architecture.pdf}
+    \vspace{-0.2in}
+    \caption{Hardware architecture of the dataflow engine.}
+    \label{fig:arch.dataflow} 
+    \vspace{-0.1in}
+  \end{center}
+\end{figure*}
+
+% Overview
+In order to perform filtering, we need to be able to identify and track flows
+of null metadata. Similarly, in order to perform partial monitoring without
+false positives, we need to be able to track the flow of invalid metadata. In
+order to efficiently handle these operations, we use a dedicated hardware
+dataflow engine. A block diagram of the dataflow engine is shown in
+Figure~\ref{fig:arch.dataflow}.
+
+% Flags and flag storage
+In order track both null metadata flows and invalid metadata flows, the
+dataflow engine tracks the flow of 2-bit flags. One bit represents is used to
+mark metadata as null while the other bit used to mark metadata as invalid.
+Monitoring schemes typically keep metadata corresponding to the main core's
+register file as well as the main core's memory space. Similarly, the dataflow
+engine uses an on-chip register file to store flags for register file metadata.
+In addition, a set of flags are stored that correspond to memory metadata.
+These memory metadata flags are accessed through a cache and backed to main
+memory. All flags are initialized as null and valid when the system starts.
+
+% Address calculation and configuration
+On a monitoring event, the dataflow engine reads in up to two flags in order to
+determine whether the event can be filtered.  The flags to be read in typically correspond to the source
+operands of the monitored instruction. For example, on a load, the dataflow
+engine will typically read in the flags corresponding the memory location
+accessed.
+However, the architecture is designed such that the address of flags to be read
+is calculated dynamically and can be configured depending on the monitoring
+scheme. Specifically, there exists a configuration table to outputs a set of
+control signals depending on the instruction type of the monitoring event.
+These control signals are used to control address calculation units which
+contain a simple ALU. The inputs to the ALU are information from the monitored
+event and a constant that is specified from the configuration table. One common
+use of this address calculation is to transform addresses from byte-addressed
+to word-addressed by right shifting the passed byte-address by 2 bits.
+
+% Filtering decision
+The decision of whether an event can be filtered out is determined using a
+lookup table. This filter decision table is configured by the monitoring scheme
+on system initialization. The lookup table determines whether an event can be
+filtered out based on the pair of input flags read and the instruction type of
+the monitored event.
+For example, on an ALU instruction, if both of the source operand flags
+indicate that the corresponding metadata is null, then the instruction can be
+filtered out. Similarly, on an ALU instruction, if either of the source operand
+flags indicate that their corresponding metadata is invalid, then the
+instruction can also be filtered out. This filter decision is sent to the
+``Filter'' block shown in Figure~\ref{fig:arch.overview} which will simply pop
+filtered entries from the FIFO and not forward them. If an entry is not
+filtered, then this Filter block will pop the monitoring event entry from the
+FIFO and forward it along the datapath.
+
+% Propagating flag information
+Finally, when an event is filtered out, it is necessary to propagate the flag
+information for the filtered event. In addition to the decision of whether to
+filter an event, the filter decision table also includes information about
+whether the destination flag should be written to and what value should be
+written to it. A third address calculation unit generates the write address for
+this destination operand flag. For example, for an ALU instruction, this will
+correspond to the destination operand of the ALU instruction. If this ALU
+instruction is filtered out due to having a pair of null source operands, then
+the filter decision table will indicate that the destination operand's
+corresponding flag should also be marked as null. In this way, the dataflow
+engine propagates the flag information forward.
+
+\subsection{Setting Dataflow Flags}
+\label{sec:arch.dropping}
+
+Although the dataflow engine automatically propagates information about
+non-null and invalid metadata, it also also requires for these flows to be
+initialized. Thus, it needs interfaces for setting flags to non-null and
+invalid states. These interfaces are marked in Figure~\ref{fig:arch.overview}.
+
+Metadata is marked as non-null when the monitoring core writes a valid metadata
+value. When the monitoring core performs these metadata update operations, it
+also writes to the corresponding dataflow flag to mark the flag as non-null.
+
+Flags are marked as invalid when a monitoring event is dropped due to
+insufficient overhead budget. Thus, whenever a monitoring event is dropped, the
+dropping hardware indicates this to the dataflow engine. The dataflow engine's
+configuration is already set for the write address to correspond to the
+monitoring event's destination operand. Thus, on a drop signal, the dataflow
+engine simply marks this destination flag as invalid.
+
+\subsection{Example Extensions}

### 2014-05-29 introduction.tex
Reorganized paper to unify discussion of dataflow engine.

@@
 parallel run-time monitoring model that this paper assumes.
-Section~\ref{sec:filter} presents our architecture that uses dataflow-guided filtering
-in order to reduce monitoring overheads. In Section~\ref{sec:drop} we present
-how the dataflow tracking engine can be extended to enable adjustable overheads.
-Section~\ref{sec:extensions} covers example implementations of three monitoring
-schemes. Section~\ref{sec:evaluation} presents our evaluation methodology and
+Section~\ref{sec:optimizations} presents the high-level approach of how we aim
+to reduce overheads through and Section~\ref{sec:arch} describes our
+hardware architecture that supports this. Section~\ref{sec:policies}
+investigates the design space for dropping decisions for partial monitoring.
+Section~\ref{sec:evaluation} presents our evaluation methodology and
 results. Finally, we discuss related work in Section~\ref{sec:related} and

### 2014-05-29 optimizations.tex
Reorganized paper to unify discussion of dataflow engine.

\section{Reducing Overheads of Run-Time Monitoring}
\label{sec:optimizations}

% Overview of overhead vs. coverage space
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/optimization_overview.pdf}
    \vspace{-0.2in}
    \caption{Filtering can be used to reduce the overheads. Partial monitoring
    allows a further trade-off between coverage and overheads.}
    \label{fig:optimizations.overview}
    \vspace{-0.1in}
  \end{center}
\end{figure}

The overheads of monitoring can be high. Figure~\ref{dummy} shows the full
monitoring overheads for three example monitoring extensions: array bounds
check (BC), uninitialized memory check (UMC), and dynamic information flow
tracking (DIFT). In order to reduce these overheads, we present two
optimizations. The first is to filter out monitoring events which relate to
null metadata (Section~\ref{sec:optimizations.filter}). Second, in order to
further reduce overheads, we propose to trade-off monitoring coverage. By
performing partial monitoring, we can further decrease the overheads incurred
due to monitoring. Figure~\ref{fig:optimizations.overview} shows an overview of
these ideas. Filtering allows us to reduce the overheads of monitoring and
partial monitoring allows us to further reduce overheads and create a trade-off
between overheads and coverage.

\subsection{Filtering Null Metadata Events}
\label{sec:optimizations.filter}

The instruction-grained run-time monitoring schemes we focus on typically set
and keep track of metadata information. This metadata information is used to
verify security or reliability probabilities. For example, for array bounds
check, the monitoring core stores metadata describing the base and bounds
addresses for array pointers. This metadata is checked on load and store
instructions. If the metadata for an event is null, then it is not relevant to
array bounds check since this represents a non-array pointer object. By
filtering out these null monitoring events, we can reduce the overheads of
monitoring by reducing the amount of work the monitoring core must do.

In order to filter out null monitoring events, we need an efficient method to
keep track of which events are null. In addition, when we filter out a null
monitoring event, we must ensure that any metadata updates that were skipped
are handled properly. Typically, these null monitoring events are just
propagating this null metadata information. Thus, when we filter out one of
these events, we must ensure that the destination operand is marked as null. In
other words, we need to be able to track the dataflow for these null metadata.

\subsection{Trading-Off Coverage for Reduced Overheads}
\label{sec:optimizations.drop}

There is a limit to the reduction in overheads by using filtering. We propose
that overheads can be further reduced by performing partial monitoring. As the
monitoring schemes we target consist of a large number of independent
monitoring checks, it can still be possible to provide partial protection by
performing a portion of the monitoring operations. Thus, we aim to enable this
trade-off between overheads and monitoring coverage.

The high-level idea is to track the run-time monitoring overheads. When the
overhead exceeds our target overhead, we will drop monitoring operations.  Of
course, dropping monitoring operations implies that some functionality of the
monitoring scheme has been lost.  This may cause either false negatives, where
an error that occurs in the main program's execution is not detected, or false
positives, where the monitor incorrectly believes an error has occurred.  For
example, a false positive can occur for array bounds check if an event is
dropped that handles copying a pointer. A subsequent array access using the new
pointer will incorrectly cause an error to be raised since the new pointer's
base and bounds will not be correctly set.  We accept false negatives as the
loss in coverage that we trade off in order to limit overheads.  However, we
must safely drop monitoring events in such a way as to avoid false positives so
that the system does not incorrectly raise an error.

In analyzing various monitoring schemes, we found that monitoring operations
are primarily of two types: \emph{checks} and \emph{metadata updates}. Monitors
\emph{check} certain properties to ensure correct main program execution and
they \emph{update} metadata for bookkeeping. Skipping a check operation can
only cause false negatives and will never cause a false positive. Therefore, we
may simply skip a check operation. Skipping an update operation can cause false
negatives but may also cause false positives. Essentially, when an update
operation is skipped, we can no longer trust the corresponding metadata. Thus,
we need to be able to mark these invalid metadata. Furthermore, metadata that
would be derived from these dropped events also will be invalid. In other
words, we need to be able to track the dataflow of invalid metadata when
monitoring operations are dropped.

### 2014-05-29 policies.tex
Reorganized paper to unify discussion of dataflow engine.

--- 2014-05-28 drop.tex
@@
-\section{Enabling Coverage/Overhead Trade-off}
-\label{sec:drop}
+\section{Dropping Policies}
+\label{sec:policies}
 
-Figure~\ref{dummy} shows the overheads of array bounds check with filtering
-based on clean metadata (see Section~\ref{sec:evaluation} for details on
-the experimental methodology). The overheads range from XX\% to XX\%. In some
-contexts, this overhead may still be considered to be too high, forcing
-monitoring to be disabled. Instead, it would be useful if partial monitoring
-can be done in order to provide partial protection with further reduced
-overheads. In this
-section, we discuss how we can selectively perform monitoring in order to
-further reduce overheads.
+In order to match an overhead budget, we need to drop a portion of the
+monitoring events. In this section, we discuss how this dropping decision is
+made. Specifically, we first discuss how to measure the run-time monitoring
+overheads and when monitoring events need to be dropped in
+Section~\ref{sec:policies.slack}. Next, we investigate how the decision of
+which events to drop can create a trade-off between closely matching the
+overhead target and the coverage that can be achieved.
 
-% Overview of dropping architecture
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\columnwidth]{figs/drop_overview.pdf}
-    \vspace{-0.3in}
-    \caption{Block diagram of architecture for monitoring with adjustable overheads.}
-    \label{fig:drop.overview}
-    \vspace{-0.1in}
-  \end{center}
-\end{figure}
-
-The goal is to be able to perform as much monitoring as possible within a specified overhead budget.
-The high-level approach is to keep track of monitoring overheads at run-time
-and drop monitoring operations when the overhead exceed the specified budget.
-A block diagram of the architecture we will present in this section is shown in
-Figure~\ref{fig:drop.overview}.  In order to limit the overheads of
-monitoring, we need to first be able to track the gap between the specified
-overhead budget and the monitoring overhead seen at run-time. We refer to this
-as slack and discuss how we measure it in Section~\ref{sec:drop.slack}. 
-Dropping monitoring operations can cause false positives which we prevent by
-extending the dataflow engine to track and filter out dropped monitoring flows (see
-Section~\ref{sec:drop.false_positives}).
-Given that we know how much slack we have for monitoring and can drop
-monitoring operations without false positives, we must decide when and which
-monitoring operations to drop.  There exists some interesting trade-offs in the
-policy for this dropping decision which we discuss in
-Section~\ref{sec:drop.policies}.
-
-\subsection{Measuring Overheads}
-\label{sec:drop.slack}
+\subsection{Measuring Monitoring Overheads}
+\label{sec:policies.slack}
 
@@
 
-% In addition to keeping track of overheads, a policy must be developed to
-% determine when monitoring event should be dropped. Complex policies involving
-% multiple conditions and safety factors could be used to finely tune
-% the trade off between monitoring and performance overhead. Similarly, introducing
-% randomness into the policy could allow for better sampling in a cooperative
-% debugging application of monitoring. Here, we present one simple
-% dropping policy that is based on the accumulated slack.
-% When the monitor is initialized, the monitor will load
-% its needed slack in order to perform a monitoring operation into a register. When
-% there is a monitoring event in the FIFO to be processed, a hardware comparator
-% checks whether the accumulated slack is greater than or equal to the necessary
-% slack for full monitoring. If enough slack exists, the monitor is signaled to
-% perform the monitoring operation. Otherwise, the monitoring event is dropped.
-
-\subsection{Preventing False Positives}
-\label{sec:drop.false_positives}
-
-Dropping a monitoring operation implies that some functionality of the monitor
-has been lost. This may cause either false negatives, where an error that
-occurs in the main program's execution is not detected, or false positives,
-where the monitor incorrectly believes an error has occurred. 
-% For example, a
-% false positive can occur for UMC if a store monitoring event is dropped. This
-% causes the store memory location to not be marked as initialized. A subsequent
-% load for the memory location will incorrectly cause an error to be raised.  
-For example, a false positive can occur for array bounds check if an event is
-dropped that handles copying a pointer. A subsequent array access using the new
-pointer will incorrectly cause an error to be raised since the new pointer's
-base and bounds will not be correctly set.
-We
-accept false negatives as the loss in coverage that we trade off in order to
-limit overheads.  However, we must safely drop monitoring events in such a way
-as to avoid false positives so that the system does not incorrectly raise an
-error.
-
-In analyzing various monitoring schemes, we found that monitoring operations
-are primarily of two types: \emph{checks} and \emph{metadata updates}. Monitors
-\emph{check} certain properties to ensure correct main program execution and
-they \emph{update} metadata for bookkeeping. Skipping a check operation can
-only cause false negatives and will never cause a false positive. Therefore, we
-may simply skip a check operation. Skipping an update operation can cause false
-negatives but may also cause false positives. Essentially, when an update
-operation is skipped, we can no longer trust the corresponding metadata.  Thus,
-by associating an invalid flag with each metadata we can mark metadata that is
-not correctly updated and use this information to prevent false positives.
-
-% % Flag propagation rules for BC
-% \begin{table}[tb]
-%   \begin{center}
-%     \begin{small}
-%     \input{tabs/propagation_rules}
-%     \end{small}
-%     \caption{Flag propagation rules for array bounds check. A dash indicates that no
-%     propagation is done while a 1 indicates that the destination flag is set.}
-%     \label{tab:drop.propagation_rules}
-%     \vspace{-0.2in}
-%   \end{center}
-% \end{table}
-
-We can track these invalid flags by extending the dataflow engine presented in
-Section~\ref{sec:filter} to keep track of an additional bit for each flag. That
-is, we use the dataflow engine to track 2-bit flags in order to keep track of
-both clean/dirty information as well as valid/invalid information.
-Similar to filtering for clean flags, we can also filter out
-monitoring operations corresponding to invalid metadata which would be useless to perform.
-However, it is important that we propagate the
-invalidation flag in these circumstances so that the entire flow is marked as invalid. 
-In the typical case, the propagation rule used is
-that if any of the source operands is marked as invalid, then the destination
-operand is marked as invalid. This is slightly different than the
-policy for clean flags which propagates a clean flag only if both
-sources are marked as clean. 
-% Table~\ref{tab:drop.propagation_rules} shows the
-% propagation rules that are used for array bounds check. 
-We note that
-the propagation rules can be customized depending on the monitor by
-reconfiguring the filter decision table.
-
-\subsection{Dropping Decision}
-\label{sec:drop.policies}
+\subsection{Dropping Decision Trade-Offs}
+\label{sec:policies.events}
 
@@
   \begin{center}
-  \subfloat[all-event dropping]{
+  \subfloat[unrestricted dropping]{
     \includegraphics[width=\columnwidth]{figs/all_drop.pdf}
@@
   \\
-  \subfloat[source-event dropping]{
+  \subfloat[source-only dropping]{
     \includegraphics[width=\columnwidth]{figs/source_drop.pdf}
@@
   \\
-  \subfloat[min-event dropping]{
+  \subfloat[split-point dropping]{
     \includegraphics[width=\columnwidth]{figs/min_drop.pdf}
@@
 not monitor an entire metadata flow. We refer to this dropping decision policy
-as \emph{source-event dropping} and we refer to the previous policy of 
-dropping any event as \emph{all-event dropping}.
+as \emph{source-only dropping} and we refer to the previous policy of 
+dropping any event as \emph{unrestricted dropping}.
 Figures~\ref{fig:drop.all_drop} and \ref{fig:drop.source_drop} shows a
@@
 Another possible method is to make the dropping decisions somewhere in between
-source-event dropping and all-event dropping. Looking at the graph from 
+source-only dropping and unrestricted dropping. Looking at the graph from 
 Figure~\ref{fig:drop.dataflow_graph}, what we would like to do is make a
@@
 no wasted work, then this information can be used at run-time to minimize the
-amount of wasted work. We refer to this policy as \emph{min-event dropping}.
+amount of wasted work. We refer to this policy as \emph{split-point dropping}.
 Figure~\ref{fig:drop.min_drop} shows the points were dropping decisions are
 made for this policy.
-Min-event dropping is a coarser-grained decision than all-event dropping, but
+Min-event dropping is a coarser-grained decision than unrestricted dropping, but
 finer-grained that source dropping. Thus, its ability to match overheads is
@@
 chance of wasted work due to strange splitting and merging metadata flow graph
-shapes, but typically we expect much less wasted work than all-event 
-dropping. Thus, we expect the coverage achieved by min-event dropping for a
-certain overhead to fall between source and all-event dropping. These three
+shapes, but typically we expect much less wasted work than unrestricted 
+dropping. Thus, we expect the coverage achieved by split-point dropping for a
+certain overhead to fall between source and unrestricted dropping. These three
 dropping policies create a trade-off space between coverage achieved and
@@
 Section~\ref{sec:evaluation}.
-
-% \subsection{Multiple-Run Coverage}
-% Coverage can be considered for a single or average run instance. It can also be
-% measured over multiple users/runs in order to capture total bug coverage of a
-% program.
-% 
-% The system presented so far operates in a deterministic manner and so each run
-% will perform the same monitoring checks. However, we can introduce randomness
-% in order to achieve different dropping patterns over runs and achieve higher
-% multiple-run coverage.
-% 
-% Due to the multiplicative effect of randomness, long dependence flows with
-% multiple points where randomness is introduced leads to a small chance that the
-% entire flow is monitored. This is a significant issue for last-minute dropping
-% where all events could be dropped. Min-point dropping reduces this effect by
-% reducing the points where dropping decisions are made while source-dropping
-% eliminates this issue entirely.
-% 
-% \subsection{Multi-Core Implementation}
-% 
-% This can be extended to mulit-core by using multiple dataflow engines and a
-% similar coherency protocol as that used for data caches.

### 2014-05-29 related.tex
Reorganized paper to unify discussion of dataflow engine.

@@
 Many monitoring schemes have been developed for various security, reliability,
-debugging, and other capabilities.  As we described in
-Section~\ref{sec:extensions}, there exist monitoring schemes for uninitialized
+debugging, and other capabilities. There exist monitoring schemes for uninitialized
 memory check \cite{mondrian-asplos02}, array bounds check
@@
 policy. Lo et al. \cite{lo-rtas14} developed a hardware architecture to limit
-monitoring in hard real-time systems. Their architecture performs all-event 
+monitoring in hard real-time systems. Their architecture performs unrestricted 
 dropping in order to guarantee strict deadlines. Our system allows for a
-trade-off between choosing source-event dropping, all-event dropping, or min-event
+trade-off between choosing source-only dropping, unrestricted dropping, or split-point 
 dropping depending on the needs of the designer or user which has not been

### 2014-05-29 introduction.tex
Edits to abstract, intro, and monitoring.

@@
 % Filtering
-Recent studies have shown that filtering out monitoring checks on clean
+Recent studies have shown that filtering out monitoring checks on clean or null
 metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}. For example, for
@@
 paper, we present an architecture that combines a hardware-based dataflow
-tracking engine with a software-based monitor in order to filter out both
-propagation and checking of clean metadata. By keeping a separate set of flags
-to track clean metadata, we are able to handle the common case of clean
-metadata without the need to read full metadata. By propagating these flags, we
-are also able to handle propagation of clean metadata efficiently in hardware.
-Thus, we are able to have the flexibility of software-based monitors with
-greatly reduced overheads.
+tracking engine with a software-based monitor in order to filter out monitoring operations 
+for null metadata. By keeping a separate set of flags
+to track null metadata, we are able to handle the common case of null
+metadata without the need to read the full metadata values. By propagating these flags, we
+are also able to handle propagation of null metadata efficiently in hardware.
+This filtering can be used to greatly reduce the overheads of monitoring.
 
 % Adjustable overhead
-With this filtering, we see overheads of XX\% - XX\% on average for
-uninitialized memory check, array bounds check, and multi-bit dynamic
-information flow tracking. 
+There is a limit to how effective filtering is at reducing monitoring overheads. 
 Traditionally, these overheads have been an all-or-nothing cost. If the
@@
 cannot be enabled.  
-Thus, we present an architecture that enables a trade-off between overheads and
-the monitoring coverage achieved. This is achieved by dynamically dropping
-monitoring operations when overheads are too high. We prevent false
-positives by extending the dataflow engine to also track these dropped
-monitoring flows. We investigate three different policies on when dropping decisions are
-made that cover a trade-off space between the ability to match a target overhead
-and the monitoring coverage achieved.
+Thus, we propose to perform partial monitoring to enable further reduction of overheads.
+ This is achieved by dynamically dropping
+monitoring operations when overheads exceed a specified overhead budget. We prevent false
+positives from partial monitoring by using the dataflow engine to track these dropped
+monitoring flows. We investigate different policies for deciding which events
+to drop for partial monitoring and show that there is a trade-off between
+closely matching the overhead budget and the monitoring coverage achieved.
 
@@
 
-Thus, the main contributions of this paper are summarized as follows.
+The main contributions of this paper are summarized as follows.
 \begin{itemize}
-  \item We use a dataflow tracking engine to efficiently filter out monitoring events in order to reduce overheads.
-  \item We present an architecture that enables adjustable monitoring, creating a trade-off between monitoring coverage and overheads.
-  \item We show how the dataflow tracking engine can be used to prevent false positives for adjustable monitoring.
-  \item We explore the design space and trade-offs in dropping policies for performing adjustable monitoring.
+  \item We propose two optimizations for reducing the overheads of monitoring: filtering null metadata and performing partial monitoring.
+  \item We present a hardware architecture using a dataflow tracking engine to efficiently enable filtering and partial monitoring.
+  \item We investigate trade-off decisions for partial monitoring on the proposed architecture.
 \end{itemize}

### 2014-05-29 monitoring.tex
Reorganized paper to unify discussion of dataflow engine.
Edits to abstract, intro, and monitoring.

@@
-\section{Parallel Run-Time Monitoring}
+\section{Parallel Programmable Run-Time Monitoring}
 \label{sec:monitoring}
@@
     \caption{Overview of run-time monitoring architecture.}
-    \label{fig:arch.overview} 
+    \label{fig:monitoring.overview} 
     \vspace{-0.1in}
@@
 
-Figure~\ref{fig:arch.overview} shows an overview of the run-time monitoring
+Figure~\ref{fig:monitoring.overview} shows an overview of the run-time monitoring
 model that is assumed in this paper.  The \emph{main program} is a computation
@@
 on the type of monitoring event, different monitoring operations are
-executed. Monitoring events are buffered in a FIFO structure to decouple the
+executed. Information about monitoring events are sent to the monitoring core and buffered in a FIFO structure to decouple the
 running of the main core and the monitoring core. If the FIFO is full, then the main
@@
 array's bounds. This can be done by associating metadata with array pointers that 
-indicates the array's base and bound addresses. On loads or stores with the
+indicates the array's base (start) and bound (end) addresses. On loads or stores with the
 array

### 2014-05-29 title.tex
Edits to abstract, intro, and monitoring.

@@
 systems. However, these run-time monitors can introduce large performance and energy
-overhead, especially when performed on programmable cores.
+overheads, especially when performed on programmable cores.
 %Traditionally, these overheads have been an
@@
 In this paper, we introduce a hardware dataflow tracking engine that can be
-used to filter out unnecessary monitoring and enable partial monitoring.
-The dataflow engine can identify events with clean metadata so that they can
+used to filter out unnecessary monitoring and enable adjustable partial monitoring.
+The dataflow engine can identify events with null metadata so that they can
 be filtered out. To further reduce the overhead of monitoring, we propose
 to enable a trade-off between monitoring coverage and overhead by dropping certain
-monitoring operations. For the partial monitoring, the dataflow engine is used
+monitoring operations. For this partial monitoring, the dataflow engine is used
 to track dropped monitoring flows so that false positives can be avoided.
@@
 made for partial monitoring and show that there exists a trade-off between closely
-matching an overhead target and coverage that can be achieved.
+matching an overhead target and the coverage that can be achieved.
 % Our experiments show that the system is able to

### 2014-05-29 optimizations.tex
Updates to optimizations section.

@@
 
-The overheads of monitoring can be high. Figure~\ref{dummy} shows the full
+Although the monitoring core operates in parallel to the main core, the
+overheads of programmable monitoring can be high due to the multi-cycle
+operation of the monitoring core.
+Figure~\ref{dummy} shows the full
 monitoring overheads for three example monitoring extensions: array bounds
 check (BC), uninitialized memory check (UMC), and dynamic information flow
-tracking (DIFT). In order to reduce these overheads, we present two
-optimizations. The first is to filter out monitoring events which relate to
-null metadata (Section~\ref{sec:optimizations.filter}). Second, in order to
-further reduce overheads, we propose to trade-off monitoring coverage. By
-performing partial monitoring, we can further decrease the overheads incurred
-due to monitoring. Figure~\ref{fig:optimizations.overview} shows an overview of
-these ideas. Filtering allows us to reduce the overheads of monitoring and
-partial monitoring allows us to further reduce overheads and create a trade-off
-between overheads and coverage.
+tracking (DIFT) (Section~\ref{sec:evaluation} details the experimental methodology). In order to reduce these overheads, we present two
+optimizations. The first optimization is to filter out monitoring events which relate to
+null metadata which do not need to processed by the monitoring core (Section~\ref{sec:optimizations.filter}). Second, in order to allow
+further reduction in overheads, we propose to perform partial monitoring. This
+allows us to further reduce the overheads of monitoring by trading-off
+monitoring coverage. Figure~\ref{fig:optimizations.overview} shows an overview
+of
+these ideas.
 
@@
 
-The instruction-grained run-time monitoring schemes we focus on typically set
-and keep track of metadata information. This metadata information is used to
+The instruction-grained run-time monitoring schemes we focus on typically
+manage a set of metadata information. This metadata information is used to
 verify security or reliability probabilities. For example, for array bounds
-check, the monitoring core stores metadata describing the base and bounds
-addresses for array pointers. This metadata is checked on load and store
-instructions. If the metadata for an event is null, then it is not relevant to
-array bounds check since this represents a non-array pointer object. By
-filtering out these null monitoring events, we can reduce the overheads of
+check, the monitoring core stores metadata describing the base and bound
+addresses for array pointers. This metadata is then checked on load and store
+instructions. However, if the metadata for an event is null, then there is no need to perform the check since this represents a non-array pointer value. By
+identifying and filtering out these null monitoring events, we can reduce the overheads of
 monitoring by reducing the amount of work the monitoring core must do.
@@
 monitoring event, we must ensure that any metadata updates that were skipped
-are handled properly. Typically, these null monitoring events are just
-propagating this null metadata information. Thus, when we filter out one of
-these events, we must ensure that the destination operand is marked as null. In
-other words, we need to be able to track the dataflow for these null metadata.
+are handled properly.  Thus, when we filter out one of
+these events, we must ensure that the destination operand is also marked as null. 
+In other words, we need to be able to track not only
+metadata which is initially null, but also the dataflow of null metadata.
 
@@
 
-There is a limit to the reduction in overheads by using filtering. We propose
-that overheads can be further reduced by performing partial monitoring. As the
+There is a limit to the reduction in overheads that can be achieved by using
+filtering alone. We propose to further reduce overheads
+by performing partial monitoring. As the
 monitoring schemes we target consist of a large number of independent
@@
 
-The high-level idea is to track the run-time monitoring overheads. When the
-overhead exceeds our target overhead, we will drop monitoring operations.  Of
-course, dropping monitoring operations implies that some functionality of the
+Our goal in using partial monitoring is to allow the user or system designer to
+specify their acceptable overheads. Given this overhead budget, monitoring is
+only performed as long as this budget is not exceeded. If the overhead budget
+is exceeded, then monitoring operations are dropped. This creates an adjustable
+trade-off between the overheads incurred and the monitoring coverage achieved.
+
+Dropping monitoring operations implies that some functionality of the
 monitoring scheme has been lost.  This may cause either false negatives, where
@@
 example, a false positive can occur for array bounds check if an event is
-dropped that handles copying a pointer. A subsequent array access using the new
-pointer will incorrectly cause an error to be raised since the new pointer's
-base and bounds will not be correctly set.  We accept false negatives as the
+dropped that handles copying a pointer. Since the new pointer does not have
+correct base and bound address metadata, it will incorrectly cause errors to be
+raised.  We accept false negatives as the
 loss in coverage that we trade off in order to limit overheads.  However, we

### 2014-05-29 arch.tex
Edits to architecture section.

@@
     \includegraphics[width=\columnwidth]{figs/architecture_overview.pdf}
-    \vspace{-0.3in}
+    \vspace{-0.2in}
     \caption{Block diagram of architecture for monitoring with reduced and adjustable overheads.}
@@
 
-In this section, we will describe our hardware architecture for reduced
-overheads by filtering out null metadata and enabling partial monitoring.
+For both filtering and partial monitoring, we need to be able to track
+dataflows. For filtering, we need to track the dataflow of null metadata and
+for partial monitoring, we need to track the dataflow of invalid metadata.
+In this section, we will describe our dataflow-guided hardware architecture for
+enabling both filtering and partial monitoring.
 Figure~\ref{fig:arch.overview} shows a high-level block diagram of the
-architecture. The basic idea is to insert hardware between the main core and
+architecture. The basic idea is to insert this dataflow-guided hardware between the main core and
 the monitoring core that will reduce the number of monitoring events that reach
@@
 invalid metadata. A hardware dataflow engine is used to track the flow of null
-and invalid metadata (Section~\ref{sec:arch.dataflow}) and decide whether
+and invalid metadata and decide whether
 specific events can be filtered out.
@@
 marked and tracked as invalid by the dataflow engine. The
-dropping decision hardware is described in Section~\ref{sec:arch.dropping}.
+dropping decision hardware is described in detail in Section~\ref{sec:policies}.
 Finally, if the event is not filtered or dropped, then it is forwarded to the
@@
 In order to perform filtering, we need to be able to identify and track flows
-of null metadata. Similarly, in order to perform partial monitoring without
-false positives, we need to be able to track the flow of invalid metadata. In
-order to efficiently handle these operations, we use a dedicated hardware
+of null metadata. Similarly, for partial monitoring 
+, we need to be able to track the flow of invalid metadata. In
+order to efficiently these flows, we use a dedicated hardware
 dataflow engine. A block diagram of the dataflow engine is shown in
@@
 In order track both null metadata flows and invalid metadata flows, the
-dataflow engine tracks the flow of 2-bit flags. One bit represents is used to
-mark metadata as null while the other bit used to mark metadata as invalid.
+dataflow engine tracks the flow of 2-bit flags. One bit is used to
+mark metadata as null while the other bit is used to mark metadata as invalid.
 Monitoring schemes typically keep metadata corresponding to the main core's
@@
 is calculated dynamically and can be configured depending on the monitoring
-scheme. Specifically, there exists a configuration table to outputs a set of
+scheme. Specifically, there exists a configuration table that outputs a set of
 control signals depending on the instruction type of the monitoring event.
-These control signals are used to control address calculation units which
+These control signals are used to control a pair of address calculation units which each
 contain a simple ALU. The inputs to the ALU are information from the monitored
@@
 use of this address calculation is to transform addresses from byte-addressed
-to word-addressed by right shifting the passed byte-address by 2 bits.
+to word-addressed by right shifting the passed memory address by 2 bits.
 
@@
 For example, on an ALU instruction, if both of the source operand flags
-indicate that the corresponding metadata is null, then the instruction can be
+indicate that the corresponding metadata is null, then typically the instruction can be
 filtered out. Similarly, on an ALU instruction, if either of the source operand
@@
 written to it. A third address calculation unit generates the write address for
-this destination operand flag. For example, for an ALU instruction, this will
-correspond to the destination operand of the ALU instruction. If this ALU
+this destination flag. For example, for an ALU instruction, this will
+correspond to the destination register of the ALU instruction. If this ALU
 instruction is filtered out due to having a pair of null source operands, then
-the filter decision table will indicate that the destination operand's
-corresponding flag should also be marked as null. In this way, the dataflow
-engine propagates the flag information forward.
+the filter decision table will indicate that the destination register's
+corresponding flag should also be marked as null.
 
@@
 Although the dataflow engine automatically propagates information about
-non-null and invalid metadata, it also also requires for these flows to be
-initialized. Thus, it needs interfaces for setting flags to non-null and
-invalid states. These interfaces are marked in Figure~\ref{fig:arch.overview}.
+non-null and invalid metadata, it also requires for these flows to be
+initialized. That is, there needs to be some way to mark metadata as non-null
+or invalid. These interfaces from the monitoring core and the dropping hardware
+are marked in Figure~\ref{fig:arch.overview}.
 
@@
 dropping hardware indicates this to the dataflow engine. The dataflow engine's
-configuration is already set for the write address to correspond to the
-monitoring event's destination operand. Thus, on a drop signal, the dataflow
-engine simply marks this destination flag as invalid.
+write address is configured for the appropriate destination flag. When it
+receives an invalidation signal from the dropping hardware, the
+dataflow engine marks this destination flag as invalid.
 

### 2014-05-29 policies.tex
Edits to dropping policies section.

@@
 
-In order to match an overhead budget, we need to drop a portion of the
-monitoring events. In this section, we discuss how this dropping decision is
+Our goal with enabling partial monitoring is to allow for a specified overhead target or budget to be met while maximizing the amount of monitoring performed.
+This is done by dynamically dropping a portion of the monitoring operations. In this section, we discuss in detail how this dropping decision is
 made. Specifically, we first discuss how to measure the run-time monitoring
 overheads and when monitoring events need to be dropped in
-Section~\ref{sec:policies.slack}. Next, we investigate how the decision of
-which events to drop can create a trade-off between closely matching the
-overhead target and the coverage that can be achieved.
+Section~\ref{sec:policies.slack}. Next, we investigate trade-offs created by
+selecting which events are dropped. Different policies in determining which
+events can be dropped create a trade-off between how closely the overhead
+budget is met and the monitoring coverage achieved.
 
@@
 {\tt B} represents that if event {\tt A} is dropped, then due to its
-invalidated metadata, it will cause event {\tt B} to be also dropped. Square
+invalidated metadata, it will cause event {\tt B} to be filtered. Square
 nodes indicate events where monitoring checks are performed. In the
@@
 dropping any event as \emph{unrestricted dropping}.
-Figures~\ref{fig:drop.all_drop} and \ref{fig:drop.source_drop} shows a
+Figures~\ref{fig:drop.all_drop} and \ref{fig:drop.source_drop} show a
 comparison of where dropping decisions are made for these two policies. Source
 dropping will
-result in no wasted work. However, because of the coarser-grained decision, it
+result in no wasted work and thus better coverage. However, because of the coarser-grained decision, it
 may be more difficult to closely match overheads. Source dropping is expected
@@
 Another possible method is to make the dropping decisions somewhere in between
-source-only dropping and unrestricted dropping. Looking at the graph from 
-Figure~\ref{fig:drop.dataflow_graph}, what we would like to do is make a
-decision at event {\tt C} whether to perform monitoring on that branch of the
-metadata flow. If we choose to perform monitoring, then we want to perform all
-monitoring on that flow. Instead, if we choose to drop event {\tt C}, then the
-entire flow is dropped with no wasted work. Similarly, dropping event {\tt F}
-drops the bottom flow with no wasted work and event {\tt A} can be dropped to
-drop the entire shown metadata flow with no wasted work. If the program can be
+source-only dropping and unrestricted dropping. From the graph in 
+Figure~\ref{fig:drop.dataflow_graph}, we can see that if the dropping
+decision is made at event {\tt C} then we can skip event {\tt E} with no wasted
+work. If we instead choose to perform monitoring for event {\tt C}, then we want to perform all
+monitoring on that flow. Similarly, event {\tt F} creates a decision point for
+the bottom flow that will result in no wasted work and dropping event {\tt A}
+allows the entire shown metadata flow to be dropped with no wasted work. If the
+program can be
 analyzed or profiled to identify these instructions which, if dropped, lead to
 no wasted work, then this information can be used at run-time to minimize the
-amount of wasted work. We refer to this policy as \emph{split-point dropping}.
+amount of wasted work. We refer to this policy as \emph{sub-flow dropping}.
 Figure~\ref{fig:drop.min_drop} shows the points were dropping decisions are
@@
 finer-grained that source dropping. Thus, its ability to match overheads is
-expected to sit between these other two policies. Similarly, there is still a
-chance of wasted work due to strange splitting and merging metadata flow graph
-shapes, but typically we expect much less wasted work than unrestricted 
+expected to sit between these other two policies. Similarly, there still exists
+edge cases where wasted work can occur, but typically we expect much less
+wasted work than unrestricted 
 dropping. Thus, we expect the coverage achieved by split-point dropping for a
@@
 dropping policies create a trade-off space between coverage achieved and
-ability to match an overhead target which we evaluate in detail in
-Section~\ref{sec:evaluation}.
+ability to match an overhead target.

### 2014-05-30 arch.tex
Edits to dropping policies section.
Add example to architecture section.

@@
 In order to perform filtering, we need to be able to identify and track flows
-of null metadata. Similarly, for partial monitoring 
-, we need to be able to track the flow of invalid metadata. In
+of null metadata. Similarly, for partial monitoring,
+ we need to be able to track the flow of invalid metadata. In
 order to efficiently these flows, we use a dedicated hardware
@@
 
-\subsection{Example Extensions}
+\subsection{Example Monitoring Scheme}
+
+% Monitoring operations
+\begin{table*}[tb]
+  \begin{center}
+    \begin{small}
+    \input{tabs/monitor}
+    \end{small}
+    \caption{Monitoring operations for array bounds check.}
+    \label{tab:arch.monitor}
+    \vspace{-0.2in}
+  \end{center}
+\end{table*}
+
+% Example dataflow operations
+\begin{table*}[tb]
+  \begin{center}
+    \begin{small}
+    \input{tabs/dataflow_operations}
+    \end{small}
+    \caption{Filtering conditions and operations for clean flags.}
+    \label{tab:arch.dataflow_operations}
+    \vspace{-0.2in}
+  \end{center}
+\end{table*}
+
+In this section, we show an example implementation of how the dataflow engine
+is configured for an array bounds check monitoring scheme.
+Table~\ref{tab:arch.monitor} shows a summary of the monitoring operations for
+array bounds check.  When an array pointer is allocated, the metadata
+associated with it set to include the base (start) address and bound (end)
+address of the corresponding array. On ALU instructions, this base and bound
+metadata is copied to the destination registers. On store instructions, the
+metadata for the source register is copied to the metadata for the store
+address. Similarly, on load instructions, the metadata value for the load
+address is copied to the destination register's metadata. In addition, on load
+and store instructions, the memory address is checked to be within the base and
+bounds metadata. If it is not, then an out-of-bounds error is detected.
+
+Table~\ref{tab:arch.dataflow_operations} summarizes when when monitoring events
+can be filtered out due to clean and invalid metadata. In addition, it shows
+what metadata needs to be modified on a filtered or dropped monitoring event.
+Note that for simplicity the policies for clean and invalid flags are shown
+separately. However, the dataflow engine actually operates on pairs of these
+flags using a combined policy.
+
+In terms of filtering using clean flags, array allocations are never filtered
+since these will always involve non-clean metadata (i.e., setting new base and
+bounds). For ALU operations, if both source operands are clean, then the event
+can be filtered by setting the destination register as clean. For load and
+store, only a single source needs to be checked to be clean in order for the
+event to be filtered. Similarly, the destination register or memory location
+needs to be marked as clean if the event is filtered.
+
+The invalid flags for array bounds check operate similarly to the clean flags.
+When ALU, load, or store events are dropped or filtered, the destination
+register or memory location is marked as invalid. In addition, we need to
+handle the possibility of dropping an array allocation. In this case, the array
+pointer is marked as invalid. In terms of filtering out invalid events, for an
+ALU event, if either of the source registers is invalid then the event can be
+filtered. For load and store instructions, if the single source operand is
+invalid, then the event can be filtered.
+
+\subsection{Applicability to Other Schemes}
+
+Our approach can be generally applied to a wide range of monitoring schemes. This
+is possible because the use of clean and invalidation flags is largely agnostic
+to the semantics of the specific monitoring scheme. That is, the dataflow
+engine must be configured based on the dataflow of the monitoring scheme, but
+it does not need detailed information of what the monitoring scheme does.
+
+We do note some limitations to the proposed approach. Previous work
+\cite{fade-hpca14} that has looked at filtering in order to reduce monitoring
+overheads has also filtered out redundant updates where the source metadata
+matches the destination metadata. Since we do not read the actual metadata, our
+architecture is not able to identify these redundant updates. However, by using
+only 2-bit flags, the bandwidth need of our dataflow engine is reduced.  
+In addition, we note that filtering using clean flags works well when a large
+portion of instructions are not relevant to the monitoring scheme (e.g., array
+bounds check only cares about array pointers). However, if most instruction
+have dirty metadata, then very little filtering will be possible.
+
+In terms of using invalidation to perform partial monitoring, this 
+approach works best for monitoring schemes where there are many pieces of
+independent metadata. Since marking a piece of metadata as invalid will cause
+future metadata with dependences to also be marked as invalid, schemes with
+large metadata dependency chains can quickly see large amounts of metadata
+become invalid.
+
+Another limitation is that the dataflow engine, as described, only marks one
+flag location on each monitoring event. Similarly, the dataflow engine can only
+read in two sets of flags in order to determine whether an event can be
+filtered or not. This was done because we found that for many monitoring
+schemes, the monitoring operations corresponded to program instructions, which
+read up to two operands and write to one destination.  For monitoring schemes
+which update multiple pieces of metadata on a monitoring event, it may be
+adequate to associate a clean and invalid flag with a single piece of metadata.  For
+example, a scheme may use a data structure that includes several pieces of
+related metadata. On a drop, one of these pieces could be marked as invalid in
+order to indicate that the entire structure is invalid.  Similarly, marking one
+specified piece as clean may indicate that the entire data structure's metadata
+is clean.  Alternatively, the read and write capabilities of the dataflow
+engine could be expanded to support more complex monitoring schemes.

### 2014-05-30 tabs/dataflow_operations.tex
Add example to architecture section.

% Operations for handling clean flags

\begin{tabular}{|l|l|l|l|}
\hline

{\bf Flag Type} & {\bf Event} & {\bf Filter Condition} & {\bf Dropping/Filter Operation} \\ \hline \hline

\multirow{4}{*}{Clean flag}  
& Array allocation & Never & - \\ \cline{2-4}
& ALU   & Both source registers are clean & Mark destination register as clean \\ \cline{2-4}
& Load  & Source memory location is clean & Mark destination register as clean \\ \cline{2-4}
& Store & Source register is clean & Mark destination memory location as clean \\
\hline\hline

\multirow{4}{*}{Invalid flag}  
& Array allocation & Never & Invalidate array pointer \\ \cline{2-4}
& ALU   & At least one invalid source register & Invalidate destination register \\ \cline{2-4}
& Load  & Invalid source memory location & Invalidate destination register \\ \cline{2-4}
& Store & Invalid source register & Invalidate destination memory location \\
\hline%\hline

% \multirow{2}{*}{UMC}  
% & Store & Never & - \\ \cline{2-4}
% & Load & Source memory location is dirty & Do nothing \\ 
% \hline\hline
% 
% \multirow{5}{*}{DIFT}  
% & Software set tag & Never & - \\ \cline{2-4}
% & ALU   & Both source registers are clean & Mark destination register as clean \\ \cline{2-4}
% & Load  & Source memory location is clean & Mark destination register as clean \\ \cline{2-4}
% & Store & Source register is clean & Mark destination memory location as clean \\ \cline{2-4}
% & Indirect jump & Source register is clean & Do nothing \\ 
% \hline

\end{tabular}

### 2014-05-30 tabs/drop.tex
Add example to architecture section.

@@
 & Store & Invalid source register & Invalidate destination memory location \\
-\hline\hline
+\hline%\hline
 
-\multirow{2}{*}{UMC}  
-& Store & Never & Invalidate destination memory location \\ \cline{2-4}
-& Load & Invalid source memory location & Do nothing \\ 
-\hline\hline
-
-\multirow{5}{*}{DIFT}  
-& Software set tag & Never & Invalidate destination memory location \\ \cline{2-4}
-& ALU   & At least one invalid source register & Invalidate destination register \\ \cline{2-4}
-& Load  & Invalid source memory location & Invalidate destination register \\ \cline{2-4}
-& Store & Invalid source register & Invalidate destination memory location \\ \cline{2-4}
-& Indirect jump & Invalid source metadata & Do nothing \\ 
-\hline
+% \multirow{2}{*}{UMC}  
+% & Store & Never & Invalidate destination memory location \\ \cline{2-4}
+% & Load & Invalid source memory location & Do nothing \\ 
+% \hline\hline
+% 
+% \multirow{5}{*}{DIFT}  
+% & Software set tag & Never & Invalidate destination memory location \\ \cline{2-4}
+% & ALU   & At least one invalid source register & Invalidate destination register \\ \cline{2-4}
+% & Load  & Invalid source memory location & Invalidate destination register \\ \cline{2-4}
+% & Store & Invalid source register & Invalidate destination memory location \\ \cline{2-4}
+% & Indirect jump & Invalid source metadata & Do nothing \\ 
+% \hline
 

### 2014-05-30 tabs/monitor.tex
Add example to architecture section.

--- 2014-05-28 tabs/monitor.tex
@@
 
-\begin{tabular}{|l|l|l|}
+\begin{tabular}{|l|l|}
 \hline
 
-{\bf Monitor} & {\bf Event} & {\bf Monitoring Operation} \\ \hline \hline
+ {\bf Event} & {\bf Monitoring Operation} \\ \hline \hline
 
-\multirow{4}{*}{BC}  
- & Array allocation & Set base and bound metadata \\ \cline{2-3}
- & ALU & Copy metadata from register to register \\  \cline{2-3}
- & Load & Copy metadata from memory to register. Check memory access address. \\ \cline{2-3}
- & Store & Copy metadata from register to memory. Check memory access address. \\ 
-\hline\hline
+ Array allocation & Set base and bound metadata \\ \hline
+ ALU & Copy metadata from register to register \\  \hline
+ Load & Copy metadata from memory to register. Check memory access address. \\ \hline
+ Store & Copy metadata from register to memory. Check memory access address. \\ 
+\hline%\hline
 
-\multirow{2}{*}{UMC}  
- & Store & Set metadata tag (data initialized) \\ \cline{2-3}
- & Load & Check if metadata tag is set \\ 
-\hline\hline
-% & Software clear tag & Clear metadata tag & Non-critical \\ \hline\hline
-
-\multirow{5}{*}{DIFT}  
- & Software set tag & Set tag to a specified value based on software command \\ \cline{2-3}
- & ALU & Output tag is the OR of the input tags \\ \cline{2-3}
- & Load & Set register tag as tag of loaded data \\ \cline{2-3}
- & Store & Set memory tag as register tag \\ \cline{2-3}
-% & Software clear tag & Clear tag & Update \\ \cline{2-4}
- & Indirect jump & Check tag \\ 
- \hline
-
-% \multirow{5}{*}{BC}  
-%  & MOV/LD/ST          & Propagate colors (Tag[dest] = Tag[src]) & Critical \\ \cline{2-4}
-%  & ADD/SUB/NOT        & Update colors (Tag[dest] = f(Tag[src])) & Critical \\ \cline{2-4}
-%  & OR/XOR             & Clear colors & Critical \\ \cline{2-4}
-%  & LD/ST              & Compare pointer and memory colors & Check \\ \cline{2-4}
-%  & Software set color & Set pointer colors & Critical \\ \hline\hline
-
-% Soft Error Check & Instruction & Check instruction computation & Check \\ \hline\hline
+% \multirow{2}{*}{UMC}  
+%  & Store & Set metadata tag (data initialized) \\ \cline{2-3}
+%  & Load & Check if metadata tag is set \\ 
+% \hline\hline
 % 
-% \multirow{2}{*}{Bank of Observers}  
-%  & System output & Update state estimation & Critical \\ \cline{2-4}
-%  & New state estimate & Check that states of all observers matches & Check \\ \hline
+% \multirow{5}{*}{DIFT}  
+%  & Software set tag & Set tag to a specified identifier value based on software command \\ \cline{2-3}
+%  & ALU & If input tags are different, output tag is a new identifier. Otherwise, output tag is the input tag. \\ \cline{2-3}
+%  & Load & Set register tag as tag of loaded data \\ \cline{2-3}
+%  & Store & Set memory tag as register tag \\ \cline{2-3}
+%  & Indirect jump & Check tag \\ 
+%  \hline
 

### 2014-05-30 optimizations.tex
Add figure for policy trade off.

@@
 manage a set of metadata information. This metadata information is used to
-verify security or reliability probabilities. For example, for array bounds
+verify security or reliability properties. For example, for array bounds
 check, the monitoring core stores metadata describing the base and bound
 addresses for array pointers. This metadata is then checked on load and store
-instructions. However, if the metadata for an event is null, then there is no need to perform the check since this represents a non-array pointer value. By
-identifying and filtering out these null monitoring events, we can reduce the overheads of
+instructions. However, if the metadata for an event is null, then there is no
+need to perform the check since this represents a non-array pointer value. By
+identifying and filtering out these null monitoring events, we can reduce the
+overheads of
 monitoring by reducing the amount of work the monitoring core must do.
@@
 keep track of which events are null. In addition, when we filter out a null
-monitoring event, we must ensure that any metadata updates that were skipped
-are handled properly.  Thus, when we filter out one of
-these events, we must ensure that the destination operand is also marked as null. 
+monitoring event, we must ensure that any metadata propagations that were
+skipped are handled properly. If this is not done, the monitoring core may
+operate on stale metadata. Thus, when we filter out a null monitoring event,
+we must ensure that the destination metadata is also marked as null. 
 In other words, we need to be able to track not only

### 2014-05-30 policies.tex
Add figure for policy trade off.

@@
     \caption{Slack and its effect on monitoring over time.}
-    \label{fig:drop.slack}
+    \label{fig:policies.slack}
     \vspace{-0.1in}
@@
     \caption{Slack tracking and drop decision hardware.}
-    \label{fig:drop.stm}
+    \label{fig:policies.stm}
     \vspace{-0.1in}
@@
 can be given in order for monitoring to be performed at the start of a program.
-Figure~\ref{fig:drop.slack} shows an example of how slack can change over time.
+Figure~\ref{fig:policies.slack} shows an example of how slack can change over time.
 If the slack falls below zero (i.e., the overhead budget is exceeded), then
@@
 
-Figure~\ref{fig:drop.stm} shows a hardware slack tracking module for
+Figure~\ref{fig:policies.stm} shows a hardware slack tracking module for
 keeping track of slack. The slack tracking module uses a counter that increments on every $k$-th
@@
     events where checks are performed.} 
-    \label{fig:drop.dataflow_graph}
+    \label{fig:policies.dataflow_graph}
     \vspace{-0.1in}
@@
 wasted work. For example, consider the metadata dependence graph shown in
-Figure~\ref{fig:drop.dataflow_graph}. Here, an edge from node {\tt A} to node
+Figure~\ref{fig:policies.dataflow_graph}. Here, an edge from node {\tt A} to node
 {\tt B} represents that if event {\tt A} is dropped, then due to its
@@
     \includegraphics[width=\columnwidth]{figs/all_drop.pdf}
-    \label{fig:drop.all_drop}
+    \label{fig:policies.all_drop}
   }
@@
     \includegraphics[width=\columnwidth]{figs/source_drop.pdf}
-    \label{fig:drop.source_drop}
+    \label{fig:policies.source_drop}
   }
@@
     \includegraphics[width=\columnwidth]{figs/min_drop.pdf}
-    \label{fig:drop.min_drop}
+    \label{fig:policies.min_drop}
   }
@@
   dropping decisions are made.}
-  \label{fig:drop.policies}
+  \label{fig:policies.policies}
 \end{figure}
@@
 dropping any event as \emph{unrestricted dropping}.
-Figures~\ref{fig:drop.all_drop} and \ref{fig:drop.source_drop} show a
+Figures~\ref{fig:policies.all_drop} and \ref{fig:policies.source_drop} show a
 comparison of where dropping decisions are made for these two policies. Source
@@
 
+% Slack tracking module
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\columnwidth]{figs/policy_trade_off.pdf}
+    \vspace{-0.3in}
+    \caption{Slack tracking and drop decision hardware.}
+    \label{fig:policies.trade_off}
+    \vspace{-0.2in}
+  \end{center}
+\end{figure}
+
 Another possible method is to make the dropping decisions somewhere in between
 source-only dropping and unrestricted dropping. From the graph in 
-Figure~\ref{fig:drop.dataflow_graph}, we can see that if the dropping
+Figure~\ref{fig:policies.dataflow_graph}, we can see that if the dropping
 decision is made at event {\tt C} then we can skip event {\tt E} with no wasted
@@
 amount of wasted work. We refer to this policy as \emph{sub-flow dropping}.
-Figure~\ref{fig:drop.min_drop} shows the points were dropping decisions are
+Figure~\ref{fig:policies.min_drop} shows the points were dropping decisions are
 made for this policy.
@@
 ability to match an overhead target.
+Figure~\ref{fig:policies.trade_off} summarizes this trade-off between coverage
+and ability to match overhead.

### 2014-05-30 introduction.tex
Revise introduction.

@@
 These run-time monitors introduce performance, power, and energy overheads.
-Implementing run-time monitoring in software using binary instrumentation or
-other similar methods has been shown to introduce high overheads. For example,
-dynamic information flow tracking (DIFT) implemented in software suffers a 3.6x
-slowdown \cite{lift-micro06}. Implementing monitors in hardware greatly decreases
-the performance impact by performing monitoring in parallel to a program's
-execution. A dedicated hardware implementation of DIFT reduces performance
-overheads to just 1.1\% \cite{dift-asplos04}. However, fixed hardware loses
-the programmability of software, and cannot be updated or changed.
+% Implementing run-time monitoring in software using binary instrumentation or
+% other similar methods has been shown to introduce high overheads. For example,
+% dynamic information flow tracking (DIFT) implemented in software suffers a 3.6x
+% slowdown \cite{lift-micro06}. 
+Implementing monitors in dedicated hardware has shown to have very small
+performance impacts. For example, a dedicated hardware implementation of
+dynamic information flow tracking (DIFT) shows performance
+overheads of only 1.1\% \cite{dift-asplos04}. However, fixed hardware loses
+the programmability of software, and cannot be updated or changed. It also
+requires for monitors to be chosen and implemented at hardware design time.
+Recent studies \cite{lba-isca08} have proposed using parallel cores in a
+multi-core system to perform monitoring. This allows monitoring to be done
+using parallel hardware while having the flexibility and programmability of
+software. However, due to the multiple cycles it may take for software to
+monitor an instruction, the overheads can easily increase execution time
+several fold.
+
+% Two optimizations
+In this paper, we present a hardware architecture that uses dynamic dataflow
+tracking in order to reduce the overheads of programmable parallel monitoring.
+This dataflow tracking engine is used to filter out unnecessary monitoring and
+enable partial monitoring.
 
@@
 Recent studies have shown that filtering out monitoring checks on clean or null
-metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}. For example, for
-an array bounds check, checking non-pointer accesses is unnecessary. In this
-paper, we present an architecture that combines a hardware-based dataflow
-tracking engine with a software-based monitor in order to filter out monitoring operations 
-for null metadata. By keeping a separate set of flags
-to track null metadata, we are able to handle the common case of null
-metadata without the need to read the full metadata values. By propagating these flags, we
-are also able to handle propagation of null metadata efficiently in hardware.
-This filtering can be used to greatly reduce the overheads of monitoring.
-
-% Adjustable overhead
-There is a limit to how effective filtering is at reducing monitoring overheads. 
-Traditionally, these overheads have been an all-or-nothing cost. If the
-designer finds that the overheads exceed their desired budget, then monitoring
-cannot be enabled.  
-Thus, we propose to perform partial monitoring to enable further reduction of overheads.
- This is achieved by dynamically dropping
-monitoring operations when overheads exceed a specified overhead budget. We prevent false
-positives from partial monitoring by using the dataflow engine to track these dropped
-monitoring flows. We investigate different policies for deciding which events
-to drop for partial monitoring and show that there is a trade-off between
-closely matching the overhead budget and the monitoring coverage achieved.
+metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}.
+For example, for
+an array bounds check, checking non-pointer accesses is unnecessary. 
+In this paper, we use a hardware-based dataflow tracking engine to track and
+filter out these null metadata monitoring operations.
+By using the dataflow engine to track a separate set of flags
+indicating null metadata, we are able to handle the common case of null
+metadata without the need to read the full metadata values.
 
 Although full monitoring coverage is always desired, trading off coverage for
-reduced overheads can be very useful. For example, this adjustable
-monitoring enables a level of protection even for systems where the full
+reduced overheads can be very useful. For example, partial
+monitoring can enable a level of protection even for systems where the full
 monitoring overheads are too high. This is especially true for energy- or
@@
 overhead should not exceed energy/power limits or real-time deadlines.
-Additionally, adjustable monitoring can be used for debugging purposes
+Additionally, partial monitoring can be used for debugging purposes
 especially in the context of sampling-based or cooperative debugging techniques
@@
 greathouse-cgo11}.
+
+% % Adjustable overhead
+% There is a limit to how effective filtering is at reducing monitoring overheads. 
+% Traditionally, these overheads have been an all-or-nothing cost. If the
+% designer finds that the overheads exceed their desired budget, then monitoring
+% cannot be enabled.  
+In this paper, we show how our dataflow-guided architecture is also able to
+enable partial monitoring. This is achieved by dynamically dropping monitoring
+operations when overheads exceed a specified overhead budget. We prevent false
+positives from partial monitoring by using the dataflow engine to track these
+dropped monitoring flows. With this architecture, we investigate different
+policies for deciding which events to drop for partial monitoring. These
+different policies show a trade-off between closely matching the overhead
+budget and the monitoring coverage achieved.
 

### 2014-05-30 introduction.tex
Minor edits to introduction.

@@
 % slowdown \cite{lift-micro06}. 
-Implementing monitors in dedicated hardware has shown to have very small
+Implementing monitors in dedicated hardware has been shown to have small
 performance impacts. For example, a dedicated hardware implementation of
@@
 overheads of only 1.1\% \cite{dift-asplos04}. However, fixed hardware loses
-the programmability of software, and cannot be updated or changed. It also
+the programmability of software, and cannot be updated or changed. It 
 requires for monitors to be chosen and implemented at hardware design time.
@@
 several fold.
-
 % Two optimizations
 In this paper, we present a hardware architecture that uses dynamic dataflow
-tracking in order to reduce the overheads of programmable parallel monitoring.
-This dataflow tracking engine is used to filter out unnecessary monitoring and
-enable partial monitoring.
+tracking in order to reduce the overheads of programmable parallel monitoring
+by filtering out unnecessary monitoring operations and by enabling partial
+monitoring.
 

### 2014-05-30 drop.tex
Started fleshing out evaluation.

@@
 
-Figure~\ref{dummy} shows the overheads of array bounds check with filtering
-based on clean metadata (see Section~\ref{sec:evaluation} for details on
-the experimental methodology). The overheads range from XX\% to XX\%. In some
+Filtering is only able to reduce overheads up to a certain point since there will still remain necessary monitoring operations. In some
 contexts, this overhead may still be considered to be too high, forcing

### 2014-05-30 evaluation.tex
Filled out extensions section.
Started fleshing out evaluation.

@@
 \subsection{Experimental Setup}
+\label{sec:evaluation.setup}
+
+We implemented our monitoring architecture for adjustable overheads by
+modifying the ARM version of the gem5 simulator \cite{gem5} to support parallel
+run-time monitoring. In order to explore the generality of the architecture for
+different monitors, we implemented three different monitors: uninitialized
+memory check (UMC), array bounds check (BC), and dynamic information flow
+tracking (DIFT).  Uninitialized memory check seeks to detect loading from
+memory locations that are not initialized first.  Array bounds check, as
+mentioned in Section~\ref{sec:monitoring} is a monitoring scheme that aims to
+detect buffer overflows where memory accesses go beyond the boundaries of an
+array.  Dynamic information flow tracking is a security monitoring scheme,
+which detects when information from untrusted sources is used to affect the
+program control flow.  We tested our system using 12 benchmarks from SPEC
+CPU2006 \cite{spec2006}. For each benchmark, we simulated for 200 million
+instructions. An initial slack of 2 million cycles, which is less than 1\% of
+the total execution time, is given for a 10\% target overhead. This initial
+slack is scaled proportionally for other overhead targets.
 
@@
 
-Effect of using initialized flag to filter out monitoring events and lower overheads.
+% Full monitoring overheads
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\columnwidth]{figs/data_filtering.pdf}
+    \vspace{-0.2in}
+    \caption{Monitoring overheads with filtering for UMC, BC, and DIFT.}
+    \label{fig:evaluation.filtering}
+    \vspace{-0.1in}
+  \end{center}
+\end{figure}
+
+Figure~\ref{fig:optimizations.full_mon} showed the execution time of benchmarks
+with monitoring normalized to without monitoring. In this case, filtering was
+not performed. UMC showed an 8x increase in execution
+time on average while BC and DIFT showed average execution time increases of
+21x and 14x respectively. Figure~\ref{fig:evaluation.filtering} shows the
+normalized execution times with filtering enabled. We see significant
+reductions in overhead for all three monitoring schemes. UMC sees normalized
+execution times of 4x on average with filtering. For BC, normalized execution
+times drop from 21x down to 3x. This is not surprising since in the baseline
+implementation all loads and stores needed to be monitored. However, with
+filtering, only loads and stores corresponding to arrays need to be forwarded.
+Finally, DIFT sees the largest reduction in overheads with only 13\% overheads
+on average. This is due to the fact that for our implementation of DIFT on SPEC
+benchmarks, we mark data read from files as tainted. For most of these
+benchmarks, this propagates to relatively few instructions. Instead, if we
+targeted network or streaming applications we would expect to see a less
+filtering
 
@@
 
-\subsection{Randomness for Multi-Run Coverage}
-
-\subsection{Custom Hardware Monitors}
-
-\subsection{Main Core Scaling}
+%\subsection{Randomness for Multi-Run Coverage}
 
 \subsection{Area and Power Overheads}
+
+% Area and Power Overheads
+\begin{table}[tb]
+  \begin{center}
+    \vspace{-0.0in}
+    \begin{footnotesize}
+    \input{tabs/area_power}
+    \end{footnotesize}
+    \caption{Average power overhead for dropping hardware at a 10\% overhead
+    budget. Percentages in parentheses are normalized to the main core
+    power.}
+    \vspace{-0.2in}
+    \label{tab:evaluation.area_power}
+  \end{center}
+\end{table}
+
+Adding the dropping hardware in order to enable adjustable overheads adds
+overheads in terms of area and power. We use McPat \cite{mcpat-micro09} to get
+a first-order estimate of these area and power overheads in a 40 nm technology
+node. McPat estimates the main core area as 1.96 mm$^2$ and the peak power usage as
+121 mW averaged across all benchmarks. The average runtime power usage was 38.2 
+mW. These area and power numbers consist of the core and
+L1 cache, but do not include L2 cache, memory controller, and other
+peripherals. The power numbers include dynamic as well as static (leakage)
+power. For the dropping hardware, the ALUs, MIRF, MIC, and
+configuration tables are modeled using the corresponding objects in McPat. We
+note that this is only a rough area and power result since components such as the
+wires connecting these modules have not been modeled. However, this gives a
+sense of the order-of-magnitude overheads involved with implementing our
+approach.
+
+For a 1 kB MIC, an additional 0.183 mm$^2$ of silicon area is needed, an
+increase of 9\% of the main core area. Table~\ref{tab:evaluation.area_power}
+shows the peak and runtime power overheads. The peak power is 13-19 mW, which is
+11-16\% of the main core's peak power usage. The average runtime power is 8-11
+mW, corresponding to 21-29\% of the main core's runtime power.

### 2014-05-30 optimizations.tex
Started fleshing out evaluation.

@@
 \label{sec:optimizations}
+
+% Full monitoring overheads
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\columnwidth]{figs/data_full_mon.pdf}
+    \vspace{-0.2in}
+    \caption{Full monitoring overheads for uninitialized memory check (UMC), array bounds check (BC), and dynamic information flow tracking (DIFT).}
+    \label{fig:optimizations.full_mon}
+    \vspace{-0.1in}
+  \end{center}
+\end{figure}
 
@@
 operation of the monitoring core.
-Figure~\ref{dummy} shows the full
+Figure~\ref{fig:optimizations.full_mon} shows the full
 monitoring overheads for three example monitoring extensions: array bounds
 check (BC), uninitialized memory check (UMC), and dynamic information flow
-tracking (DIFT) (Section~\ref{sec:evaluation} details the experimental methodology). In order to reduce these overheads, we present two
+tracking (DIFT) (Section~\ref{sec:evaluation} details the experimental
+methodology). We see overheads from 4-28x. In order to reduce these overheads, we present two
 optimizations. The first optimization is to filter out monitoring events which relate to

### 2014-05-30 evaluation.tex
Fleshed out evaluation section with 200m inst data and some fake data.

@@
 \subsection{Coverage with Adjustable Overheads}
-Coverage and overhead budget accuracy for source, last-minute, and min-point dropping.
 
-%\subsection{Randomness for Multi-Run Coverage}
+% BC sweep
+\begin{figure*}
+  \begin{center}
+    \includegraphics[width=\linewidth]{figs/data_bc_sweep.pdf}
+    \vspace{-0.4in}
+    \caption{Coverage versus varying overhead budget for array bounds check.}
+    \label{fig:evaluation.bc_sweep}
+    \vspace{-0.2in}
+  \end{center}
+\end{figure*}
+
+% UMC sweep
+\begin{figure*}
+  \begin{center}
+    \includegraphics[width=\linewidth]{figs/data_umc_sweep.pdf}
+    \vspace{-0.4in}
+    \caption{Coverage versus varying overhead budget for uninitialized memory check.}
+    \label{fig:evaluation.umc_sweep}
+    \vspace{-0.1in}
+  \end{center}
+\end{figure*}
+
+Although the overheads of DIFT are quite low after filtering, BC and UMC still
+show significant overheads. In this section, we evaluate the effectiveness of
+using partial monitoring to trade-off coverage for further reduced overheads.
+Figure~\ref{fig:evaluation.bc_sweep} shows the monitoring coverage achieved by
+array bounds check as we vary the overhead budget. We measure monitoring
+coverage monitoring check operations that are still performed as a percentage
+of the total number of monitoring check operations. This number does not
+include the number metadata update operations that are done and serves as a
+better metric for the actual coverage of the security or reliability technique.
+The overhead is measured as a percentage of the program without monitoring. An
+unrestricted dropping policy is used here.
+
+We see that by varying the overhead budget, the coverage achieved also varies.
+Furthermore, with only a 50\% overhead budget, array bounds check still
+achieves 87\% monitoring coverage on average. This overhead budget corresponds
+to 26\% of the average overheads seen by BC after filtering. The high coverage
+achieved with such low overheads is due to two main effects.  The first is that
+monitoring can be done in parallel, providing monitoring coverage without
+introducing overheads. The second effect is that, although we aggressively
+filter out monitoring events, there may still exist a large number of
+monitoring events that do not lead to security or reliability checks. As a
+result, dropping these events can reduce overheads without a large impact on
+monitoring coverage.
+
+Figure~\ref{fig:evaluation.umc_sweep} shows the analogous graph for UMC. Again
+we see that varying overhead budgets enables partial monitoring. We see that
+with a 50\% overhead budget, UMC achieves 50\% monitoring coverage on average.
+This overhead budget corresponds to only 15\% of the overheads seen by UMC
+after filtering.
+
+\subsection{Comparing Dropping Policies}
+
+In this section we evaluate the trade-offs of the three dropping policies
+presented in Section~\ref{sec:policies.events}: unrestricted dropping, source
+dropping, and sub-flow dropping.
+
+% BC exec time
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\columnwidth]{figs/data_bc_exec_time.pdf}
+    \vspace{-0.2in}
+    \caption{Error in meeting overhead budget for array bounds check.}
+    \label{fig:evaluation.bc_exec_time}
+    \vspace{-0.2in}
+  \end{center}
+\end{figure}
+
+% UMC exec time
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\columnwidth]{figs/data_umc_exec_time.pdf}
+    \vspace{-0.2in}
+    \caption{Error in meeting overhead budget for uninitialized memory check.}
+    \label{fig:evaluation.umc_exec_time}
+    \vspace{-0.2in}
+  \end{center}
+\end{figure}
+
+First, we evaluate how well the different policies are able to meet a specified
+overhead budget. Figures~\ref{fig:evaluation.bc_exec_time} and
+\ref{fig:evaluation.umc_exec_time} show the difference between the run-time
+overheads seen and the overhead specified. These errors are averaged over all
+benchmarks. A positive value means that the
+overhead target was overshot while a negative value indicates that the
+overheads were below the specified budget.
+
+% BC coverage across policies
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\columnwidth]{figs/data_bc_coverage.pdf}
+    \vspace{-0.2in}
+    \caption{Coverage for array bounds check for different dropping policies.}
+    \label{fig:evaluation.bc_coverage}
+    \vspace{-0.2in}
+  \end{center}
+\end{figure}
+
+% UMC coverage across policies
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\columnwidth]{figs/data_umc_coverage.pdf}
+    \vspace{-0.2in}
+    \caption{Coverage for uninitialized memory check for different dropping policies.}
+    \label{fig:evaluation.umc_coverage}
+    \vspace{-0.1in}
+  \end{center}
+\end{figure}
+
+Next, we evaluate the coverage achieved by the different policies.
+Figure~\ref{fig:evaluation.bc_coverage} shows the average coverage across
+benchmarks for each of the three policies with varying overhead budgets.
+Figure~\ref{fig:evaluation.umc_coverage} shows this data for uninitialized
+memory check.
 

### 2014-05-30 evaluation.tex
Revise evaluation section.

@@
 modifying the ARM version of the gem5 simulator \cite{gem5} to support parallel
-run-time monitoring. In order to explore the generality of the architecture for
+run-time monitoring. We model the main and monitoring cores as running at 2.5
+GHz with 4-way set-associative private L1 I/D caches and a shared 8-way 2 MB L2
+cache. This setup is similar to the Snapdragon 801 processor commonly found in
+mobile systems. The dataflow engine uses a 1 kB cache for flags.
+
+In order to explore the generality of the architecture for
 different monitors, we implemented three different monitors: uninitialized
@@
 detect buffer overflows where memory accesses go beyond the boundaries of an
-array.  Dynamic information flow tracking is a security monitoring scheme,
+array. We modify the implementation of {\tt malloc} to set base and bound
+metadata information. Dynamic information flow tracking is a security
+monitoring scheme,
 which detects when information from untrusted sources is used to affect the
-program control flow.  We tested our system using 12 benchmarks from SPEC
-CPU2006 \cite{spec2006}. For each benchmark, we simulated for 200 million
+program control flow. For our DIFT implementation, we set mark data read from
+files as untrusted. This data is marked with a 32-bit metadata identifier so
+that if an error is detected, it is possible to have information about where
+the information originated from.
+
+We tested our system using all C benchmarks from SPECint
+CPU2006 \cite{spec2006}. We focus on the C benchmarks because of our
+modification of {\tt malloc} to support array bounds check. Although we do not
+show results for the C++ benchmarks, we note that the results for UMC and DIFT
+for these benchmarks are similar to the other results shown. For each
+benchmark, we simulated for 200 million
 instructions. An initial slack of 2 million cycles, which is less than 1\% of
@@
 slack is scaled proportionally for other overhead targets.
+
+\subsection{Baseline Monitoring Overheads}
+% Full monitoring overheads
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\columnwidth]{figs/data_full_mon.pdf}
+    \vspace{-0.2in}
+    \caption{Full monitoring overheads for UMC, BC, and DIFT.}
+    \label{fig:evaluation.full_mon}
+    \vspace{-0.1in}
+  \end{center}
+\end{figure} Figure~\ref{fig:evaluation.full_mon} shows the execution times of
+performing full monitoring normalized to the execution times of the benchmarks
+without monitoring. In these results, no filtering or partial monitoring is
+done. UMC shows normalized execution times from 5x to 11x with an average
+of 8x. BC shows normalized execution times of 15-28x with an average of
+22x while DIFT shows normalized execution times of 11x-17x with an average of
+14x. One of the reasons for these high overheads is that our implementations of
+these monitors dynamically allocate memory for new metadata. This can be an
+expensive, multi-cycle operation. Statically allocating memory for metadata
+will reduce the execution time but requires a large upfront memory footprint.
 
@@
 
-Figure~\ref{fig:optimizations.full_mon} showed the execution time of benchmarks
-with monitoring normalized to without monitoring. In this case, filtering was
-not performed. UMC showed an 8x increase in execution
-time on average while BC and DIFT showed average execution time increases of
-21x and 14x respectively. Figure~\ref{fig:evaluation.filtering} shows the
+Figure~\ref{fig:evaluation.filtering} shows the
 normalized execution times with filtering enabled. We see significant

### 2014-05-30 evaluation.tex
Updated some of data with 1bn instruction runs.

@@
 for these benchmarks are similar to the other results shown. For each
-benchmark, we simulated for 200 million
-instructions. An initial slack of 2 million cycles, which is less than 1\% of
+benchmark, we simulated for 1 billion instructions. An initial slack of 1
+million cycles, which is less than 1\% of
 the total execution time, is given for a 10\% target overhead. This initial
@@
 without monitoring. In these results, no filtering or partial monitoring is
-done. UMC shows normalized execution times from 5x to 11x with an average
-of 8x. BC shows normalized execution times of 15-28x with an average of
-22x while DIFT shows normalized execution times of 11x-17x with an average of
-14x. One of the reasons for these high overheads is that our implementations of
+done. UMC shows normalized execution times from 4x to 11x with an average
+of 7x. BC shows normalized execution times of 9-28x with an average of
+18x while DIFT shows normalized execution times of 6x-17x with an average of
+11x. One of the reasons for these high overheads is that our implementations of
 these monitors dynamically allocate memory for new metadata. This can be an
-expensive, multi-cycle operation. Statically allocating memory for metadata
+expensive, many-cycle operation. Statically allocating memory for metadata
 will reduce the execution time but requires a large upfront memory footprint.
@@
     \vspace{-0.2in}
-    \caption{Monitoring overheads with filtering for UMC, BC, and DIFT.}
+    \caption{Monitoring overheads with filtering null metadata for UMC, BC, and DIFT.}
     \label{fig:evaluation.filtering}
@@
 reductions in overhead for all three monitoring schemes. UMC sees normalized
-execution times of 4x on average with filtering. For BC, normalized execution
-times drop from 21x down to 3x. This is not surprising since in the baseline
+execution times 2-6x with an average of 4x with null metadata filtering. For BC, normalized execution
+times drop from 18x down to 4x on average. The range of normalized execution
+times for filtered BC is 1.1-17x. This is not surprising since in the baseline
 implementation all loads and stores needed to be monitored. However, with
@@
 Finally, DIFT sees the largest reduction in overheads with only 13\% overheads
-on average. This is due to the fact that for our implementation of DIFT on SPEC
-benchmarks, we mark data read from files as tainted. For most of these
+on average after filtering with a maximum of 42\%. This is due, in part, to the fact that
+for our implementation of DIFT on SPEC
+benchmarks, we only mark data read from files as tainted. For most of these
 benchmarks, this propagates to relatively few instructions. Instead, if we
-targeted network or streaming applications we would expect to see a less
-filtering
+targeted network or streaming applications, which have larger amounts in
+untrusted input data, we would expect to see less filtering.
 
@@
 show significant overheads. In this section, we evaluate the effectiveness of
-using partial monitoring to trade-off coverage for further reduced overheads.
+using partial monitoring (in addition to filtering) to trade-off coverage for further reduced overheads.
 Figure~\ref{fig:evaluation.bc_sweep} shows the monitoring coverage achieved by
@@
 better metric for the actual coverage of the security or reliability technique.
-The overhead is measured as a percentage of the program without monitoring. An
-unrestricted dropping policy is used here.
+The overhead is measured as a percentage of total number of checks performed
+with full monitoring. The results shown here are using an unrestricted dropping
+policy.
 
 We see that by varying the overhead budget, the coverage achieved also varies.
-Furthermore, with only a 50\% overhead budget, array bounds check still
-achieves 87\% monitoring coverage on average. This overhead budget corresponds
-to 26\% of the average overheads seen by BC after filtering. The high coverage
+With only a 50\% overhead budget, array bounds check still
+achieves 83\% monitoring coverage on average. This overhead budget corresponds
+to just 20\% of the average overheads seen by BC after filtering only. The high coverage
 achieved with such low overheads is due to two main effects.  The first is that
@@
 we see that varying overhead budgets enables partial monitoring. We see that
-with a 50\% overhead budget, UMC achieves 50\% monitoring coverage on average.
-This overhead budget corresponds to only 15\% of the overheads seen by UMC
-after filtering.
+with a 50\% overhead budget, UMC achieves 51\% monitoring coverage on average.
+Again, this overhead budget corresponds to only 20\% of the overheads seen by UMC
+with filtering only.
 

### 2014-05-30 arch.tex
Minor edits based on Ed's comments.

@@
     \end{small}
-    \caption{Filtering conditions and operations for clean flags.}
+    \caption{Filtering conditions and dropping/filtering operations for null and invalid flags.}
     \label{tab:arch.dataflow_operations}
@@
 
-Table~\ref{tab:arch.dataflow_operations} summarizes when when monitoring events
-can be filtered out due to clean and invalid metadata. In addition, it shows
+Table~\ref{tab:arch.dataflow_operations} summarizes when monitoring events
+can be filtered out due to null and invalid metadata. In addition, it shows
 what metadata needs to be modified on a filtered or dropped monitoring event.
-Note that for simplicity the policies for clean and invalid flags are shown
+Note that for simplicity the policies for null and invalid flags are shown
 separately. However, the dataflow engine actually operates on pairs of these
@@
 
-In terms of filtering using clean flags, array allocations are never filtered
-since these will always involve non-clean metadata (i.e., setting new base and
-bounds). For ALU operations, if both source operands are clean, then the event
-can be filtered by setting the destination register as clean. For load and
-store, only a single source needs to be checked to be clean in order for the
+In terms of filtering using null flags, array allocations are never filtered
+since these will always involve non-null metadata (i.e., setting new base and
+bounds). For ALU operations, if both source operands are null, then the event
+can be filtered by setting the destination register as null. For load and
+store, only a single source needs to be checked to be null in order for the
 event to be filtered. Similarly, the destination register or memory location
-needs to be marked as clean if the event is filtered.
+needs to be marked as null if the event is filtered.
 
-The invalid flags for array bounds check operate similarly to the clean flags.
+The invalid flags for array bounds check operate similarly to the null flags.
 When ALU, load, or store events are dropped or filtered, the destination
@@
 Our approach can be generally applied to a wide range of monitoring schemes. This
-is possible because the use of clean and invalidation flags is largely agnostic
+is possible because the use of null and invalidation flags is largely agnostic
 to the semantics of the specific monitoring scheme. That is, the dataflow
@@
 only 2-bit flags, the bandwidth need of our dataflow engine is reduced.  
-In addition, we note that filtering using clean flags works well when a large
+In addition, we note that filtering using null flags works well when a large
 portion of instructions are not relevant to the monitoring scheme (e.g., array
 bounds check only cares about array pointers). However, if most instruction
-have dirty metadata, then very little filtering will be possible.
+have non-null metadata, then very little filtering will be possible.
 
@@
 which update multiple pieces of metadata on a monitoring event, it may be
-adequate to associate a clean and invalid flag with a single piece of metadata.  For
+adequate to associate a null and invalid flag with a single piece of metadata.  For
 example, a scheme may use a data structure that includes several pieces of
@@
 order to indicate that the entire structure is invalid.  Similarly, marking one
-specified piece as clean may indicate that the entire data structure's metadata
-is clean.  Alternatively, the read and write capabilities of the dataflow
+specified piece as null may indicate that the entire data structure's metadata
+is null.  Alternatively, the read and write capabilities of the dataflow
 engine could be expanded to support more complex monitoring schemes.

### 2014-05-30 evaluation.tex
Minor edits based on Ed's comments.

@@
 which detects when information from untrusted sources is used to affect the
-program control flow. For our DIFT implementation, we set mark data read from
-files as untrusted. This data is marked with a 32-bit metadata identifier so
+program control flow. For the benchmarks we consider, we mark data read from
+files as untrusted. We implemented a multi-bit DIFT scheme which marks
+untrusted data with a 32-bit metadata identifier so
 that if an error is detected, it is possible to have information about where
-the information originated from.
+the data originated from.
 
@@
 
-Adding the dropping hardware in order to enable adjustable overheads adds
+Adding the dataflow engine in order to enable filtering and adjustable adjustable adds
 overheads in terms of area and power. We use McPat \cite{mcpat-micro09} to get
@@
 peripherals. The power numbers include dynamic as well as static (leakage)
-power. For the dropping hardware, the ALUs, MIRF, MIC, and
-configuration tables are modeled using the corresponding objects in McPat. We
+power. For the dataflow engine, we modeled the ALUs used for address
+calculation, the dataflow flag register file and cache, the configuration
+tables, and the filter decision table. These were modeled using the
+corresponding objects in McPat. We
 note that this is only a rough area and power result since components such as the
@@
 
-For a 1 kB MIC, an additional 0.183 mm$^2$ of silicon area is needed, an
+Our results show that an additional 0.183 mm$^2$ of silicon area is needed, an
 increase of 9\% of the main core area. Table~\ref{tab:evaluation.area_power}

### 2014-05-30 optimizations.tex
Revise evaluation section.
Minor edits based on Ed's comments.

--- 2014-05-30 optimizations.tex
@@
 operation of the monitoring core.
-Figure~\ref{dummy} shows the full
-monitoring overheads for three example monitoring extensions: array bounds
-check (BC), uninitialized memory check (UMC), and dynamic information flow
-tracking (DIFT) (Section~\ref{sec:evaluation} details the experimental methodology). In order to reduce these overheads, we present two
+ In order to reduce these overheads, we present two
 optimizations. The first optimization is to filter out monitoring events which relate to
@@
 
-The instruction-grained run-time monitoring schemes we focus on typically
+The instruction-grained run-time monitoring schemes that we focus on typically
 manage a set of metadata information. This metadata information is used to
@@
 identifying and filtering out these null monitoring events, we can reduce the
-overheads of
-monitoring by reducing the amount of work the monitoring core must do.
+mount of work done by the monitoring core and thus reduce the overheads of
+monitoring.
 

### 2014-05-30 policies.tex
Minor edits based on Ed's comments.

@@
   \\
-  \subfloat[split-point dropping]{
+  \subfloat[sub-flow dropping]{
     \includegraphics[width=\columnwidth]{figs/min_drop.pdf}
@@
 made for this policy.
-Min-event dropping is a coarser-grained decision than unrestricted dropping, but
+Sub-flow dropping is a coarser-grained decision than unrestricted dropping, but
 finer-grained that source dropping. Thus, its ability to match overheads is
@@
 wasted work than unrestricted 
-dropping. Thus, we expect the coverage achieved by split-point dropping for a
+dropping. Thus, we expect the coverage achieved by sub-flow dropping for a
 certain overhead to fall between source and unrestricted dropping. These three

### 2014-05-30 related.tex
Minor edits based on Ed's comments.

@@
 FADE \cite{fade-hpca14} presents a general filtering engine in order to reduce
-the overheads of monitoring. It filters out redundant updates and clean checks
+the overheads of monitoring. It filters out redundant updates and null checks
 by reading metadata values. We do not handle filtering redundant updates, but
-we do filter out clean checks as well as clean (i.e., uninitialized)
+we do filter out null checks as well as null (i.e., uninitialized)
 propagations.  We are able to do this because the dataflow engine can also
-propagate the clean information.  In addition, our dataflow engine only
+propagate the null information.  In addition, our dataflow engine only
 requires reading and writing 2-bit flags rather than the entire metadata.  As
@@
 parallel monitoring architectures. In addition, these works prevented
-false positives by enabling and disabling monitoring using a source-dropping
+false positives by enabling and disabling monitoring using only a source-dropping
 policy. Lo et al. \cite{lo-rtas14} developed a hardware architecture to limit
@@
 dropping in order to guarantee strict deadlines. Our system allows for a
-trade-off between choosing source-only dropping, unrestricted dropping, or split-point 
+trade-off between choosing source-only dropping, unrestricted dropping, or sub-flow 
 dropping depending on the needs of the designer or user which has not been

### 2014-05-30 tabs/dataflow_operations.tex
Minor edits based on Ed's comments.

@@
-% Operations for handling clean flags
+% Operations for handling null flags
 
@@
 & Array allocation & Never & - \\ \cline{2-4}
-& ALU   & Both source registers are clean & Mark destination register as clean \\ \cline{2-4}
-& Load  & Source memory location is clean & Mark destination register as clean \\ \cline{2-4}
-& Store & Source register is clean & Mark destination memory location as clean \\
+& ALU   & Both source registers are null & Mark destination register as null \\ \cline{2-4}
+& Load  & Source memory location is null & Mark destination register as null \\ \cline{2-4}
+& Store & Source register is null & Mark destination memory location as null \\
 \hline\hline
@@
 % & Software set tag & Never & - \\ \cline{2-4}
-% & ALU   & Both source registers are clean & Mark destination register as clean \\ \cline{2-4}
-% & Load  & Source memory location is clean & Mark destination register as clean \\ \cline{2-4}
-% & Store & Source register is clean & Mark destination memory location as clean \\ \cline{2-4}
-% & Indirect jump & Source register is clean & Do nothing \\ 
+% & ALU   & Both source registers are null & Mark destination register as null \\ \cline{2-4}
+% & Load  & Source memory location is null & Mark destination register as null \\ \cline{2-4}
+% & Store & Source register is null & Mark destination memory location as null \\ \cline{2-4}
+% & Indirect jump & Source register is null & Do nothing \\ 
 % \hline

### 2014-05-30 arch.tex
More edits based on Ed's comments.

@@
 
-Although the dataflow engine automatically propagates information about
+Although the dataflow engine handles the propagation of information about
 non-null and invalid metadata, it also requires for these flows to be
@@
 register or memory location is marked as invalid. In addition, we need to
-handle the possibility of dropping an array allocation. In this case, the array
+handle the possibility of dropping an array allocation. In this case, the metadata for the array
 pointer is marked as invalid. In terms of filtering out invalid events, for an
-ALU event, if either of the source registers is invalid then the event can be
-filtered. For load and store instructions, if the single source operand is
-invalid, then the event can be filtered.
+ALU event, if \emph{either} of the source registers is invalid then the event can be
+filtered. This differs from the policy for null flags where the event can only
+filtered if \emph{both} flags are marked as null. This highlights the need to
+store and track separate bits for tracking null and invalid information. For
+load and store instructions, if the single source operand is invalid, then the
+event can be filtered.
 
@@
 is null.  Alternatively, the read and write capabilities of the dataflow
-engine could be expanded to support more complex monitoring schemes.
+engine could be expanded to support these more complex monitoring schemes.

### 2014-05-30 introduction.tex
More edits based on Ed's comments.

@@
 requires for monitors to be chosen and implemented at hardware design time.
-Recent studies \cite{lba-isca08} have proposed using parallel cores in a
+Recent studies \cite{lba-isca08, ruwase-spaa08, paralog-asplos10, oases-sigops09} have proposed using parallel cores in a
 multi-core system to perform monitoring. This allows monitoring to be done
@@
 
-Although full monitoring coverage is always desired, trading off coverage for
-reduced overheads can be very useful. For example, partial
+There is a limit to how effective filtering is at reducing monitoring overheads
+as it only eliminates unnecessary monitoring operations. The ability
+to also skip relevant monitoring operations can be used to further reduce overheads.
+Although this comes at the cost of reduced monitoring coverage (i.e.,
+effectiveness), this partial monitoring can still be very useful.
+For example, partial
 monitoring can enable a level of protection even for systems where the full
@@
 % % Adjustable overhead
-% There is a limit to how effective filtering is at reducing monitoring overheads. 
-% Traditionally, these overheads have been an all-or-nothing cost. If the
-% designer finds that the overheads exceed their desired budget, then monitoring
-% cannot be enabled.  
 In this paper, we show how our dataflow-guided architecture is also able to
@@
   \item We propose two optimizations for reducing the overheads of monitoring: filtering null metadata and performing partial monitoring.
-  \item We present a hardware architecture using a dataflow tracking engine to efficiently enable filtering and partial monitoring.
-  \item We investigate trade-off decisions for partial monitoring on the proposed architecture.
+  \item We present a hardware architecture using a dataflow tracking engine to
+  efficiently enable filtering and partial monitoring. This hardware
+  architecture is designed to be generally applicable to a wide-range of run-time monitoring
+  techniques.
+  \item We investigate interesting trade-off decisions for partial monitoring on the proposed architecture.
 \end{itemize}

### 2014-05-30 related.tex
More edits based on Ed's comments.

@@
 our system is able to do. In addition, with the exception of Testudo, these
-methods have not considered parallel monitoring. Finally, we note that the
-concept of distributing the monitoring over many users and runs to increase
-coverage can also be enabled using our architecture.
+methods have not considered parallel monitoring. Our techniques for generally
+reducing the overheads of monitoring can also be used to enable these
+low-overhead, multi-user/run debugging techniques.
 
 % Limited Monitoring
-In terms of controlling overheads, the Quality Virtual Machine (QVM) is a
+In terms of partial monitoring, the Quality Virtual Machine (QVM) is a
 modification of the Java Virtual Machine (JVM) that supports run-time

### 2014-05-30 evaluation.tex
Updated area and power results.

@@
 
-We implemented our monitoring architecture for adjustable overheads by
+We implemented dataflow-guided monitoring architecture by
 modifying the ARM version of the gem5 simulator \cite{gem5} to support parallel
@@
     \end{footnotesize}
-    \caption{Average power overhead for dropping hardware at a 10\% overhead
+    \caption{Average power overhead for dropping hardware at a 50\% overhead
     budget. Percentages in parentheses are normalized to the main core
@@
 a first-order estimate of these area and power overheads in a 40 nm technology
-node. McPat estimates the main core area as 1.96 mm$^2$ and the peak power usage as
-121 mW averaged across all benchmarks. The average runtime power usage was 38.2 
+node. McPat estimates the main core area as 2.71 mm$^2$ and the peak power usage as
+965 mW averaged across all benchmarks. The average runtime power usage was 385 
 mW. These area and power numbers consist of the core and
@@
 tables, and the filter decision table. These were modeled using the
-corresponding objects in McPat. We
+corresponding memory and ALU objects in McPat. We
 note that this is only a rough area and power result since components such as the
@@
 
-Our results show that an additional 0.183 mm$^2$ of silicon area is needed, an
-increase of 9\% of the main core area. Table~\ref{tab:evaluation.area_power}
-shows the peak and runtime power overheads. The peak power is 13-19 mW, which is
-11-16\% of the main core's peak power usage. The average runtime power is 8-11
-mW, corresponding to 21-29\% of the main core's runtime power.
+Our results show that an additional 0.197 mm$^2$ of silicon area is needed, an
+increase of 7\% of the main core area. Table~\ref{tab:evaluation.area_power}
+shows the peak and runtime power overheads averaged across all benchmarks with
+a 50\% monitoring overhead target. The peak power is 5-7 mW, which is
+5-7\% of the main core's peak power usage. The average runtime power is 3-4 
+mW, corresponding to 8-11\% of the main core's runtime power.

### 2014-05-30 evaluation.tex
Multi-run coverage results.

@@
 
+\subsection{Multiple-Run Coverage}
+
+% Multi-run coverage with unrestricted dropping
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\columnwidth]{figs/data_multirun_coverage.pdf}
+    \vspace{-0.2in}
+    \caption{Coverage over multiple runs for BC at 20\% overhead.}
+    \label{fig:evaluation.multirun}
+    \vspace{-0.1in}
+  \end{center}
+\end{figure}
+
+One possible application of partial monitoring with low overheads is to enable
+cooperative debugging. The idea with cooperative debugging is to enable partial
+monitoring with very low overheads across a large number of users or runs. By
+varying the pattern of partial monitoring done on each run, the goal is to
+achieve high or total coverage across all runs. Varying the monitoring that is
+done for different runs can be achieved by introducing some randomness into the
+dropping decision. Figure~\ref{fig:evaluation.multirun} shows the total
+coverage achieved over multiple runs for array bounds check running with a 20\%
+overhead budget and an unrestricted dropping policy. Here, when events are
+dropped probabilistically with the probability scaled proportionally to slack.
+That is, when slack is high, the probability of dropping is low and the
+probability increases as we run out of slack. We see that the coverage achieved
+quickly asymptotes for an unrestricted dropping policy. This is due to the fact
+that since there are multiple dependent monitoring operations that must be done
+in order for a final check to be successful, a small amount of randomness at
+each event quickly compounds to reduce the probability of certain checks to
+near zero. Instead, source dropping is much better suited for achieve high
+coverage over multiple runs.
+
 \subsection{Area and Power Overheads}

### 2014-05-30 conclusion.tex
Added conclusion.

\section{Conclusion}
\label{sec:conclusion}

Parallel run-time monitoring techniques are attractive solutions for improving
the reliability, security, and debugging capabilities of systems. In this
paper, we have presented an architecture that reduces the overheads of
monitoring through filtering and adjustable partial monitoring. This is done by
using a hardware dataflow tracking engine in order to track and filter out
monitoring for null metadata. In addition, the dataflow engine tracks invalid
metdata flows in order to enable partial monitoring without false positives.
Our experimental results show that filtering can greatly reduce the average
overheads of monitoring from 7x, 18x, and 11x for UMC, BC, and DIFT down to 4x
for UMC and BC and just 13\% for DIFT. With partial monitoring, we see that
with a 50\% overhead budget, BC can still achieve 83\% coverage and UMC can
still achieve 51\% coverage. Finally, we show that although source dropping can
achieve better coverage than unrestricted dropping when it works, it is not
able to match overhead targets as well as unrestricted dropping.

### 2014-05-30 conclusion.tex
Added conclusion.

@@
 with a 50\% overhead budget, BC can still achieve 83\% coverage and UMC can
-still achieve 51\% coverage. Finally, we show that although source dropping can
+still achieve 51\% coverage. Finally, we show that although source-only dropping can
 achieve better coverage than unrestricted dropping when it works, it is not

### 2014-05-30 evaluation.tex
Added conclusion.

@@
 In this section we evaluate the trade-offs of the three dropping policies
-presented in Section~\ref{sec:policies.events}: unrestricted dropping, source
+presented in Section~\ref{sec:policies.events}: unrestricted dropping, source-only
 dropping, and sub-flow dropping.
@@
 each event quickly compounds to reduce the probability of certain checks to
-near zero. Instead, source dropping is much better suited for achieve high
+near zero. Instead, source-only dropping is much better suited for achieve high
 coverage over multiple runs.

### 2014-05-30 policies.tex
More edits based on Ed's comments.
Added conclusion.

@@
 accumulated slack. Whenever the main core is stalled due to the monitor, slack
-is decremented.
+is decremented. This can be difficult to determine precisely due to difficult
+to measure overheads such as contention for shared memory. However, we have
+found that using only the stalls due to FIFO back pressure works well in
+practice.
 
@@
 Sub-flow dropping is a coarser-grained decision than unrestricted dropping, but
-finer-grained that source dropping. Thus, its ability to match overheads is
+finer-grained that source-only dropping. Thus, its ability to match overheads is
 expected to sit between these other two policies. Similarly, there still exists
@@
 dropping. Thus, we expect the coverage achieved by sub-flow dropping for a
-certain overhead to fall between source and unrestricted dropping. These three
+certain overhead to fall between source-only and unrestricted dropping. These three
 dropping policies create a trade-off space between coverage achieved and

### 2014-05-30 introduction.tex
Update numbers in intro/abstract.

@@
 size and semantics of their metadata, and the operations performed on metadata.
-Our experiments show that with 10\% overheads, we can still provide 49-90\% of
-coverage on average depending on the monitoring technique. By increasing the
-overhead budget, the coverage rate can be increased. Our architecture is able
-to closely meet the specified budgets. For all but one benchmark, the overheads
-seen were within 2\% of the target overhead. Our results show that X policy is
-able to achieve very close overhead matching (within X\% on average) with an
-average coverage of X\%. On the other hand, by using Y policy, the overhead
-matching is not as high (within Y\% on average) but average coverage increases
-to (Y\%).
+Our experiments show that filtering reduces the average overhead of these
+monitors from 7-18x down to 13-300\%.
+In addition, with partial monitoring, for the monitoring schemes which still
+see average overheads of 300\%, we show that with just 50\% overheads, we can still
+provide 51-83\% of coverage on average. By increasing the
+overhead budget, this coverage rate can be increased.
 

### 2014-05-30 title.tex
Update numbers in intro/abstract.

@@
 The experimental results show that filtering can significantly reduce monitoring
-overheads from XX\%-XX\% on average, depending on the monitoring scheme, to
-XX\%-XX\%. Additionally, the partial monitoring is shown to still achieve XX\% coverage 
-only with a 10\% overhead target.
+overheads from 7x-18x on average, depending on the monitoring scheme, to
+13\%-300\%. Additionally, for the monitoring schemes which still show 300\%
+average overhead, partial monitoring can reduce this overhead to 50\% while
+still achieving 51-83\% average monitoring coverage.
 

### 2014-05-30 evaluation.tex
Added data for comparing policies.

@@
 
-In this section we evaluate the trade-offs of the three dropping policies
-presented in Section~\ref{sec:policies.events}: unrestricted dropping, source-only
-dropping, and sub-flow dropping.
-
-% BC exec time
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\columnwidth]{figs/data_bc_exec_time.pdf}
-    \vspace{-0.2in}
-    \caption{Error in meeting overhead budget for array bounds check.}
-    \label{fig:evaluation.bc_exec_time}
-    \vspace{-0.2in}
-  \end{center}
-\end{figure}
-
 % UMC exec time
@@
     \label{fig:evaluation.umc_exec_time}
-    \vspace{-0.2in}
-  \end{center}
-\end{figure}
-
-First, we evaluate how well the different policies are able to meet a specified
-overhead budget. Figures~\ref{fig:evaluation.bc_exec_time} and
-\ref{fig:evaluation.umc_exec_time} show the difference between the run-time
-overheads seen and the overhead specified. These errors are averaged over all
-benchmarks. A positive value means that the
-overhead target was overshot while a negative value indicates that the
-overheads were below the specified budget.
-
-% BC coverage across policies
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\columnwidth]{figs/data_bc_coverage.pdf}
-    \vspace{-0.2in}
-    \caption{Coverage for array bounds check for different dropping policies.}
-    \label{fig:evaluation.bc_coverage}
     \vspace{-0.2in}
@@
 \end{figure}
+In this section we evaluate the trade-offs of unrestricted dropping and source-only dropping.
+Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
+overhead budget and the run-time monitoring overheads for UMC. A positive value
+means that the overhead target was overshot while a negative value indicates
+that the overhead budget was met. We see similar results for unrestricted
+dropping and source-only dropping due to the fact that UMC consists of a large
+number of independent monitoring flows.
+Figure~\ref{fig:evaluation.umc_coverage} shows the coverage of UMC for
+unrestricted dropping and source-only dropping. We see that source-dropping
+consistently achieves higher coverage than unrestricted dropping. This is due
+to the fact that some of the overheads of monitoring for unrestricted dropping
+are being spent on wasted work as discussed in Section~\ref{sec:policies.events}.
 
-Next, we evaluate the coverage achieved by the different policies.
-Figure~\ref{fig:evaluation.bc_coverage} shows the average coverage across
-benchmarks for each of the three policies with varying overhead budgets.
-Figure~\ref{fig:evaluation.umc_coverage} shows this data for uninitialized
-memory check.
+% BC exec time
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\columnwidth]{figs/data_bc_exec_time.pdf}
+    \vspace{-0.2in}
+    \caption{Error in meeting overhead budget for array bounds check.}
+    \label{fig:evaluation.bc_exec_time}
+    \vspace{-0.2in}
+  \end{center}
+\end{figure}
+
+% BC coverage across policies
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\columnwidth]{figs/data_bc_coverage.pdf}
+    \vspace{-0.2in}
+    \caption{Coverage for array bounds check for different dropping policies.}
+    \label{fig:evaluation.bc_coverage}
+    \vspace{-0.2in}
+  \end{center}
+\end{figure}
+
+Next, we evaluate these trade-offs between source-only dropping and unrestricted dropping for BC.
+Figure~\ref{fig:evaluation.bc_exec_time} shows the overhead differences for BC
+and Figure~\ref{fig:evaluation.bc_coverage} shows the coverage for BC.
+From Figure~\ref{fig:evaluation.bc_exec_time}, we see that for several
+benchmarks, source-only drop fails to meet the specified overhead target.
+The overshoot of the overhead target is quite high with an overhead difference
+of 420\% for {\tt bzip2}, 120\% for {\tt gcc}, and 1500\% for {\tt hmmer}.
+Since only array allocations are considered source events for BC, it is
+difficult for source-only dropping to match overhead targets. Although
+Figure~\ref{fig:evaluation.bc_coverage} again shows higher coverage for
+source-only dropping, this is in part due to the fact that is running with
+higher overheads.
 

### 2014-05-30 policies.tex
Remove sub-flow policy.

@@
   \vspace{-0.1in}
-  \\
-  \subfloat[sub-flow dropping]{
-    \includegraphics[width=\columnwidth]{figs/min_drop.pdf}
-    \label{fig:policies.min_drop}
-  }
+%   \\
+%   \subfloat[sub-flow dropping]{
+%     \includegraphics[width=\columnwidth]{figs/min_drop.pdf}
+%     \label{fig:policies.min_drop}
+%   }
   \end{center}
-  \vspace{-0.2in}
+  \vspace{-0.1in}
   \caption{Comparison of dropping policies. Blue (dark) nodes indicate where
@@
 
-% Slack tracking module
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\columnwidth]{figs/policy_trade_off.pdf}
-    \vspace{-0.3in}
-    \caption{Slack tracking and drop decision hardware.}
-    \label{fig:policies.trade_off}
-    \vspace{-0.2in}
-  \end{center}
-\end{figure}
-
-Another possible method is to make the dropping decisions somewhere in between
-source-only dropping and unrestricted dropping. From the graph in 
-Figure~\ref{fig:policies.dataflow_graph}, we can see that if the dropping
-decision is made at event {\tt C} then we can skip event {\tt E} with no wasted
-work. If we instead choose to perform monitoring for event {\tt C}, then we want to perform all
-monitoring on that flow. Similarly, event {\tt F} creates a decision point for
-the bottom flow that will result in no wasted work and dropping event {\tt A}
-allows the entire shown metadata flow to be dropped with no wasted work. If the
-program can be
-analyzed or profiled to identify these instructions which, if dropped, lead to
-no wasted work, then this information can be used at run-time to minimize the
-amount of wasted work. We refer to this policy as \emph{sub-flow dropping}.
-Figure~\ref{fig:policies.min_drop} shows the points were dropping decisions are
-made for this policy.
-Sub-flow dropping is a coarser-grained decision than unrestricted dropping, but
-finer-grained that source-only dropping. Thus, its ability to match overheads is
-expected to sit between these other two policies. Similarly, there still exists
-edge cases where wasted work can occur, but typically we expect much less
-wasted work than unrestricted 
-dropping. Thus, we expect the coverage achieved by sub-flow dropping for a
-certain overhead to fall between source-only and unrestricted dropping. These three
-dropping policies create a trade-off space between coverage achieved and
-ability to match an overhead target.
-Figure~\ref{fig:policies.trade_off} summarizes this trade-off between coverage
-and ability to match overhead.
+% Trade-off between source-drop and unrestricted dropping
+% \begin{figure}
+%   \begin{center}
+%     \includegraphics[width=\columnwidth]{figs/policy_trade_off.pdf}
+%     \vspace{-0.3in}
+%     \caption{Slack tracking and drop decision hardware.}
+%     \label{fig:policies.trade_off}
+%     \vspace{-0.2in}
+%   \end{center}
+% \end{figure}
+% 
+% Another possible method is to make the dropping decisions somewhere in between
+% source-only dropping and unrestricted dropping. From the graph in 
+% Figure~\ref{fig:policies.dataflow_graph}, we can see that if the dropping
+% decision is made at event {\tt C} then we can skip event {\tt E} with no wasted
+% work. If we instead choose to perform monitoring for event {\tt C}, then we want to perform all
+% monitoring on that flow. Similarly, event {\tt F} creates a decision point for
+% the bottom flow that will result in no wasted work and dropping event {\tt A}
+% allows the entire shown metadata flow to be dropped with no wasted work. If the
+% program can be
+% analyzed or profiled to identify these instructions which, if dropped, lead to
+% no wasted work, then this information can be used at run-time to minimize the
+% amount of wasted work. We refer to this policy as \emph{sub-flow dropping}.
+% Figure~\ref{fig:policies.min_drop} shows the points were dropping decisions are
+% made for this policy.
+% Sub-flow dropping is a coarser-grained decision than unrestricted dropping, but
+% finer-grained that source-only dropping. Thus, its ability to match overheads is
+% expected to sit between these other two policies. Similarly, there still exists
+% edge cases where wasted work can occur, but typically we expect much less
+% wasted work than unrestricted 
+% dropping. Thus, we expect the coverage achieved by sub-flow dropping for a
+% certain overhead to fall between source-only and unrestricted dropping. These three
+% dropping policies create a trade-off space between coverage achieved and
+% ability to match an overhead target.
+% Figure~\ref{fig:policies.trade_off} summarizes this trade-off between coverage
+% and ability to match overhead.

### 2014-05-30 related.tex
Remove sub-flow policy.

@@
 monitoring in hard real-time systems. Their architecture performs unrestricted 
-dropping in order to guarantee strict deadlines. Our system allows for a
-trade-off between choosing source-only dropping, unrestricted dropping, or sub-flow 
-dropping depending on the needs of the designer or user which has not been
-explored by these previous works.
+dropping in order to guarantee strict deadlines. 
+Our system allows either source-only dropping or unrestricted dropping to be performed.
+This allows the designer or user to choose the appropriate policy depending on
+the monitoring scheme which has not been previously explored.

### 2014-05-30 evaluation.tex
Added 10% overhead data.

@@
 
-\subsection{Coverage with Adjustable Overheads}
+\subsection{Coverage with Adjustable Partial Monitoring}
 
@@
 We see that by varying the overhead budget, the coverage achieved also varies.
-With only a 50\% overhead budget, array bounds check still
-achieves 83\% monitoring coverage on average. This overhead budget corresponds
-to just 20\% of the average overheads seen by BC after filtering only. The high coverage
+With only a 10\% overhead budget, array bounds check still
+achieves 82\% monitoring coverage on average. The high coverage
 achieved with such low overheads is due to two main effects.  The first is that
@@
 we see that varying overhead budgets enables partial monitoring. We see that
-with a 50\% overhead budget, UMC achieves 51\% monitoring coverage on average.
-Again, this overhead budget corresponds to only 20\% of the overheads seen by UMC
-with filtering only.
+with a 10\% overhead budget, UMC achieves 32\% monitoring coverage on average
+and with a 50\% budget, UMC achieves 51\% coverage.
 

### 2014-05-30 introduction.tex
Added 10% overhead data.

@@
 In addition, with partial monitoring, for the monitoring schemes which still
-see average overheads of 300\%, we show that with just 50\% overheads, we can still
-provide 51-83\% of coverage on average. By increasing the
+see average overheads of 300\%, we show that with just 10\% overheads, we can still
+provide 32-82\% of coverage on average. By increasing the
 overhead budget, this coverage rate can be increased.

### 2014-05-30 title.tex
Added 10% overhead data.

@@
 13\%-300\%. Additionally, for the monitoring schemes which still show 300\%
-average overhead, partial monitoring can reduce this overhead to 50\% while
-still achieving 51-83\% average monitoring coverage.
+average overhead, partial monitoring can reduce this overhead to 10\% while
+still achieving 32-82\% average monitoring coverage.
 

### 2014-05-31 arch.tex
Various typos/fixes.

@@
 architecture. The basic idea is to insert this dataflow-guided hardware between the main core and
-the monitoring core that will reduce the number of monitoring events that reach
-the monitoring core and thus reduce the overheads of monitoring.
+the monitoring core. This hardware reduces the number of monitoring events that reach
+the monitoring core in order to reduce monitoring overheads.
 
@@
  we need to be able to track the flow of invalid metadata. In
-order to efficiently these flows, we use a dedicated hardware
+order to efficiently track these flows, we use a dedicated hardware
 dataflow engine. A block diagram of the dataflow engine is shown in
@@
 % Flags and flag storage
-In order track both null metadata flows and invalid metadata flows, the
+In order to track both null metadata flows and invalid metadata flows, the
 dataflow engine tracks the flow of 2-bit flags. One bit is used to
@@
 value. When the monitoring core performs these metadata update operations, it
-also writes to the corresponding dataflow flag to mark the flag as non-null.
+also writes to the corresponding dataflow flag to mark the flag as non-null. In
+addition, it also re-validates the flag if it was set as invalid due to a
+dropped monitoring event.
 
@@
 bounds). For ALU operations, if both source operands are null, then the event
-can be filtered by setting the destination register as null. For load and
-store, only a single source needs to be checked to be null in order for the
+can be filtered by setting the destination register as null. For loads and
+stores, only a single source needs to be checked to be null in order for the
 event to be filtered. Similarly, the destination register or memory location
@@
 example, a scheme may use a data structure that includes several pieces of
-related metadata. On a drop, one of these pieces could be marked as invalid in
+related metadata. On a drop, one portion of the data structure could be marked as invalid in
 order to indicate that the entire structure is invalid.  Similarly, marking one

### 2014-05-31 evaluation.tex
Various typos/fixes.

@@
 
-We implemented dataflow-guided monitoring architecture by
+We implemented our dataflow-guided monitoring architecture by
 modifying the ARM version of the gem5 simulator \cite{gem5} to support parallel
@@
 memory locations that are not initialized first.  Array bounds check, as
-mentioned in Section~\ref{sec:monitoring} is a monitoring scheme that aims to
+mentioned in Section~\ref{sec:monitoring}, is a monitoring scheme that aims to
 detect buffer overflows where memory accesses go beyond the boundaries of an
@@
 metadata information. Dynamic information flow tracking is a security
-monitoring scheme,
+monitoring scheme
 which detects when information from untrusted sources is used to affect the
-program control flow. For the benchmarks we consider, we mark data read from
+program control flow (i.e., indirect control instructions). For the benchmarks we consider, we mark data read from
 files as untrusted. We implemented a multi-bit DIFT scheme which marks
@@
 We tested our system using all C benchmarks from SPECint
-CPU2006 \cite{spec2006}. We focus on the C benchmarks because of our
-modification of {\tt malloc} to support array bounds check. Although we do not
+CPU2006 \cite{spec2006}. Since our implementation of BC depends on the
+modification of {\tt malloc} to set array bounds information, we focus on the C
+SPECint benchmarks. Although we do not
 show results for the C++ benchmarks, we note that the results for UMC and DIFT
@@
 expensive, many-cycle operation. Statically allocating memory for metadata
-will reduce the execution time but requires a large upfront memory footprint.
+will reduce the execution time but requires a large upfront memory footprint
+which caused issues with our simulator.
 
@@
 reductions in overhead for all three monitoring schemes. UMC sees normalized
-execution times 2-6x with an average of 4x with null metadata filtering. For BC, normalized execution
+execution times of 2-6x with an average of 4x with null metadata filtering. For BC, normalized execution
 times drop from 18x down to 4x on average. The range of normalized execution
@@
 Finally, DIFT sees the largest reduction in overheads with only 13\% overheads
-on average after filtering with a maximum of 42\%. This is due, in part, to the fact that
+on average after filtering, with a maximum of 42\%. This is due, in part, to the fact that
 for our implementation of DIFT on SPEC
@@
 benchmarks, this propagates to relatively few instructions. Instead, if we
-targeted network or streaming applications, which have larger amounts in
+targeted network or streaming applications, which have larger amounts of
 untrusted input data, we would expect to see less filtering.
@@
 Figure~\ref{fig:evaluation.bc_sweep} shows the monitoring coverage achieved by
-array bounds check as we vary the overhead budget. We measure monitoring
-coverage monitoring check operations that are still performed as a percentage
-of the total number of monitoring check operations. This number does not
-include the number metadata update operations that are done and serves as a
-better metric for the actual coverage of the security or reliability technique.
-The overhead is measured as a percentage of total number of checks performed
-with full monitoring. The results shown here are using an unrestricted dropping
-policy.
+array bounds check as we vary the overhead budget. 
+We define \emph{monitoring coverage} as the 
+percentage of checks that are performed 
+(indirect jumps in DIFT, loads in UMC, and memory accesses in BC). 
+The metric is chosen to understand
+how likely an error/attack instance is to be detected on an individual system. 
+While we could not evaluate real errors/attacks due to difficulty in setting up
+a large number of bugs and exploits, we believe that the percentage of checks
+provides a good estimate of detection probability when errors/attacks are 
+uniformly distributed across checks.
+Note that this metric is different from the
+percentage of monitoring events that are not dropped, which includes
+non-check instructions.
 
@@
 we see that varying overhead budgets enables partial monitoring. We see that
-with a 10\% overhead budget, UMC achieves 32\% monitoring coverage on average
-and with a 50\% budget, UMC achieves 51\% coverage.
+with a 10\% overhead budget, UMC achieves 32\% monitoring coverage on average.
+By increasing this overhead budget to 50\%, UMC achieves 51\% coverage. Even
+higher coverage can be achieved by allowing higher overheads.
 
@@
     \vspace{-0.2in}
-    \caption{Error in meeting overhead budget for uninitialized memory check.}
+    \caption{Error in meeting overhead budget for uninitialized memory check for different dropping policies.}
     \label{fig:evaluation.umc_exec_time}
@@
 \end{figure}
-In this section we evaluate the trade-offs of unrestricted dropping and source-only dropping.
+In this section we evaluate the trade-offs between unrestricted dropping and source-only dropping.
 Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
@@
     \vspace{-0.2in}
-    \caption{Error in meeting overhead budget for array bounds check.}
+    \caption{Error in meeting overhead budget for array bounds check for different dropping policies.}
     \label{fig:evaluation.bc_exec_time}
@@
     \vspace{-0.2in}
-    \caption{Coverage over multiple runs for BC at 20\% overhead.}
+    \caption{Coverage over multiple runs for BC at 10\% overhead.}
     \label{fig:evaluation.multirun}
@@
 One possible application of partial monitoring with low overheads is to enable
-cooperative debugging. The idea with cooperative debugging is to enable partial
+cooperative debugging. The idea with cooperative debugging is to use partial
 monitoring with very low overheads across a large number of users or runs. By
 varying the pattern of partial monitoring done on each run, the goal is to
-achieve high or total coverage across all runs. Varying the monitoring that is
+achieve high coverage across multiple runs. Varying the monitoring that is
 done for different runs can be achieved by introducing some randomness into the
 dropping decision. Figure~\ref{fig:evaluation.multirun} shows the total
-coverage achieved over multiple runs for array bounds check running with a 20\%
-overhead budget and an unrestricted dropping policy. Here, when events are
+coverage achieved over multiple runs for array bounds check running with a 10\%
+overhead budget and an unrestricted dropping policy. Each run was simulated for
+100 million instructions. For these simulations, events were
 dropped probabilistically with the probability scaled proportionally to slack.
@@
 each event quickly compounds to reduce the probability of certain checks to
-near zero. Instead, source-only dropping is much better suited for achieve high
+near zero. Instead, source-only dropping is much better suited for achieving high
 coverage over multiple runs.
@@
 
-Adding the dataflow engine in order to enable filtering and adjustable adjustable adds
+Adding the dataflow engine in order to enable filtering and partial monitoring adds
 overheads in terms of area and power. We use McPat \cite{mcpat-micro09} to get
@@
 corresponding memory and ALU objects in McPat. We
-note that this is only a rough area and power result since components such as the
+note that this is only a rough area and power estimate since components such as the
 wires connecting these modules have not been modeled. However, this gives a
@@
 a 50\% monitoring overhead target. The peak power is 5-7 mW, which is
-5-7\% of the main core's peak power usage. The average runtime power is 3-4 
-mW, corresponding to 8-11\% of the main core's runtime power.
+less than 1\% of the main core's peak power usage. Similarly, the average runtime power is 3-4 
+mW, corresponding to about 1\% of the main core's runtime power.

### 2014-05-31 introduction.tex
Various typos/fixes.

@@
 Section~\ref{sec:optimizations} presents the high-level approach of how we aim
-to reduce overheads through and Section~\ref{sec:arch} describes our
+to reduce overheads through filtering and partial monitoring and Section~\ref{sec:arch} describes our
 hardware architecture that supports this. Section~\ref{sec:policies}

### 2014-05-31 optimizations.tex
Various typos/fixes.

@@
 optimizations. The first optimization is to filter out monitoring events which relate to
-null metadata which do not need to processed by the monitoring core (Section~\ref{sec:optimizations.filter}). Second, in order to allow
+null metadata which do not need to be processed by the monitoring core (Section~\ref{sec:optimizations.filter}). Second, in order to allow
 further reduction in overheads, we propose to perform partial monitoring. This
@@
 raised.  We accept false negatives as the
-loss in coverage that we trade off in order to limit overheads.  However, we
+loss in coverage that we trade-off in order to limit overheads.  However, we
 must safely drop monitoring events in such a way as to avoid false positives so

### 2014-05-31 policies.tex
Various typos/fixes.

@@
 Section~\ref{sec:policies.slack}. Next, we investigate trade-offs created by
-selecting which events are dropped. Different policies in determining which
+selecting which events are dropped (Section~\ref{sec:policies.events}). Different policies in determining which
 events can be dropped create a trade-off between how closely the overhead
@@
 accumulated slack. Whenever the main core is stalled due to the monitor, slack
-is decremented. This can be difficult to determine precisely due to difficult
-to measure overheads such as contention for shared memory. However, we have
-found that using only the stalls due to FIFO back pressure works well in
-practice.
+is decremented. This can be difficult to determine precisely due to the
+difficulty in measuring certain overheads such as contention for shared memory.
+However, we have found that using only the stalls due to FIFO back pressure
+works well in practice.
 
@@
 
-One method to eliminate this wasted work is to only make dropping decisions at
+An alternative dropping policy which eliminates this wasted work is to only make dropping decisions at
 the root of these metadata flows. That is, we will decide to either monitor or
@@
 result in no wasted work and thus better coverage. However, because of the coarser-grained decision, it
-may be more difficult to closely match overheads. Source dropping is expected
+may be more difficult to closely match overheads. We expect source dropping 
 to work well when there are a large number of independent metadata flows.

### 2014-05-31 evaluation.tex
Minor edits to evalutation.
Include reduction percentage for filtering.

@@
 reductions in overhead for all three monitoring schemes. UMC sees normalized
-execution times of 2-6x with an average of 4x with null metadata filtering. For BC, normalized execution
-times drop from 18x down to 4x on average. The range of normalized execution
+execution times of 2-6x with an average of 4x with null metadata filtering.
+This is an average reduction in overheads of 52\%. For BC, normalized execution
+times drop from 18x down to 4x on average (an 82\% reduction). The range of normalized execution
 times for filtered BC is 1.1-17x. This is not surprising since in the baseline
@@
 Finally, DIFT sees the largest reduction in overheads with only 13\% overheads
-on average after filtering, with a maximum of 42\%. This is due, in part, to the fact that
+on average after filtering, a 90\% reduction in monitoring overheads, with a
+maximum overhead of 42\% after filtering. This is due, in part, to the fact
+that
 for our implementation of DIFT on SPEC
@@
     \vspace{-0.2in}
-    \caption{Coverage over multiple runs for BC at 10\% overhead.}
+    \caption{Coverage over multiple runs for BC at 10\% overhead with an unrestricted dropping policy.}
     \label{fig:evaluation.multirun}

### 2014-08-20 micro14/rebuttal.txt
Draft of MICRO 2014 rebuttal.

Review-A

-How is metadata dependence graph encoded and processed?
The full graph structure does not need to be stored. The dataflow engine
dynamically propagates the information about whether metadata is null or
invalid. Source-only dropping is enabled by only making dropping decisions on
explicit instructions which modify metadata. This does not require information
about the full dependence graph to identify.

Review-B

-Previous work has removed null-metadata (e.g., FlexiTaint).
As we mention in the paper, we are not the first work to filter out null
metadata. However, we show how a single dataflow engine can be used to provide
both this filtering mechanism as well a dropping mechanism to further reduce
overheads.

-FlexCore allows an instruction type to be not forwarded or not forwarded if
the FIFO is full.
Our implementation also filters out instruction types which are not
relevant to the monitor. Not forwarding instructions if the FIFO is full can
lead to false positives which FlexCore does not prevent.

-Baseline overheads are high. In addition, FlexCore shows much lower overheads.
One reason for our high baseline overheads is due to the need to dynamically
allocate memory space for metadata due to the large metadata we are using.
FlexCore shows much lower overheads by performing monitoring on a
reconfigurable fabric. This requires dedicated on-chip reconfigurable
fabric for this purpose. Our evaluation focuses on a core-based monitor but our
techniques could also be applied to FlexCore to further reduce its overheads.

-Information flow tracking can have false positives and negatives (e.g.,
implicit flows). Is the effect of these examined?
In this work we have focused on the false positives and negatives that arise
from dropping a portion of the information flow tracking operations. We have
not looked at the inherent false positives and negatives that arise with
specific information flow tracking schemes, such as handling implicit flows.

-Why use extra HW + extra core as opposed to a programmable HW unit?
We focused our evaluation on a core-based monitor as this has greater
flexibility and requires less hardware changes to implement.
However, our techniques could also be applied to hardware monitors such as
FlexiTaint and FlexCore to enable reduced overheads by filtering and
dropping monitoring events.

Review-C

-Why choose security monitoring instead of applications like profiling which
 can tolerate partial coverage?
Profiling would also serve as a good application for our system and we
could study such a scheme on our system in the future. We chose to focus
on security monitoring as the issues with targeting security monitoring
were a superset of those for most profiling monitors (e.g., dealing with false
positives).

Review-D

-Core idea of Section 2 (monitoring) is same as Heapmon.
Section 2 describes our baseline run-time monitoring architecture.  Our work
extends previous work on run-time monitoring, such as Heapmon and others, by
filtering out monitoring on null metadata and dropping monitoring events to
further reduce overheads.

-Claiming false negatives can be ignored undermines the soundness of the
 monitoring. Can you afford for tainted variable to be used by untrusted
 function?
Although full monitoring is always desired, if the performance overheads
of full monitoring cannot be tolerated, then false negatives are a
necessary trade off. The alternative would be to completely disable
monitoring (i.e., 100% false negatives).

-Claim that work can emulate all prior works on memory debugging and control
 flow integrity check. How can your scheme emulate every single one of them?
As the monitor is implemented in software on a core, it can support a wide
range of monitoring operations. Specifically, any monitoring technique
which performs operations in response to instruction executions should be
implementable.

Review-E

-How much additional benefit does data-flow filtering provide over
 instruction-type filtering (FlexCore)?
Like FlexCore, our system also filters out monitoring for instruction
types which are not relevant to a monitoring technique. The results we
show are the additional benefit of our technique on top of
instruction-type filtering.

-Baseline selected based on what runs on simulator instead of
 state-of-the-art (dynamic allocation due to memory issues with simulator). 
Monitoring tags can require a large memory space. This is a problem not
unique to our baseline, though it is not always addressed. We have chosen to
deal with this by performing dynamic allocation of memory for metadata rather
than statically allocating a large memory space for metadata (up to 4 GB).
This comes at a trade-off in overheads. UMC, which has smaller metadata, was
implemented by statically allocating the metadata memory and thus does not
see as high overheads as BC and DIFT. A dedicated metadata TLB could be used
to reduce these overheads.

-Processor with hardware support for information flow tracking and support
 for configurable policies can be re-purposed to provide same functionality as
 the technique proposed in the paper. 
Our technique for using a dataflow engine for filtering and dropping metadata
can be applied to both core-based monitors and dedicated hardware monitors such
as a hardware-supported information flow tracking mechanism. For a sufficiently
configurable information flow track module, it may be possible to incorporate
this into the information flow tracking policy.  However, as far as we know, we
are the first to propose using this flow tracking mechanism to provide both
filtering and dropping of monitoring events to reduce monitoring overheads. In
addition, there are monitoring schemes that may not be possible to be mapped to
information flow tracking (e.g., control flow integrity checking, hardware
error detection, etc.) in which case a separate dataflow engine, such as the
one we describe, would be needed to enable filtering and dropping.

### 2014-08-21 micro14/rebuttal.txt
Update of MICRO14 rebuttal.

@@
+General
+
+-How does this work compared to previous work (FlexiTaint, FlexCore, Heapmon,
+etc.)?
+Run-time monitoring is a well studied problem. However, our work is the first
+that proposes a general hardware platform that allows for an adjustable
+overhead target by trading off coverage for overhead.
+
 Review-A
@@
 -How is metadata dependence graph encoded and processed?
-The full graph structure does not need to be stored. The dataflow engine
-dynamically propagates the information about whether metadata is null or
-invalid. Source-only dropping is enabled by only making dropping decisions on
-explicit instructions which modify metadata. This does not require information
-about the full dependence graph to identify.
+Sources are marked statically in the program by the compiler or programmer
+annotation. The dataflow engine stores 2 bits per metadata indicating whether
+it is null or invalid. It dynamically propagates this information as
+instructions execute.
 
@@
 
--Previous work has removed null-metadata (e.g., FlexiTaint).
-As we mention in the paper, we are not the first work to filter out null
-metadata. However, we show how a single dataflow engine can be used to provide
-both this filtering mechanism as well a dropping mechanism to further reduce
+-How does this compare to FlexCore's dropping of events?
+We ensure that false positives do not occur when events are dropped. FlexCore
+does not prevent these false positives.
+
+-Baseline overheads are high.
+Our overheads are comparable to previous work such as [Newsome NDSS'05]
+(24-37x) and [Chen ASID'06] (4-10x). More recent works have aggressively
+optimized these monitors to reduce overheads, though these overheads could
+still be considered too high to enable monitoring in certain contexts. These
+optimizations could be applied to our system as well, resulting in higher
+coverage at the same target overhead or comparable coverage at even lower
 overheads.
-
--FlexCore allows an instruction type to be not forwarded or not forwarded if
-the FIFO is full.
-Our implementation also filters out instruction types which are not
-relevant to the monitor. Not forwarding instructions if the FIFO is full can
-lead to false positives which FlexCore does not prevent.
-
--Baseline overheads are high. In addition, FlexCore shows much lower overheads.
-One reason for our high baseline overheads is due to the need to dynamically
-allocate memory space for metadata due to the large metadata we are using.
-FlexCore shows much lower overheads by performing monitoring on a
-reconfigurable fabric. This requires dedicated on-chip reconfigurable
-fabric for this purpose. Our evaluation focuses on a core-based monitor but our
-techniques could also be applied to FlexCore to further reduce its overheads.
 
@@
 implicit flows). Is the effect of these examined?
-In this work we have focused on the false positives and negatives that arise
-from dropping a portion of the information flow tracking operations. We have
-not looked at the inherent false positives and negatives that arise with
-specific information flow tracking schemes, such as handling implicit flows.
+The information flow tracking scheme that we examined only tracked explicit
+flows. Monitoring schemes which track implicit flows may require additional
+support to ensure no false positives on our hardware platform.
 
 -Why use extra HW + extra core as opposed to a programmable HW unit?
-We focused our evaluation on a core-based monitor as this has greater
-flexibility and requires less hardware changes to implement.
-However, our techniques could also be applied to hardware monitors such as
-FlexiTaint and FlexCore to enable reduced overheads by filtering and
-dropping monitoring events.
+We evaluated our idea on a core-based monitor as this has greater flexibility
+and requires less hardware changes to implement. However, our techniques could
+also be applied to hardware monitors, such as FlexiTaint and FlexCore, to enable
+adjustable overheads.
 
@@
  can tolerate partial coverage?
-Profiling would also serve as a good application for our system and we
-could study such a scheme on our system in the future. We chose to focus
-on security monitoring as the issues with targeting security monitoring
-were a superset of those for most profiling monitors (e.g., dealing with false
-positives).
+We focused on security monitoring due to the challenges that arise with false
+positives for these monitoring techniques. Our platform could also be used with
+profiling applications in order to enable reduced and adjustable overheads.
 
@@
 
--Core idea of Section 2 (monitoring) is same as Heapmon.
-Section 2 describes our baseline run-time monitoring architecture.  Our work
-extends previous work on run-time monitoring, such as Heapmon and others, by
-filtering out monitoring on null metadata and dropping monitoring events to
-further reduce overheads.
-
--Claiming false negatives can be ignored undermines the soundness of the
- monitoring. Can you afford for tainted variable to be used by untrusted
- function?
+-Claiming false negatives can be ignored.
 Although full monitoring is always desired, if the performance overheads
@@
 
--Claim that work can emulate all prior works on memory debugging and control
- flow integrity check. How can your scheme emulate every single one of them?
+-How can your scheme emulate other monitoring techniques?
 As the monitor is implemented in software on a core, it can support a wide
-range of monitoring operations. Specifically, any monitoring technique
-which performs operations in response to instruction executions should be
-implementable.
+range of monitoring operations. Specifically, any instruction-grained
+monitoring technique should be implementable.
 
@@
 -Baseline selected based on what runs on simulator instead of
- state-of-the-art (dynamic allocation due to memory issues with simulator). 
-Monitoring tags can require a large memory space. This is a problem not
-unique to our baseline, though it is not always addressed. We have chosen to
-deal with this by performing dynamic allocation of memory for metadata rather
-than statically allocating a large memory space for metadata (up to 4 GB).
-This comes at a trade-off in overheads. UMC, which has smaller metadata, was
-implemented by statically allocating the metadata memory and thus does not
-see as high overheads as BC and DIFT. A dedicated metadata TLB could be used
-to reduce these overheads.
+ state-of-the-art (dynamic allocation of memory).
+Dynamic allocation of metadata memory is required for the large metadata used
+by the monitors we evaluated (up to 4 GB total memory required). Static
+allocation may be possible for systems with large memory space, but was not
+reasonable for our target system.
 
--Processor with hardware support for information flow tracking and support
- for configurable policies can be re-purposed to provide same functionality as
- the technique proposed in the paper. 
-Our technique for using a dataflow engine for filtering and dropping metadata
-can be applied to both core-based monitors and dedicated hardware monitors such
-as a hardware-supported information flow tracking mechanism. For a sufficiently
-configurable information flow track module, it may be possible to incorporate
-this into the information flow tracking policy.  However, as far as we know, we
-are the first to propose using this flow tracking mechanism to provide both
-filtering and dropping of monitoring events to reduce monitoring overheads. In
-addition, there are monitoring schemes that may not be possible to be mapped to
-information flow tracking (e.g., control flow integrity checking, hardware
-error detection, etc.) in which case a separate dataflow engine, such as the
-one we describe, would be needed to enable filtering and dropping.
+-Processor with hardware support for information flow tracking can provide the
+same functionality.
+The main focus of our platform is to enable reduced and adjustable overheads
+which has not been studied by previous works on hardware support for
+information flow tracking. The monitoring techniques we evaluated could be
+implemented with a sufficiently configurable information flow tracking
+mechanism. However, there are other monitors (e.g., control flow integrity
+checking, hardware error detection, etc.) which our technique could also be
+applied to.

### 2014-08-22 micro14/rebuttal.txt
updated rebuttal

@@
 overhead target by trading off coverage for overhead.
+While we use a core-based monitor to demonstrate techniques, our platform
+is applicable to other types of monitors such as FlexCore.
 
@@
 We ensure that false positives do not occur when events are dropped. FlexCore
-does not prevent these false positives.
+does not prevent these false positives. Our dropping is also more intelligent
+selectively dropping based on the overhead target and importance.
 
@@
 (24-37x) and [Chen ASID'06] (4-10x). More recent works have aggressively
-optimized these monitors to reduce overheads, though these overheads could
-still be considered too high to enable monitoring in certain contexts. These
-optimizations could be applied to our system as well, resulting in higher
-coverage at the same target overhead or comparable coverage at even lower
-overheads.
+optimized these monitors to reduce overheads. Yet, the overhead especially
+for monitors with large tags can still be significant.
+The dataflow engine can be applied to the optimized monitors to further
+reduce the overhead or enable coverage at a lower overhead target.
 
--Information flow tracking can have false positives and negatives (e.g.,
-implicit flows). Is the effect of these examined?
-The information flow tracking scheme that we examined only tracked explicit
-flows. Monitoring schemes which track implicit flows may require additional
+-The dataflow engine does not track implicit flows. Can this lead to
+false positives or negatives?
+The monitors that we examined only tracked explicit flows. As a results,
+the dataflow engine only needed to track explicit flows.
+Monitoring schemes which track implicit flows may require additional
 support to ensure no false positives on our hardware platform.
@@
 -Why use extra HW + extra core as opposed to a programmable HW unit?
-We evaluated our idea on a core-based monitor as this has greater flexibility
-and requires less hardware changes to implement. However, our techniques could
+We evaluated our idea on a core-based monitor as this represents the most
+flexible design and leverages multi-cores with less hardware changes.
+However, our techniques could
 also be applied to hardware monitors, such as FlexiTaint and FlexCore, to enable
@@
  can tolerate partial coverage?
-We focused on security monitoring due to the challenges that arise with false
-positives for these monitoring techniques. Our platform could also be used with
-profiling applications in order to enable reduced and adjustable overheads.
+We agree that profiling is a promising application for our platform with
+adjustable overhead.
+We focused on monitors for program bugs because they address important problems
+and represent a general case where false positives need to be handled.
 
@@
 As the monitor is implemented in software on a core, it can support a wide
-range of monitoring operations. Specifically, any instruction-grained
-monitoring technique should be implementable.
+range of monitoring operations. Specifically, any technique that monitors
+a trace of instructions should be implementable.
 
@@
  instruction-type filtering (FlexCore)?
-Like FlexCore, our system also filters out monitoring for instruction
+Like FlexCore, our baseline system filters out monitoring for instruction
 types which are not relevant to a monitoring technique. The results we
@@
  state-of-the-art (dynamic allocation of memory).
-Dynamic allocation of metadata memory is required for the large metadata used
-by the monitors we evaluated (up to 4 GB total memory required). Static
-allocation may be possible for systems with large memory space, but was not
-reasonable for our target system.
+We implemented dynamic metadata allocation because it is required for monitors
+with large metadata (for example, 64-bits per word in bound checking or
+32-bit per byte for multi-bit DIFT). We also tested static allocation for 
+smaller metadata.
 
@@
 mechanism. However, there are other monitors (e.g., control flow integrity
-checking, hardware error detection, etc.) which our technique could also be
-applied to.
+checking, hardware error detection, race detection, etc.) that cannot be 
+implemented as DIFT but can use our technique.

### 2014-08-22 micro14/rebuttal.txt
Minor edits to rebuttal.

@@
 that proposes a general hardware platform that allows for an adjustable
-overhead target by trading off coverage for overhead.
-While we use a core-based monitor to demonstrate techniques, our platform
-is applicable to other types of monitors such as FlexCore.
+overhead target by trading off coverage for overhead.  While we use a
+core-based monitor to demonstrate techniques, our platform is applicable to
+other types of monitors such as FlexCore.
 
@@
 does not prevent these false positives. Our dropping is also more intelligent
-selectively dropping based on the overhead target and importance.
+by selectively dropping based on the overhead target and importance.
 
@@
 (24-37x) and [Chen ASID'06] (4-10x). More recent works have aggressively
-optimized these monitors to reduce overheads. Yet, the overhead especially
-for monitors with large tags can still be significant.
-The dataflow engine can be applied to the optimized monitors to further
-reduce the overhead or enable coverage at a lower overhead target.
+optimized these monitors to reduce overheads. Yet, the overhead, especially for
+monitors with large tags, can still be significant. Our dataflow engine can be
+applied to the optimized monitors to further reduce the overhead or enable
+coverage at a lower overhead target.
 
@@
 We evaluated our idea on a core-based monitor as this represents the most
-flexible design and leverages multi-cores with less hardware changes.
-However, our techniques could
-also be applied to hardware monitors, such as FlexiTaint and FlexCore, to enable
-adjustable overheads.
+flexible design and leverages multi-cores with less hardware changes. However,
+our techniques could also be applied to hardware monitors, such as FlexiTaint
+and FlexCore, to enable adjustable overheads.
 
@@
 We agree that profiling is a promising application for our platform with
-adjustable overhead.
-We focused on monitors for program bugs because they address important problems
-and represent a general case where false positives need to be handled.
+adjustable overhead. We focused on monitors for program bugs because they
+address important problems and represent a general case where false positives
+need to be handled.
 
@@
 with large metadata (for example, 64-bits per word in bound checking or
-32-bit per byte for multi-bit DIFT). We also tested static allocation for 
+32-bits per byte for multi-bit DIFT). We also tested static allocation for 
 smaller metadata.

### 2014-09-05 dropping.tex
Reorganized paper for HPCA 2015 submission.

\section{Adjustable Monitoring Infrastructure}
\label{sec:dropping}

In this section, we present an architecture that can limit the amount of
monitoring performed in order to meet a target overhead budget. We consider the
overhead budget as a soft constraint that we seek to stay below or close to,
but do not require strict guarantees that it is met. For this paper, we focus on
the performance overhead of monitoring, though other overheads could also be
considered.

% Overview of full architecture
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/architecture_overview.pdf}
    \vspace{-0.2in}
    \caption{Block diagram of architecture for monitoring with reduced and adjustable overheads.}
    \label{fig:dropping.overview}
    \vspace{-0.1in}
  \end{center}
\end{figure}

Figure~\ref{fig:dropping.overview} shows a high-level block diagram of our
proposed architecture for enabling partial monitoring. The basic idea is to
insert hardware between the main core and the monitoring core to transparently
drop monitoring events.

There were three main challenges in designing this dropping hardware:
\begin{enumerate}
  \item \textbf{General:} The hardware needs to be applicable to a wide range of monitoring schemes.
  \item \textbf{No false positives:} False positives should never occur as a result of dropping.
  \item \textbf{Intelligent dropping:} The hardware needs to maximize the amount of useful monitoring done while staying within the overhead budget.
\end{enumerate}

In order to reduce overheads, the ``Drop'' module decides when monitoring event
should be skipped.  This decision of when to drop and which monitoring events
to dropped is discussed in more detail in Section~\ref{sec:policies}.

When an event is dropped, metadata needs to be invalidated in order to prevent
false positives. This is done using a hardware dataflow engine.
Sections~\ref{sec:dropping.false_neg_pos} and
\ref{sec:dropping.prevent_false_pos} discuss these false positive issues and
the dataflow engine in detail.

In addition, we can use the dataflow engine to filter out monitoring operations
that would be operating on invalid metadata. This maximizes the useful
monitoring that is done by the monitoring core.
Section~\ref{sec:dropping.filter} talks about how this is done using
information from the dataflow engine.

\subsection{False Negatives and False Positives}
\label{sec:dropping.false_neg_pos}

% Example of dropping
\begin{figure*}
  \begin{center}
  \subfloat[no error] {
    \includegraphics[trim=1.6in 0 0 0,clip=true]{figs/example_no_error.pdf}
    \label{fig:dropping.example_no_error}
  }
  \hspace{0.2in}
  \subfloat[false negative] {
    \includegraphics[trim=1.6in 0 0 0,clip=true]{figs/example_false_negative.pdf}
    \label{fig:dropping.example_false_negative}
  }
  \hspace{0.2in}
  \subfloat[false positive] {
    \includegraphics[trim=1.6in 0 0 0,clip=true]{figs/example_false_positive.pdf}
    \label{fig:dropping.example_false_positive}
  }
  \caption{Dropping}
  \label{fig:dropping.example_dropping}
  \end{center}
\end{figure*}

Our goal is to drop some monitoring operations in order to reduce the overheads
of run-time monitoring. This dropping can affect the functionality of the
monitoring schemes. There are three possible outcomes for dropping a
monitoring operation. The first is that there is no difference in operation
from the original execution. For example, if we drop line 3 from our array
bounds check example (see Figure~\ref{fig:dropping.example_no_error}), then the
check on accessing {\tt x+7} is skipped. However, this is a valid access and so
skipping the check does not change anything.

On the other hand, if the monitoring for accessing memory location {\tt y+7} on
line 4 is skipped (see Figure~\ref{fig:dropping.example_false_negative}), then
a false negative occurs. Originally, the monitor would catch this access as an
out-of-bounds access and raise an error. However, if the monitoring operation
for this is dropped, then the error is not detected. This is the trade-off that
we make in order to reduce overheads. That is, instead of either 100\% coverage
with all the associated overheads or no coverage and no overheads, we 
enable middle points of partial coverage with some fraction of the full
overheads.

The final possible outcome of dropping a monitoring operation is shown in
Figure~\ref{fig:dropping.example_dropping}. In this case, the monitoring for
line 1 is dropped, causing the bound information for pointer {\tt x} to never
be set. The result is that when the access using {\tt x} is checked on line 3,
an error will be raised. This creates a false positive where an error is
incorrectly raised. Although false negatives are part of the trade-off we make
to reduce overheads, we need to prevent false positives.

\subsection{Invalidation for Preventing False Positives}
\label{sec:dropping.prevent_false_pos}

In analyzing various monitoring schemes, we found that monitoring operations
are primarily of two types: \emph{checks} and \emph{metadata updates}. Monitors
\emph{check} certain properties to ensure correct main program execution (lines
3-4 of Figure~\ref{fig:monitoring.example_full}) and they \emph{update} metadata
for bookkeeping (lines 1-2 of Figure~\ref{fig:monitoring.example_full}). Skipping
a check operation can only cause false negatives and will never cause a false
positive (see Figure~\ref{fig:dropping.example_no_error} and
Figure~\ref{fig:dropping.example_false_negative}). Therefore, we may simply
skip a check operation. Skipping an update operation can cause false
negatives but may also cause false positives (see
Figure~\ref{fig:dropping.example_false_positive}).

% Example with invalidation
\begin{figure*}
  \begin{center}
    \includegraphics[]{figs/example_invalid.pdf}
    \caption{Invalidation flags are used to prevent false positives.}
    \label{fig:dropping.example_invalid}
    \vspace{-0.1in}
  \end{center}
\end{figure*}

Essentially, when an update operation is skipped, we can no longer trust the
corresponding metadata.  Thus, our approach is to mark these metadata as
invalid. Furthermore, metadata that would be derived from these dropped events
should also be marked as invalid. Figure~\ref{fig:dropping.example_invalid}
shows an example of how associating an invalid flag with metadata can be used
to avoid false positives.  In this case, when the monitoring operation to set
the metadata on {\tt x} is dropped, the metadata for {\tt x} is also marked as
invalid. The result is that when the metadata for {\tt x} is checked on line 3,
no error is raised since the metadata is invalid. Note that in this case, the
check on line 4 is also skipped. This is necessary because an error would have
been raised even if the access was in bounds.

\subsection{Dataflow Engine for Preventing False Positives}
\label{sec:dropping.arch}

% Dataflow engine high-level
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/dataflow_overview.pdf}
    \caption{Hardware support for dropping.}
    \label{fig:dropping.dataflow_overview}
    \vspace{-0.1in}
  \end{center}
\end{figure}

In order to efficiently support dropping monitoring events and to prevent false
positives, we propose to insert a hardware module between the main and
monitoring core (see Figure~\ref{fig:dropping.dataflow_overview}). 
This module performs dataflow tracking on invalidation information in
order to prevent false positives. In addition, checks on invalid metadata are not useful and
thus are filtered out by the hardware module before reaching the monitor.
Similarly, updating metadata based on metadata that is invalid is also not a
productive use of the monitor. Thus, in order to save the use of the monitor
for useful operations, the hardware module also filters out update operations
based on invalid metadata.

% Detailed architecture of dataflow engine
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/dataflow_architecture.pdf}
    \vspace{-0.2in}
    \caption{Hardware architecture of the dataflow engine.}
    \label{fig:dropping.dataflow} 
    \vspace{-0.1in}
  \end{center}
\end{figure*}

Figure~\ref{fig:dropping.dataflow} shows a detailed block diagram of the
hardware module inserted between the main core and the monitoring. 
Monitoring schemes typically keep metadata corresponding to the main core's
register file as well as the main core's memory space. Similarly, the dataflow
engine uses an on-chip register file to store flags for register file metadata.
In addition, a set of flags are stored that correspond to memory metadata.
These memory metadata flags are accessed through a cache and backed to main
memory. All flags are initialized as valid when the system starts.

% Address calculation and configuration
On a monitoring event, the dataflow engine reads in up to two flags in order to
determine whether the event can be filtered.  The flags to be read in typically correspond to the source
operands of the monitored instruction. For example, on a load, the dataflow
engine will typically read in the flags corresponding the memory location
accessed.
However, the architecture is designed such that the address of flags to be read
is calculated dynamically and can be configured depending on the monitoring
scheme. Specifically, there exists a configuration table that outputs a set of
control signals depending on the instruction type of the monitoring event.
These control signals are used to control a pair of address calculation units which each
contain a simple ALU. The inputs to the ALU are information from the monitored
event and a constant that is specified from the configuration table. One common
use of this address calculation is to transform addresses from byte-addressed
to word-addressed by right shifting the passed memory address by 2 bits.

\subsection{Filtering Invalid Events}
\label{sec:dropping.filter}

% Filtering decision
The decision of whether an event can be filtered out is determined using a
lookup table. This filter decision table is configured by the monitoring scheme
on system initialization. The lookup table determines whether an event can be
filtered out based on the pair of input flags read and the instruction type of
the monitored event.
For example, on an ALU instruction, if both of the source operand flags
indicate that the corresponding metadata is null, then typically the instruction can be
filtered out. Similarly, on an ALU instruction, if either of the source operand
flags indicate that their corresponding metadata is invalid, then the
instruction can also be filtered out. This filter decision is sent to the
``Filter'' block shown in Figure~\ref{fig:dropping.overview} which will simply pop
filtered entries from the FIFO and not forward them. If an entry is not
filtered, then this Filter block will pop the monitoring event entry from the
FIFO and forward it along the datapath.

% Propagating flag information
Finally, when an event is filtered out, it is necessary to propagate the flag
information for the filtered event. In addition to the decision of whether to
filter an event, the filter decision table also includes information about
whether the destination flag should be written to and what value should be
written to it. A third address calculation unit generates the write address for
this destination flag. For example, for an ALU instruction, this will
correspond to the destination register of the ALU instruction. If this ALU
instruction is filtered out due to having a pair of null source operands, then
the filter decision table will indicate that the destination register's
corresponding flag should also be marked as null.

% Flags are marked as invalid when a monitoring event is dropped due to
% insufficient overhead budget. Thus, whenever a monitoring event is dropped, the
% dropping hardware indicates this to the dataflow engine. The dataflow engine's
% write address is configured for the appropriate destination flag. When it
% receives an invalidation signal from the dropping hardware, the
% dataflow engine marks this destination flag as invalid.

### 2014-09-05 evaluation.tex
Reorganized paper for HPCA 2015 submission.

@@
 \subsection{Baseline Monitoring Overheads}
+
 % Full monitoring overheads
@@
   \end{center}
-\end{figure} Figure~\ref{fig:evaluation.full_mon} shows the execution times of
+\end{figure}
+
+Figure~\ref{fig:evaluation.full_mon} shows the execution times of
 performing full monitoring normalized to the execution times of the benchmarks

### 2014-09-05 filtering.tex
Reorganized paper for HPCA 2015 submission.

\section{Filtering Null Metadata}
\label{sec:filtering}

One way to reduce the number of monitoring events that must be handled by the
monitor is to filter out monitoring events that correspond to operating on null
or uninitialized metadata. Null metadata typically corresponds to events that
are not relevant to the monitor. For example, Hardbound
\cite{hardbound-asplos08} filters out operations on non-pointer (i.e., no base
and bounds metadata) instructions since it is not relevant to array bounds
checking. More recently, FADE \cite{fade-hpca14} has been proposed a general
hardware module to perform this null metadata filtering for a variety of
monitoring schemes.

We enable this null metadata filtering by
simply extending the dataflow flags to be two bits wide in order to keep track
of both validity due to dropping and null information.
All flags are initialized to valid and null. Whenever the monitor sets metadata
to a non-null value, it will also mark the corresponding flag in the dataflow
engine as non-null. In addition to checking validity when reading flags, the
flags are also checked for whether the metadata is null. An event only
continues to the monitor if all source flags are valid and any source flag is
marked as non-null.

### 2014-09-05 introduction.tex
Reorganized paper for HPCA 2015 submission.

--- 2014-05-30 introduction.tex
@@
 % Two optimizations
-In this paper, we present a hardware architecture that uses dynamic dataflow
-tracking in order to reduce the overheads of programmable parallel monitoring
-by filtering out unnecessary monitoring operations and by enabling partial
-monitoring.
+Since these overheads can be prohibitive for run-time use, we propose to enable
+partial monitoring in order to reduce the overheads.
 
-% Filtering
-Recent studies have shown that filtering out monitoring checks on clean or null
-metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}.
-For example, for
-an array bounds check, checking non-pointer accesses is unnecessary. 
-In this paper, we use a hardware-based dataflow tracking engine to track and
-filter out these null metadata monitoring operations.
-By using the dataflow engine to track a separate set of flags
-indicating null metadata, we are able to handle the common case of null
-metadata without the need to read the full metadata values.
-
-There is a limit to how effective filtering is at reducing monitoring overheads
-as it only eliminates unnecessary monitoring operations. The ability
-to also skip relevant monitoring operations can be used to further reduce overheads.
-Although this comes at the cost of reduced monitoring coverage (i.e.,
-effectiveness), this partial monitoring can still be very useful.
+Although partial monitoring comes at the cost of reduced monitoring coverage (i.e.,
+effectiveness), this can still be very useful.
 For example, partial
@@
 
-% % Adjustable overhead
-In this paper, we show how our dataflow-guided architecture is also able to
-enable partial monitoring. This is achieved by dynamically dropping monitoring
-operations when overheads exceed a specified overhead budget. We prevent false
-positives from partial monitoring by using the dataflow engine to track these
-dropped monitoring flows. With this architecture, we investigate different
-policies for deciding which events to drop for partial monitoring. These
-different policies show a trade-off between closely matching the overhead
-budget and the monitoring coverage achieved.
+% Adjustable overhead
+Partial monitoring with adjustable overheads is achieved by dynamically
+dropping monitoring operations when overheads exceed a specified overhead
+budget.  In order to prevent dropped monitoring operations from leading to a
+false positive, we need to track these dropped flows.  We present a
+hardware-based dataflow tracking engine that keeps track of these dropped
+flows. With this architecture, we investigate different policies for deciding
+which events to drop for partial monitoring. These different policies show a
+trade-off between closely matching the overhead budget and the monitoring
+coverage achieved.
 
-The main contributions of this paper are summarized as follows.
+% Filtering
+Recent studies have shown that filtering out monitoring checks on clean or null
+metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}.
+For example, for an array bounds check, checking non-pointer accesses is
+unnecessary. In addition to enabling partial monitoring, we show how a simple
+extension to our dataflow engine enables this null metadata filtering.
+
+The main contributions of this paper are summarized as follows:
 \begin{itemize}
-  \item We propose two optimizations for reducing the overheads of monitoring: filtering null metadata and performing partial monitoring.
   \item We present a hardware architecture using a dataflow tracking engine to
-  efficiently enable filtering and partial monitoring. This hardware
-  architecture is designed to be generally applicable to a wide-range of run-time monitoring
-  techniques.
-  \item We investigate interesting trade-off decisions for partial monitoring on the proposed architecture.
+  efficiently enable partial monitoring while preventing false positives. This
+  hardware architecture is designed to be generally applicable to a wide-range
+  of run-time monitoring techniques.
+  \item We investigate interesting trade-offs in deciding what to skip for
+  partial monitoring on the proposed architecture.
+  \item We extend the dataflow engine to enable filtering null metadata to
+  reduce overheads.
 \end{itemize}
@@
 This paper is organized as follows. Section~\ref{sec:monitoring} presents the
-parallel run-time monitoring model that this paper assumes.
-Section~\ref{sec:optimizations} presents the high-level approach of how we aim
-to reduce overheads through and Section~\ref{sec:arch} describes our
-hardware architecture that supports this. Section~\ref{sec:policies}
+baseline parallel run-time monitoring model that this paper assumes.
+Section~\ref{sec:dropping} discusses enabling partial monitoring using our dataflow tracking engine.
+Section~\ref{sec:policies}
 investigates the design space for dropping decisions for partial monitoring.
+Section~\ref{sec:filtering} shows how the dataflow tracking engine can be extended to filter null metadata.
 Section~\ref{sec:evaluation} presents our evaluation methodology and

### 2014-09-05 monitoring.tex
More edits based on Ed's comments.
Reorganized paper for HPCA 2015 submission.

@@
 \label{sec:monitoring}
+
+\subsection{Overheads of Run-Time Monitoring}
+
+% Previous work overheads
+\begin{table*}[tb]
+  \begin{center}
+    \vspace{-0.0in}
+    \begin{footnotesize}
+    \input{tabs/previous_overheads}
+    \end{footnotesize}
+    \caption{Overheads of various proposed run-time monitoring systems.}
+    \vspace{-0.2in}
+    \label{tab:monitoring.previous_overheads}
+  \end{center}
+\end{table*}
+
+There have been numerous previous works exploring various implementations of
+run-time monitoring systems. Table~\ref{tab:monitoring.previous_overheads}
+summarizes some of these implementations and the average and worst-case
+overheads reported. These systems vary in their hardware complexity,
+flexibility, monitoring schemes, and overheads. Dedicated hardware can provide low 
+overheads for specific monitoring schemes. However, more pronounced overheads are seen for 
+flexible monitoring systems that implement monitoring on a processor core
+\cite{lba-asid06, lba-isca08, nagarajan-interact08} or using an FPGA fabric
+\cite{flexcore-micro10}. These systems allow a wide range of monitoring schemes to
+be implemented on the same hardware architecture and also allow monitors to be modified and updated after deployment.
+However, in order to support this flexibility, they can easily incur tens of percent of
+overhead in the average-case and severalfold slowdowns in the worst-case. Partial monitoring
+can be a useful technique for reducing overheads for any of these systems.
+In this paper, we will focus our discussion on using a parallel
+processing core to implement the monitor. Our evaluation includes results
+for both the core-based monitor as well as a higher performance FPGA-based monitor.
+
+\subsection{Baseline Architecture}
 
@@
 bound addresses. In addition, this base and bound metadata
-is propagated on certain ALU instructions to track the creation of new array pointers.
+is propagated on ALU and memory instructions to track the corresponding array pointers.
 % One example is an uninitialized memory check (UMC) where monitoring is used to
@@
 % \cite{flexcore-micro10} instead uses an FPGA-fabric to implement the
-% monitor. The approach we describe in this paper applies to any of these
+% monitor. 
+% 
+% The approach we describe in this paper applies to any of these
 % parallel monitors. However, for experiments, we model an FPGA-based monitor
@@
 % main core and metadata cache are implemented as ASICs.
+
+% Example monitor
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\columnwidth]{figs/example_full.pdf}
+    \vspace{-0.2in}
+    \caption{Example of array bounds check monitor.}
+    \label{fig:monitoring.example_full}
+    \vspace{-0.1in}
+  \end{center}
+\end{figure}
+
+Figure~\ref{fig:monitoring.example_full} shows an example pseudo-code segment and
+the corresponding monitoring operations for an array bounds check monitor. 
+First, an array {\tt x} is allocated using {\tt malloc}. When this happens, the
+monitor associates the array with metadata corresponding to its base and bounds
+address. Next, {\tt y} is set to point to the middle of {\tt y}. The metadata
+for {\tt y} is the same as the metadata {\tt x} to ensure that accesses using
+{\tt y} do not exceed the bounds of {\tt x}. Finally, both pointers {\tt x} and
+{\tt y} are used to perform a write. In both cases, the monitor checks whether
+the address that is written to is within the bounds stored in the metadata. In
+the first case, {\tt x+7} is within the original array's bounds. In the second
+case, {\tt y+7} is not within the array's bounds and the monitor will raise an
+error.

### 2014-09-05 tabs/previous_overheads.tex
Reorganized paper for HPCA 2015 submission.

\begin{tabular}{|l|p{2in}|r|r|}

\hline
{\bf Name} & {\bf Description} & {\bf Average Overhead} & {\bf Worst-case Overhead} \\ \hline\hline

DIFT \cite{dift-asplos04} & Dedicated hardware for information flow tracking & 1.1\% & 23\% \\ \hline
FlexiTaint \cite{flexitaint-hpca08} & Dedicated hardware for configurable information flow tracking & 1.1\%-3.7\% & 8.7\% \\ \hline
Hardbound \cite{hardbound-asplos08} & Dedicated hardware for array bounds check & 5\%-9\% & 22\% \\ \hline
Harmoni \cite{harmoni-dsn12} & Configurable hardware for tag-based monitoring & 1\%-10\% & 20\% \\ \hline\hline

FlexCore \cite{flexcore-micro10} & FPGA-based monitor & 1.05x-1.44x & 1.84x \\ \hline
Log-Based Architecture \cite{lba-asid06} & Core-based monitor & 3.23x-7.80x & 11x \\ \hline
Log-Based Architecture Optimized \cite{lba-isca08} & Core-based monitor with hardware acceleration  & 1.02x-3.27x & 5x \\ \hline
Multi-core DIFT \cite{nagarajan-interact08} & Core-based with per-application optimized dynamic information flow tracking monitor & 1.48x & 2.2x \\ \hline\hline

LIFT \cite{lift-micro06} & Dynamic binary instrumentation for information flow tracking & 3.6x & 7.9x \\ \hline
Purfiy \cite{purify-usenix92} & Dynamic binary instrumentation for memory leaks & 2.3x & 5.5x \\ \hline
TaintCheck \cite{taintcheck-05} & Dynamic binary instrumentation for taint analysis & 10x & 27x \\ \hline

\end{tabular}

### 2014-09-09 dropping.tex
move the baseline arch to Sec 3.

@@
-\section{Adjustable Monitoring Infrastructure}
+\section{Architecture for Adjustable Monitoring}
 \label{sec:dropping}
@@
 
-% Overview of full architecture
+\subsection{Baseline Architecture}
+
+% Run-time monitoring overview
 \begin{figure}
   \begin{center}
-    \includegraphics[width=\columnwidth]{figs/architecture_overview.pdf}
+    \includegraphics[width=\columnwidth]{figs/monitoring_architecture.pdf}
     \vspace{-0.2in}
-    \caption{Block diagram of architecture for monitoring with reduced and adjustable overheads.}
-    \label{fig:dropping.overview}
+    \caption{Overview of run-time monitoring architecture.}
+    \label{fig:monitoring.overview} 
     \vspace{-0.1in}
@@
 \end{figure}
+
+Figure~\ref{fig:monitoring.overview} shows an overview of the run-time monitoring
+model that is assumed in this paper.  The \emph{main program} is a computation
+task that performs the original function of the system and is run on the
+\emph{main core}.  On certain events during the main program, such as the
+execution of certain types of instructions, the \emph{monitoring core} performs a
+series of \emph{monitoring operations}. The monitoring core operates in parallel to the
+main core. These events are referred to as \emph{monitoring events}. Depending
+on the type of monitoring event, different monitoring operations are
+executed. Information about monitoring events are sent to the monitoring core and buffered in a FIFO structure to decouple the
+running of the main core and the monitoring core. If the FIFO is full, then the main
+core is forced to stall on a monitoring event until a FIFO entry becomes
+available. These stalls are a major source of
+overhead because the monitoring core may take several cycles to process a single event
+from the main core. We refer to these stalls and other overheads, such as
+contention for shared resources, as \emph{monitoring overheads}. If the monitoring core
+detects an inconsistent or undesired behavior in the monitoring events, then
+an error is detected.
+
+There are many possible monitoring schemes that can be implemented on this type
+of fine-grained parallel monitoring architecture such as memory protection
+\cite{mondrian-asplos02}, information flow tracking \cite{dift-asplos04,
+testudo-micro08}, soft error detection \cite{argus-micro07}, data-race
+detection \cite{cord-hpca06}, etc.  For example, an array bounds check (BC)
+\cite{hardbound-asplos08} can be implemented in order to detect
+when software attempts to read or write to a memory location outside of an
+array's bounds. This can be done by associating metadata with array pointers that 
+indicates the array's base (start) and bound (end) addresses. On loads or stores with the
+array
+pointer, the monitoring core checks that the memory address accessed is within the base and
+bound addresses. In addition, this base and bound metadata
+is propagated on ALU and memory instructions to track the corresponding array pointers.
+% One example is an uninitialized memory check (UMC) where monitoring is used to
+% detect when software attempts to read from a memory location that was not
+% previously initialized. This can be done by forwarding each load and store
+% instruction from the main core to the monitor. For every memory location, the
+% monitor keeps one bit of metadata. On a store to a memory location, the monitor
+% marks the corresponding metadata bit to indicate that the memory location has
+% been initialized. On a load, the monitor checks that the corresponding metadata
+% bit has been previously marked as initialized.
+
+% There are multiple options for implementing programmable parallel monitors. For
+% example, the Log-Based Architecture \cite{lba-isca08} uses processor cores in a
+% multi-core system as monitors. The FlexCore architecture
+% \cite{flexcore-micro10} instead uses an FPGA-fabric to implement the
+% monitor. 
+% 
+% The approach we describe in this paper applies to any of these
+% parallel monitors. However, for experiments, we model an FPGA-based monitor
+% similar to FlexCore.  The on-chip FPGA fabric is used to implement the
+% ``Monitor'' block in Figure~\ref{fig:arch.overview} while the FIFO from the
+% main core and metadata cache are implemented as ASICs.
+
+% Example monitor
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\columnwidth]{figs/example_full.pdf}
+    \vspace{-0.2in}
+    \caption{Example of array bounds check monitor.}
+    \label{fig:monitoring.example_full}
+    \vspace{-0.1in}
+  \end{center}
+\end{figure}
+
+Figure~\ref{fig:monitoring.example_full} shows an example pseudo-code segment and
+the corresponding monitoring operations for an array bounds check monitor. 
+First, an array {\tt x} is allocated using {\tt malloc}. When this happens, the
+monitor associates the array with metadata corresponding to its base and bounds
+address. Next, {\tt y} is set to point to the middle of {\tt y}. The metadata
+for {\tt y} is the same as the metadata {\tt x} to ensure that accesses using
+{\tt y} do not exceed the bounds of {\tt x}. Finally, both pointers {\tt x} and
+{\tt y} are used to perform a write. In both cases, the monitor checks whether
+the address that is written to is within the bounds stored in the metadata. In
+the first case, {\tt x+7} is within the original array's bounds. In the second
+case, {\tt y+7} is not within the array's bounds and the monitor will raise an
+error.
+
+\subsection{Architecture Overview}
+
+% Overview of full architecture
+%\begin{figure}
+%  \begin{center}
+%    \includegraphics[width=\columnwidth]{figs/architecture_overview.pdf}
+%    \vspace{-0.2in}
+%    \caption{Block diagram of architecture for monitoring with reduced and adjustable overheads.}
+%    \label{fig:dropping.overview}
+%    \vspace{-0.1in}
+%  \end{center}
+%\end{figure}
 

### 2014-09-09 monitoring.tex
move the baseline arch to Sec 3.

@@
-\section{Parallel Programmable Run-Time Monitoring}
+\section{Adjustable Run-Time Monitoring}
 \label{sec:monitoring}
 
-\subsection{Overheads of Run-Time Monitoring}
-
 % Previous work overheads
-\begin{table*}[tb]
+\begin{table*}[t]
   \begin{center}
@@
     \end{footnotesize}
-    \caption{Overheads of various proposed run-time monitoring systems.}
+    \caption{Trade-off between performance and flexibility/complexity of run-time monitoring systems.}
     \vspace{-0.2in}
@@
 \end{table*}
+
+\subsection{Run-Time Monitoring Overhead}
 
@@
 
-\subsection{Baseline Architecture}
+\subsection{Adjustable and Programmable Monitoring}
 
-% Run-time monitoring overview
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\columnwidth]{figs/monitoring_architecture.pdf}
-    \vspace{-0.2in}
-    \caption{Overview of run-time monitoring architecture.}
-    \label{fig:monitoring.overview} 
-    \vspace{-0.1in}
-  \end{center}
-\end{figure}
+There were three main challenges in designing this dropping hardware:
+\begin{enumerate}
+  \item \textbf{General:} The hardware needs to be applicable to a wide range of monitoring schemes.
+  \item \textbf{No false positives:} False positives should never occur as a result of dropping.
+  \item \textbf{Intelligent dropping:} The hardware needs to maximize the amount of useful monitoring done while staying within the overhead budget.
+\end{enumerate}
 
-Figure~\ref{fig:monitoring.overview} shows an overview of the run-time monitoring
-model that is assumed in this paper.  The \emph{main program} is a computation
-task that performs the original function of the system and is run on the
-\emph{main core}.  On certain events during the main program, such as the
-execution of certain types of instructions, the \emph{monitoring core} performs a
-series of \emph{monitoring operations}. The monitoring core operates in parallel to the
-main core. These events are referred to as \emph{monitoring events}. Depending
-on the type of monitoring event, different monitoring operations are
-executed. Information about monitoring events are sent to the monitoring core and buffered in a FIFO structure to decouple the
-running of the main core and the monitoring core. If the FIFO is full, then the main
-core is forced to stall on a monitoring event until a FIFO entry becomes
-available. These stalls are a major source of
-overhead because the monitoring core may take several cycles to process a single event
-from the main core. We refer to these stalls and other overheads, such as
-contention for shared resources, as \emph{monitoring overheads}. If the monitoring core
-detects an inconsistent or undesired behavior in the monitoring events, then
-an error is detected.
-
-There are many possible monitoring schemes that can be implemented on this type
-of fine-grained parallel monitoring architecture such as memory protection
-\cite{mondrian-asplos02}, information flow tracking \cite{dift-asplos04,
-testudo-micro08}, soft error detection \cite{argus-micro07}, data-race
-detection \cite{cord-hpca06}, etc.  For example, an array bounds check (BC)
-\cite{hardbound-asplos08} can be implemented in order to detect
-when software attempts to read or write to a memory location outside of an
-array's bounds. This can be done by associating metadata with array pointers that 
-indicates the array's base (start) and bound (end) addresses. On loads or stores with the
-array
-pointer, the monitoring core checks that the memory address accessed is within the base and
-bound addresses. In addition, this base and bound metadata
-is propagated on ALU and memory instructions to track the corresponding array pointers.
-% One example is an uninitialized memory check (UMC) where monitoring is used to
-% detect when software attempts to read from a memory location that was not
-% previously initialized. This can be done by forwarding each load and store
-% instruction from the main core to the monitor. For every memory location, the
-% monitor keeps one bit of metadata. On a store to a memory location, the monitor
-% marks the corresponding metadata bit to indicate that the memory location has
-% been initialized. On a load, the monitor checks that the corresponding metadata
-% bit has been previously marked as initialized.
-
-% There are multiple options for implementing programmable parallel monitors. For
-% example, the Log-Based Architecture \cite{lba-isca08} uses processor cores in a
-% multi-core system as monitors. The FlexCore architecture
-% \cite{flexcore-micro10} instead uses an FPGA-fabric to implement the
-% monitor. 
-% 
-% The approach we describe in this paper applies to any of these
-% parallel monitors. However, for experiments, we model an FPGA-based monitor
-% similar to FlexCore.  The on-chip FPGA fabric is used to implement the
-% ``Monitor'' block in Figure~\ref{fig:arch.overview} while the FIFO from the
-% main core and metadata cache are implemented as ASICs.
-
-% Example monitor
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\columnwidth]{figs/example_full.pdf}
-    \vspace{-0.2in}
-    \caption{Example of array bounds check monitor.}
-    \label{fig:monitoring.example_full}
-    \vspace{-0.1in}
-  \end{center}
-\end{figure}
-
-Figure~\ref{fig:monitoring.example_full} shows an example pseudo-code segment and
-the corresponding monitoring operations for an array bounds check monitor. 
-First, an array {\tt x} is allocated using {\tt malloc}. When this happens, the
-monitor associates the array with metadata corresponding to its base and bounds
-address. Next, {\tt y} is set to point to the middle of {\tt y}. The metadata
-for {\tt y} is the same as the metadata {\tt x} to ensure that accesses using
-{\tt y} do not exceed the bounds of {\tt x}. Finally, both pointers {\tt x} and
-{\tt y} are used to perform a write. In both cases, the monitor checks whether
-the address that is written to is within the bounds stored in the metadata. In
-the first case, {\tt x+7} is within the original array's bounds. In the second
-case, {\tt y+7} is not within the array's bounds and the monitor will raise an
-error.
+\subsection{Applications and Metrics}

### 2014-09-09 monitoring.tex
reorganized the overhead table. I will continue working on Sec 2 later in the evening.

@@
     \end{footnotesize}
-    \caption{Trade-off between performance and flexibility/complexity of run-time monitoring systems.}
+    \caption{Trade-off between performance overhead and flexibility/complexity of run-time monitoring systems.}
     \vspace{-0.2in}
@@
 
-\subsection{Run-Time Monitoring Overhead}
+\subsection{Overhead of Run-Time Monitoring}
 
-There have been numerous previous works exploring various implementations of
-run-time monitoring systems. Table~\ref{tab:monitoring.previous_overheads}
+There have been a number of proposals for run-time monitoring systems, exploring various
+design points in the trade-off space between efficiency and flexibility.
+Table~\ref{tab:monitoring.previous_overheads} summarizes some of representative designs
+and their reported performance overhead.
+
 summarizes some of these implementations and the average and worst-case

### 2014-09-09 tabs/previous_overheads.tex
reorganized the overhead table. I will continue working on Sec 2 later in the evening.

@@
-\begin{tabular}{|l|p{2in}|r|r|}
+%\begin{tabular}{|l|p{2in}|r|r|}
+\begin{tabular}{|l|l|l|r|}
 
 \hline
-{\bf Name} & {\bf Description} & {\bf Average Overhead} & {\bf Worst-case Overhead} \\ \hline\hline
+{\bf Name} & {\bf Type} & {\bf Monitoring scheme and flexibility} & {\bf Slowdown (avg./worst)} \\ \hline\hline
 
-DIFT \cite{dift-asplos04} & Dedicated hardware for information flow tracking & 1.1\% & 23\% \\ \hline
-FlexiTaint \cite{flexitaint-hpca08} & Dedicated hardware for configurable information flow tracking & 1.1\%-3.7\% & 8.7\% \\ \hline
-Hardbound \cite{hardbound-asplos08} & Dedicated hardware for array bounds check & 5\%-9\% & 22\% \\ \hline
-Harmoni \cite{harmoni-dsn12} & Configurable hardware for tag-based monitoring & 1\%-10\% & 20\% \\ \hline\hline
+DIFT \cite{dift-asplos04} & Custom HW & DIFT only & 1.1\% / 23\% \\ \hline
+FlexiTaint \cite{flexitaint-hpca08} & Custom HW & DIFT w/ programmable policies & 1.1\%-3.7\% / 8.7\% \\ \hline
+Hardbound \cite{hardbound-asplos08} & Custom HW & Array bounds checks only & 5\%-9\% / 22\% \\ \hline
+Harmoni \cite{harmoni-dsn12} & Custom HW & Tag-based monitors & 1\%-10\% / 20\% \\ \hline\hline
 
-FlexCore \cite{flexcore-micro10} & FPGA-based monitor & 1.05x-1.44x & 1.84x \\ \hline
-Log-Based Architecture \cite{lba-asid06} & Core-based monitor & 3.23x-7.80x & 11x \\ \hline
-Log-Based Architecture Optimized \cite{lba-isca08} & Core-based monitor with hardware acceleration  & 1.02x-3.27x & 5x \\ \hline
-Multi-core DIFT \cite{nagarajan-interact08} & Core-based with per-application optimized dynamic information flow tracking monitor & 1.48x & 2.2x \\ \hline\hline
+FlexCore \cite{flexcore-micro10} & Dedicated FPGA & Instruction-trace monitoring & 1.05x-1.44x / 1.84x \\ \hline
+FADE \cite{FIXME} & Core+Custom HW & Instruction-trace monitoring (effective when HW filters work) & \\ \hline
+LBA-accelerated \cite{lba-isca08} & Multi-core+Custom HW & Instruction-trace monitoring (effective when accelerators work) & 1.02x-3.27x / 5x \\ \hline
+LBA \cite{lba-asid06} & Multi-core+Custom HW & Instruction trace monitoring & 3.23x-7.80x / 11x \\ \hline \hline
 
-LIFT \cite{lift-micro06} & Dynamic binary instrumentation for information flow tracking & 3.6x & 7.9x \\ \hline
-Purfiy \cite{purify-usenix92} & Dynamic binary instrumentation for memory leaks & 2.3x & 5.5x \\ \hline
-TaintCheck \cite{taintcheck-05} & Dynamic binary instrumentation for taint analysis & 10x & 27x \\ \hline
+Multi-core DIFT \cite{nagarajan-interact08} & SW (multithreaded) & DIFT (compiled for each application) & 1.48x / 2.2x \\ \hline
+
+LIFT \cite{lift-micro06} & SW (DBI) & DIFT (fully flexible) & 3.6x / 7.9x \\ \hline
+Purfiy \cite{purify-usenix92} & SW (DBI) & Memory leak checks (fully flexible) & 2.3x / 5.5x \\ \hline
+TaintCheck \cite{taintcheck-05} & SW (DBI) & DIFT (fully flexible) & 10x / 27x \\ \hline
 

### 2014-09-10 monitoring.tex
outline upate

@@
 
-There have been a number of proposals for run-time monitoring systems, exploring various
-design points in the trade-off space between efficiency and flexibility.
+There have been a number of proposals for run-time monitoring systems exploring various
+design points. % in the trade-off space between efficiency and flexibility.
 Table~\ref{tab:monitoring.previous_overheads} summarizes some of representative designs
-and their reported performance overhead.
+and their reported performance overhead. The previous studies clearly show that there
+exist trade-offs among efficiency, flexibility, and hardware costs. 
+For example, a run-time monitoring scheme can often be realized with fairly low
+performance overhead (less than 20\%) if implemented with custom hardware that is
+designed only for one monitor or a narrow set of monitors. However, the custom
+hardware monitors cannot be modified or updated, and require dedicated silicon area. 
+One the other hand, flexible systems that support a wide range of monitors lead 
+to noticeable performance overhead, often too high for wide deployment in practice.
+Software-only implementations \cite{FIXME} or multi-core monitors with minimal
+hardware changes \cite{lba-asid06} are reported to have severalfold slowdowns.
+On-chip FPGA monitors \cite{flexcore-micro10} and cores with monitoring accelerators
+\cite{lba-isca08, fade-hpca14} can reduce overhead significantly, but still show
+slowdowns of several tens of percents in some cases.
 
-summarizes some of these implementations and the average and worst-case
-overheads reported. These systems vary in their hardware complexity,
-flexibility, monitoring schemes, and overheads. Dedicated hardware can provide low 
-overheads for specific monitoring schemes. However, more pronounced overheads are seen for 
-flexible monitoring systems that implement monitoring on a processor core
-\cite{lba-asid06, lba-isca08, nagarajan-interact08} or using an FPGA fabric
-\cite{flexcore-micro10}. These systems allow a wide range of monitoring schemes to
-be implemented on the same hardware architecture and also allow monitors to be modified and updated after deployment.
-However, in order to support this flexibility, they can easily incur tens of percent of
-overhead in the average-case and severalfold slowdowns in the worst-case. Partial monitoring
-can be a useful technique for reducing overheads for any of these systems.
-In this paper, we will focus our discussion on using a parallel
+\subsection{Partial Monitoring for Adjustable Overhead}
+
+In this paper, we investigate adding a new dimension to the trade-off space of 
+run-time monitors by allowing coverage or accuracy of a monitor to be traded
+off for lower performance and/or hardware overhead.
+
+namely coverage/accuracy of monitors.
+
+will focus our discussion on using a parallel
 processing core to implement the monitor. Our evaluation includes results
@@
 
-\subsection{Adjustable and Programmable Monitoring}
+Benefits - low overhead or simpler HW
+
+\subsection{Applications and Metrics}
+
+Three main app
+- Partial protection for run-time attacks or bugs: you care about the dynamic coverage 
+- Soft real-time systems: tight deadline required
+- Sampling for Profiling: target metric - accuracy
+
+\subsection{Challenges}
 
@@
 \end{enumerate}
-
-\subsection{Applications and Metrics}

### 2014-09-11 monitoring.tex
revised Section 2 to better motivate adjustable monitoring

@@
 slowdowns of several tens of percents in some cases.
+In today's monitoring systems, the overhead is often unpredictable because they
+depend heavily on the characteristics of applications and monitoring operations.
+In summary, users need to either pay noticeable overhead or the cost of custom
+hardware in order to use fine-grained run-time monitoring in deployed systems.
 
@@
 
-In this paper, we investigate adding a new dimension to the trade-off space of 
-run-time monitors by allowing coverage or accuracy of a monitor to be traded
-off for lower performance and/or hardware overhead.
+In this paper, we aim to develop a general framework that enables monitoring
+overhead to be dynamically adjusted by dropping a portion of monitoring 
+operations if necessary. In essence, this framework adds a new dimension to
+the monitor design space by allowing coverage or accuracy to be traded off
+for lower overhead. For example, the capability to adjust overhead allows users
+to use monitoring with partial coverage in order to reduce performance
+overhead. Alternatively, partial monitoring can allow users to use simple
+one may choose to use simple hardware such as a
+multi-core without elaborate acceleration features to provide partial monitoring
+with the performance overhead that is comparable to custom hardware.
 
-namely coverage/accuracy of monitors.
-
-will focus our discussion on using a parallel
-processing core to implement the monitor. Our evaluation includes results
-for both the core-based monitor as well as a higher performance FPGA-based monitor.
-
-Benefits - low overhead or simpler HW
+In this framework, a user can specify how much monitoring should be done such as a target 
+overhead budget, a target coverage, a percentage of monitoring operations to be performed, etc.
+Then, the framework dynamically drops a portion of monitoring operations to match
+the target. In particular, this paper focuses on matching a performance overhead target
+while maximizing the coverage. Given that the overhead of custom hardware monitors can
+already be quite low, the focus will be enabling the trade-off in {\em general-purpose} 
+monitoring systems that support a wide range of monitors.
+We also consider the target overhead as a soft constraint and do not aim to
+provide a strict worst-case guarantee.
 
@@
 
-Three main app
-- Partial protection for run-time attacks or bugs: you care about the dynamic coverage 
-- Soft real-time systems: tight deadline required
-- Sampling for Profiling: target metric - accuracy
+While it is ideal if run-time monitoring can be performed in full, we believe that
+the capability to trade-off coverage/accuracy for lower performance/hardware overhead
+will be useful in many application scenarios where full monitoring is not a viable option.
+Here, we briefly discuss example applications of partial monitoring and the metrics
+that are important in each case.
 
-\subsection{Challenges}
+{\bf Cooperative testing, debugging, and protection:}
+Recently studies \cite{FIXME} have shown that software testing, debugging,
+or attack detection may be done in a cooperative fashion across
+a large number of systems. In this case, each system is only willing to pay very low
+overhead, say a few percents, and only performs a small subset of checks. 
+A high coverage is achieved by having different systems check different parts of a program.
+The partial monitoring framework allows high-overhead monitoring to be used in a
+cooperative fashion. 
+The main metric that represents the effectiveness of monitoring in this case is
+the coverage (the percentage of checks there were performed) over multiple runs.
 
-There were three main challenges in designing this dropping hardware:
+{\bf Monitoring of soft real-time systems:}
+Soft real-time systems or interactive systems often need to meet deadlines or response
+time requirements. Unfortunately, the overhead of run-time monitoring is often 
+unpredictable and varies significantly depending on the application and monitoring
+characteristics. The partial monitoring framework allows monitoring to be performed
+while providing a level of guarantee on its impact on the execution time. 
+In this case, it is important that the system can closely match the desired
+overhead target while maximizing the effectiveness of monitoring.
+
+{\bf Partial protection for low overhead:}
+Even without real-time constraints, systems may have tight budgets for performance
+or hardware overhead that they can tolerate for run-time monitoring. In such cases,
+full monitoring cannot be enabled unless its overhead is low enough. The adjustable
+monitoring allows partial protection on such systems. For example, array bounds
+may be checked for a portion of memory accesses, or a subset of information flows
+may be tracked in DIFT for attack detection. In this scenario, the effectiveness of 
+partial monitoring can be measured as the percentage of run-time checks that are
+performed on each program run, which reflects how likely for a bug or an attack 
+to be detected for a system.
+
+{\bf Profiling:} 
+The run-time monitoring system can be used to implement various profiling tools
+to collect statistics on program behaviors for performance optimizations as well
+as security protection. For example, a recent study showed that an instruction
+mix can be used to identify malware from normal programs \cite{FIXME}. 
+In such profiling tools, the partial monitoring framework can be used to obtain
+the statics on samples rather than all program events, essentially trading off
+accuracy for lower overhead.
+
+\subsection{Design Challenges}
+
+While conceptually simple, designing a general framework to dynamically adjust monitoring
+overhead introduces new challenges that need to be addressed.  
+The following summarizes the main design goals and associated design challenges.
+
 \begin{enumerate}
-  \item \textbf{General:} The hardware needs to be applicable to a wide range of monitoring schemes.
-  \item \textbf{No false positives:} False positives should never occur as a result of dropping.
-  \item \textbf{Intelligent dropping:} The hardware needs to maximize the amount of useful monitoring done while staying within the overhead budget.
+  \item \textbf{General:} The framework needs to be general enough to be applicable to 
+  a wide range of monitoring schemes. Because we target flexible, general-purpose
+  run-time monitoring systems, a solution that only works for one monitoring scheme will
+  not work.
+
+  \item \textbf{No false positive:} The framework needs to ensure that dropping a portion
+  of monitoring operations does not lead to a false positive. We solve this problem using
+  a data-flow engine that tracks invalid monitoring meta-data (Section~\ref{sec:dropping}).
+
+  \item \textbf{Intelligent dropping:} The framework needs to match the overhead budget
+  while maximizing the effectiveness of monitoring. To this end, adjustable overhead
+  monitoring needs to carefully choose which operations to drop and when. We address this
+  problem by studying different dropping policies (Section~\ref{sec:policies}) and their
+  trade-offs.
+
 \end{enumerate}

### 2014-09-11 dropping.tex
Updates to Section 3.

@@
 
-In this section, we present an architecture that can limit the amount of
-monitoring performed in order to meet a target overhead budget. We consider the
-overhead budget as a soft constraint that we seek to stay below or close to,
-but do not require strict guarantees that it is met. For this paper, we focus on
-the performance overhead of monitoring, though other overheads could also be
-considered.
+In this section, we present our hardware architecture that enables partial
+monitoring for adjustable overheads. Section~\ref{sec:dropping.baseline} first
+describes our baseline model for run-time monitoring while the rest of the
+Section describes our design.
 
 \subsection{Baseline Architecture}
+\label{sec:dropping.baseline}
 
@@
 overhead because the monitoring core may take several cycles to process a single event
-from the main core. We refer to these stalls and other overheads, such as
-contention for shared resources, as \emph{monitoring overheads}. If the monitoring core
+from the main core. 
+%We refer to these stalls and other overheads, such as
+%contention for shared resources, as \emph{monitoring overheads}. 
+If the monitoring core
 detects an inconsistent or undesired behavior in the monitoring events, then
@@
 % Example monitor
-\begin{figure}
+\begin{figure*}
   \begin{center}
-    \includegraphics[width=\columnwidth]{figs/example_full.pdf}
+    \includegraphics[]{figs/example_full.pdf}
     \vspace{-0.2in}
@@
   \end{center}
-\end{figure}
+\end{figure*}
 
-Figure~\ref{fig:monitoring.example_full} shows an example pseudo-code segment and
-the corresponding monitoring operations for an array bounds check monitor. 
-First, an array {\tt x} is allocated using {\tt malloc}. When this happens, the
-monitor associates the array with metadata corresponding to its base and bounds
-address. Next, {\tt y} is set to point to the middle of {\tt y}. The metadata
-for {\tt y} is the same as the metadata {\tt x} to ensure that accesses using
-{\tt y} do not exceed the bounds of {\tt x}. Finally, both pointers {\tt x} and
-{\tt y} are used to perform a write. In both cases, the monitor checks whether
-the address that is written to is within the bounds stored in the metadata. In
-the first case, {\tt x+7} is within the original array's bounds. In the second
-case, {\tt y+7} is not within the array's bounds and the monitor will raise an
-error.
+Figure~\ref{fig:monitoring.example_full} shows an example pseudo-code segment,
+its assembly level instructions, and the corresponding monitoring operations
+for an array bounds check monitor. 
+First, an array {\tt x} is allocated using {\tt malloc}. 
+As {\tt malloc} returns the array's address in a register, the monitor
+associates base and bounds metadata with the corresponding metadata. 
+Next, pointer {\tt y} is set to point to the middle of array {\tt x}. 
+At the assembly level, a register {\tt r2} is set to array {\tt x}'s address plus an offset.
+The monitor propagates the metadata of the original pointer in {\tt r1} to the
+metadata of {\tt r2}. This is to ensure that pointer {\tt y} is not used to exceed the array's bounds.
+Finally, both pointers {\tt x} and {\tt y} are used to perform a write. 
+This requires setting register {\tt r3} to the write value. When this happens,
+{\tt r3}'s metadata is also reset to NULL in case it previously stored a
+pointer. In both cases, the monitor checks whether the address that is written
+to is within the bounds stored in the metadata. In the first case, {\tt x+12}
+is within the original array's bounds. No error is raised and the metadata of
+{\tt r3} is propagated to the metadata of the store address. If {\tt r3} were a
+pointer, than this would allow a future instruction to load the pointer and use
+it to access the corresponding array or data structure. In the second case,
+{\tt y+12} corresponds to {\tt x+20} which is not within the array's bounds and
+the monitor will raise an error.
 
-\subsection{Architecture Overview}
+Here, we have shown the dynamic sequence of operations of the monitoring core.
+The actual static code consists of a set of instructions to be run for each
+monitored instruction type. Register and memory addresses are read from
+information sent by the main core.
 
-% Overview of full architecture
-%\begin{figure}
-%  \begin{center}
-%    \includegraphics[width=\columnwidth]{figs/architecture_overview.pdf}
-%    \vspace{-0.2in}
-%    \caption{Block diagram of architecture for monitoring with reduced and adjustable overheads.}
-%    \label{fig:dropping.overview}
-%    \vspace{-0.1in}
-%  \end{center}
-%\end{figure}
+% \subsection{Architecture Overview}
+% 
+% % Overview of full architecture
+% %\begin{figure}
+% %  \begin{center}
+% %    \includegraphics[width=\columnwidth]{figs/architecture_overview.pdf}
+% %    \vspace{-0.2in}
+% %    \caption{Block diagram of architecture for monitoring with reduced and adjustable overheads.}
+% %    \label{fig:dropping.overview}
+% %    \vspace{-0.1in}
+% %  \end{center}
+% %\end{figure}
+% 
+% Figure~\ref{fig:dropping.overview} shows a high-level block diagram of our
+% proposed architecture for enabling partial monitoring. The basic idea is to
+% insert hardware between the main core and the monitoring core to transparently
+% drop monitoring events. 
+% 
+% There were three main challenges in designing this dropping hardware:
+% \begin{enumerate}
+%   \item \textbf{General:} The hardware needs to be applicable to a wide range of monitoring schemes.
+%   \item \textbf{No false positives:} False positives should never occur as a result of dropping.
+%   \item \textbf{Intelligent dropping:} The hardware needs to maximize the amount of useful monitoring done while staying within the overhead budget.
+% \end{enumerate}
+% 
+% In order to reduce overheads, the ``Drop'' module decides when monitoring event
+% should be skipped.  This decision of when to drop and which monitoring events
+% to dropped is discussed in more detail in Section~\ref{sec:policies}.
+% 
+% When an event is dropped, metadata needs to be invalidated in order to prevent
+% false positives. This is done using a hardware dataflow engine.
+% Sections~\ref{sec:dropping.false_neg_pos} and
+% \ref{sec:dropping.prevent_false_pos} discuss these false positive issues and
+% the dataflow engine in detail. 
+% 
+% In addition, we can use the dataflow engine to filter out monitoring operations
+% that would be operating on invalid metadata. This maximizes the useful
+% monitoring that is done by the monitoring core.
+% Section~\ref{sec:dropping.filter} talks about how this is done using
+% information from the dataflow engine.
 
-Figure~\ref{fig:dropping.overview} shows a high-level block diagram of our
-proposed architecture for enabling partial monitoring. The basic idea is to
-insert hardware between the main core and the monitoring core to transparently
-drop monitoring events.
-
-There were three main challenges in designing this dropping hardware:
-\begin{enumerate}
-  \item \textbf{General:} The hardware needs to be applicable to a wide range of monitoring schemes.
-  \item \textbf{No false positives:} False positives should never occur as a result of dropping.
-  \item \textbf{Intelligent dropping:} The hardware needs to maximize the amount of useful monitoring done while staying within the overhead budget.
-\end{enumerate}
-
-In order to reduce overheads, the ``Drop'' module decides when monitoring event
-should be skipped.  This decision of when to drop and which monitoring events
-to dropped is discussed in more detail in Section~\ref{sec:policies}.
-
-When an event is dropped, metadata needs to be invalidated in order to prevent
-false positives. This is done using a hardware dataflow engine.
-Sections~\ref{sec:dropping.false_neg_pos} and
-\ref{sec:dropping.prevent_false_pos} discuss these false positive issues and
-the dataflow engine in detail.
-
-In addition, we can use the dataflow engine to filter out monitoring operations
-that would be operating on invalid metadata. This maximizes the useful
-monitoring that is done by the monitoring core.
-Section~\ref{sec:dropping.filter} talks about how this is done using
-information from the dataflow engine.
-
-\subsection{False Negatives and False Positives}
+\subsection{Effects of Dropping Monitoring}
 \label{sec:dropping.false_neg_pos}
-
-% Example of dropping
-\begin{figure*}
-  \begin{center}
-  \subfloat[no error] {
-    \includegraphics[trim=1.6in 0 0 0,clip=true]{figs/example_no_error.pdf}
-    \label{fig:dropping.example_no_error}
-  }
-  \hspace{0.2in}
-  \subfloat[false negative] {
-    \includegraphics[trim=1.6in 0 0 0,clip=true]{figs/example_false_negative.pdf}
-    \label{fig:dropping.example_false_negative}
-  }
-  \hspace{0.2in}
-  \subfloat[false positive] {
-    \includegraphics[trim=1.6in 0 0 0,clip=true]{figs/example_false_positive.pdf}
-    \label{fig:dropping.example_false_positive}
-  }
-  \caption{Dropping}
-  \label{fig:dropping.example_dropping}
-  \end{center}
-\end{figure*}
 
@@
 monitoring operation. The first is that there is no difference in operation
-from the original execution. For example, if we drop line 3 from our array
-bounds check example (see Figure~\ref{fig:dropping.example_no_error}), then the
-check on accessing {\tt x+7} is skipped. However, this is a valid access and so
+from the original execution. For example, if we drop line 3b from our array
+bounds check example (Figure~\ref{fig:dropping.example_full}), then the
+check on accessing {\tt x+12} is skipped. However, this is a valid access and so
 skipping the check does not change anything.
 
-On the other hand, if the monitoring for accessing memory location {\tt y+7} on
-line 4 is skipped (see Figure~\ref{fig:dropping.example_false_negative}), then
+On the other hand, if the monitoring for accessing memory location {\tt y+12} on
+line 4 is skipped, then
 a false negative occurs. Originally, the monitor would catch this access as an
@@
 
-The final possible outcome of dropping a monitoring operation is shown in
-Figure~\ref{fig:dropping.example_dropping}. In this case, the monitoring for
+The final possible outcome of dropping a monitoring operation is a false positive. 
+For example, suppose the monitoring for
 line 1 is dropped, causing the bound information for pointer {\tt x} to never
-be set. The result is that when the access using {\tt x} is checked on line 3,
+be set. The result is that when the access using {\tt x} is checked on line 3b,
 an error will be raised. This creates a false positive where an error is
@@
 
-In analyzing various monitoring schemes, we found that monitoring operations
-are primarily of two types: \emph{checks} and \emph{metadata updates}. Monitors
-\emph{check} certain properties to ensure correct main program execution (lines
-3-4 of Figure~\ref{fig:monitoring.example_full}) and they \emph{update} metadata
-for bookkeeping (lines 1-2 of Figure~\ref{fig:monitoring.example_full}). Skipping
-a check operation can only cause false negatives and will never cause a false
-positive (see Figure~\ref{fig:dropping.example_no_error} and
-Figure~\ref{fig:dropping.example_false_negative}). Therefore, we may simply
-skip a check operation. Skipping an update operation can cause false
-negatives but may also cause false positives (see
-Figure~\ref{fig:dropping.example_false_positive}).
+% In analyzing various monitoring schemes, we found that monitoring operations
+% are primarily of two types: \emph{checks} and \emph{metadata updates}. Monitors
+% \emph{check} certain properties to ensure correct main program execution (lines
+% 3b-4 of Figure~\ref{fig:monitoring.example_full}) and they \emph{update} metadata
+% for bookkeeping (lines 1-3a of Figure~\ref{fig:monitoring.example_full}). Skipping
+% a check operation can only cause false negatives and will never cause a false
+% positive (see Figure~\ref{fig:dropping.example_no_error} and
+% Figure~\ref{fig:dropping.example_false_negative}). Therefore, we may simply
+% skip a check operation. Skipping an update operation can cause false
+% negatives but may also cause false positives (see
+% Figure~\ref{fig:dropping.example_false_positive}).
 
@@
     \includegraphics[]{figs/example_invalid.pdf}
-    \caption{Invalidation flags are used to prevent false positives.}
+    \caption{Example of using invalidation information to prevent false positives.}
     \label{fig:dropping.example_invalid}
@@
 
+The key cause of false positives is dropping monitoring operations that update
+metadata. Dropping monitoring operations that check for an error (such as check
+found for line 4 in Figure~\ref{fig:monitoring.example_full} can only cause
+false negatives but will never cause false positives. On the other hand,
+skipping monitoring operations that update metadata can lead to false positives
+and false negatives.
 Essentially, when an update operation is skipped, we can no longer trust the
-corresponding metadata.  Thus, our approach is to mark these metadata as
-invalid. Furthermore, metadata that would be derived from these dropped events
-should also be marked as invalid. Figure~\ref{fig:dropping.example_invalid}
-shows an example of how associating an invalid flag with metadata can be used
-to avoid false positives.  In this case, when the monitoring operation to set
-the metadata on {\tt x} is dropped, the metadata for {\tt x} is also marked as
-invalid. The result is that when the metadata for {\tt x} is checked on line 3,
-no error is raised since the metadata is invalid. Note that in this case, the
-check on line 4 is also skipped. This is necessary because an error would have
-been raised even if the access was in bounds.
+corresponding metadata. Thus, our general approach for preventing false positives is to mark these metadata as
+invalid. Furthermore, this invalidation information is propagated in the same
+way that metadata is. Figure~\ref{fig:dropping.example_invalid}
+shows an example of associating invalid flags with metadata. In the example, no
+monitoring operations are dropped. However, suppose that the monitoring for
+line 1 is dropped. In this case, {\tt rf\_invalid[1]} will be marked as {\tt
+true}. Thus, when line 3 is reached, the monitoring event is dropped
+since {\tt rf\_invalid[1]} is marked as true. Note that in this case, line 2
+also propagates this invalid flag to {\tt rf\_invalid[2]} and causes the check
+performed on line 4 to also be dropped. This is necessary because an error
+would have been raised even if the access on line 4 was in bounds.
 
@@
 monitoring core (see Figure~\ref{fig:dropping.dataflow_overview}). 
-This module performs dataflow tracking on invalidation information in
-order to prevent false positives. In addition, checks on invalid metadata are not useful and
-thus are filtered out by the hardware module before reaching the monitor.
-Similarly, updating metadata based on metadata that is invalid is also not a
-productive use of the monitor. Thus, in order to save the use of the monitor
-for useful operations, the hardware module also filters out update operations
-based on invalid metadata.
+This module handles the invalidation operations shown in the middle column of
+Figure~\ref{fig:dropping.example_invalid}.
+There are two operations that are done for handling invalidation information:
+
+\begin{enumerate}
+  \item Propagate invalid flags, following the dataflow of metadata.
+  \item Filter out monitoring operations based on invalid flags.
+\end{enumerate}
 
@@
 
-Figure~\ref{fig:dropping.dataflow} shows a detailed block diagram of the
-hardware module inserted between the main core and the monitoring. 
-Monitoring schemes typically keep metadata corresponding to the main core's
-register file as well as the main core's memory space. Similarly, the dataflow
-engine uses an on-chip register file to store flags for register file metadata.
-In addition, a set of flags are stored that correspond to memory metadata.
-These memory metadata flags are accessed through a cache and backed to main
-memory. All flags are initialized as valid when the system starts.
+Thus, the hardware acts effectively as a dataflow tracking engine in order to
+track invalid flags. Figure~\ref{fig:dropping.dataflow} shows a detailed block
+diagram of this hardware module. The dataflow engine uses two address
+calculation units to read in up two invalid flags. These source invalid flags
+are used to decide whether a monitoring event should filtered. A third address
+calculation unit is used to optionally specify a target to propagate the
+invalidation information. The module includes a register file to store invalid
+flags corresponding to register file metadata. In addition, it uses a small
+memory-backed cache to handle invalid flags corresponding to memory metadata.
+
+Since different monitoring operations are performed based on instruction type,
+the dataflow engine is also configured based on instruction type. The source
+and operation of the address generation units are set based on instruction
+type. Note that the address calculation units also take information from the
+monitored event as inputs. Thus, in the same way that the monitor selects
+metadata based on register indices or memory address of the specific monitoring
+event, the dataflow engine also reads the appropriate flags. In addition, the
+filter decision table is configured based on instruction type to decide what
+combination of input flags will lead to a filtered event and whether
+propagation is required.
+
+% Monitoring schemes typically keep metadata corresponding to the main core's
+% register file as well as the main core's memory space. Similarly, the dataflow
+% engine uses an on-chip register file to store flags for register file metadata.
+% In addition, a set of flags are stored that correspond to memory metadata.
+% These memory metadata flags are accessed through a cache and backed to main
+% memory. All flags are initialized as valid when the system starts.
 
 % Address calculation and configuration
-On a monitoring event, the dataflow engine reads in up to two flags in order to
-determine whether the event can be filtered.  The flags to be read in typically correspond to the source
-operands of the monitored instruction. For example, on a load, the dataflow
-engine will typically read in the flags corresponding the memory location
-accessed.
-However, the architecture is designed such that the address of flags to be read
-is calculated dynamically and can be configured depending on the monitoring
-scheme. Specifically, there exists a configuration table that outputs a set of
-control signals depending on the instruction type of the monitoring event.
-These control signals are used to control a pair of address calculation units which each
-contain a simple ALU. The inputs to the ALU are information from the monitored
-event and a constant that is specified from the configuration table. One common
-use of this address calculation is to transform addresses from byte-addressed
-to word-addressed by right shifting the passed memory address by 2 bits.
+% On a monitoring event, the dataflow engine reads in up to two flags in order to
+% determine whether the event can be filtered.  The flags to be read in typically correspond to the source
+% operands of the monitored instruction. For example, on a load, the dataflow
+% engine will typically read in the flags corresponding the memory location
+% accessed.
+% However, the architecture is designed such that the address of flags to be read
+% is calculated dynamically and can be configured depending on the monitoring
+% scheme. Specifically, there exists a configuration table that outputs a set of
+% control signals depending on the instruction type of the monitoring event.
+% These control signals are used to control a pair of address calculation units which each
+% contain a simple ALU. The inputs to the ALU are information from the monitored
+% event and a constant that is specified from the configuration table. One common
+% use of this address calculation is to transform addresses from byte-addressed
+% to word-addressed by right shifting the passed memory address by 2 bits.
 
-\subsection{Filtering Invalid Events}
-\label{sec:dropping.filter}
-
-% Filtering decision
-The decision of whether an event can be filtered out is determined using a
-lookup table. This filter decision table is configured by the monitoring scheme
-on system initialization. The lookup table determines whether an event can be
-filtered out based on the pair of input flags read and the instruction type of
-the monitored event.
-For example, on an ALU instruction, if both of the source operand flags
-indicate that the corresponding metadata is null, then typically the instruction can be
-filtered out. Similarly, on an ALU instruction, if either of the source operand
-flags indicate that their corresponding metadata is invalid, then the
-instruction can also be filtered out. This filter decision is sent to the
-``Filter'' block shown in Figure~\ref{fig:dropping.overview} which will simply pop
-filtered entries from the FIFO and not forward them. If an entry is not
-filtered, then this Filter block will pop the monitoring event entry from the
-FIFO and forward it along the datapath.
-
-% Propagating flag information
-Finally, when an event is filtered out, it is necessary to propagate the flag
-information for the filtered event. In addition to the decision of whether to
-filter an event, the filter decision table also includes information about
-whether the destination flag should be written to and what value should be
-written to it. A third address calculation unit generates the write address for
-this destination flag. For example, for an ALU instruction, this will
-correspond to the destination register of the ALU instruction. If this ALU
-instruction is filtered out due to having a pair of null source operands, then
-the filter decision table will indicate that the destination register's
-corresponding flag should also be marked as null.
+% \subsection{Filtering Invalid Events}
+% \label{sec:dropping.filter}
+% 
+% % Filtering decision
+% The decision of whether an event can be filtered out is determined using a
+% lookup table. This filter decision table is configured by the monitoring scheme
+% on system initialization. The lookup table determines whether an event can be
+% filtered out based on the pair of input flags read and the instruction type of
+% the monitored event.
+% For example, on an ALU instruction, if both of the source operand flags
+% indicate that the corresponding metadata is null, then typically the instruction can be
+% filtered out. Similarly, on an ALU instruction, if either of the source operand
+% flags indicate that their corresponding metadata is invalid, then the
+% instruction can also be filtered out. This filter decision is sent to the
+% ``Filter'' block shown in Figure~\ref{fig:dropping.overview} which will simply pop
+% filtered entries from the FIFO and not forward them. If an entry is not
+% filtered, then this Filter block will pop the monitoring event entry from the
+% FIFO and forward it along the datapath.
+% 
+% % Propagating flag information
+% Finally, when an event is filtered out, it is necessary to propagate the flag
+% information for the filtered event. In addition to the decision of whether to
+% filter an event, the filter decision table also includes information about
+% whether the destination flag should be written to and what value should be
+% written to it. A third address calculation unit generates the write address for
+% this destination flag. For example, for an ALU instruction, this will
+% correspond to the destination register of the ALU instruction. If this ALU
+% instruction is filtered out due to having a pair of null source operands, then
+% the filter decision table will indicate that the destination register's
+% corresponding flag should also be marked as null.
 

### 2014-09-11 evaluation.tex
Update Evaluation.

@@
 different monitors, we implemented three different monitors: uninitialized
-memory check (UMC), array bounds check (BC), and dynamic information flow
-tracking (DIFT).  Uninitialized memory check seeks to detect loading from
+memory check (UMC), array bounds check (BC), dynamic information flow
+tracking (DIFT), and instruction-mix profiling (IMP).  Uninitialized memory
+check seeks to detect loading from
 memory locations that are not initialized first. Array bounds check, as
@@
 that if an error is detected, it is possible to have information about where
-the data originated from.
+the data originated from. Instruction-mix profiling counts the number of ALU,
+load, store, and control instructions that are executed. This profiling
+information can be useful for performing optimizations. Additionally, recent
+work has shown that this information can be used to detect malicious software \cite{tang-raid14}.
 
@@
 SPECint benchmarks. Although we do not
-show results for the C++ benchmarks, we note that the results for UMC and DIFT
+show results for the C++ benchmarks, we note that the results for UMC, DIFT, and IMP
 for these benchmarks are similar to the other results shown. For each
-benchmark, we simulated for 1 billion instructions. An initial slack of 1
+benchmark, we simulated for 2 billion instructions. An initial slack of 2
 million cycles, which is less than 1\% of
@@
     \vspace{-0.2in}
-    \caption{Full monitoring overheads for UMC, BC, and DIFT.}
+    \caption{Full monitoring overheads for UMC, BC, DIFT, and IMP.}
     \label{fig:evaluation.full_mon}
@@
 without monitoring. In these results, no filtering or partial monitoring is
-done. UMC shows normalized execution times from 4x to 11x with an average
-of 7x. BC shows normalized execution times of 9-28x with an average of
-18x while DIFT shows normalized execution times of 6x-17x with an average of
-11x. One of the reasons for these high overheads is that our implementations of
-these monitors dynamically allocate memory for new metadata. This can be an
-expensive, many-cycle operation. Statically allocating memory for metadata
-will reduce the execution time but requires a large upfront memory footprint
-which caused issues with our simulator.
+done. UMC shows normalized execution times from 5x to 16x with an average
+of 10x. BC shows normalized execution times of 8-24x with an average of
+16x while DIFT shows normalized execution times of 4x-14x with an average of
+9x. IMP shows normalized execution times from 2-6x with an average of 4x. 
+These monitors show high overheads because they require several
+instructions to process each monitored instruction from the main core. Even
+small increases in the number of instructions needed by the monitor lead to
+high overheads when multiplied across the main program's execution. For
+example, because of the large metadata requirements of UMC, BC, and DIFT, these
+monitors must dynamically allocate memory for new metadata. On the other hand,
+IMP does not require this dynamic allocation and is overall a simple monitor.
+Thus, its overheads are much lower.
 
@@
 reductions in overhead for all three monitoring schemes. UMC sees normalized
-execution times of 2-6x with an average of 4x with null metadata filtering.
-This is an average reduction in overheads of 52\%. For BC, normalized execution
-times drop from 18x down to 4x on average (an 82\% reduction). The range of normalized execution
-times for filtered BC is 1.1-17x. This is not surprising since in the baseline
+execution times of 1.7-7x with an average of 4x with null metadata filtering.
+This is an average reduction in overheads of 65\%. For BC, normalized execution
+times drop from 16x down to FIXMEx on average (an FIXME\% reduction). The range of normalized execution
+times for filtered BC is FIXME-FIXMEx. This is not surprising since in the baseline
 implementation all loads and stores needed to be monitored. However, with
 filtering, only loads and stores corresponding to arrays need to be forwarded.
-Finally, DIFT sees the largest reduction in overheads with only 13\% overheads
-on average after filtering, a 90\% reduction in monitoring overheads, with a
-maximum overhead of 42\% after filtering. This is due, in part, to the fact
+Finally, DIFT sees the largest reduction in overheads with only 10\% overheads
+on average after filtering, an 88\% reduction in monitoring overheads, with a
+maximum overhead of 31\% after filtering. This is due, in part, to the fact
 that
@@
 targeted network or streaming applications, which have larger amounts of
-untrusted input data, we would expect to see less filtering.
+untrusted input data, we would expect to see less filtering. IMP does not have
+null metadata and thus is not affected by null metadata filtering.
 
@@
     \includegraphics[width=\linewidth]{figs/data_bc_sweep.pdf}
-    \vspace{-0.4in}
+    \vspace{-0.2in}
     \caption{Coverage versus varying overhead budget for array bounds check.}
@@
     \includegraphics[width=\linewidth]{figs/data_umc_sweep.pdf}
-    \vspace{-0.4in}
+    \vspace{-0.2in}
     \caption{Coverage versus varying overhead budget for uninitialized memory check.}
@@
 
+% IMP sweep
+\begin{figure*}
+  \begin{center}
+    \includegraphics[width=\linewidth]{figs/data_imp_sweep.pdf}
+    \vspace{-0.2in}
+    \caption{Error versus varying overhead budget for instruction-mix profiling.}
+    \label{fig:evaluation.imp_sweep}
+    \vspace{-0.1in}
+  \end{center}
+\end{figure*}
+
 Although the overheads of DIFT are quite low after filtering, BC and UMC still
-show significant overheads. In this section, we evaluate the effectiveness of
+show significant overheads. Similarly, the overheads of IMP are not improved by filtering. In this section, we evaluate the effectiveness of
 using partial monitoring (in addition to filtering) to trade-off coverage for further reduced overheads.
@@
 we see that varying overhead budgets enables partial monitoring. We see that
-with a 10\% overhead budget, UMC achieves 32\% monitoring coverage on average.
-By increasing this overhead budget to 50\%, UMC achieves 51\% coverage. Even
+with a 10\% overhead budget, UMC achieves 30\% monitoring coverage on average.
+By increasing this overhead budget to 50\%, UMC achieves 48\% coverage. Even
 higher coverage can be achieved by allowing higher overheads.
+
+The instruction-mix profiling monitor does not perform check operations. Thus,
+the concept of coverage is not applicable here. Instead, for each instruction
+type profiled, we calculate its count as a percentage of the total instructions
+monitored. We compare these percentages against the case when full monitoring
+is performed. Figure~\ref{fig:evaluation.imp_sweep} shows the max error across
+instruction types for each benchmark and for varying overheads. We see that
+with a 10\% overhead a max error of 9\% is observed and an average of less than
+4\%. This error decreases as overhead is increased.
 
@@
     \vspace{-0.2in}
-    \caption{Coverage over multiple runs for BC at 10\% overhead with an unrestricted dropping policy.}
+    \caption{Coverage over multiple runs for BC with a 10\% probability to not drop events.}
     \label{fig:evaluation.multirun}
@@
 achieve high coverage across multiple runs. Varying the monitoring that is
-done for different runs can be achieved by introducing some randomness into the
-dropping decision. Figure~\ref{fig:evaluation.multirun} shows the total
-coverage achieved over multiple runs for array bounds check running with a 10\%
-overhead budget and an unrestricted dropping policy. Each run was simulated for
-100 million instructions. For these simulations, events were
-dropped probabilistically with the probability scaled proportionally to slack.
-That is, when slack is high, the probability of dropping is low and the
-probability increases as we run out of slack. We see that the coverage achieved
-quickly asymptotes for an unrestricted dropping policy. This is due to the fact
-that since there are multiple dependent monitoring operations that must be done
-in order for a final check to be successful, a small amount of randomness at
-each event quickly compounds to reduce the probability of certain checks to
-near zero. Instead, source-only dropping is much better suited for achieving high
-coverage over multiple runs.
+done for different runs can be achieved by randomly deciding whether events
+should be dropped.
+Figure~\ref{fig:evaluation.multirun} shows the total
+coverage achieved over multiple runs for array bounds check running for
+unrestricted and source-only dropping policies. Here, we use a 10\% probability
+that events are not dropped. Since the goal is code coverage, coverage here is
+measured as the percentage of static instructions which are monitored. Each run
+was simulated for 200 million instructions. 
+We see that for the unrestricted dropping policy there is almost no increase in
+coverage.  This is due to the fact that since there are multiple dependent
+monitoring operations that must be done in order for a final check to be
+successful, a small amount of randomness at each event quickly compounds to
+reduce the probability of certain checks to near zero. Instead, source-only
+dropping is much better suited for achieving high coverage over multiple runs.
+Source-only dropping shows a 5\% increase in coverage with only seven runs.
+Over hundreds or thousands of runs, source-only dropping should be able to
+reach 100\% code coverage.
 
@@
 shows the peak and runtime power overheads averaged across all benchmarks with
-a 50\% monitoring overhead target. The peak power is 5-7 mW, which is
-less than 1\% of the main core's peak power usage. Similarly, the average runtime power is 3-4 
-mW, corresponding to about 1\% of the main core's runtime power.
+a 50\% monitoring overhead target. The peak power is 41-72 mW, which is
+less than 8\% of the main core's peak power usage. Similarly, the average runtime power is 20-43
+mW, corresponding to 5-11\% of the main core's runtime power.

### 2014-09-11 dropping.tex
typo fixes. added comments that need to be addressed.

@@
 describes our baseline model for run-time monitoring while the rest of the
-Section describes our design.
+section describes our design.
 
@@
 \end{figure}
+
+ED: monitoring core vs. monitor?
 
@@
 an error is detected.
+
+ED: describe reg/mem meta-data both in the figure and in the text.
 

### 2014-09-11 evaluation.tex
typo fixes. added comments that need to be addressed.

@@
 In order to explore the generality of the architecture for
-different monitors, we implemented three different monitors: uninitialized
+different monitors, we implemented four different monitors: uninitialized
 memory check (UMC), array bounds check (BC), dynamic information flow
@@
 work has shown that this information can be used to detect malicious software \cite{tang-raid14}.
+
+ED: We need to point out that these monitors cannot be implemented with a simple
+1-bit dataflow engine (DIFT) that we use for filtering. 
+Point: address potential reviewer questions on why don't we just use a 
+programmable dataflow engine for monitoring.
 
@@
 Thus, its overheads are much lower.
+
+ED: we need to better justify the high overhead. 
+Key points: 1) the overhead is in line with LBA, which represent multi-core
+monitoring w/ low HW costs.
+2) If we are going to mention dynamic memory allocation. We need to say why
+it's really necessary -- it's not our implementation problem. It's something
+that you need to have for large meta-data. Also, I don't think UMC would need
+dynamic allocation.
 
@@
 We define \emph{monitoring coverage} as the 
-percentage of checks that are performed 
+percentage of dynamic checks that are performed 
 (indirect jumps in DIFT, loads in UMC, and memory accesses in BC). 
 The metric is chosen to understand
-how likely an error/attack instance is to be detected on an individual system. 
-While we could not evaluate real errors/attacks due to difficulty in setting up
-a large number of bugs and exploits, we believe that the percentage of checks
-provides a good estimate of detection probability when errors/attacks are 
-uniformly distributed across checks.
+how likely an error/attack instance is to be detected on an individual system, 
+assuming that errors/attacks are uniformly distributed across checks. 
+%While we could not evaluate real errors/attacks due to difficulty in setting up
+%a large number of bugs and exploits, we believe that the percentage of checks
+%provides a good estimate of detection probability when errors/attacks are 
+%uniformly distributed across checks.
 Note that this metric is different from the

### 2014-09-11 monitoring.tex
typo fixes. added comments that need to be addressed.

@@
 hardware monitors cannot be modified or updated, and require dedicated silicon area. 
-One the other hand, flexible systems that support a wide range of monitors lead 
+On the other hand, flexible systems that support a wide range of monitors lead 
 to noticeable performance overhead, often too high for wide deployment in practice.
@@
 slowdowns of several tens of percents in some cases.
-In today's monitoring systems, the overhead is often unpredictable because they
+In today's monitoring systems, the overhead is also unpredictable because they
 depend heavily on the characteristics of applications and monitoring operations.
-In summary, users need to either pay noticeable overhead or the cost of custom
+In summary, users currently need to either pay noticeable overhead or the cost of custom
 hardware in order to use fine-grained run-time monitoring in deployed systems.
@@
 to use monitoring with partial coverage in order to reduce performance
-overhead. Alternatively, partial monitoring can allow users to use simple
-one may choose to use simple hardware such as a
-multi-core without elaborate acceleration features to provide partial monitoring
-with the performance overhead that is comparable to custom hardware.
+or energy overhead. Alternatively, partial monitoring allows designers to use
+less expensive hardware for a given performance overhead budget.
+%simple hardware such as a
+%multi-core without elaborate acceleration features to provide partial monitoring
+%with the performance overhead that is comparable to custom hardware.
 
-In this framework, a user can specify how much monitoring should be done such as a target 
+In this framework, a user specifies how much monitoring should be done in a form of
+a target 
 overhead budget, a target coverage, a percentage of monitoring operations to be performed, etc.
@@
 while maximizing the coverage. Given that the overhead of custom hardware monitors can
-already be quite low, the focus will be enabling the trade-off in {\em general-purpose} 
+already be quite low, the focus is on enabling the trade-off in {\em general-purpose} 
 monitoring systems that support a wide range of monitors.
@@
 {\bf Monitoring of soft real-time systems:}
-Soft real-time systems or interactive systems often need to meet deadlines or response
+Soft real-time systems or interactive systems need to meet deadlines or response
 time requirements. Unfortunately, the overhead of run-time monitoring is often 
-unpredictable and varies significantly depending on the application and monitoring
+unpredictable and varies significantly depending on the application and monitor
 characteristics. The partial monitoring framework allows monitoring to be performed
@@
 {\bf Partial protection for low overhead:}
-Even without real-time constraints, systems may have tight budgets for performance
-or hardware overhead that they can tolerate for run-time monitoring. In such cases,
+Even without real-time constraints, systems may have tight budgets for performance,
+energy, or hardware overhead that they can tolerate for run-time monitoring. In such cases,
 full monitoring cannot be enabled unless its overhead is low enough. The adjustable
 monitoring allows partial protection on such systems. For example, array bounds
-may be checked for a portion of memory accesses, or a subset of information flows
-may be tracked in DIFT for attack detection. In this scenario, the effectiveness of 
+may be checked for a subset of memory accesses. For DIFT, a subset of information flows
+may be tracked for partial attack detection. In this scenario, the effectiveness of 
 partial monitoring can be measured as the percentage of run-time checks that are
@@
 \begin{enumerate}
-  \item \textbf{General:} The framework needs to be general enough to be applicable to 
-  a wide range of monitoring schemes. Because we target flexible, general-purpose
-  run-time monitoring systems, a solution that only works for one monitoring scheme will
-  not work.
+  \item \textbf{General:} Because we mainly target flexible run-time monitoring systems,
+  which often have high overhead, the framework also needs to be general enough to be
+  applicable to a wide range of monitoring scheme.
 
   \item \textbf{No false positive:} The framework needs to ensure that dropping a portion
-  of monitoring operations does not lead to a false positive. We solve this problem using
-  a data-flow engine that tracks invalid monitoring meta-data (Section~\ref{sec:dropping}).
+  of monitoring operations does not lead to a false positive. We found that a data-flow
+  that tracks invalid meta-data can serve as a general solution to this problem 
+  (Section~\ref{sec:dropping}).
 
   \item \textbf{Intelligent dropping:} The framework needs to match the overhead budget
-  while maximizing the effectiveness of monitoring. To this end, adjustable overhead
+  while maximizing the effectiveness of monitoring. To this end, partial 
   monitoring needs to carefully choose which operations to drop and when. We address this

### 2014-09-11 evaluation.tex
imported an authcomments package from Andrew Myers

@@
 
-ED: We need to point out that these monitors cannot be implemented with a simple
+\ED{We need to point out that these monitors cannot be implemented with a simple
 1-bit dataflow engine (DIFT) that we use for filtering. 
 Point: address potential reviewer questions on why don't we just use a 
-programmable dataflow engine for monitoring.
+programmable dataflow engine for monitoring.}
 
@@
 
-ED: we need to better justify the high overhead. 
+\ED{we need to better justify the high overhead. 
 Key points: 1) the overhead is in line with LBA, which represent multi-core
@@
 that you need to have for large meta-data. Also, I don't think UMC would need
-dynamic allocation.
+dynamic allocation.}
 

### 2014-09-11 dropping.tex
imported an authcomments package from Andrew Myers
Moved null filtering into Section 3.

@@
     \caption{Overview of run-time monitoring architecture.}
-    \label{fig:monitoring.overview} 
+    \label{fig:dropping.overview} 
     \vspace{-0.1in}
@@
 
-ED: monitoring core vs. monitor?
+\ED{Should we use monitoring core or just monitor?}
 
-Figure~\ref{fig:monitoring.overview} shows an overview of the run-time monitoring
+Figure~\ref{fig:dropping.overview} shows an overview of the run-time monitoring
 model that is assumed in this paper.  The \emph{main program} is a computation
@@
 
-ED: describe reg/mem meta-data both in the figure and in the text.
+\ED{describe reg/mem meta-data both in the figure and in the text.}
 
@@
     \caption{Example of array bounds check monitor.}
-    \label{fig:monitoring.example_full}
+    \label{fig:dropping.example_full}
     \vspace{-0.1in}
@@
 
-Figure~\ref{fig:monitoring.example_full} shows an example pseudo-code segment,
+Figure~\ref{fig:dropping.example_full} shows an example pseudo-code segment,
 its assembly level instructions, and the corresponding monitoring operations
@@
 metadata. Dropping monitoring operations that check for an error (such as check
-found for line 4 in Figure~\ref{fig:monitoring.example_full} can only cause
+found for line 4 in Figure~\ref{fig:dropping.example_full} can only cause
 false negatives but will never cause false positives. On the other hand,
@@
 % dataflow engine marks this destination flag as invalid.
+
+\subsection{Filtering Null Metadata}
+\label{sec:dropping.null_filtering}
+
+% Null filtering code example
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\columnwidth]{figs/example_null_filtering.pdf}
+    \vspace{-0.2in}
+    \caption{Example of using information about null metadata to filter monitoring events.}
+    \label{fig:dropping.example_null} 
+    \vspace{-0.1in}
+  \end{center}
+\end{figure}
+
+One way to reduce the number of monitoring events that must be handled by the
+monitor is to filter out monitoring events that correspond to operating on null
+or uninitialized metadata. Null metadata typically corresponds to events that
+are not relevant to the monitor. For example, Hardbound
+\cite{hardbound-asplos08} filters out operations on non-pointer (i.e., no base
+and bounds metadata) instructions since it is not relevant to array bounds
+checking. More recently, FADE \cite{fade-hpca14} has been proposed a general
+hardware module to perform this null metadata filtering for a variety of
+monitoring schemes.
+
+Figure~\ref{fig:dropping.example_null} shows an example of how this null filtering operates.
+Here, the main core's code has been slightly modified from
+Figure~\ref{fig:dropping.example_full} and {\tt r2} on line 2 is no longer set as an array pointer.
+Without null metadata filtering, the instructions for lines 2 and 4 would be
+forwarded to the monitor since the system does not know whether {\tt r2}
+contains a pointer address or not. However, if we use a flag to mark {\tt r2}
+as null when it is loaded with a constant, then we can propagate this null
+information and filter out monitoring for lines 2 and 4.
+
+The operations performed by null filtering are almost identical to the
+operations needed for invalidation shown in
+Figure~\ref{fig:dropping.example_invalid}.
+Thus, we can easily enable this null metadata filtering on our architecture by
+simply extending the dataflow flags to be two bits wide. One bit is used to
+keep track of invalidation information while the second bit is used to keep
+track of null information. All flags are initialized to null and the filter
+decision table is extended with the propagation and filtering decision rules
+for null metadata filtering.

### 2014-09-12 evaluation.tex
Add results for FPGA-based monitor.
Also filled in missing data for filtered/bc sweep.

@@
 
+% Flex UMC sweep
+\begin{figure*}
+  \begin{center}
+    \includegraphics[width=\linewidth]{figs/data_fpga_umc_sweep.pdf}
+    \vspace{-0.2in}
+    \caption{Coverage versus varying overhead budget for UMC running on an FPGA-based monitor.}
+    \label{fig:evaluation.fpga_umc_sweep}
+    \vspace{-0.1in}
+  \end{center}
+\end{figure*}
+
+In addition to evaluating our system for a core-based monitor, we also
+evaluated partial monitoring on a higher performance FPGA-based monitor. This
+setup is based on the FlexCore \cite{flexcore-micro10} setup and uses a
+fully-pipelined monitor running on an FPGA fabric at half the frequency of the
+main core. We show results only for UMC. In this case, full overheads of
+monitoring range from 10-70\%. Figure~\ref{fig:evaluation.fpga_umc_sweep} shows
+the coverage achieved as we sweep overhead from 1-70\%. We see that partial
+monitoring can allow the overheads of such a system to be pushed below 10\%
+while still providing partial coverage.
+
 \subsection{Comparing Dropping Policies}

### 2014-09-12 evaluation.tex
Added discussion on random dropping to policies.

@@
 the total execution time, is given for a 10\% target overhead. This initial
-slack is scaled proportionally for other overhead targets.
+slack is scaled proportionally for other overhead targets. Unless otherwise
+specified, an unrestricted, slack-based dropping policy is used.
 
@@
 to the fact that some of the overheads of monitoring for unrestricted dropping
-are being spent on wasted work as discussed in Section~\ref{sec:policies.events}.
+are being spent on wasted work as discussed in Section~\ref{sec:policies.which}.
 
@@
 One possible application of partial monitoring with low overheads is to enable
-cooperative debugging. The idea with cooperative debugging is to use partial
+cooperative debugging. 
+The idea with cooperative debugging is to use partial
 monitoring with very low overheads across a large number of users or runs. By
@@
 achieve high coverage across multiple runs. Varying the monitoring that is
-done for different runs can be achieved by randomly deciding whether events
-should be dropped.
+done for different runs can be achieved by using a probabilistic dropping policy.
 Figure~\ref{fig:evaluation.multirun} shows the total
 coverage achieved over multiple runs for array bounds check running for
-unrestricted and source-only dropping policies. Here, we use a 10\% probability
+unrestricted and source-only dropping policies with probabilistic dropping. Here, we use a 10\% probability
 that events are not dropped. Since the goal is code coverage, coverage here is
@@
 monitoring operations that must be done in order for a final check to be
-successful, a small amount of randomness at each event quickly compounds to
-reduce the probability of certain checks to near zero. Instead, source-only
+successful as discussed in Section~\ref{sec:policies.when}.
+Instead, source-only
 dropping is much better suited for achieving high coverage over multiple runs.

### 2014-09-12 policies.tex
Added discussion on random dropping to policies.

@@
 
-Our goal with enabling partial monitoring is to allow for a specified overhead target or budget to be met while maximizing the amount of monitoring performed.
-This is done by dynamically dropping a portion of the monitoring operations. In this section, we discuss in detail how this dropping decision is
-made. Specifically, we first discuss how to measure the run-time monitoring
-overheads and when monitoring events need to be dropped in
-Section~\ref{sec:policies.slack}. Next, we investigate trade-offs created by
-selecting which events are dropped (Section~\ref{sec:policies.events}). Different policies in determining which
-events can be dropped create a trade-off between how closely the overhead
-budget is met and the monitoring coverage achieved.
+In order to use partial monitoring to enable adjustable overheads, we must also
+specify a policy for when and which monitoring events are dropped. In this
+section, we discuss some of the options and trade-offs for dropping policies.
+We split this decision into two components:
+\begin{enumerate}
+  \item When do we need to drop events in order to enable reduced overheads?
+  \item Which events should be valid for dropping?
+\end{enumerate}
 
-\subsection{Measuring Monitoring Overheads}
-\label{sec:policies.slack}
+Section~\ref{sec:policies.when} discusses options for deciding when events are
+dropped while Section~\ref{sec:policies.which} discusses trade-offs for which
+events are capable of being dropped.
+
+% Our goal with enabling partial monitoring is to allow for a specified overhead target or budget to be met while maximizing the amount of monitoring performed.
+% This is done by dynamically dropping a portion of the monitoring operations. In this section, we discuss in detail how this dropping decision is
+% made. Specifically, we first discuss how to decide when dropping should be performed 
+% Section~\ref{sec:policies.slack}. Next, we investigate trade-offs created by
+% selecting which events are dropped (Section~\ref{sec:policies.events}). Different policies in determining which
+% events can be dropped create a trade-off between how closely the overhead
+% budget is met and the monitoring coverage achieved.
+
+\subsection{Deciding When to Drop}
+\label{sec:policies.when}
 
@@
 
-% Slack tracking module
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\columnwidth]{figs/stm.pdf}
-    \vspace{-0.2in}
-    \caption{Slack tracking and drop decision hardware.}
-    \label{fig:policies.stm}
-    \vspace{-0.1in}
-  \end{center}
-\end{figure}
+In this section, we discuss two possible ways to determine when events should be dropped.
+The first possibility is to probabilistically drop events. By setting the probability of
+dropping events appropriately, overheads can be reduced. This is works well for
+enabling partial monitoring for cooperative testing and debugging since the
+randomness allows different users and runs to monitor different portions of the
+program. However, using a probabilistic dropping policy can make it difficult
+to meet a target overhead without prior profiling.
 
-In order to decide whether a monitoring event should be dropped or not, we need
-a way to keep track of overheads. In this section, we present a specific method
+Alternatively, we can specify a target overhead and estimate, at run-time, the
+overheads of monitoring in order to decide whether dropping is needed.
+% In order to decide whether a monitoring event should be dropped or not, we need
+% a way to keep track of overheads. 
+We present a specific method
 for measuring execution time overheads. However, we note that other methods
-could be used in order to target budgets on IPC, power, energy, etc.  
+could also be used in order to target budgets on IPC, power, energy, etc.  
 % For example, one way to measure execution time overheads would be to insert
@@
 Figure~\ref{fig:policies.slack} shows an example of how slack can change over time.
-If the slack falls below zero (i.e., the overhead budget is exceeded), then
+In this slack-based policy, if the slack falls below zero (i.e., the overhead budget is exceeded), then
 monitoring events should be dropped.
 
-Figure~\ref{fig:policies.stm} shows a hardware slack tracking module for
-keeping track of slack. The slack tracking module uses a counter that increments on every $k$-th
+% % Slack tracking module
+% \begin{figure}
+%   \begin{center}
+%     \includegraphics[width=\columnwidth]{figs/stm.pdf}
+%     \vspace{-0.2in}
+%     \caption{Slack tracking and drop decision hardware.}
+%     \label{fig:policies.stm}
+%     \vspace{-0.1in}
+%   \end{center}
+% \end{figure}
+
+% Figure~\ref{fig:policies.stm} shows a hardware slack tracking module for
+% keeping track of slack. 
+Slack can be easily measured in hardware by using a counter that increments on every $k$-th
 cycle of the main core. This $k$ can be calculated by taking the reciprocal of
@@
 increments on every 5th clock cycle. The value of this counter is the
-accumulated slack. Whenever the main core is stalled due to the monitor, slack
+accumulated slack. Whenever the main core is stalled due to the monitor, the measured slack
 is decremented. This can be difficult to determine precisely due to the
@@
 
-\subsection{Dropping Decision Trade-Offs}
-\label{sec:policies.events}
+\subsection{Deciding Which Events to Drop} 
+\label{sec:policies.which}
 
@@
 
-The simplest policy for when to drop monitoring events is to drop all monitoring events as
+In addition to deciding when dropping is required, trade-offs also exist in
+deciding which dropping events can be dropped.
+The simplest policy is to drop all monitoring events as
 long as slack is less than or equal to zero.  However, this can result in
@@
 to work well when there are a large number of independent metadata flows.
+In addition, using a probabilistic dropping policy works poorly with
+unrestricted dropping since each event in a chain needs to be randomly selected
+for monitoring in order for the check to be performed. Instead, probabilistic
+dropping works better when coupled with source-only dropping.
 

### 2014-09-12 evaluation.tex
Fix up citations.

@@
 This is an average reduction in overheads of 65\%. For BC, normalized execution
-times drop from 16x down to FIXMEx on average (an FIXME\% reduction). The range of normalized execution
-times for filtered BC is FIXME-FIXMEx. This is not surprising since in the baseline
+times drop from 16x down to 3x on average (an 80\% reduction). The range of normalized execution
+times for filtered BC is 1.1-8.3x. This is not surprising since in the baseline
 implementation all loads and stores needed to be monitored. However, with

### 2014-09-12 monitoring.tex
Fix up citations.

@@
 to noticeable performance overhead, often too high for wide deployment in practice.
-Software-only implementations \cite{FIXME} or multi-core monitors with minimal
+Software-only implementations \cite{nagarajan-interact08, lift-micro06,
+purify-usenix92, taintcheck-ndsss05} or multi-core monitors with minimal
 hardware changes \cite{lba-asid06} are reported to have severalfold slowdowns.
@@
 {\bf Cooperative testing, debugging, and protection:}
-Recently studies \cite{FIXME} have shown that software testing, debugging,
+Recently studies \cite{liblit-pldi05, chilimbi-asplos04, greathouse-cgo11, testudo-micro08} have shown that software testing, debugging,
 or attack detection may be done in a cooperative fashion across
@@
 as security protection. For example, a recent study showed that an instruction
-mix can be used to identify malware from normal programs \cite{FIXME}. 
+mix can be used to identify malware from normal programs \cite{demme-isca13, tang-raid14}. 
 In such profiling tools, the partial monitoring framework can be used to obtain

### 2014-09-12 tabs/previous_overheads.tex
Fix up citations.

@@
 FlexCore \cite{flexcore-micro10} & Dedicated FPGA & Instruction-trace monitoring & 1.05x-1.44x / 1.84x \\ \hline
-FADE \cite{FIXME} & Core+Custom HW & Instruction-trace monitoring (effective when HW filters work) & \\ \hline
+FADE \cite{fade-hpca14} & Core+Custom HW & Instruction-trace monitoring (effective when HW filters work) & 1.2x-1.8x / 3.3x \\ \hline
 LBA-accelerated \cite{lba-isca08} & Multi-core+Custom HW & Instruction-trace monitoring (effective when accelerators work) & 1.02x-3.27x / 5x \\ \hline
@@
 Purfiy \cite{purify-usenix92} & SW (DBI) & Memory leak checks (fully flexible) & 2.3x / 5.5x \\ \hline
-TaintCheck \cite{taintcheck-05} & SW (DBI) & DIFT (fully flexible) & 10x / 27x \\ \hline
+TaintCheck \cite{taintcheck-ndsss05} & SW (DBI) & DIFT (fully flexible) & 10x / 27x \\ \hline
 

### 2014-09-12 dropping.tex
Edits to Section 3 based on Ed's comments.

@@
 
-\ED{Should we use monitoring core or just monitor?}
-
 Figure~\ref{fig:dropping.overview} shows an overview of the run-time monitoring
@@
 \emph{main core}.  On certain events during the main program, such as the
-execution of certain types of instructions, the \emph{monitoring core} performs a
-series of \emph{monitoring operations}. The monitoring core operates in parallel to the
+execution of certain types of instructions, the \emph{monitor} performs a
+series of \emph{monitoring operations}. The monitor operates in parallel to the
 main core. These events are referred to as \emph{monitoring events}. Depending
 on the type of monitoring event, different monitoring operations are
-executed. Information about monitoring events are sent to the monitoring core and buffered in a FIFO structure to decouple the
-running of the main core and the monitoring core. If the FIFO is full, then the main
+executed. Information about monitoring events is sent to the monitor and buffered in a FIFO structure to decouple the
+running of the main core and the monitor. If the FIFO is full, then the main
 core is forced to stall on a monitoring event until a FIFO entry becomes
 available. These stalls are a major source of
-overhead because the monitoring core may take several cycles to process a single event
+overhead because the monitor may take several cycles to process a single event
 from the main core. 
+Monitors typically maintain a set of metadata corresponding to the main
+program's memory. Similarly, monitors also typically maintain a set of metadata
+corresponding to the main core's register file.
 %We refer to these stalls and other overheads, such as
 %contention for shared resources, as \emph{monitoring overheads}. 
-If the monitoring core
+If the monitor
 detects an inconsistent or undesired behavior in the monitoring events, then
 an error is detected.
-
-\ED{describe reg/mem meta-data both in the figure and in the text.}
 
@@
 array
-pointer, the monitoring core checks that the memory address accessed is within the base and
+pointer, the monitor checks that the memory address accessed is within the base and
 bound addresses. In addition, this base and bound metadata
@@
 
-Here, we have shown the dynamic sequence of operations of the monitoring core.
+Here, we have shown the dynamic sequence of operations of the monitor.
 The actual static code consists of a set of instructions to be run for each
@@
 positives, we propose to insert a hardware module between the main and
-monitoring core (see Figure~\ref{fig:dropping.dataflow_overview}). 
+monitor (see Figure~\ref{fig:dropping.dataflow_overview}). 
 This module handles the invalidation operations shown in the middle column of

### 2014-09-12 evaluation.tex
Revised Evaluation.

@@
 work has shown that this information can be used to detect malicious software \cite{tang-raid14}.
+Note that, although our dataflow engine performs similar operations to the DIFT
+monitor, the dataflow engine operates on smaller, 2-bit flags.
 
@@
     \vspace{-0.2in}
-    \caption{Full monitoring overheads for UMC, BC, DIFT, and IMP.}
+    \caption{Full monitoring overheads.}
     \label{fig:evaluation.full_mon}
@@
 9x. IMP shows normalized execution times from 2-6x with an average of 4x. 
-These monitors show high overheads because they require several
-instructions to process each monitored instruction from the main core. Even
-small increases in the number of instructions needed by the monitor lead to
-high overheads when multiplied across the main program's execution. For
-example, because of the large metadata requirements of UMC, BC, and DIFT, these
-monitors must dynamically allocate memory for new metadata. On the other hand,
-IMP does not require this dynamic allocation and is overall a simple monitor.
-Thus, its overheads are much lower.
+% These monitors show high overheads because they require several
+% instructions to process each monitored instruction from the main core. Even
+% small increases in the number of instructions needed by the monitor lead to
+% high overheads when multiplied across the main program's execution.
 
@@
 
-\subsection{Filtering of Monitoring Events}
-
-% Full monitoring overheads
+% Filter monitoring overheads
 \begin{figure}
@@
     \vspace{-0.2in}
-    \caption{Monitoring overheads with filtering null metadata for UMC, BC, and DIFT.}
+    \caption{Monitoring overheads with filtering null metadata.}
     \label{fig:evaluation.filtering}
@@
 times drop from 16x down to 3x on average (an 80\% reduction). The range of normalized execution
-times for filtered BC is 1.1-8.3x. This is not surprising since in the baseline
-implementation all loads and stores needed to be monitored. However, with
-filtering, only loads and stores corresponding to arrays need to be forwarded.
+times for filtered BC is 1.1-8.3x. 
+% This is not surprising since in the baseline
+% implementation all loads and stores needed to be monitored. However, with
+% filtering, only loads and stores corresponding to arrays need to be forwarded.
 Finally, DIFT sees the largest reduction in overheads with only 10\% overheads
@@
 maximum overhead of 31\% after filtering. This is due, in part, to the fact
-that
-for our implementation of DIFT on SPEC
+that for our implementation of DIFT on SPEC
 benchmarks, we only mark data read from files as tainted. For most of these
@@
 null metadata and thus is not affected by null metadata filtering.
+These overheads are similar to previous results for multi-core monitoring with
+minimal hardware costs such as LBA \cite{lba-asid06} (see
+Table~\ref{tab:monitoring.previous_overheads}). In
+Section~\ref{sec:evaluation.fpga}, we evaluate a higher performance, FPGA-based
+monitor that shows low tens of percent of overhead.
 
@@
 % BC sweep
-\begin{figure*}
+\begin{figure}
   \begin{center}
@@
   \end{center}
-\end{figure*}
+\end{figure}
 
 % UMC sweep
-\begin{figure*}
+\begin{figure}
   \begin{center}
@@
   \end{center}
-\end{figure*}
+\end{figure}
 
@@
 We see that by varying the overhead budget, the coverage achieved also varies.
-With only a 10\% overhead budget, array bounds check still
-achieves 82\% monitoring coverage on average. The high coverage
+With only a 1.1x overhead target, array bounds check still
+achieves 84\% monitoring coverage on average. The high coverage
 achieved with such low overheads is due to two main effects.  The first is that
@@
 we see that varying overhead budgets enables partial monitoring. We see that
-with a 10\% overhead budget, UMC achieves 30\% monitoring coverage on average.
-By increasing this overhead budget to 50\%, UMC achieves 48\% coverage. Even
-higher coverage can be achieved by allowing higher overheads.
+with a 1.1x overhead target, UMC achieves 25\% monitoring coverage on average.
+By increasing this overhead budget to 1.5x, UMC achieves 43\% coverage on
+average. Even higher coverage can be achieved by allowing higher overheads.
 
@@
 instruction types for each benchmark and for varying overheads. We see that
-with a 10\% overhead a max error of 9\% is observed and an average of less than
-4\%. This error decreases as overhead is increased.
+with a 1.1x overhead a max error of 9\% is observed and an average of less than
+4\%. This error decreases as overhead is increased and is 0\% for all
+benchmarks once a 6x overhead target is used.
+
+\subsection{FPGA-based Monitor}
+\label{sec:evaluation.fpga}
 
 % Flex UMC sweep
-\begin{figure*}
+\begin{figure}
   \begin{center}
@@
 
-% UMC exec time
+% % UMC exec time
+% \begin{figure}
+%   \begin{center}
+%     \includegraphics[width=\columnwidth]{figs/data_umc_exec_time.pdf}
+%     \vspace{-0.2in}
+%     \caption{Error in meeting overhead budget for uninitialized memory check for different dropping policies.}
+%     \label{fig:evaluation.umc_exec_time}
+%     \vspace{-0.2in}
+%   \end{center}
+% \end{figure}
+% 
+% % UMC coverage across policies
+% \begin{figure}
+%   \begin{center}
+%     \includegraphics[width=\columnwidth]{figs/data_umc_coverage.pdf}
+%     \vspace{-0.2in}
+%     \caption{Coverage for uninitialized memory check for different dropping policies.}
+%     \label{fig:evaluation.umc_coverage}
+%     \vspace{-0.1in}
+%   \end{center}
+% \end{figure}
+
 \begin{figure}
   \begin{center}
+  \vspace{-0.2in}
+  \subfloat[Overhead budget error.]{
     \includegraphics[width=\columnwidth]{figs/data_umc_exec_time.pdf}
+    \label{fig:evaluation.umc_exec_time}
+  }
   \vspace{-0.2in}
-    \caption{Error in meeting overhead budget for uninitialized memory check for different dropping policies.}
-    \label{fig:evaluation.umc_exec_time}
+  \subfloat[Coverage.]{
+    \includegraphics[width=\columnwidth]{figs/data_umc_coverage.pdf}
+    \label{fig:evaluation.umc_coverage}
+  }
+  \caption{Comparing dropping policies for UMC.}
   \vspace{-0.2in}
   \end{center}
+  \label{fig:evaluation.umc_policies}
 \end{figure}
 
-% UMC coverage across policies
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\columnwidth]{figs/data_umc_coverage.pdf}
-    \vspace{-0.2in}
-    \caption{Coverage for uninitialized memory check for different dropping policies.}
-    \label{fig:evaluation.umc_coverage}
-    \vspace{-0.1in}
-  \end{center}
-\end{figure}
 In this section we evaluate the trade-offs between unrestricted dropping and source-only dropping.
+The results are shown for an overhead target of 1.5x.
 Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
@@
 Figure~\ref{fig:evaluation.umc_coverage} shows the coverage of UMC for
-unrestricted dropping and source-only dropping. We see that source-dropping
-consistently achieves higher coverage than unrestricted dropping. This is due
+unrestricted dropping and source-only dropping. We see that, in several cases, source-only dropping
+achieves higher coverage than unrestricted dropping. On average, unrestricted
+dropping achieves 46\% coverage while source-only dropping achieves 49\%
+coverage. This is due
 to the fact that some of the overheads of monitoring for unrestricted dropping
@@
 
-% BC exec time
+% % BC exec time
+% \begin{figure}
+%   \begin{center}
+%     \includegraphics[width=\columnwidth]{figs/data_bc_exec_time.pdf}
+%     \vspace{-0.2in}
+%     \caption{Error in meeting overhead budget for array bounds check for different dropping policies.}
+%     \label{fig:evaluation.bc_exec_time}
+%     \vspace{-0.2in}
+%   \end{center}
+% \end{figure}
+% 
+% % BC coverage across policies
+% \begin{figure}
+%   \begin{center}
+%     \includegraphics[width=\columnwidth]{figs/data_bc_coverage.pdf}
+%     \vspace{-0.2in}
+%     \caption{Coverage for array bounds check for different dropping policies.}
+%     \label{fig:evaluation.bc_coverage}
+%     \vspace{-0.2in}
+%   \end{center}
+% \end{figure}
+
 \begin{figure}
   \begin{center}
+  \vspace{-0.2in}
+  \subfloat[Overhead budget error.]{
     \includegraphics[width=\columnwidth]{figs/data_bc_exec_time.pdf}
+    \label{fig:evaluation.bc_exec_time}
+  }
   \vspace{-0.2in}
-    \caption{Error in meeting overhead budget for array bounds check for different dropping policies.}
-    \label{fig:evaluation.bc_exec_time}
+  \subfloat[Coverage.]{
+    \includegraphics[width=\columnwidth]{figs/data_bc_coverage.pdf}
+    \label{fig:evaluation.bc_coverage}
+  }
+  \caption{Comparing dropping policies for BC.}
   \vspace{-0.2in}
   \end{center}
-\end{figure}
-
-% BC coverage across policies
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\columnwidth]{figs/data_bc_coverage.pdf}
-    \vspace{-0.2in}
-    \caption{Coverage for array bounds check for different dropping policies.}
-    \label{fig:evaluation.bc_coverage}
-    \vspace{-0.2in}
-  \end{center}
+  \label{fig:evaluation.bc_policies}
 \end{figure}
@@
 benchmarks, source-only drop fails to meet the specified overhead target.
-The overshoot of the overhead target is quite high with an overhead difference
+The overshoot of the overhead target is quite high with an overhead differences
 of 420\% for {\tt bzip2}, 120\% for {\tt gcc}, and 1500\% for {\tt hmmer}.
@@
 difficult for source-only dropping to match overhead targets. Although
-Figure~\ref{fig:evaluation.bc_coverage} again shows higher coverage for
-source-only dropping, this is in part due to the fact that is running with
+Figure~\ref{fig:evaluation.bc_coverage} shows higher coverage for
+source-only dropping, this is largely due to the fact that is running with
 higher overheads.
@@
     \end{footnotesize}
-    \caption{Average power overhead for dropping hardware at a 50\% overhead
-    budget. Percentages in parentheses are normalized to the main core
-    power.}
+    \caption{Average power overhead for dropping hardware. Percentages 
+    are normalized to the main core power.}
     \vspace{-0.2in}
@@
 shows the peak and runtime power overheads averaged across all benchmarks with
-a 50\% monitoring overhead target. The peak power is 41-72 mW, which is
+a 1.5x monitoring overhead target. The peak power is 41-72 mW, which is
 less than 8\% of the main core's peak power usage. Similarly, the average runtime power is 20-43

### 2014-09-12 related.tex
Revised related work.

@@
 be generally applicable to any of these instruction-grained monitoring schemes.
-In particular, our proposed architecture may be especially useful for some of
-these more complex schemes which also have higher overheads.  Similarly,
-although many of these schemes showed low overheads in custom hardware, our
-architecture can be used to reduce the overheads when applying these schemes on
-a multi-core platform.
+% In particular, our proposed architecture may be especially useful for some of
+% these more complex schemes which also have higher overheads.  Similarly,
+% although many of these schemes showed low overheads in custom hardware, our
+% architecture can be used to reduce the overheads when applying these schemes on
+% a multi-core platform.
 
@@
 the overheads of monitoring. It filters out redundant updates and null checks
-by reading metadata values. We do not handle filtering redundant updates, but
-we do filter out null checks as well as null (i.e., uninitialized)
-propagations.  We are able to do this because the dataflow engine can also
-propagate the null information.  In addition, our dataflow engine only
-requires reading and writing 2-bit flags rather than the entire metadata.  As
-we have shown, our dataflow engine also enables adjustable overheads without
-false positives which is not a capability described by FADE.
+by reading metadata values. 
+By extending our dataflow engine for partial monitoring, we are able to also
+filter out null checks as well as null metadata propagations, but we do not
+handle filtering redundant updates.
+% We are able to do this because the dataflow engine can also
+% propagate the null information.  
+In contrast to FADE, our dataflow engine only
+requires reading and writing 2-bit flags rather than the entire metadata. 
+% As
+% we have shown, our dataflow engine also enables adjustable overheads without
+% false positives which is not a capability described by FADE.
 

### 2014-09-12 conclusion.tex
Added 10% overhead data.
Updates to abstract, intro, and conclusion.

--- 2014-05-30 conclusion.tex
@@
 the reliability, security, and debugging capabilities of systems. In this
-paper, we have presented an architecture that reduces the overheads of
-monitoring through filtering and adjustable partial monitoring. This is done by
-using a hardware dataflow tracking engine in order to track and filter out
-monitoring for null metadata. In addition, the dataflow engine tracks invalid
-metdata flows in order to enable partial monitoring without false positives.
-Our experimental results show that filtering can greatly reduce the average
-overheads of monitoring from 7x, 18x, and 11x for UMC, BC, and DIFT down to 4x
-for UMC and BC and just 13\% for DIFT. With partial monitoring, we see that
-with a 50\% overhead budget, BC can still achieve 83\% coverage and UMC can
-still achieve 51\% coverage. Finally, we show that although source dropping can
-achieve better coverage than unrestricted dropping when it works, it is not
-able to match overhead targets as well as unrestricted dropping.
+paper, we have presented an architecture that enables adjustable overheads through partial monitoring. 
+The architecture uses a dataflow tracking engine to prevent false positives
+through tracking invalidation information. In addition, we show how the
+architecture can be extended to enable null metadata filtering.
+Given this architecture, we explore design choices related to the dropping policy.
+With partial monitoring, we see that with a 1.5x overhead target, BC can still
+achieve 82\% average coverage and UMC can still achieve 43\% average coverage.
+In addition, with a 1.5x overhead target, IMP shows less than 3\% error on
+average.
+
+% This is done by
+% using a hardware dataflow tracking engine in order to track and filter out
+% monitoring for null metadata. In addition, the dataflow engine tracks invalid
+% metdata flows in order to enable partial monitoring without false positives.
+% Our experimental results show that filtering can greatly reduce the average
+% overheads of monitoring from 7x, 18x, and 11x for UMC, BC, and DIFT down to 4x
+% for UMC and BC and just 13\% for DIFT. With partial monitoring, we see that
+% with a 10\% overhead budget, BC can still achieve 82\% coverage and UMC can
+% still achieve 32\% coverage. Finally, we show that although source-only dropping can
+% achieve better coverage than unrestricted dropping when it works, it is not
+% able to match overhead targets as well as unrestricted dropping.

### 2014-09-12 introduction.tex
Fix up citations.
Updates to abstract, intro, and conclusion.

@@
 % Programmable hardware monitors
-These run-time monitors introduce performance, power, and energy overheads.
-% Implementing run-time monitoring in software using binary instrumentation or
-% other similar methods has been shown to introduce high overheads. For example,
-% dynamic information flow tracking (DIFT) implemented in software suffers a 3.6x
-% slowdown \cite{lift-micro06}. 
-Implementing monitors in dedicated hardware has been shown to have small
-performance impacts. For example, a dedicated hardware implementation of
-dynamic information flow tracking (DIFT) shows performance
-overheads of only 1.1\% \cite{dift-asplos04}. However, fixed hardware loses
-the programmability of software, and cannot be updated or changed. It 
-requires for monitors to be chosen and implemented at hardware design time.
-Recent studies \cite{lba-isca08, ruwase-spaa08, paralog-asplos10, oases-sigops09} have proposed using parallel cores in a
-multi-core system to perform monitoring. This allows monitoring to be done
-using parallel hardware while having the flexibility and programmability of
-software. However, due to the multiple cycles it may take for software to
-monitor an instruction, the overheads can easily increase execution time
-several fold.
-% Two optimizations
-Since these overheads can be prohibitive for run-time use, we propose to enable
-partial monitoring in order to reduce the overheads.
+% These run-time monitors introduce performance, power, and energy overheads.
+% Implementing monitors in dedicated hardware has been shown to have small
+% performance impacts. For example, a dedicated hardware implementation of
+% dynamic information flow tracking (DIFT) shows performance
+% overheads of only 1.1\% \cite{dift-asplos04}. However, fixed hardware loses
+% the programmability of software, and cannot be updated or changed. It 
+% requires for monitors to be chosen and implemented at hardware design time.
+% Recent studies \cite{lba-isca08, ruwase-spaa08, paralog-asplos10, oases-sigops09} have proposed using parallel cores in a
+% multi-core system to perform monitoring. This allows monitoring to be done
+% using parallel hardware while having the flexibility and programmability of
+% software. However, due to the multiple cycles it may take for software to
+% monitor an instruction, the overheads can easily increase execution time
+% several fold.
+% % Two optimizations
+% Since these overheads can be prohibitive for run-time use, we propose to enable
+% partial monitoring in order to reduce the overheads.
+
+Previous studies have explored various design points in terms of efficiency,
+flexibility, and hardware costs for implementing run-time monitors. Custom
+hardware designs that target a single or narrow-range of monitors have been
+shown to have low performance overheads. On the other hand, flexible systems
+that support a range of monitors have shown much higher overheads. In this
+paper, we explore a new trade-off dimension for run-time monitoring designs. We
+propose to enable low overhead monitors without sacrificing flexibility by
+using partial monitoring. This instead allows a trade-off between overheads and
+monitoring coverage or accuracy.
 
@@
 % Evaluation
-In order to evaluate our approach, we applied it to 3 different monitoring
+In order to evaluate our approach, we applied it to 4 different monitoring
 techniques. These monitoring techniques vary in what events they monitor, the
 size and semantics of their metadata, and the operations performed on metadata.
-Our experiments show that filtering reduces the average overhead of these
-monitors from 7-18x down to 13-300\%.
-In addition, with partial monitoring, for the monitoring schemes which still
-see average overheads of 300\%, we show that with just 10\% overheads, we can still
-provide 32-82\% of coverage on average. By increasing the
-overhead budget, this coverage rate can be increased.
+% Our experiments show that filtering reduces the average overhead of these
+% monitors from 7-18x down to 13-300\%.
+% In addition, with partial monitoring, for the monitoring schemes which still
+% see average overheads of 300\%, we show that with just 10\% overheads, we can still
+% provide 32-82\% of coverage on average. By increasing the
+% overhead budget, this coverage rate can be increased. 
+These monitors show baseline overheads of 4-16x without null metadata filtering
+and 1.1-4.3x after null metadata filtering.  Our results show that with 1.5x
+overhead, coverage of 43-82\% on average can still be achieved and that
+overhead and coverage can be varied to create a trade-off.
 
@@
 investigates the design space for dropping decisions for partial monitoring.
-Section~\ref{sec:filtering} shows how the dataflow tracking engine can be extended to filter null metadata.
 Section~\ref{sec:evaluation} presents our evaluation methodology and

### 2014-09-12 title.tex
revised Section 2 to better motivate adjustable monitoring
Updates to abstract, intro, and conclusion.

--- 2014-05-29 title.tex
@@
 \title{
-Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring
+%Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring
+%Efficient and Adjustable Run-Time Monitoring Using Dataflow-Guided Filtering
+Run-Time Monitoring with Adjustable Overheads Using Dataflow-Guided Filtering
 }
@@
 %too large.  
-In this paper, we introduce a hardware dataflow tracking engine that can be
-used to filter out unnecessary monitoring and enable adjustable partial monitoring.
-The dataflow engine can identify events with null metadata so that they can
-be filtered out. To further reduce the overhead of monitoring, we propose
-to enable a trade-off between monitoring coverage and overhead by dropping certain
-monitoring operations. For this partial monitoring, the dataflow engine is used
-to track dropped monitoring flows so that false positives can be avoided.
-Given the architecture, we also investigate how the dropping decisions should be
+In this paper, we introduce a hardware dataflow tracking engine that enables
+adjustable overheads through partial monitoring. This allows a trade-off to be
+made between monitoring coverage and overhead. We show how this dataflow engine
+can be extended to filter out monitoring operations associated with null
+metadata in order to reduce overheads.
+% In this paper, we introduce a hardware dataflow tracking engine that can be
+% used to filter out unnecessary monitoring and enable adjustable partial monitoring.
+% The dataflow engine can identify events with null metadata so that they can
+% be filtered out. To further reduce the overhead of monitoring, we propose
+% to enable a trade-off between monitoring coverage and overhead by dropping certain
+% monitoring operations. For this partial monitoring, the dataflow engine is used
+% to track dropped monitoring flows so that false positives can be avoided.
+Given this architecture, we also investigate how the dropping decisions should be
 made for partial monitoring and show that there exists a trade-off between closely
 matching an overhead target and the coverage that can be achieved.
-% Our experiments show that the system is able to
-% closely meet the target overhead and achieve significant monitoring coverage
-% for reasonable overhead targets.
-The experimental results show that filtering can significantly reduce monitoring
-overheads from XX\%-XX\% on average, depending on the monitoring scheme, to
-XX\%-XX\%. Additionally, the partial monitoring is shown to still achieve XX\% coverage 
-only with a 10\% overhead target.
+Our experimental results show that overheads can be reduced from 4-16x on
+average to 1.5x while still achieving 43-82\% coverage.
 

### 2014-09-12 introduction.tex
a quick pass on the abstract and the intro

@@
 that support a range of monitors have shown much higher overheads. In this
-paper, we explore a new trade-off dimension for run-time monitoring designs. We
+paper, we explore a new trade-off dimension for run-time monitor designs. We
 propose to enable low overhead monitors without sacrificing flexibility by
-using partial monitoring. This instead allows a trade-off between overheads and
+using partial monitoring. In essence, this enables a trade-off between overheads and
 monitoring coverage or accuracy.
@@
 Although partial monitoring comes at the cost of reduced monitoring coverage (i.e.,
-effectiveness), this can still be very useful.
+effectiveness), this can still be useful in many scenarios.
 For example, partial
@@
 % Filtering
-Recent studies have shown that filtering out monitoring checks on clean or null
+Recent studies have shown that filtering out monitoring operations on clean or null
 metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}.
@@
   of run-time monitoring techniques.
-  \item We investigate interesting trade-offs in deciding what to skip for
-  partial monitoring on the proposed architecture.
+  \item We investigate multiple policies on when and which operations to drop 
+  for partial monitoring, and show the trade-offs in the design space.
   \item We extend the dataflow engine to enable filtering null metadata to
@@
 % overhead budget, this coverage rate can be increased. 
-These monitors show baseline overheads of 4-16x without null metadata filtering
-and 1.1-4.3x after null metadata filtering.  Our results show that with 1.5x
-overhead, coverage of 43-82\% on average can still be achieved and that
-overhead and coverage can be varied to create a trade-off.
+These monitors show slowdowns of 1.1-4.3x with the null metadata filtering when
+implemented on a multi-core platform.
+%4-16x without null metadata filtering
+%and 1.1-4.3x after null metadata filtering.  
+Our results show that the partial monitoring can still achieve coverage of 43-82\%
+on average at the reduced slowdown of 1.5x. 
+The results also show that overhead and coverage can be varied to create a trade-off.
 
 % Section summary
-This paper is organized as follows. Section~\ref{sec:monitoring} presents the
-baseline parallel run-time monitoring model that this paper assumes.
-Section~\ref{sec:dropping} discusses enabling partial monitoring using our dataflow tracking engine.
+This paper is organized as follows. Section~\ref{sec:monitoring} introduces the notion
+of adjustable run-time monitoring with partial monitoring.
+Section~\ref{sec:dropping} discusses the hardware architecture that enables 
+partial monitoring using our dataflow tracking engine.
 Section~\ref{sec:policies}
-investigates the design space for dropping decisions for partial monitoring.
+investigates the design space for dropping policies that determine when and which
+monitoring operations to drop.
 Section~\ref{sec:evaluation} presents our evaluation methodology and

### 2014-09-12 title.tex
a quick pass on the abstract and the intro

@@
 adjustable overheads through partial monitoring. This allows a trade-off to be
-made between monitoring coverage and overhead. We show how this dataflow engine
-can be extended to filter out monitoring operations associated with null
+made between monitoring coverage and overhead. This dataflow engine
+can also be extended to filter out monitoring operations associated with null
 metadata in order to reduce overheads.
@@
 % to track dropped monitoring flows so that false positives can be avoided.
-Given this architecture, we also investigate how the dropping decisions should be
+Given this architecture, we investigate how the dropping decisions should be
 made for partial monitoring and show that there exists a trade-off between closely
 matching an overhead target and the coverage that can be achieved.
-Our experimental results show that overheads can be reduced from 4-16x on
-average to 1.5x while still achieving 43-82\% coverage.
+Our experimental results show that overheads can be reduced significantly 
+by trading off coverage. For example, the proposed architecture can still
+achieve 43-82\% coverage when the slowdown is reduced from 1.1-4.3x to 1.5x.
+\Ed{It's a bit strange to say reduced from 1.1x to 1.5x. Revise?}
 

### 2014-09-12 related.tex
comments for the related work.

@@
 
+\Ed{I don't think we need to have a separate discussion on FADE. It's mentioned 
+in Section 2.}
 % Filtering
@@
 other run-time monitoring techniques. The Testudo project
-\cite{testudo-micro08} specifically targets hardware-based run-time monitors.
+%\cite{testudo-micro08} specifically targets hardware-based run-time monitors.
+\cite{testudo-micro08} applies the sampling specifically to dynamic information
+flow tracking. %targets hardware-based run-time monitors.
 These sampling-based techniques all limit the debugging capabilities in order
@@
 However, these approaches are not able to explicitly set an overhead budget, as
-our system is able to do. In addition, with the exception of Testudo, these
+our system is able to do. In addition, with the exception of Testudo that is
+designed for one specific monitor, these
 methods have not considered parallel monitoring. Our techniques for generally
@@
 low-overhead, multi-user/run debugging techniques.
+\Ed{clarify how our approach is different. Is it more general - supporting
+a large class of monitors? Is it different because it is for parallel monitoring?
+If so, what's the significance of the parallel monitors?}
 
@@
 architecture provides a general hardware mechanism to control overheads of
-parallel monitoring architectures. In addition, these works prevented
+parallel monitoring architectures, largely transparent to the original monitor. 
+In addition, these works prevented
 false positives by enabling and disabling monitoring using only a source-dropping
@@
 the monitoring scheme which has not been previously explored.
+\Ed{Is the use of source vs. unrestricted dropping the only major difference for
+this work? That sounds rather weak.}

### 2014-09-12 evaluation.tex
Updated evaluation and figures.
Changed figures to use size 8 Arial.
Added LockSet data.
Added min/max whiskers for IMP accuracy.

@@
 memory check (UMC), array bounds check (BC), dynamic information flow
-tracking (DIFT), and instruction-mix profiling (IMP).  Uninitialized memory
+tracking (DIFT), instruction-mix profiling (IMP), and LockSet-based race detection (LS).  Uninitialized memory
 check seeks to detect loading from
@@
 work has shown that this information can be used to detect malicious software \cite{tang-raid14}.
+LockSet attempts to detect possible race conditions in multi-threaded programs
+by associating information about which locks have access to shared memory
+locations. If a shared memory location is accessed without the use of a lock
+then a race condition may exist.
 Note that, although our dataflow engine performs similar operations to the DIFT
@@
 
-\ED{We need to point out that these monitors cannot be implemented with a simple
-1-bit dataflow engine (DIFT) that we use for filtering. 
-Point: address potential reviewer questions on why don't we just use a 
-programmable dataflow engine for monitoring.}
-
-We tested our system using all C benchmarks from SPECint
+We tested our system using benchmarks from SPECint
 CPU2006 \cite{spec2006}. Since our implementation of BC depends on the
@@
 for these benchmarks are similar to the other results shown. For each
-benchmark, we simulated for 2 billion instructions. An initial slack of 2
-million cycles, which is less than 1\% of
-the total execution time, is given for a 10\% target overhead. This initial
-slack is scaled proportionally for other overhead targets. Unless otherwise
-specified, an unrestricted, slack-based dropping policy is used.
+benchmark, we simulated for 2 billion instructions. 
+% An initial slack of 2
+% million cycles, which is less than 1\% of
+% the total execution time, is given for a 1.1x target overhead. This initial
+% slack is scaled proportionally for other overhead targets. 
+Since LS race detection requires multi-threaded programs, we tested LS using
+the SPLASH-2 benchmark set \cite{splash-isca95}. We ran these benchmarks for
+200 million instructions.
+Unless otherwise specified, an unrestricted, slack-based dropping policy is
+used.
 
@@
 
-% Full monitoring overheads
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\columnwidth]{figs/data_full_mon.pdf}
-    \vspace{-0.2in}
-    \caption{Full monitoring overheads.}
-    \label{fig:evaluation.full_mon}
-    \vspace{-0.1in}
-  \end{center}
-\end{figure}
+% % Full monitoring overheads
+% \begin{figure}
+%   \begin{center}
+%     \includegraphics[width=\columnwidth]{figs/data_full_mon.pdf}
+%     \vspace{-0.2in}
+%     \caption{Full monitoring overheads.}
+%     \label{fig:evaluation.full_mon}
+%     \vspace{-0.1in}
+%   \end{center}
+% \end{figure}
 
-Figure~\ref{fig:evaluation.full_mon} shows the execution times of
-performing full monitoring normalized to the execution times of the benchmarks
-without monitoring. In these results, no filtering or partial monitoring is
-done. UMC shows normalized execution times from 5x to 16x with an average
+% Figure~\ref{fig:evaluation.full_mon} shows the execution times of
+% performing full monitoring normalized to the execution times of the benchmarks
+% without monitoring. In these results, no filtering or partial monitoring is
+% done. 
+Our initial results for performing full core-based monitoring show severalfold slowdowns.
+UMC shows normalized execution times from 5x to 16x with an average
 of 10x. BC shows normalized execution times of 8-24x with an average of
 16x while DIFT shows normalized execution times of 4x-14x with an average of
-9x. IMP shows normalized execution times from 2-6x with an average of 4x. 
-% These monitors show high overheads because they require several
-% instructions to process each monitored instruction from the main core. Even
-% small increases in the number of instructions needed by the monitor lead to
-% high overheads when multiplied across the main program's execution.
-
-\ED{we need to better justify the high overhead. 
-Key points: 1) the overhead is in line with LBA, which represent multi-core
-monitoring w/ low HW costs.
-2) If we are going to mention dynamic memory allocation. We need to say why
-it's really necessary -- it's not our implementation problem. It's something
-that you need to have for large meta-data. Also, I don't think UMC would need
-dynamic allocation.}
+9x. IMP shows normalized execution times from 2-6x with an average of 4x and LS
+shows slowdowns of 1.8-3.3x with an average of 2.4x.
 
@@
   \begin{center}
-    \includegraphics[width=\columnwidth]{figs/data_filtering.pdf}
+    \includegraphics[width=\columnwidth, clip=true, trim=0 0.5in 0 0]{figs/data_filtering.pdf}
     \vspace{-0.2in}
@@
 null metadata and thus is not affected by null metadata filtering.
+LS (not shown) had 1.03-1.72x after null metadata filtering and average
+overhead observed was 1.33x for the SPLASH-2 benchmarks.
 These overheads are similar to previous results for multi-core monitoring with
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{figs/data_bc_sweep.pdf}
+    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0]{figs/data_bc_sweep.pdf}
     \vspace{-0.2in}
-    \caption{Coverage versus varying overhead budget for array bounds check.}
+    \caption{Coverage vs. varying overhead budget for BC.}
     \label{fig:evaluation.bc_sweep}
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{figs/data_umc_sweep.pdf}
+    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0]{figs/data_umc_sweep.pdf}
     \vspace{-0.2in}
-    \caption{Coverage versus varying overhead budget for uninitialized memory check.}
+    \caption{Coverage vs. varying overhead budget for UMC.}
     \label{fig:evaluation.umc_sweep}
+    \vspace{-0.1in}
+  \end{center}
+\end{figure}
+
+% LS sweep
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0]{figs/data_ls_sweep.pdf}
+    \vspace{-0.2in}
+    \caption{Coverage vs. varying overhead budget for LS.}
+    \label{fig:evaluation.ls_sweep}
     \vspace{-0.1in}
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{figs/data_imp_sweep.pdf}
+    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0]{figs/data_imp_sweep.pdf}
     \vspace{-0.2in}
-    \caption{Error versus varying overhead budget for instruction-mix profiling.}
+    \caption{Average error vs. varying overhead budget for instruction-mix profiling. Whisker bars show min and max errors.}
     \label{fig:evaluation.imp_sweep}
@@
 average. Even higher coverage can be achieved by allowing higher overheads.
+Similarly, Figure~\ref{fig:evaluation.ls_sweep} shows the coverage achieved by LS.
+With a 1.1x overhead, LS achieves 53\% coverage on average. This increases to
+71\% coverage with a 1.2x overhead. 
+\DAN{Try to explain water\_nsquared results not reaching 100\%?}
 
@@
 monitored. We compare these percentages against the case when full monitoring
-is performed. Figure~\ref{fig:evaluation.imp_sweep} shows the max error across
+is performed. Figure~\ref{fig:evaluation.imp_sweep} shows the min, max, and average error across
 instruction types for each benchmark and for varying overheads. We see that
-with a 1.1x overhead a max error of 9\% is observed and an average of less than
-4\%. This error decreases as overhead is increased and is 0\% for all
-benchmarks once a 6x overhead target is used.
+with a 1.1x overhead a max error of 9\% is observed across benchmarks and the
+average error across instruction types and benchmarks is less than 2\%.
 
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{figs/data_fpga_umc_sweep.pdf}
+    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0]{figs/data_fpga_umc_sweep.pdf}
     \vspace{-0.2in}
@@
   \subfloat[Overhead budget error.]{
-    \includegraphics[width=\columnwidth]{figs/data_umc_exec_time.pdf}
+    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0]{figs/data_umc_exec_time.pdf}
     \label{fig:evaluation.umc_exec_time}
@@
   \subfloat[Coverage.]{
-    \includegraphics[width=\columnwidth]{figs/data_umc_coverage.pdf}
+    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0]{figs/data_umc_coverage.pdf}
     \label{fig:evaluation.umc_coverage}
   }
+  \vspace{-0.1in}
   \caption{Comparing dropping policies for UMC.}
@@
   \subfloat[Overhead budget error.]{
-    \includegraphics[width=\columnwidth]{figs/data_bc_exec_time.pdf}
+    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0]{figs/data_bc_exec_time.pdf}
     \label{fig:evaluation.bc_exec_time}
@@
   \subfloat[Coverage.]{
-    \includegraphics[width=\columnwidth]{figs/data_bc_coverage.pdf}
+    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0]{figs/data_bc_coverage.pdf}
     \label{fig:evaluation.bc_coverage}
   }
+  \vspace{-0.1in}
   \caption{Comparing dropping policies for BC.}
@@
   \begin{center}
-    \includegraphics[width=\columnwidth]{figs/data_multirun_coverage.pdf}
+    \includegraphics[width=\columnwidth, clip=true, trim=0 0.2in 0 0]{figs/data_multirun_coverage.pdf}
     \vspace{-0.2in}
@@
 coverage achieved over multiple runs for array bounds check running for
-unrestricted and source-only dropping policies with probabilistic dropping. Here, we use a 10\% probability
+unrestricted and source-only dropping policies with probabilistic dropping.
+These numbers are averaged across all benchmarks. Here, we use a 10\%
+probability
 that events are not dropped. Since the goal is code coverage, coverage here is

### 2014-09-12 dropping.tex
Double-check for consistent figure fonts.
Various edits based on Ed's comments.

@@
 %contention for shared resources, as \emph{monitoring overheads}. 
+Monitors typically maintain metadata corresponding to each memory location
+({\tt mem\_metadata[addr]}) and register ({\tt rf\_metadata[reg\_id]}) of the
+main core.
 If the monitor
@@
 for an array bounds check monitor. 
-First, an array {\tt x} is allocated using {\tt malloc}. 
+First, an array {\tt x} is allocated using {\tt malloc} (line 1). 
 As {\tt malloc} returns the array's address in a register, the monitor
 associates base and bounds metadata with the corresponding metadata. 
-Next, pointer {\tt y} is set to point to the middle of array {\tt x}. 
+Next, pointer {\tt y} is set to point to the middle of array {\tt x} (line 2). 
 At the assembly level, a register {\tt r2} is set to array {\tt x}'s address plus an offset.
@@
 metadata of {\tt r2}. This is to ensure that pointer {\tt y} is not used to exceed the array's bounds.
-Finally, both pointers {\tt x} and {\tt y} are used to perform a write. 
-This requires setting register {\tt r3} to the write value. When this happens,
+Line 3a shows setting register {\tt r3} to a constant value of 1.
+When this happens,
 {\tt r3}'s metadata is also reset to NULL in case it previously stored a
-pointer. In both cases, the monitor checks whether the address that is written
-to is within the bounds stored in the metadata. In the first case, {\tt x+12}
+pointer. 
+Finally, the value of {\tt r3} is written using both pointers {\tt x} and {\tt y}.
+In both cases, the monitor checks whether the address that is written
+to is within the bounds stored in the metadata. In the first case (line 3b), {\tt x+12}
 is within the original array's bounds. No error is raised and the metadata of
@@
 it to access the corresponding array or data structure. In the second case,
-{\tt y+12} corresponds to {\tt x+20} which is not within the array's bounds and
+{\tt y+12} (line 4) corresponds to {\tt x+20} which is not within the array's bounds and
 the monitor will raise an error.
@@
 Essentially, when an update operation is skipped, we can no longer trust the
-corresponding metadata. Thus, our general approach for preventing false positives is to mark these metadata as
+corresponding metadata. Thus, our general approach for preventing false
+positives is to add a 1-bit flag to each metadata in order to mark these
+metadata as
 invalid. Furthermore, this invalidation information is propagated in the same
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{figs/dataflow_architecture.pdf}
+    \includegraphics[]{figs/dataflow_architecture.pdf}
     \vspace{-0.2in}
@@
 Thus, the hardware acts effectively as a dataflow tracking engine in order to
-track invalid flags. Figure~\ref{fig:dropping.dataflow} shows a detailed block
+track a 1-bit invalid flag per metadata. Figure~\ref{fig:dropping.dataflow} shows a detailed block
 diagram of this hardware module. The dataflow engine uses two address
@@
 monitor is to filter out monitoring events that correspond to operating on null
-or uninitialized metadata. Null metadata typically corresponds to events that
+metadata. Null metadata corresponds to events that
 are not relevant to the monitor. For example, Hardbound
@@
 operations needed for invalidation shown in
-Figure~\ref{fig:dropping.example_invalid}.
+Figure~\ref{fig:dropping.example_invalid}. Instead of taking a logical OR of the
+source invalid flags to determine whether monitoring can be skipped, null
+metadata filtering takes a logical AND of the source null flags.
 Thus, we can easily enable this null metadata filtering on our architecture by
-simply extending the dataflow flags to be two bits wide. One bit is used to
+extending the dataflow flags to be two bits wide. One bit is used to
 keep track of invalidation information while the second bit is used to keep
@@
 for null metadata filtering.
+The results is a single hardware design that provides enables both partial
+monitoring and null metadata filtering.

### 2014-09-12 evaluation.tex
Various edits based on Ed's comments.

@@
 modifying the ARM version of the gem5 simulator \cite{gem5} to support parallel
-run-time monitoring. We model the main and monitoring cores as running at 2.5
+run-time monitoring. We implement the monitor as a software-based monitor
+running on a parallel processor core. We model the main and monitoring cores as
+running at 2.5
 GHz with 4-way set-associative private L1 I/D caches and a shared 8-way 2 MB L2
@@
 In order to explore the generality of the architecture for
-different monitors, we implemented four different monitors: uninitialized
+different monitors, we implemented five different monitors: uninitialized
 memory check (UMC), array bounds check (BC), dynamic information flow
@@
 locations. If a shared memory location is accessed without the use of a lock
-then a race condition may exist.
-Note that, although our dataflow engine performs similar operations to the DIFT
-monitor, the dataflow engine operates on smaller, 2-bit flags.
+then a race condition may exist. Since we implement the monitor using a
+processor core and our dataflow engine was designed to be generally applicable,
+the same hardware platform supports all of these monitoring techniques.
 
@@
 execution times of 1.7-7x with an average of 4x with null metadata filtering.
-This is an average reduction in overheads of 65\%. For BC, normalized execution
-times drop from 16x down to 3x on average (an 80\% reduction). The range of normalized execution
+For BC, normalized execution
+times drop from 16x down to 3x on average. The range of normalized execution
 times for filtered BC is 1.1-8.3x. 
@@
 Finally, DIFT sees the largest reduction in overheads with only 10\% overheads
-on average after filtering, an 88\% reduction in monitoring overheads, with a
+on average after filtering, with a
 maximum overhead of 31\% after filtering. This is due, in part, to the fact
@@
 null metadata and thus is not affected by null metadata filtering.
-LS (not shown) had 1.03-1.72x after null metadata filtering and average
-overhead observed was 1.33x for the SPLASH-2 benchmarks.
+Overheads for LS (not shown) are 1.03-1.72x after null metadata filtering and the average
+overhead observed is 1.33x for the SPLASH-2 benchmarks.
 These overheads are similar to previous results for multi-core monitoring with
@@
 
-Although the overheads of DIFT are quite low after filtering, BC and UMC still
-show significant overheads. Similarly, the overheads of IMP are not improved by filtering. In this section, we evaluate the effectiveness of
+In this section, we evaluate the effectiveness of
 using partial monitoring (in addition to filtering) to trade-off coverage for further reduced overheads.
@@
 We see that by varying the overhead budget, the coverage achieved also varies.
-With only a 1.1x overhead target, array bounds check still
+With only a 1.1x overhead target (10\% slowdown), array bounds check still
 achieves 84\% monitoring coverage on average. The high coverage
 achieved with such low overheads is due to two main effects.  The first is that
-monitoring can be done in parallel, providing monitoring coverage without
+monitoring can be done in parallel, providing a certain level of monitoring coverage without
 introducing overheads. The second effect is that, although we aggressively
@@
 71\% coverage with a 1.2x overhead. 
+Since the overheads of DIFT are low to begin with, setting an overhead target
+1.1x is enough for all benchmarks tested to reach 100\% coverage.
 \DAN{Try to explain water\_nsquared results not reaching 100\%?}
@@
 fully-pipelined monitor running on an FPGA fabric at half the frequency of the
-main core. We show results only for UMC. In this case, full overheads of
-monitoring range from 10-70\%. Figure~\ref{fig:evaluation.fpga_umc_sweep} shows
-the coverage achieved as we sweep overhead from 1-70\%. We see that partial
+main core. We show results for UMC due to space constraints though other
+monitors show similar trends. In this case, full overheads of
+monitoring range from 1.1-1.7x. Figure~\ref{fig:evaluation.fpga_umc_sweep} shows
+the coverage achieved as we sweep overhead from 1.01-1.7x. We see that partial
 monitoring can allow the overheads of such a system to be pushed below 10\%
-while still providing partial coverage.
+while still providing over 50\% coverage.
 
@@
 dropping and source-only dropping due to the fact that UMC consists of a large
-number of independent monitoring flows.
+number of short, independent monitoring chains.
 Figure~\ref{fig:evaluation.umc_coverage} shows the coverage of UMC for
@@
 of 420\% for {\tt bzip2}, 120\% for {\tt gcc}, and 1500\% for {\tt hmmer}.
-Since only array allocations are considered source events for BC, it is
+Since only array allocations are considered as source events for BC, it is
 difficult for source-only dropping to match overhead targets. Although
@@
 We see that for the unrestricted dropping policy there is almost no increase in
-coverage.  This is due to the fact that since there are multiple dependent
-monitoring operations that must be done in order for a final check to be
-successful as discussed in Section~\ref{sec:policies.when}.
+coverage.  This is due to the fact that it is likely that at least one
+monitoring event in a chain will be dropped with the unrestricted dropping
+policy. 
 Instead, source-only

### 2014-09-12 policies.tex
Various edits based on Ed's comments.

@@
   \item When do we need to drop events in order to enable reduced overheads?
-  \item Which events should be valid for dropping?
+  \item Which events should be dropped?
 \end{enumerate}
@@
 dropped while Section~\ref{sec:policies.which} discusses trade-offs for which
-events are capable of being dropped.
+events can be dropped.
 
@@
 The first possibility is to probabilistically drop events. By setting the probability of
-dropping events appropriately, overheads can be reduced. This is works well for
+dropping events appropriately, overheads can be reduced. This works well for
 enabling partial monitoring for cooperative testing and debugging since the
@@
 accumulated slack. Whenever the main core is stalled due to the monitor, the measured slack
-is decremented. This can be difficult to determine precisely due to the
+is decremented. It is difficult to perfectly determine when the main core is impacted by monitoring due to the
 difficulty in measuring certain overheads such as contention for shared memory.
@@
 In addition to deciding when dropping is required, trade-offs also exist in
-deciding which dropping events can be dropped.
-The simplest policy is to drop all monitoring events as
-long as slack is less than or equal to zero.  However, this can result in
+deciding which events should be dropped.
+The simplest policy is to drop any monitoring events when slack is less than or equal to zero.  However, this can result in
 wasted work. For example, consider the metadata dependence graph shown in
@@
 {\tt B} represents that if event {\tt A} is dropped, then due to its
-invalidated metadata, it will cause event {\tt B} to be filtered. Square
+invalidated metadata, it will cause event {\tt B} to be dropped. Square
 nodes indicate events where monitoring checks are performed. In the
@@
 since their results were not used for any monitoring checks.
+That is, by the time we decide to drop event {\tt E}, we have already 
+updated metadata due to events {\tt C} and {\tt D} which is no longer needed.
 
@@
 dropping any event as \emph{unrestricted dropping}.
+
 Figures~\ref{fig:policies.all_drop} and \ref{fig:policies.source_drop} show a
@@
 result in no wasted work and thus better coverage. However, because of the coarser-grained decision, it
-may be more difficult to closely match overheads. We expect source dropping 
+may be more difficult to closely match overheads. We expect source-only dropping 
 to work well when there are a large number of independent metadata flows.
 In addition, using a probabilistic dropping policy works poorly with
-unrestricted dropping since each event in a chain needs to be randomly selected
-for monitoring in order for the check to be performed. Instead, probabilistic
-dropping works better when coupled with source-only dropping.
+unrestricted dropping. Since each event in a dependent chain (e.g., events {\tt
+A} through {\tt E}) needs to be monitored in order for the monitoring check to
+be useful, randomly dropping any single event will cause the final check to be
+invalid.
+
+For applications where closely matching an overhead target is important, a
+slack-based, unrestricted dropping policy is appropriate. However, if matching
+the overhead target is not as important, then a slack-based, source-only dropping
+policy could provide better coverage. 
+Finally, if the goal is to use partial monitoring to enable cooperative
+debugging and testing with very low overheads, then a probabilistic,
+source-only dropping policy can be used to provide good total coverage over
+multiple runs.
 

### 2014-09-12 dropping.tex
Minor edits from Mohamed's comments.

@@
 The key cause of false positives is dropping monitoring operations that update
-metadata. Dropping monitoring operations that check for an error (such as check
-found for line 4 in Figure~\ref{fig:dropping.example_full} can only cause
+metadata. Dropping monitoring operations that check for an error (such as the check
+for line 4 in Figure~\ref{fig:dropping.example_full}) can only cause
 false negatives but will never cause false positives. On the other hand,
@@
 In order to efficiently support dropping monitoring events and to prevent false
-positives, we propose to insert a hardware module between the main and
+positives, we propose to insert a hardware module between the main core and
 monitor (see Figure~\ref{fig:dropping.dataflow_overview}). 
@@
 and bounds metadata) instructions since it is not relevant to array bounds
-checking. More recently, FADE \cite{fade-hpca14} has been proposed a general
+checking. More recently, FADE \cite{fade-hpca14} has been proposed as a general
 hardware module to perform this null metadata filtering for a variety of

### 2014-09-12 evaluation.tex
Minor edits from Mohamed's comments.

@@
 main core. We show results for UMC due to space constraints though other
-monitors show similar trends. In this case, full overheads of
+monitors show similar trends. In this case, the full overheads of
 monitoring range from 1.1-1.7x. Figure~\ref{fig:evaluation.fpga_umc_sweep} shows
-the coverage achieved as we sweep overhead from 1.01-1.7x. We see that partial
+the coverage achieved as we sweep the overhead target from 1.01-1.7x. We see that partial
 monitoring can allow the overheads of such a system to be pushed below 10\%

### 2014-09-12 monitoring.tex
Minor edits from Mohamed's comments.

@@
 design points. % in the trade-off space between efficiency and flexibility.
-Table~\ref{tab:monitoring.previous_overheads} summarizes some of representative designs
+Table~\ref{tab:monitoring.previous_overheads} summarizes some the of representative designs
 and their reported performance overhead. The previous studies clearly show that there
-exist trade-offs among efficiency, flexibility, and hardware costs. 
+exist trade-offs between efficiency, flexibility, and hardware costs. 
 For example, a run-time monitoring scheme can often be realized with fairly low
@@
 In such profiling tools, the partial monitoring framework can be used to obtain
-the statics on samples rather than all program events, essentially trading off
+statistical samples rather than complete counts of all program events, essentially trading off
 accuracy for lower overhead.
@@
   which often have high overhead, the framework also needs to be general enough to be
-  applicable to a wide range of monitoring scheme.
+  applicable to a wide range of monitoring schemes.
 
   \item \textbf{No false positive:} The framework needs to ensure that dropping a portion
-  of monitoring operations does not lead to a false positive. We found that a data-flow
-  that tracks invalid meta-data can serve as a general solution to this problem 
+  of monitoring operations does not lead to a false positive. We found that a dataflow engine
+  that tracks invalid metadata can serve as a general solution to this problem 
   (Section~\ref{sec:dropping}).

### 2014-09-12 dropping.tex
Edits based on Ji's comments.

@@
 
-In order to efficiently support dropping monitoring events and to prevent false
+Although the functionality of dropping and invalidation could be implemented on
+the monitor, this is unlikely to be much faster than performing the full monitoring
+operations.
+Instead, in order to efficiently support dropping monitoring events and to prevent false
 positives, we propose to insert a hardware module between the main core and

### 2014-09-12 title.tex
Various edits based on Ed's comments.
Edits based on Ji's comments.

@@
 Given this architecture, we investigate how the dropping decisions should be
-made for partial monitoring and show that there exists a trade-off between closely
+made for partial monitoring and show that there exist interesting policy
+decisions which create a trade-off between closely
 matching an overhead target and the coverage that can be achieved.
@@
 achieve 43-82\% coverage when the slowdown is reduced from 1.1-4.3x to 1.5x.
-\Ed{It's a bit strange to say reduced from 1.1x to 1.5x. Revise?}
+\ED{It's a bit strange to say reduced from 1.1x to 1.5x. Revise?}
 

### 2014-09-12 related.tex
Various edits based on Ed's comments.
Updates to abstract and related work.

--- 2014-09-12 related.tex
@@
 
-% Filtering
-FADE \cite{fade-hpca14} presents a general filtering engine in order to reduce
-the overheads of monitoring. It filters out redundant updates and null checks
-by reading metadata values. 
-By extending our dataflow engine for partial monitoring, we are able to also
-filter out null checks as well as null metadata propagations, but we do not
-handle filtering redundant updates.
-% We are able to do this because the dataflow engine can also
-% propagate the null information.  
-In contrast to FADE, our dataflow engine only
-requires reading and writing 2-bit flags rather than the entire metadata. 
-% As
-% we have shown, our dataflow engine also enables adjustable overheads without
-% false positives which is not a capability described by FADE.
+% % Filtering
+% FADE \cite{fade-hpca14} presents a general filtering engine in order to reduce
+% the overheads of monitoring. It filters out redundant updates and null checks
+% by reading metadata values. 
+% By extending our dataflow engine for partial monitoring, we are able to also
+% filter out null checks as well as null metadata propagations, but we do not
+% handle filtering redundant updates.
+% In contrast to FADE, our dataflow engine only
+% requires reading and writing 2-bit flags rather than the entire metadata.
 
@@
 \cite{liblit-pldi05}. This idea has also been applied to detecting memory
-leaks~\cite{chilimbi-asplos04}, dataflow analysis~\cite{greathouse-cgo11}, and
-other run-time monitoring techniques. The Testudo project
-\cite{testudo-micro08} specifically targets hardware-based run-time monitors.
+leaks~\cite{chilimbi-asplos04}, dataflow analysis~\cite{greathouse-cgo11}, and dynamic information flow tracking\cite{testudo-micro08} and
+other run-time monitoring techniques. 
 These sampling-based techniques all limit the debugging capabilities in order
 to provide lower performance overheads, similar to our architecture's goals.
-However, these approaches are not able to explicitly set an overhead budget, as
-our system is able to do. In addition, with the exception of Testudo, these
-methods have not considered parallel monitoring. Our techniques for generally
-reducing the overheads of monitoring can also be used to enable these
-low-overhead, multi-user/run debugging techniques.
-
 % Limited Monitoring
-In terms of partial monitoring, the Quality Virtual Machine (QVM) is a
+In terms of partial monitoring with adjustable overheads, the Quality Virtual Machine (QVM) is a
 modification of the Java Virtual Machine (JVM) that supports run-time
@@
 al. created a framework for controlling the overheads of software-based
-monitoring \cite{huang-sttt12}. These projects are targeting a similar problem
-to the one we address in this paper. However, both projects are focused on
-software-based monitoring and thus their mechanisms modify the software
-monitoring framework in order to enable or disable monitoring. Instead, our
-architecture provides a general hardware mechanism to control overheads of
-parallel monitoring architectures. In addition, these works prevented
-false positives by enabling and disabling monitoring using only a source-dropping
-policy. Lo et al. \cite{lo-rtas14} developed a hardware architecture to limit
-monitoring in hard real-time systems. Their architecture performs unrestricted 
-dropping in order to guarantee strict deadlines. 
-Our system allows either source-only dropping or unrestricted dropping to be performed.
-This allows the designer or user to choose the appropriate policy depending on
-the monitoring scheme which has not been previously explored.
+monitoring \cite{huang-sttt12}. 
+Lo et al. \cite{lo-rtas14} developed a hardware architecture to limit
+monitoring specifically for hard real-time systems.
+
+Our work presents a general and efficient hardware platform to enable
+partial monitoring. The architecture is designed to be applicable to a wide range of monitoring
+techniques as well as monitor implementations. By studying various potential
+uses of partial monitoring, including sampling for cooperative debugging and
+partial protection for low overheads, we have determined important dropping
+policy decisions to consider in order to best meet the requirements of the
+application.
+
+% However, these approaches are not able to explicitly set an overhead budget, as
+% our system is able to do. In addition, with the exception of Testudo that is
+% designed for one specific monitor, these
+% methods have not considered parallel monitoring. Our techniques for generally
+% reducing the overheads of monitoring can also be used to enable these
+% low-overhead, multi-user/run debugging techniques.
+% \ED{clarify how our approach is different. Is it more general - supporting
+% a large class of monitors? Is it different because it is for parallel monitoring?
+% If so, what's the significance of the parallel monitors?}
+
+% These projects are targeting a similar problem
+% to the one we address in this paper. However, both projects are focused on
+% software-based monitoring and thus their mechanisms modify the software
+% monitoring framework in order to enable or disable monitoring. 
+% Instead, our
+% architecture provides a general hardware mechanism to control overheads of
+% parallel monitoring architectures, largely transparent to the original monitor. 
+% In addition, these works prevented
+% false positives by enabling and disabling monitoring using only a source-dropping
+% policy. 
+%  Their architecture performs unrestricted 
+% dropping in order to guarantee strict deadlines. 
+% Our system allows either source-only dropping or unrestricted dropping to be performed.
+% This allows the designer or user to choose the appropriate policy depending on
+% the monitoring scheme which has not been previously explored.
+% \ED{Is the use of source vs. unrestricted dropping the only major difference for
+% this work? That sounds rather weak.}

### 2014-09-12 title.tex
Updates to abstract and related work.

@@
 %Efficient and Adjustable Run-Time Monitoring Using Dataflow-Guided Filtering
-Run-Time Monitoring with Adjustable Overheads Using Dataflow-Guided Filtering
+Run-Time Monitoring with Adjustable Overheads \\Using Dataflow-Guided Filtering
 }
@@
 Our experimental results show that overheads can be reduced significantly 
-by trading off coverage. For example, the proposed architecture can still
-achieve 43-82\% coverage when the slowdown is reduced from 1.1-4.3x to 1.5x.
-\ED{It's a bit strange to say reduced from 1.1x to 1.5x. Revise?}
+by trading off coverage. For example, although average overheads for full
+monitoring range between 3-4x, the proposed architecture can still
+achieve 43-82\% coverage when the slowdown is reduced to 1.5x.
 

### 2014-09-12 dropping.tex
Various edits.

@@
 on the type of monitoring event, different monitoring operations are
-executed. Information about monitoring events is sent to the monitor and buffered in a FIFO structure to decouple the
+executed. Information about monitoring events is sent to the monitor and
+buffered in a FIFO structure to decouple the
 running of the main core and the monitor. If the FIFO is full, then the main
@@
 
-Here, we have shown the dynamic sequence of operations of the monitor.
-The actual static code consists of a set of instructions to be run for each
-monitored instruction type. Register and memory addresses are read from
-information sent by the main core.
+Note that this monitoring is performed automatically and transparently with
+almost no modification needed to the main program. The only modifications to
+the main program that are needed are to initially set metadata, such as setting
+the base and bounds addresses on a {\tt malloc} call for array bounds check.
+Otherwise, on the commit
+of instruction types relevant to the specific monitor, the needed address and
+register information is forwarded to the monitor automatically.
+Here, we have shown the dynamic sequence of operations executed by the monitor but
+the actual static code consists of a set of instructions to be run for each
+possible instruction type (load, store, etc.).
 

### 2014-09-12 evaluation.tex
Edits based on Ji's comments.
Various edits.

@@
 200 million instructions.
-Unless otherwise specified, an unrestricted, slack-based dropping policy is
-used.
 
@@
 % done. 
-Our initial results for performing full core-based monitoring show severalfold slowdowns.
-UMC shows normalized execution times from 5x to 16x with an average
-of 10x. BC shows normalized execution times of 8-24x with an average of
-16x while DIFT shows normalized execution times of 4x-14x with an average of
-9x. IMP shows normalized execution times from 2-6x with an average of 4x and LS
-shows slowdowns of 1.8-3.3x with an average of 2.4x.
+% Our initial results for performing full core-based monitoring show severalfold slowdowns.
+% UMC shows normalized execution times from 5x to 16x with an average
+% of 10x. BC shows normalized execution times of 8-24x with an average of
+% 16x while DIFT shows normalized execution times of 4x-14x with an average of
+% 9x. IMP shows normalized execution times from 2-6x with an average of 4x and LS
+% shows slowdowns of 1.8-3.3x with an average of 2.4x.
 
@@
     \vspace{-0.2in}
-    \caption{Monitoring overheads with filtering null metadata.}
+    \caption{Monitoring overheads with null metadata filtering.}
     \label{fig:evaluation.filtering}
@@
 Figure~\ref{fig:evaluation.filtering} shows the
-normalized execution times with filtering enabled. We see significant
-reductions in overhead for all three monitoring schemes. UMC sees normalized
-execution times of 1.7-7x with an average of 4x with null metadata filtering.
+execution times of performing monitoring with null filtering enabled normalized
+to the execution time of the benchmarks without monitoring. UMC sees normalized
+execution times of 1.7-7x with an average of 4x.
 For BC, normalized execution
-times drop from 16x down to 3x on average. The range of normalized execution
-times for filtered BC is 1.1-8.3x. 
+are 3x on average and range from 1.1x to 8.3x.
 % This is not surprising since in the baseline
@@
 % filtering, only loads and stores corresponding to arrays need to be forwarded.
-Finally, DIFT sees the largest reduction in overheads with only 10\% overheads
-on average after filtering, with a
-maximum overhead of 31\% after filtering. This is due, in part, to the fact
+DIFT sees an average of 10\% overhead with a maximum overhead of 31\% with null
+metadata filtering. This is due, in part, to the fact
 that for our implementation of DIFT on SPEC
@@
 targeted network or streaming applications, which have larger amounts of
-untrusted input data, we would expect to see less filtering. IMP does not have
-null metadata and thus is not affected by null metadata filtering.
-Overheads for LS (not shown) are 1.03-1.72x after null metadata filtering and the average
+untrusted input data, we would higher overheads. IMP shows normalized
+executions from 2-6x with an average of 4x. Overheads for LS (not pictured) are
+1.03-1.72x after null metadata filtering and the average
 overhead observed is 1.33x for the SPLASH-2 benchmarks.
-These overheads are similar to previous results for multi-core monitoring with
-minimal hardware costs such as LBA \cite{lba-asid06} (see
+These overheads are similar to the overheads seen by previous multi-core
+monitoring systems such as LBA-accelerated \cite{lba-isca08} and FADE
+\cite{fade-hpca14} (see
 Table~\ref{tab:monitoring.previous_overheads}). In
@@
 In this section, we evaluate the effectiveness of
-using partial monitoring (in addition to filtering) to trade-off coverage for further reduced overheads.
+using partial monitoring to trade-off coverage for reduced overheads.
+For these results, we use a slack-based, unrestricted dropping policy.
 Figure~\ref{fig:evaluation.bc_sweep} shows the monitoring coverage achieved by
@@
 
+From these results, we see that for monitors which many short, independent
+chains, such as UMC, source-only dropping can provide better coverage than
+unrestricted dropping with the same overhead. For other monitors, such as BC,
+source-only dropping is prone to overshoot the overhead target.
+Thus, for applications where staying with an overhead target is especially
+important, such as soft real-time systems, a slack-based, unrestricted dropping
+policy is more appropriate. On the other hand, if occasional slowdowns are
+acceptable, then source-only dropping can provide better coverage on average.
+
 \subsection{Multiple-Run Coverage}
@@
 that events are not dropped. Since the goal is code coverage, coverage here is
-measured as the percentage of static instructions which are monitored. Each run
+measured as the percentage of static instructions which are monitored in at
+least one of the runs. Each run
 was simulated for 200 million instructions. 

### 2014-09-12 introduction.tex
Edits based on Ji's comments.
Various edits.

@@
 dropping monitoring operations when overheads exceed a specified overhead
-budget.  In order to prevent dropped monitoring operations from leading to a
-false positive, we need to track these dropped flows.  We present a
+budget. Although this results in reduced monitoring coverage (i.e., false
+negatives), false positives can also occur due to out-of-date metadata information.
+In order to prevent these
+false positive, we need to track the dropped metadata flows.  We present a
 hardware-based dataflow tracking engine that keeps track of these dropped

### 2014-09-13 dropping.tex
Various edits based on a full pass.

@@
 from the main core. 
-Monitors typically maintain a set of metadata corresponding to the main
-program's memory. Similarly, monitors also typically maintain a set of metadata
-corresponding to the main core's register file.
-%We refer to these stalls and other overheads, such as
-%contention for shared resources, as \emph{monitoring overheads}. 
 Monitors typically maintain metadata corresponding to each memory location
@@
 As {\tt malloc} returns the array's address in a register, the monitor
-associates base and bounds metadata with the corresponding metadata. 
+associates base and bounds metadata with the corresponding register. 
 Next, pointer {\tt y} is set to point to the middle of array {\tt x} (line 2). 
@@
 pointer. 
-Finally, the value of {\tt r3} is written using both pointers {\tt x} and {\tt y}.
-In both cases, the monitor checks whether the address that is written
-to is within the bounds stored in the metadata. In the first case (line 3b), {\tt x+12}
+Finally, the value of {\tt r3} is written to memory using both pointers {\tt x} and {\tt y}.
+In both cases, the monitor checks whether the store address is within the bounds of the register metadata. In the first case (line 3b), {\tt x+12}
 is within the original array's bounds. No error is raised and the metadata of
@@
 the base and bounds addresses on a {\tt malloc} call for array bounds check.
-Otherwise, on the commit
-of instruction types relevant to the specific monitor, the needed address and
-register information is forwarded to the monitor automatically.
+Later propagation and checks of the metadata occur automatically as
+instructions are forwarded.
 Here, we have shown the dynamic sequence of operations executed by the monitor but
 the actual static code consists of a set of instructions to be run for each
-possible instruction type (load, store, etc.).
+possible instruction type (load, store, etc.). Only instruction types relevant
+for the particular monitoring scheme are forwarded.
 
@@
 corresponding metadata. Thus, our general approach for preventing false
-positives is to add a 1-bit flag to each metadata in order to mark these
-metadata as
+positives is to associate a 1-bit flag with each metadata in order to mark these
+metadata as valid or
 invalid. Furthermore, this invalidation information is propagated in the same
@@
 monitoring operations are dropped. However, suppose that the monitoring for
-line 1 is dropped. In this case, {\tt rf\_invalid[1]} will be marked as {\tt
+line 1 is dropped. When a monitoring event is dropped, metadata that would have
+been updated is marked as invalid. In this case, {\tt rf\_invalid[1]} will be
+marked as {\tt
 true}. Thus, when line 3 is reached, the monitoring event is dropped
@@
 forwarded to the monitor since the system does not know whether {\tt r2}
-contains a pointer address or not. However, if we use a flag to mark {\tt r2}
+contains a pointer address or not. However, if we use a 1-bit flag to mark {\tt r2}
 as null when it is loaded with a constant, then we can propagate this null
@@
 for null metadata filtering.
-The results is a single hardware design that provides enables both partial
+The result is a single hardware design that enables both partial
 monitoring and null metadata filtering.

### 2014-09-13 evaluation.tex
Various edits based on a full pass.

@@
 cache. This setup is similar to the Snapdragon 801 processor commonly found in
-mobile systems. The dataflow engine uses a 1 kB cache for flags.
+mobile systems. The dataflow engine uses a 1 kB cache for invalid and null flags.
 
@@
 memory locations that are not initialized first. Array bounds check, as
-mentioned in Section~\ref{sec:monitoring}, is a monitoring scheme that aims to
+described in Section~\ref{sec:dropping}, is a monitoring scheme that aims to
 detect buffer overflows where memory accesses go beyond the boundaries of an
@@
 load, store, and control instructions that are executed. This profiling
-information can be useful for performing optimizations. Additionally, recent
-work has shown that this information can be used to detect malicious software \cite{tang-raid14}.
+information can be useful for performing optimizations or to detect malicious
+software \cite{tang-raid14}.
 LockSet attempts to detect possible race conditions in multi-threaded programs
-by associating information about which locks have access to shared memory
-locations. If a shared memory location is accessed without the use of a lock
+by tracking metadata about which locks are protecting shared memory
+locations. If a shared memory location is accessed while unprotected
 then a race condition may exist. Since we implement the monitor using a
@@
 show results for the C++ benchmarks, we note that the results for UMC, DIFT, and IMP
-for these benchmarks are similar to the other results shown. For each
-benchmark, we simulated for 2 billion instructions. 
+for these benchmarks are similar to the other results shown.
+We simulated each benchmark for 2 billion instructions. 
 % An initial slack of 2
@@
 % slack is scaled proportionally for other overhead targets.
-Since LS race detection requires multi-threaded programs, we tested LS using
-the SPLASH-2 benchmark set \cite{splash-isca95}. We ran these benchmarks for
-200 million instructions.
+
+Since LockSet race detection requires multi-threaded programs, we tested LS using
+the applications from the SPLASH-2 benchmark set \cite{splash-isca95}. We ran these benchmarks for
+200 million instructions with 2 main cores running the benchmark application.
+Each main core was connected to a dataflow engine and a monitoring core.
 
@@
 % filtering, only loads and stores corresponding to arrays need to be forwarded.
-DIFT sees an average of 10\% overhead with a maximum overhead of 31\% with null
-metadata filtering. This is due, in part, to the fact
+DIFT sees an average of 1.1x slowdown with a maximum slowdown of 1.3x with null
+metadata filtering. This low overhead is due, in part, to the fact
 that for our implementation of DIFT on SPEC
-benchmarks, we only mark data read from files as tainted. For most of these
-benchmarks, this propagates to relatively few instructions. Instead, if we
+benchmarks, we only mark data read from files as tainted. Instead, if we
 targeted network or streaming applications, which have larger amounts of
 untrusted input data, we would higher overheads. IMP shows normalized
-executions from 2-6x with an average of 4x. Overheads for LS (not pictured) are
+execution times of 2-6x with an average of 4x. Overheads for LS (not pictured) are
 1.03-1.72x after null metadata filtering and the average
 overhead observed is 1.33x for the SPLASH-2 benchmarks.
-These overheads are similar to the overheads seen by previous multi-core
+Our baseline system and overheads are similar previous multi-core
 monitoring systems such as LBA-accelerated \cite{lba-isca08} and FADE
@@
 Table~\ref{tab:monitoring.previous_overheads}). In
-Section~\ref{sec:evaluation.fpga}, we evaluate a higher performance, FPGA-based
+Section~\ref{sec:evaluation.fpga}, we also evaluate a higher performance, FPGA-based
 monitor that shows low tens of percent of overhead.
@@
 percentage of dynamic checks that are performed 
-(indirect jumps in DIFT, loads in UMC, and memory accesses in BC). 
+(indirect jumps in DIFT, loads in UMC, and memory accesses in BC and LS). 
 The metric is chosen to understand
@@
 monitoring can be done in parallel, providing a certain level of monitoring coverage without
-introducing overheads. The second effect is that, although we aggressively
-filter out monitoring events, there may still exist a large number of
+introducing overheads. The second effect is that there may still exist a large number of
 monitoring events that do not lead to security or reliability checks. As a
@@
 1.1x is enough for all benchmarks tested to reach 100\% coverage.
-\DAN{Try to explain water\_nsquared results not reaching 100\%?}
 
@@
 type profiled, we calculate its count as a percentage of the total instructions
-monitored. We compare these percentages against the case when full monitoring
-is performed. Figure~\ref{fig:evaluation.imp_sweep} shows the min, max, and average error across
+monitored. We take the difference of these percentages compared to the case
+when full monitoring
+is performed to calculate the error. Figure~\ref{fig:evaluation.imp_sweep}
+shows the min, max, and average error across
 instruction types for each benchmark and for varying overheads. We see that
@@
 fully-pipelined monitor running on an FPGA fabric at half the frequency of the
-main core. We show results for UMC due to space constraints though other
+main core. We only show results for UMC due to space constraints though other
 monitors show similar trends. In this case, the full overheads of
@@
 important, such as soft real-time systems, a slack-based, unrestricted dropping
-policy is more appropriate. On the other hand, if occasional slowdowns are
-acceptable, then source-only dropping can provide better coverage on average.
+policy is more appropriate. On the other hand, if maximum coverage is desired
+and occasional slowdowns are acceptable, then source-only dropping can provide
+better coverage on average.
 
@@
 
-One possible application of partial monitoring with low overheads is to enable
+One application of partial monitoring with low overheads is to enable
 cooperative debugging. 
@@
 increase of 7\% of the main core area. Table~\ref{tab:evaluation.area_power}
-shows the peak and runtime power overheads averaged across all benchmarks with
-a 1.5x monitoring overhead target. The peak power is 41-72 mW, which is
-less than 8\% of the main core's peak power usage. Similarly, the average runtime power is 20-43
-mW, corresponding to 5-11\% of the main core's runtime power.
+shows the peak and runtime power overheads averaged across all benchmarks running with
+a 1.5x monitoring overhead target. The peak power is 31-72 mW, which is
+less than 8\% of the main core's peak power usage. Similarly, the average runtime power is 13-43
+mW, corresponding to 4-11\% of the main core's runtime power.

### 2014-09-13 introduction.tex
Various edits based on a full pass.

@@
 coverage achieved.
-
-% Filtering
-Recent studies have shown that filtering out monitoring operations on clean or null
-metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}.
+In addition to enabling partial monitoring, we show how a simple
+extension to our dataflow engine enables this null metadata filtering.
 For example, for an array bounds check, checking non-pointer accesses is
-unnecessary. In addition to enabling partial monitoring, we show how a simple
-extension to our dataflow engine enables this null metadata filtering.
+unnecessary and can be filtered out using the dataflow engine.
 
@@
   hardware architecture is designed to be generally applicable to a wide-range
-  of run-time monitoring techniques.
+  of run-time monitoring techniques and monitor implementations.
   \item We investigate multiple policies on when and which operations to drop 
@@
 % Evaluation
-In order to evaluate our approach, we applied it to 4 different monitoring
+In order to evaluate our approach, we applied it to five different monitoring
 techniques. These monitoring techniques vary in what events they monitor, the
@@
 implemented on a multi-core platform.
-%4-16x without null metadata filtering
-%and 1.1-4.3x after null metadata filtering.  
-Our results show that the partial monitoring can still achieve coverage of 43-82\%
-on average at the reduced slowdown of 1.5x. 
-The results also show that overhead and coverage can be varied to create a trade-off.
+Our results show that the partial monitoring can still achieve significant
+coverage with reduced overheads. For example, for monitors which showed 3-4x
+average overheads, slowdown could be reduced to 1.5x while still achieving a
+coverage of 43-82\% on average.
 

### 2014-09-13 monitoring.tex
Various edits based on a full pass.

@@
 
-In this framework, a user specifies how much monitoring should be done in a form of
+In this framework, a user specifies how much monitoring should be done in the form of
 a target 
@@
 {\bf Cooperative testing, debugging, and protection:}
-Recently studies \cite{liblit-pldi05, chilimbi-asplos04, greathouse-cgo11, testudo-micro08} have shown that software testing, debugging,
+Recently, studies \cite{liblit-pldi05, chilimbi-asplos04, greathouse-cgo11, testudo-micro08} have shown that software testing, debugging,
 or attack detection may be done in a cooperative fashion across
 a large number of systems. In this case, each system is only willing to pay very low
-overhead, say a few percents, and only performs a small subset of checks. 
-A high coverage is achieved by having different systems check different parts of a program.
+overhead (e.g., a few percent) and only performs a small subset of checks. 
+High coverage is achieved by having different systems check different parts of a program.
 The partial monitoring framework allows high-overhead monitoring to be used in a
@@
 {\bf Partial protection for low overhead:}
-Even without real-time constraints, systems may have tight budgets for performance,
+Even without real-time constraints, systems may have tight budgets for the performance,
 energy, or hardware overhead that they can tolerate for run-time monitoring. In such cases,
-full monitoring cannot be enabled unless its overhead is low enough. The adjustable
+full monitoring cannot be enabled unless its overhead is low enough. Adjustable
 monitoring allows partial protection on such systems. For example, array bounds
@@
 partial monitoring can be measured as the percentage of run-time checks that are
-performed on each program run, which reflects how likely for a bug or an attack 
+performed on each program run, which reflects how likely it is for a bug or an attack 
 to be detected for a system.

### 2014-09-13 policies.tex
Various edits based on a full pass.

@@
 \begin{enumerate}
-  \item When do we need to drop events in order to enable reduced overheads?
-  \item Which events should be dropped?
+  \item When do we need to drop events in order to enable reduced overheads? (Section~\ref{sec:policies.when})
+  \item Which events should be dropped? (Section~\ref{sec:policies.which})
 \end{enumerate}
-
-Section~\ref{sec:policies.when} discusses options for deciding when events are
-dropped while Section~\ref{sec:policies.which} discusses trade-offs for which
-events can be dropped.
 
@@
 overheads of monitoring in order to decide whether dropping is needed.
-% In order to decide whether a monitoring event should be dropped or not, we need
-% a way to keep track of overheads. 
-We present a specific method
-for measuring execution time overheads. However, we note that other methods
-could also be used in order to target budgets on IPC, power, energy, etc.  
-% For example, one way to measure execution time overheads would be to insert
-% checkpoints in the program. By recording the execution time of these
-% checkpoints without monitoring, the overheads incurred while running with
-% monitoring could be determined. Although this would give an accurate
-% measurement of overheads, it would require modifying the program binary. In
-% addition, the overhead budget for monitoring would only be updated at
-% checkpoints.
-
-% Instead, we develop a model to estimate the execution time overhead budget in a
-% gradual manner and without the need to modify the main program. Specifically,
-We specify the overhead budget as a percentage of the main program's execution
+The overhead budget is specified as a percentage of the main program's execution
 cycles without monitoring. We define \emph{slack} as the number of cycles of
@@
 Slack can be easily measured in hardware by using a counter that increments on every $k$-th
-cycle of the main core. This $k$ can be calculated by taking the reciprocal of
-the target budget. For example, if the target budget is 20\%, then the counter
-increments on every 5th clock cycle. The value of this counter is the
-accumulated slack. Whenever the main core is stalled due to the monitor, the measured slack
-is decremented. It is difficult to perfectly determine when the main core is impacted by monitoring due to the
+cycle of the main core (e.g., every 5th cycle for a 20\% target budget). 
+% This $k$ can be calculated by taking the reciprocal of
+% the target budget. For example, if the target budget is 20\%, then the counter
+% increments on every 5th clock cycle. 
+The value of this counter is the
+accumulated slack. Whenever the main core is stalled due to the monitor, the
+measured slack is decremented. It is difficult to precisely determine the
+entire impact of monitoring on the main core due to the
 difficulty in measuring certain overheads such as contention for shared memory.
@@
 
-% Varying slack's impact on coverage for UMC
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\columnwidth]{figs/dataflow_graph.pdf}
-    \vspace{-0.3in}
-    \caption{Example dependence graph for metadata. Square nodes represent
-    events where checks are performed.} 
-    \label{fig:policies.dataflow_graph}
-    \vspace{-0.1in}
-  \end{center}
-\end{figure}
-
-In addition to deciding when dropping is required, trade-offs also exist in
-deciding which events should be dropped.
-The simplest policy is to drop any monitoring events when slack is less than or equal to zero.  However, this can result in
-wasted work. For example, consider the metadata dependence graph shown in
-Figure~\ref{fig:policies.dataflow_graph}. Here, an edge from node {\tt A} to node
-{\tt B} represents that if event {\tt A} is dropped, then due to its
-invalidated metadata, it will cause event {\tt B} to be dropped. Square
-nodes indicate events where monitoring checks are performed. In the
-example, suppose that event {\tt E} is meant to perform a check operation but is dropped.
-In this situation, the
-monitoring operations that were done for events {\tt C} and {\tt D} were wasted
-since their results were not used for any monitoring checks.
-That is, by the time we decide to drop event {\tt E}, we have already 
-updated metadata due to events {\tt C} and {\tt D} which is no longer needed.
+% % Varying slack's impact on coverage for UMC
+% \begin{figure}
+%   \begin{center}
+%     \includegraphics[width=\columnwidth]{figs/dataflow_graph.pdf}
+%     \vspace{-0.3in}
+%     \caption{Example dependence graph for metadata. Square nodes represent
+%     events where checks are performed.} 
+%     \label{fig:policies.dataflow_graph}
+%     \vspace{-0.1in}
+%   \end{center}
+% \end{figure}
 
@@
   \vspace{-0.1in}
-  \caption{Comparison of dropping policies. Blue (dark) nodes indicate where
-  dropping decisions are made.}
+  \caption{Comparison of dropping policies using metadata dependence graphs.
+  Square nodes represent events where check are performed. Blue (dark) nodes
+  indicate which nodes can be dropped.}
   \label{fig:policies.policies}
 \end{figure}
+
+In addition to deciding when dropping is required, trade-offs also exist in
+deciding which events should be dropped.
+The simplest policy is to drop any monitoring events when slack is less than or equal to zero.  However, this can result in
+wasted work. For example, consider the metadata dependence graph shown in
+Figure~\ref{fig:policies.all_drop}. Here, an edge from node {\tt A} to node
+{\tt B} represents that if event {\tt A} is dropped, then due to its
+invalidated metadata, it will cause event {\tt B} to be dropped. Square
+nodes indicate events where monitoring checks are performed. In the
+example, suppose that event {\tt E} is meant to perform a check operation but is dropped.
+In this situation, the
+monitoring operations that were done for events {\tt C} and {\tt D} were wasted
+since their results were not used for any monitoring checks.
+That is, by the time we decide to drop event {\tt E}, we have already 
+updated metadata due to events {\tt C} and {\tt D} which is no longer needed.
 
@@
 
+Depending on the target application of partial monitoring, different policies
+are more applicable.
 For applications where closely matching an overhead target is important, a

### 2014-09-13 related.tex
Various edits based on a full pass.

@@
 % Monitoring
-Many monitoring schemes have been developed for various security, reliability,
-debugging, and other capabilities. There exist monitoring schemes for uninitialized
-memory check \cite{mondrian-asplos02}, array bounds check
-\cite{hardbound-asplos08, clause-ase07}, and dynamic information flow
-tracking \cite{dift-asplos04, raksha-isca07, loki-osdi08}.
-Memtracker \cite{memtracker-hpca07} uses monitoring to check for various
-memory-related bugs and errors.  Control flow verification techniques perform
-monitoring to check that program execution follows expected control flows
-\cite{schuette-comp87, impres-dac06,
-kayaalp-isca12}.  Our monitoring architecture is designed to
-be generally applicable to any of these instruction-grained monitoring schemes.
+% Many monitoring schemes have been developed for various security, reliability,
+% debugging, and other capabilities. There exist monitoring schemes for uninitialized
+% memory check \cite{mondrian-asplos02}, array bounds check
+% \cite{hardbound-asplos08, clause-ase07}, and dynamic information flow
+% tracking \cite{dift-asplos04, raksha-isca07, loki-osdi08}.
+% Memtracker \cite{memtracker-hpca07} uses monitoring to check for various
+% memory-related bugs and errors.  Control flow verification techniques perform
+% monitoring to check that program execution follows expected control flows
+% \cite{schuette-comp87, impres-dac06,
+% kayaalp-isca12}.  Our monitoring architecture is designed to
+% be generally applicable to any of these instruction-grained monitoring schemes.
+
 % In particular, our proposed architecture may be especially useful for some of
@@
 These sampling-based techniques all limit the debugging capabilities in order
-to provide lower performance overheads, similar to our architecture's goals.
+to provide lower performance overheads, which is one of the target applications of our architecture.
 % Limited Monitoring
@@
 
-Our work presents a general and efficient hardware platform to enable
+In contrast, our work is the first that we know of to present a general hardware platform that enables
 partial monitoring. The architecture is designed to be applicable to a wide range of monitoring
-techniques as well as monitor implementations. By studying various potential
+techniques as well as monitor implementations (core, FPGA, etc.). By studying various potential
 uses of partial monitoring, including sampling for cooperative debugging and

### 2014-09-13 tabs/previous_overheads.tex
Various edits based on a full pass.

@@
 
-DIFT \cite{dift-asplos04} & Custom HW & DIFT only & 1.1\% / 23\% \\ \hline
-FlexiTaint \cite{flexitaint-hpca08} & Custom HW & DIFT w/ programmable policies & 1.1\%-3.7\% / 8.7\% \\ \hline
-Hardbound \cite{hardbound-asplos08} & Custom HW & Array bounds checks only & 5\%-9\% / 22\% \\ \hline
-Harmoni \cite{harmoni-dsn12} & Custom HW & Tag-based monitors & 1\%-10\% / 20\% \\ \hline\hline
+% DIFT \cite{dift-asplos04} & Custom HW & DIFT only & 1.1\% / 23\% \\ \hline
+% FlexiTaint \cite{flexitaint-hpca08} & Custom HW & DIFT w/ programmable policies & 1.1\%-3.7\% / 8.7\% \\ \hline
+% Hardbound \cite{hardbound-asplos08} & Custom HW & Array bounds checks only & 5\%-9\% / 22\% \\ \hline
+% Harmoni \cite{harmoni-dsn12} & Custom HW & Tag-based monitors & 1\%-10\% / 20\% \\ \hline\hline
+
+DIFT \cite{dift-asplos04} & Custom HW & DIFT only & 1.01x / 1.23x \\ \hline
+FlexiTaint \cite{flexitaint-hpca08} & Custom HW & DIFT w/ programmable policies & 1.01x-1.04x / 1.09x \\ \hline
+Hardbound \cite{hardbound-asplos08} & Custom HW & Array bounds checks only & 1.05x-1.09x / 1.22x \\ \hline
+Harmoni \cite{harmoni-dsn12} & Custom HW & Tag-based monitors & 1.01x-1.10x / 1.20x \\ \hline\hline
 

### 2014-09-13 title.tex
Various edits based on a full pass.

@@
 systems. However, these run-time monitors can introduce large performance and energy
-overheads, especially when performed on programmable cores.
+overheads, especially for flexible systems that support a range of monitors.
 %Traditionally, these overheads have been an
@@
 made for partial monitoring and show that there exist interesting policy
-decisions which create a trade-off between closely
-matching an overhead target and the coverage that can be achieved.
+decisions depending on the target application of partial monitoring.
 Our experimental results show that overheads can be reduced significantly 
-by trading off coverage. For example, although average overheads for full
-monitoring range between 3-4x, the proposed architecture can still
+by trading off coverage. For example, with average overheads of 3-4x for full
+monitoring range, the proposed architecture can still
 achieve 43-82\% coverage when the slowdown is reduced to 1.5x.

### 2014-09-13 conclusion.tex
Various edits based on a full pass.
Cut to 11 pages.

@@
 architecture can be extended to enable null metadata filtering.
-Given this architecture, we explore design choices related to the dropping policy.
+Given this architecture, we explore the design choices related to the dropping policy.
 With partial monitoring, we see that with a 1.5x overhead target, BC can still
 achieve 82\% average coverage and UMC can still achieve 43\% average coverage.
-In addition, with a 1.5x overhead target, IMP shows less than 3\% error on
-average.
+At this overhead target, IMP shows less than 3\% error on average.
+With a 1.2x overhead target, LS achieves over 70\% average coverage.
 

### 2014-09-13 evaluation.tex
Cut to 11 pages.

@@
   \begin{center}
-    \includegraphics[width=\columnwidth, clip=true, trim=0 0.5in 0 0]{figs/data_filtering.pdf}
+    \includegraphics[width=\columnwidth, clip=true, trim=0 0.5in 0 0.1in]{figs/data_filtering.pdf}
     \vspace{-0.2in}
@@
   \begin{center}
-    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0]{figs/data_bc_sweep.pdf}
+    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_bc_sweep.pdf}
     \vspace{-0.2in}
@@
   \begin{center}
-    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0]{figs/data_umc_sweep.pdf}
+    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_umc_sweep.pdf}
     \vspace{-0.2in}
@@
   \begin{center}
-    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0]{figs/data_ls_sweep.pdf}
+    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_ls_sweep.pdf}
     \vspace{-0.2in}
@@
   \begin{center}
-    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0]{figs/data_imp_sweep.pdf}
+    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_imp_sweep.pdf}
     \vspace{-0.2in}
@@
   \begin{center}
-    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0]{figs/data_fpga_umc_sweep.pdf}
+    \vspace{-0.1in}
+    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_fpga_umc_sweep.pdf}
     \vspace{-0.2in}
@@
   \begin{center}
-  \vspace{-0.2in}
+  \vspace{-0.1in}
   \subfloat[Overhead budget error.]{
-    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0]{figs/data_umc_exec_time.pdf}
+    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_umc_exec_time.pdf}
     \label{fig:evaluation.umc_exec_time}
   }
-  \vspace{-0.2in}
+  \vspace{-0.1in}
   \subfloat[Coverage.]{
-    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0]{figs/data_umc_coverage.pdf}
+    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_umc_coverage.pdf}
     \label{fig:evaluation.umc_coverage}
@@
   \begin{center}
-  \vspace{-0.2in}
+  \vspace{-0.1in}
   \subfloat[Overhead budget error.]{
-    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0]{figs/data_bc_exec_time.pdf}
+    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_bc_exec_time.pdf}
     \label{fig:evaluation.bc_exec_time}
   }
-  \vspace{-0.2in}
+  \vspace{-0.1in}
   \subfloat[Coverage.]{
-    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0]{figs/data_bc_coverage.pdf}
+    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_bc_coverage.pdf}
     \label{fig:evaluation.bc_coverage}
@@
   \begin{center}
-    \includegraphics[width=\columnwidth, clip=true, trim=0 0.2in 0 0]{figs/data_multirun_coverage.pdf}
+    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_multirun_coverage.pdf}
     \vspace{-0.2in}

### 2014-09-13 evaluation.tex
typo fixes

@@
 run-time monitoring. We implement the monitor as a software-based monitor
-running on a parallel processor core. We model the main and monitoring cores as
+running on a parallel processor core, similar to LBA \cite{lbaa-asid06}.
+We model the main and monitoring cores as
 running at 2.5
@@
 
-Since LockSet race detection requires multi-threaded programs, we tested LS using
+Since LockSet race detection requires multi-threaded programs, we tested it using
 the applications from the SPLASH-2 benchmark set \cite{splash-isca95}. We ran these benchmarks for
@@
 overhead observed is 1.33x for the SPLASH-2 benchmarks.
-Our baseline system and overheads are similar previous multi-core
+Our baseline system and overheads are similar to previous multi-core
 monitoring systems such as LBA-accelerated \cite{lba-isca08} and FADE
@@
 71\% coverage with a 1.2x overhead. 
-Since the overheads of DIFT are low to begin with, setting an overhead target
-1.1x is enough for all benchmarks tested to reach 100\% coverage.
+We do not show a graph for DIFT because the overhead target of 1.1x is enough for
+all benchmark tested to reach 100\% coverage.
+%Since the overheads of DIFT are low to begin with, setting an overhead target
+%1.1x is enough for all benchmarks tested to reach 100\% coverage.
 
@@
 
-From these results, we see that for monitors which many short, independent
+From these results, we see that for monitors with many short, independent
 chains, such as UMC, source-only dropping can provide better coverage than
@@
 coverage.  This is due to the fact that it is likely that at least one
-monitoring event in a chain will be dropped with the unrestricted dropping
+monitoring event in a metadata dependence chain will be dropped with the unrestricted dropping
 policy. 
@@
 Source-only dropping shows a 5\% increase in coverage with only seven runs.
-Over hundreds or thousands of runs, source-only dropping should be able to
+Over many runs, source-only dropping should be able to
 reach 100\% code coverage.

### 2014-09-13 related.tex
typo fixes

@@
 Lo et al. \cite{lo-rtas14} developed a hardware architecture to limit
-monitoring specifically for hard real-time systems.
+monitoring to meet deadlines for hard real-time systems.
 
@@
 uses of partial monitoring, including sampling for cooperative debugging and
-partial protection for low overheads, we have determined important dropping
+partial protection for low overheads, we also identified important trade-offs in dropping
 policy decisions to consider in order to best meet the requirements of the

### 2014-09-13 dropping.tex
Cut to 11 pages.
Edits from Tao's comments.

@@
 testudo-micro08}, soft error detection \cite{argus-micro07}, data-race
-detection \cite{cord-hpca06}, etc.  For example, an array bounds check (BC)
+detection \cite{cord-hpca06, eraser-tocs97}, etc.  For example, an array bounds check (BC)
 \cite{hardbound-asplos08} can be implemented in order to detect
@@
 {\tt r3} is propagated to the metadata of the store address. If {\tt r3} were a
-pointer, than this would allow a future instruction to load the pointer and use
+pointer, then this would allow a future instruction to load the pointer and use
 it to access the corresponding array or data structure. In the second case,
@@
   \begin{center}
+    \vspace{-0.1in}
     \includegraphics[width=\columnwidth]{figs/dataflow_overview.pdf}
+    \vspace{-0.2in}
     \caption{Hardware support for dropping.}
     \label{fig:dropping.dataflow_overview}
-    \vspace{-0.1in}
+    \vspace{-0.2in}
   \end{center}
@@
     \includegraphics[]{figs/dataflow_architecture.pdf}
-    \vspace{-0.2in}
+    \vspace{-0.3in}
     \caption{Hardware architecture of the dataflow engine.}

### 2014-09-13 evaluation.tex
Edits from Tao's comments.

@@
 run-time monitoring. We implement the monitor as a software-based monitor
-running on a parallel processor core, similar to LBA \cite{lbaa-asid06}.
+running on a parallel processor core, similar to LBA \cite{lba-asid06}.
 We model the main and monitoring cores as
@@
 Figure~\ref{fig:evaluation.bc_sweep} shows the monitoring coverage achieved by
-array bounds check as we vary the overhead budget. 
+array bounds check as we vary the overhead budget. The bottom portion of each
+bar shows the coverage for a 1.1x overhead target and the additional
+coverage for increased overhead targets are stacked above this.
 We define \emph{monitoring coverage} as the 
@@
 of 420\% for {\tt bzip2}, 120\% for {\tt gcc}, and 1500\% for {\tt hmmer}.
-Since only array allocations are considered as source events for BC, it is
+Since only array allocations, which are infrequent, are considered as source events for BC, it is
 difficult for source-only dropping to match overhead targets. Although

### 2014-09-13 evaluation.tex
Minor edits, cut to 11 pages.

@@
 software \cite{tang-raid14}.
-LockSet attempts to detect possible race conditions in multi-threaded programs
+LockSet \cite{eraser-tocs97} attempts to detect possible race conditions in multi-threaded programs
 by tracking metadata about which locks are protecting shared memory

### 2014-09-13 introduction.tex
Minor edits, cut to 11 pages.

@@
 Our results show that the partial monitoring can still achieve significant
-coverage with reduced overheads. For example, for monitors which showed 3-4x
-average overheads, slowdown could be reduced to 1.5x while still achieving a
+coverage with reduced overheads. For example, for the monitors which showed over 3x
+average overheads, slowdown could be cut to 1.5x while still achieving a
 coverage of 43-82\% on average.

### 2014-09-13 title.tex
Minor edits, cut to 11 pages.

@@
 Our experimental results show that overheads can be reduced significantly 
-by trading off coverage. For example, with average overheads of 3-4x for full
-monitoring range, the proposed architecture can still
-achieve 43-82\% coverage when the slowdown is reduced to 1.5x.
+by trading off coverage. For example, for monitoring techniques with average
+overheads of 3-4x, the proposed architecture is able to reduce slowdowns to
+1.5x while still achieving 43-82\% average coverage.
 

### 2014-09-13 conclusion.tex
Minor edits.

@@
 With partial monitoring, we see that with a 1.5x overhead target, BC can still
-achieve 82\% average coverage and UMC can still achieve 43\% average coverage.
+achieve 82\% average coverage and UMC and LS can still achieve 43\%  and 94\%
+average coverage respectively.
 At this overhead target, IMP shows less than 3\% error on average.
-With a 1.2x overhead target, LS achieves over 70\% average coverage.
 

### 2014-09-13 dropping.tex
Minor edits.

@@
 When this happens,
-{\tt r3}'s metadata is also reset to NULL in case it previously stored a
+{\tt r3}'s metadata is reset to NULL in case it previously stored a
 pointer. 
@@
 way that metadata is. Figure~\ref{fig:dropping.example_invalid}
-shows an example of associating invalid flags with metadata. In the example, no
-monitoring operations are dropped. However, suppose that the monitoring for
+shows an example of associating invalid flags with metadata. Suppose that the monitoring for
 line 1 is dropped. When a monitoring event is dropped, metadata that would have
-been updated is marked as invalid. In this case, {\tt rf\_invalid[1]} will be
+been updated is marked as invalid. In this case, instead of the normal
+operation of marking {\tt rf\_invalid[1]} as {\tt false}, it is instead
 marked as {\tt
@@
 Figure~\ref{fig:dropping.example_full} and {\tt r2} on line 2 is no longer set as an array pointer.
-Without null metadata filtering, the instructions for lines 2 and 4 would be
+Without null metadata filtering, the instruction for line 4 would be
 forwarded to the monitor since the system does not know whether {\tt r2}
@@
 as null when it is loaded with a constant, then we can propagate this null
-information and filter out monitoring for lines 2 and 4.
+information and filter out monitoring for line 4.
 

### 2014-09-13 evaluation.tex
Minor edits.

@@
 the applications from the SPLASH-2 benchmark set \cite{splash-isca95}. We ran these benchmarks for
-200 million instructions with 2 main cores running the benchmark application.
+200 million instructions eachi using 2 main cores to run the benchmark application.
 Each main core was connected to a dataflow engine and a monitoring core.
@@
 targeted network or streaming applications, which have larger amounts of
-untrusted input data, we would higher overheads. IMP shows normalized
-execution times of 2-6x with an average of 4x. Overheads for LS (not pictured) are
+untrusted input data, we would observe higher overheads. IMP shows normalized
+execution times of 2-6x with an average of 4x. Overheads for LS (not pictured)
+applied to the SPLASH-2 benchmarks are
 1.03-1.72x after null metadata filtering and the average
-overhead observed is 1.33x for the SPLASH-2 benchmarks.
+overhead observed is 1.33x.
 Our baseline system and overheads are similar to previous multi-core
@@
 Figure~\ref{fig:evaluation.bc_sweep} shows the monitoring coverage achieved by
-array bounds check as we vary the overhead budget. The bottom portion of each
-bar shows the coverage for a 1.1x overhead target and the additional
-coverage for increased overhead targets are stacked above this.
+array bounds check as we vary the overhead budget. 
 We define \emph{monitoring coverage} as the 
@@
 non-check instructions.
+The bottom portion of each
+bar shows the coverage for a 1.1x overhead target and the additional
+coverage for increased overhead targets are stacked above this.
 
@@
 instruction types for each benchmark and for varying overheads. We see that
-with a 1.1x overhead a max error of 9\% is observed across benchmarks and the
+with a 1.1x overhead, a max error of 9\% is observed across the benchmarks and the
 average error across instruction types and benchmarks is less than 2\%.
@@
 From Figure~\ref{fig:evaluation.bc_exec_time}, we see that for several
-benchmarks, source-only drop fails to meet the specified overhead target.
-The overshoot of the overhead target is quite high with an overhead differences
+benchmarks, source-only dropping fails to meet the specified overhead target.
+The overshoot of the overhead target is quite high with overhead differences
 of 420\% for {\tt bzip2}, 120\% for {\tt gcc}, and 1500\% for {\tt hmmer}.
-Since only array allocations, which are infrequent, are considered as source events for BC, it is
+Since only infrequently occurring array allocations are considered as source events for BC, it can be 
 difficult for source-only dropping to match overhead targets. Although
@@
 source-only dropping is prone to overshoot the overhead target.
-Thus, for applications where staying with an overhead target is especially
+Thus, for applications where staying within an overhead target is especially
 important, such as soft real-time systems, a slack-based, unrestricted dropping
@@
 Figure~\ref{fig:evaluation.multirun} shows the total
-coverage achieved over multiple runs for array bounds check running for
+coverage achieved over multiple runs for array bounds check using
 unrestricted and source-only dropping policies with probabilistic dropping.

### 2014-09-13 monitoring.tex
Edits from Tao's comments.
Minor edits.

@@
 design points. % in the trade-off space between efficiency and flexibility.
-Table~\ref{tab:monitoring.previous_overheads} summarizes some the of representative designs
+Table~\ref{tab:monitoring.previous_overheads} summarizes some of the representative designs
 and their reported performance overhead. The previous studies clearly show that there
@@
 slowdowns of several tens of percents in some cases.
-In today's monitoring systems, the overhead is also unpredictable because they
+In today's monitoring systems, the overheads are also unpredictable because they
 depend heavily on the characteristics of applications and monitoring operations.
@@
 The main metric that represents the effectiveness of monitoring in this case is
-the coverage (the percentage of checks there were performed) over multiple runs.
+the coverage (the percentage of checks that were performed) over multiple runs.
 
@@
 The run-time monitoring system can be used to implement various profiling tools
-to collect statistics on program behaviors for performance optimizations as well
+to collect statistics on program behavior for performance optimizations as well
 as security protection. For example, a recent study showed that an instruction

### 2014-09-13 policies.tex
Cut to 11 pages.
Minor edits.

@@
   \subfloat[unrestricted dropping]{
-    \includegraphics[width=\columnwidth]{figs/all_drop.pdf}
+    \includegraphics[width=\columnwidth, clip=true, trim=0 0.1in 0 0]{figs/all_drop.pdf}
     \label{fig:policies.all_drop}
@@
   \subfloat[source-only dropping]{
-    \includegraphics[width=\columnwidth]{figs/source_drop.pdf}
+    \includegraphics[width=\columnwidth, clip=true, trim=0 0.1in 0 0]{figs/source_drop.pdf}
     \label{fig:policies.source_drop}
@@
 That is, by the time we decide to drop event {\tt E}, we have already 
-updated metadata due to events {\tt C} and {\tt D} which is no longer needed.
+updated metadata due to events {\tt C} and {\tt D} even though they are no longer needed.
 

### 2014-09-13 related.tex
Minor edits.

@@
 \cite{liblit-pldi05}. This idea has also been applied to detecting memory
-leaks~\cite{chilimbi-asplos04}, dataflow analysis~\cite{greathouse-cgo11}, and dynamic information flow tracking\cite{testudo-micro08} and
-other run-time monitoring techniques. 
+leaks~\cite{chilimbi-asplos04}, dataflow analysis~\cite{greathouse-cgo11}, and
+dynamic information flow tracking\cite{testudo-micro08}.
 These sampling-based techniques all limit the debugging capabilities in order

### 2014-09-16 evaluation.tex
Minor edit to eval.

@@
 non-check instructions.
-The bottom portion of each
+In the figure, the bottom portion of each
 bar shows the coverage for a 1.1x overhead target and the additional

### 2014-10-29 hpca15/hpca2015-rebuttal.txt
Initial summary of reviews.

Review-A

Overhead compared to previous literature (see review for references)

Consistency between checking coprocessor and processor.

1. How is your monitoring architecture (minus the partial monitoring support) different than one 
proposed in "Decoupling Dynamic Information Flow Tracking with a Dedicated Coprocessor"
by Kozyrakis et al.?

2. Do you guarantee to detect violations at instruction boundaries?

3. What is your exact definition of monitoring overhead?

4. Have you assessed the physical memory metadata requirements for monitors you designed?

5. Have you considered a design where the monitor uses a RAW microprocessor like fabric 
as opposed to a FIFO for communication?

Compare to Raksha.

What about false negatives?

Previous SW results that show low overheads.

Review-B

False negatives

Design space choice: Is it possible to use another core for monitoring tasks.

Review-C

For run time monitoring can you provide any precise guarantees regarding dropped requests (w/ the slack policy)?

Is hardware support really necessary? Partial monitoring with software
instrumentation (Sec 5.6 or [15])?

[7] has similarities to our approach.

Review-D

How does this compare to FADE?

How much better does this do than ad-hoc filtering in terms of error? (I believe Flexitaint also did some amount of drop-on-overhead policy)

Review-E

Advances over FADE

Measure coverage in terms of errors that were detected (i.e. false negative rate)

Review-F

Could you compare energy consumed by full and partial monitors?

Track invalid tags through different threads in multi-processor system? Coherence for dataflow monitor?

False negatives
Paper assumes that error/attacks are uniformly distributed across checks. Why is this a valid assumption?  It is possible that portions of execution that incur high monitoring overhead (say the ones with frequent monitoring events) would always not get checked. So you might never find bugs in such portions even if you monitor several executions.

How are the sources identified for the source-only dropping scheme? How and when do you decide drop the source?

Describe the hardware used to estimate slack.

Software tools have done sampling (see review for references).

### 2014-10-30 hpca15/hpca2015-rebuttal.txt
Initial draft of rebuttals for HPCA 2015.

@@
+Common Questions
+
+-Review-D,E: How does this compare to FADE?
+Our main contribution in this work is an architecture which enables partial
+monitoring (i.e., trade-off between coverage and overheads). We are not the
+first to introduce null metadata filtering as a method for reducing monitoring
+overheads. For example, FADE [11] introduces hardware to do this. However, we
+show that our hardware support for partial monitoring is able to support null
+metadata filtering with minor modifications. Thus, it is not necessary to
+implement the full FADE hardware in order to add null metadata filtering to our
+architecture for partial monitoring.
+
+-Review-A,B,E,F: Evaluation of false negatives/error detection rate.
+While we could not evaluate real errors/attacks due to difficulty in setting up
+a large number of bugs and exploits and/or running real-world applications on
+our simulation infrastructure, we believe that the percentage of checks
+provides a good initial estimate of detection probability when errors/attacks
+are uniformly distributed across checks. It will be interesting as future work
+to evaluate the system with real-world bugs to see if this distribution holds
+or if there exists correlations between bug locations and slack.
+
 Review-A
 
-Overhead compared to previous literature (see review for references)
+-Comparison to Raksha and [Kannan DSN 2009].
+Raksha and the architecture proposed in [Kannan DSN 2009]
+provide hardware architectures for information flow tracking. These
+architectures have very low overheads because they are designed for a specific
+monitoring technique.  Although Raskha provides some programmability in the
+information flow tracking policy, it does not support the wide range of
+monitors supported by our architecture (e.g., race detection).
 
-Consistency between checking coprocessor and processor.
+-Do you guarantee to detect violations at instruction boundaries?
+We do not guarantee when violations are detected. However, previous work by
+[Deng et al. ICCD 2011] has described how precise exceptions can be supported
+on a monitoring architecture similar to ours.
 
-1. How is your monitoring architecture (minus the partial monitoring support) different than one 
-proposed in "Decoupling Dynamic Information Flow Tracking with a Dedicated Coprocessor"
-by Kozyrakis et al.?
+-What is your exact definition of monitoring overhead?
+Monitoring overhead is the increase in execution time of the system running
+monitoring compared to without monitoring.
 
-2. Do you guarantee to detect violations at instruction boundaries?
+-Have you assessed the physical memory metadata requirements for monitors you designed?
+The worst-case memory usage can be determined based on the size of metadata per
+memory location. We have not evaluated the actual memory usage, but previous
+work on the implemented monitoring techniques have looked at this.
 
-3. What is your exact definition of monitoring overhead?
-
-4. Have you assessed the physical memory metadata requirements for monitors you designed?
-
-5. Have you considered a design where the monitor uses a RAW microprocessor like fabric 
+-Have you considered a design where the monitor uses a RAW microprocessor like fabric 
 as opposed to a FIFO for communication?
-
-Compare to Raksha.
-
-What about false negatives?
-
-Previous SW results that show low overheads.
+A FIFO provides a simple solution for providing buffered communication between
+the main core and the monitor.
 
@@
 
-False negatives
-
-Design space choice: Is it possible to use another core for monitoring tasks.
+-Design space choice: Is it possible to use another core for monitoring tasks.
+The monitoring task is performed by another core (with the exception of the
+FPGA-based monitor that was evaluated). However, in order to minimize the
+performance impact of dropping a monitoring event, this was implemented using
+custom hardware. In the past, we have explore performing these dropping
+operations in software. However, this was almost as expensive as performing the
+full monitoring operations and so was not effective at creating a trade-off
+between coverage and performance.
 
@@
 
-For run time monitoring can you provide any precise guarantees regarding dropped requests (w/ the slack policy)?
+-For run time monitoring can you provide any precise guarantees regarding
+ dropped requests (w/ the slack policy)?
+Dropping is decided based on run-time overheads (i.e., slack). This is
+difficult to predict a priori and so we are not able to give precise guarantees
+on what is dropped or the coverage achieved.
 
-Is hardware support really necessary? Partial monitoring with software
-instrumentation (Sec 5.6 or [15])?
+-Is hardware support really necessary? Partial monitoring with software
+instrumentation (Sec 5.6 or [15])? Higher overhead/coverage ratio, but with
+scaling out to many users, does it matter?
+A software solution is cheaper to implement. However, if the overhead/coverage
+ratio is too high, then the number of users needed to detect an error could be
+prohibitively high. Table 1 shows that the gap in overhead between software and
+hardware solutions for full monitoring can be an order of magnitude.
 
-[7] has similarities to our approach.
+-[7] has similarities to our approach.
+Our baseline architecture is similar to the one proposed by [7] and other
+related work. Our work shows how partial monitoring can be supported on these
+architectures using a dataflow engine.
 
@@
 
-How does this compare to FADE?
-
-How much better does this do than ad-hoc filtering in terms of error? (I believe Flexitaint also did some amount of drop-on-overhead policy)
+-How much better does this do than ad-hoc filtering in terms of error?
+Previous work uses filtering to reduce the overheads of monitoring. However,
+they are not able to say anything about the overheads seen, especially with new
+applications. The overheads can vary wildly (see Figure 9). Our architecture
+allows a specified overhead to be targeted.
 
 Review-E
-
-Advances over FADE
-
-Measure coverage in terms of errors that were detected (i.e. false negative rate)
+See common questions.
 
@@
 
-Could you compare energy consumed by full and partial monitors?
+-Could you compare energy consumed by full and partial monitors? 
+We have not evaluated the difference in energy consumed by full and partial
+monitors. It is not clear which uses more energy and it largely depends on the
+number of events dropped and thus the overhead target. If everything is
+dropped, then less energy is used since the dataflow engine operates on
+single-bit flags and is more energy-efficient than the monitor. However, if no
+events are dropped or filtered out, then the dataflow engine adds energy usage
+above the baseline full monitoring because it adds additional operations to be
+performed.
 
-Track invalid tags through different threads in multi-processor system? Coherence for dataflow monitor?
+-Tracking invalid tags through different threads in multi-processor system?
+ Coherence for dataflow monitor?
+Invalid tags for different threads would be tracked by the dataflow engine
+through shared memory, similar to shared data for multi-threaded programs. A
+multi-processor implementation of the architecture would require coherence for
+the caches of the dataflow engine and monitor, similar to the coherence needed
+for the cores.
 
-False negatives
-Paper assumes that error/attacks are uniformly distributed across checks.
-Why is this a valid assumption?
-It is possible that portions of execution that incur high monitoring overhead (say the ones with frequent monitoring events) would always not get checked.
-So you might never find bugs in such portions even if you monitor several executions.
-
-How are the sources identified for the source-only dropping scheme? How and when do you decide drop the source?
-
-Describe the hardware used to estimate slack.
-
-Software tools have done sampling (see review for references).
+-How are the sources identified for the source-only dropping scheme? How and
+when do you decide drop the source?
+Sources can be identified by instruction type or by the programmer/compiler.
+For example, for uninitialized memory check, all store instructions are
+configured to be marked as sources. Alternatively, instructions can be marked
+as source events by the programmer or compiler. For example, this is done for
+array bounds check when pointers are initialized to mark these points as
+sources. Dropping sources can be done based on slack or using a probabilistic
+approach.

### 2014-10-30 hpca15/hpca2015-rebuttal.txt
Rebuttal revision.

@@
 first to introduce null metadata filtering as a method for reducing monitoring
-overheads. For example, FADE [11] introduces hardware to do this. However, we
-show that our hardware support for partial monitoring is able to support null
-metadata filtering with minor modifications. Thus, it is not necessary to
-implement the full FADE hardware in order to add null metadata filtering to our
-architecture for partial monitoring.
+overheads (e.g., FADE). However, we show that our hardware support for
+partial monitoring is able to support null metadata filtering with minor
+modifications. Thus, it is not necessary to implement the full FADE hardware in
+order to add null metadata filtering to our architecture for partial
+monitoring.
 
@@
 
--Comparison to Raksha and [Kannan DSN 2009].
-Raksha and the architecture proposed in [Kannan DSN 2009]
+-Comparison to Raksha and [Kannan DSN2009].
+Raksha and the architecture proposed in [Kannan DSN2009]
 provide hardware architectures for information flow tracking. These
@@
 We do not guarantee when violations are detected. However, previous work by
-[Deng et al. ICCD 2011] has described how precise exceptions can be supported
+[Deng ICCD2011] has described how precise exceptions can be supported
 on a monitoring architecture similar to ours.
 
--What is your exact definition of monitoring overhead?
-Monitoring overhead is the increase in execution time of the system running
+-Definition of monitoring overhead?
+Monitoring overhead is the increase in execution time of a program running
 monitoring compared to without monitoring.
 
--Have you assessed the physical memory metadata requirements for monitors you designed?
+-Have you assessed the physical memory metadata requirements?
 The worst-case memory usage can be determined based on the size of metadata per
@@
 
--Have you considered a design where the monitor uses a RAW microprocessor like fabric 
-as opposed to a FIFO for communication?
-A FIFO provides a simple solution for providing buffered communication between
-the main core and the monitor.
+-Have you considered using a RAW-like fabric as opposed to a FIFO for
+communication?
+A FIFO provided a simple solution for buffered communication between the main
+core and the monitor.
 
@@
 
--Design space choice: Is it possible to use another core for monitoring tasks.
-The monitoring task is performed by another core (with the exception of the
-FPGA-based monitor that was evaluated). However, in order to minimize the
-performance impact of dropping a monitoring event, this was implemented using
-custom hardware. In the past, we have explore performing these dropping
-operations in software. However, this was almost as expensive as performing the
-full monitoring operations and so was not effective at creating a trade-off
+-Use another core for monitoring tasks.
+The monitoring task is performed by another core. However, dropping operations
+were implemented using custom hardware (i.e., dataflow engine) in order to
+minimize their performance impact. We have explored performing these dropping
+operations in software. However, this can be almost as expensive as performing
+the full monitoring operations and so was not effective at creating a trade-off
 between coverage and performance.
@@
 
--For run time monitoring can you provide any precise guarantees regarding
- dropped requests (w/ the slack policy)?
+-Can you provide any precise guarantees regarding dropped requests?
 Dropping is decided based on run-time overheads (i.e., slack). This is
@@
 
--Is hardware support really necessary? Partial monitoring with software
-instrumentation (Sec 5.6 or [15])? Higher overhead/coverage ratio, but with
+-Is hardware support really necessary? Higher overhead/coverage ratio, but with
 scaling out to many users, does it matter?
-A software solution is cheaper to implement. However, if the overhead/coverage
-ratio is too high, then the number of users needed to detect an error could be
-prohibitively high. Table 1 shows that the gap in overhead between software and
-hardware solutions for full monitoring can be an order of magnitude.
+Previous work that performs partial monitoring in software [15, 1] applied it
+to more coarse-grained monitoring techniques instead of the fine-grained
+instruction-level monitoring we target.  Table 1 shows that the gap in
+overhead between software and hardware solutions for fine-grained monitoring
+can be an order of magnitude.  Although a software solution is easier to
+deploy, if the overhead/coverage ratio is too high, then the number of users
+needed to detect an error or the minimum overhead to achieve any coverage could
+be prohibitively high.
 
--[7] has similarities to our approach.
+-[7] has similarities to your approach.
 Our baseline architecture is similar to the one proposed by [7] and other
-related work. Our work shows how partial monitoring can be supported on these
-architectures using a dataflow engine.
+related work. However, our work is the first to show how partial monitoring can
+be supported on these architectures using a dataflow engine.
 
@@
 they are not able to say anything about the overheads seen, especially with new
-applications. The overheads can vary wildly (see Figure 9). Our architecture
-allows a specified overhead to be targeted.
+applications. These overheads can vary wildly (see Figure 9). Our architecture
+allows a specified overhead to be targeted. FlexiTaint [Venkataramani HPCA2008]
+does not have a drop-on-overhead policy.
 
 Review-E
-See common questions.
+
+-Should we be measuring coverage in terms of errors detected, false positives
+that were generated?
+In terms of errors detected, see the response in Common Questions. Our
+architecture does not generate any false positives.
 
@@
 
--Could you compare energy consumed by full and partial monitors? 
+-Energy consumed by full and partial monitors.
 We have not evaluated the difference in energy consumed by full and partial
-monitors. It is not clear which uses more energy and it largely depends on the
-number of events dropped and thus the overhead target. If everything is
-dropped, then less energy is used since the dataflow engine operates on
-single-bit flags and is more energy-efficient than the monitor. However, if no
-events are dropped or filtered out, then the dataflow engine adds energy usage
-above the baseline full monitoring because it adds additional operations to be
-performed.
+monitors. Dropping an event is less energy than performing full monitoring, but
+the dataflow engine also adds energy overhead when not dropping. Thus, it is
+not clear which uses more energy and it largely depends on the overhead
+target and the number of events dropped.
 
@@
 Invalid tags for different threads would be tracked by the dataflow engine
-through shared memory, similar to shared data for multi-threaded programs. A
-multi-processor implementation of the architecture would require coherence for
-the caches of the dataflow engine and monitor, similar to the coherence needed
-for the cores.
+through shared memory. A multi-processor implementation of the architecture
+would require coherence for the caches of the dataflow engine and monitor,
+similar to the coherence needed for the core caches.
 
--How are the sources identified for the source-only dropping scheme? How and
-when do you decide drop the source?
-Sources can be identified by instruction type or by the programmer/compiler.
+-How are the sources identified? How do you decide to drop the source?
+Sources are identified by instruction type.
 For example, for uninitialized memory check, all store instructions are
-configured to be marked as sources. Alternatively, instructions can be marked
-as source events by the programmer or compiler. For example, this is done for
-array bounds check when pointers are initialized to mark these points as
+configured to be marked as sources. In addition, custom instructions used for
+setting metadata (e.g., setting array bounds metadata) are always marked as
 sources. Dropping sources can be done based on slack or using a probabilistic

### 2014-10-31 hpca15/hpca2015-rebuttal.txt
Cut down to 800 words.

@@
 -Comparison to Raksha and [Kannan DSN2009].
-Raksha and the architecture proposed in [Kannan DSN2009]
-provide hardware architectures for information flow tracking. These
-architectures have very low overheads because they are designed for a specific
-monitoring technique.  Although Raskha provides some programmability in the
-information flow tracking policy, it does not support the wide range of
-monitors supported by our architecture (e.g., race detection).
+Raksha and the architecture proposed in [Kannan DSN2009] provide hardware
+architectures for information flow tracking. These architectures show low
+overheads because they are designed for a specific monitoring technique.
+Although Raskha provides some programmability in the information flow tracking
+policy, it does not support the wide range of monitors supported by our
+architecture.
 
 -Do you guarantee to detect violations at instruction boundaries?
-We do not guarantee when violations are detected. However, previous work by
-[Deng ICCD2011] has described how precise exceptions can be supported
-on a monitoring architecture similar to ours.
-
--Definition of monitoring overhead?
-Monitoring overhead is the increase in execution time of a program running
-monitoring compared to without monitoring.
-
--Have you assessed the physical memory metadata requirements?
-The worst-case memory usage can be determined based on the size of metadata per
-memory location. We have not evaluated the actual memory usage, but previous
-work on the implemented monitoring techniques have looked at this.
-
--Have you considered using a RAW-like fabric as opposed to a FIFO for
-communication?
-A FIFO provided a simple solution for buffered communication between the main
-core and the monitor.
+We do not guarantee when violations are detected. However, previous work [Deng
+ICCD2011] has described how precise exceptions can be supported on such an
+architecture.
 
@@
 
--Use another core for monitoring tasks.
-The monitoring task is performed by another core. However, dropping operations
-were implemented using custom hardware (i.e., dataflow engine) in order to
-minimize their performance impact. We have explored performing these dropping
-operations in software. However, this can be almost as expensive as performing
-the full monitoring operations and so was not effective at creating a trade-off
-between coverage and performance.
+-Using another core for monitoring tasks.
+The monitoring task is performed by another core. Dropping operations were
+implemented using custom hardware in order to minimize their performance
+impact. We have explored performing these dropping operations in software.
+However, this can lead to high overheads even when all events are dropped.
 
@@
 -Can you provide any precise guarantees regarding dropped requests?
-Dropping is decided based on run-time overheads (i.e., slack). This is
-difficult to predict a priori and so we are not able to give precise guarantees
-on what is dropped or the coverage achieved.
+Dropping is decided based on run-time overheads. This is difficult to predict a
+priori and so we are not able to give precise guarantees on what is dropped or
+the coverage achieved.
 
@@
 scaling out to many users, does it matter?
-Previous work that performs partial monitoring in software [15, 1] applied it
-to more coarse-grained monitoring techniques instead of the fine-grained
-instruction-level monitoring we target.  Table 1 shows that the gap in
-overhead between software and hardware solutions for fine-grained monitoring
-can be an order of magnitude.  Although a software solution is easier to
-deploy, if the overhead/coverage ratio is too high, then the number of users
-needed to detect an error or the minimum overhead to achieve any coverage could
-be prohibitively high.
+Table 1 shows that the gap in overhead between software and hardware solutions
+for fine-grained monitoring can be an order of magnitude. Although a software
+solution is easier to deploy, if the overhead/coverage ratio is too high, then
+the number of users needed to detect an error or the minimum overhead to
+achieve any coverage can be prohibitively high.
 
@@
 
--How much better does this do than ad-hoc filtering in terms of error?
+-How much better does this do than ad-hoc filtering?
 Previous work uses filtering to reduce the overheads of monitoring. However,
-they are not able to say anything about the overheads seen, especially with new
-applications. These overheads can vary wildly (see Figure 9). Our architecture
-allows a specified overhead to be targeted. FlexiTaint [Venkataramani HPCA2008]
-does not have a drop-on-overhead policy.
+the overheads seen can vary wildly (see Figure 9). Our architecture allows a
+target overhead to be specified. FlexiTaint [Venkataramani HPCA2008] does not
+mention a drop-on-overhead policy.
 
@@
 -Should we be measuring coverage in terms of errors detected, false positives
-that were generated?
-In terms of errors detected, see the response in Common Questions. Our
-architecture does not generate any false positives.
+ generated?
+For errors detected, see Common Questions. Our architecture does not generate
+any false positives.
 
@@
 -Energy consumed by full and partial monitors.
-We have not evaluated the difference in energy consumed by full and partial
-monitors. Dropping an event is less energy than performing full monitoring, but
+Dropping an event is less energy than performing full monitoring, but
 the dataflow engine also adds energy overhead when not dropping. Thus, it is
@@
 
--Tracking invalid tags through different threads in multi-processor system?
- Coherence for dataflow monitor?
+-Tracking invalid tags through different threads? Coherence for dataflow
+ monitor?
 Invalid tags for different threads would be tracked by the dataflow engine
@@
 
--How are the sources identified? How do you decide to drop the source?
-Sources are identified by instruction type.
-For example, for uninitialized memory check, all store instructions are
-configured to be marked as sources. In addition, custom instructions used for
-setting metadata (e.g., setting array bounds metadata) are always marked as
-sources. Dropping sources can be done based on slack or using a probabilistic
-approach.
+-How are sources identified? How do you decide to drop the source?
+Sources are identified by instruction type. For example, for uninitialized
+memory check, all store instructions are configured to be marked as sources. In
+addition, custom instructions used for setting metadata (e.g., setting array
+bounds metadata) are always marked as sources. Dropping sources can be done
+based on slack or using a probabilistic approach.
+
+-Sample-based software tools.
+Sampling-based methods typically can reduce overheads but overheads can vary
+across applications. Results typically show that average overheads are low but
+can include outliers. This is not surprising as the overheads without sampling
+can vary (see Figure 9). Our architecture allows an overhead target to be
+specified.

### 2014-10-31 hpca15/hpca2015-rebuttal.txt
minor revisions

@@
 While we could not evaluate real errors/attacks due to difficulty in setting up
-a large number of bugs and exploits and/or running real-world applications on
-our simulation infrastructure, we believe that the percentage of checks
+a large number of bugs for multiple monitoring schemes,
+we believe that the percentage of checks
 provides a good initial estimate of detection probability when errors/attacks
@@
 
--Do you guarantee to detect violations at instruction boundaries?
-We do not guarantee when violations are detected. However, previous work [Deng
-ICCD2011] has described how precise exceptions can be supported on such an
-architecture.
+-Guarantee to detect violations at instruction boundaries?
+We do not enforce a check to be done before the monitored instruction finishes.
+However, previous work [Deng ICCD2011] described how precise exceptions can be 
+supported on such decoupled architecture.
 
@@
 
--Using another core for monitoring tasks.
-The monitoring task is performed by another core. Dropping operations were
+-Using another core for monitoring?
+We explored performing monitoring on another core and FPGA fabric. Dropping operations were
 implemented using custom hardware in order to minimize their performance
-impact. We have explored performing these dropping operations in software.
-However, this can lead to high overheads even when all events are dropped.
+impact. We explored performing the dropping operations in software, but
+found that that design leads to high overheads even when all events are dropped.
 
@@
 
--Can you provide any precise guarantees regarding dropped requests?
-Dropping is decided based on run-time overheads. This is difficult to predict a
-priori and so we are not able to give precise guarantees on what is dropped or
-the coverage achieved.
+-Provide any precise guarantees regarding dropped requests?
+The system does not provide strict guarantees on the overhead, but can closely
+match the target in practice. Because run-time overheads are difficult to
+predict a priori, there is no precise guarantee on what is dropped.
 
--Is hardware support really necessary? Higher overhead/coverage ratio, but with
-scaling out to many users, does it matter?
-Table 1 shows that the gap in overhead between software and hardware solutions
-for fine-grained monitoring can be an order of magnitude. Although a software
-solution is easier to deploy, if the overhead/coverage ratio is too high, then
-the number of users needed to detect an error or the minimum overhead to
-achieve any coverage can be prohibitively high.
+-Is hardware support really necessary? 
+We found that hardware support is necessary to closely match the overhead
+target (source dropping in software is insufficient) and provide meaningful coverage 
+with low overhead. High per-system coverage is also important when the goal
+is to provide a partial protection for individual systems.
 
@@
 
--How much better does this do than ad-hoc filtering?
+-How much better compared to ad-hoc filtering?
 Previous work uses filtering to reduce the overheads of monitoring. However,
@@
 
--Should we be measuring coverage in terms of errors detected, false positives
- generated?
+-Measuring coverage in errors detected? False positives?
 For errors detected, see Common Questions. Our architecture does not generate
@@
 -Energy consumed by full and partial monitors.
-Dropping an event is less energy than performing full monitoring, but
-the dataflow engine also adds energy overhead when not dropping. Thus, it is
-not clear which uses more energy and it largely depends on the overhead
-target and the number of events dropped.
+In our baseline monitoring architecture, monitoring uses another core and 
+has the power consumption comparable to that of the main core. On the other hand,
+the power consumption of the dropping hardware is often less than 10% of
+the main core power consumption. Therefore, our architecture certainly saves
+energy when monitoring operations are dropped. However, when almost no monitoring
+operation is dropped, we found that the dropping hardware incurs a small
+energy overhead.
+We will add more explicit energy comparisons.
 
@@
 through shared memory. A multi-processor implementation of the architecture
-would require coherence for the caches of the dataflow engine and monitor,
-similar to the coherence needed for the core caches.
+would require coherence for the meta-data, similar to what have been studied
+for LBA and Raksha.
 
 -How are sources identified? How do you decide to drop the source?
-Sources are identified by instruction type. For example, for uninitialized
+Sources are identified by instruction type or explicit SW instructions. 
+For example, for uninitialized
 memory check, all store instructions are configured to be marked as sources. In
 addition, custom instructions used for setting metadata (e.g., setting array
-bounds metadata) are always marked as sources. Dropping sources can be done
+bounds metadata) are marked as sources. Dropping sources can be done
 based on slack or using a probabilistic approach.

### 2014-10-31 hpca15/hpca2015-rebuttal.txt
Minor edits.

@@
 While we could not evaluate real errors/attacks due to difficulty in setting up
-a large number of bugs for multiple monitoring schemes,
-we believe that the percentage of checks
-provides a good initial estimate of detection probability when errors/attacks
-are uniformly distributed across checks. It will be interesting as future work
-to evaluate the system with real-world bugs to see if this distribution holds
-or if there exists correlations between bug locations and slack.
+a large number of bugs for multiple monitoring schemes, we believe that the
+percentage of checks provides a good initial estimate of detection probability
+when errors/attacks are uniformly distributed across checks. It will be
+interesting as future work to evaluate the system with real-world bugs to see
+if this distribution holds or if there exists correlations between bug
+locations and slack.
 
@@
 However, previous work [Deng ICCD2011] described how precise exceptions can be 
-supported on such decoupled architecture.
+supported on such decoupled architectures.
 
@@
 -Using another core for monitoring?
-We explored performing monitoring on another core and FPGA fabric. Dropping operations were
-implemented using custom hardware in order to minimize their performance
-impact. We explored performing the dropping operations in software, but
-found that that design leads to high overheads even when all events are dropped.
+We explored performing monitoring on another core and FPGA fabric. Dropping
+operations were implemented using custom hardware in order to minimize their
+performance impact. We explored performing the dropping operations in software,
+but found that that design leads to high overheads even when all events are
+dropped.
 
@@
 
--How much better compared to ad-hoc filtering?
+-How does this compare to ad-hoc filtering?
 Previous work uses filtering to reduce the overheads of monitoring. However,
@@
 operation is dropped, we found that the dropping hardware incurs a small
-energy overhead.
-We will add more explicit energy comparisons.
+energy overhead. We will add more explicit energy comparisons.
 
@@
 through shared memory. A multi-processor implementation of the architecture
-would require coherence for the meta-data, similar to what have been studied
+would require coherence for the metadata, similar to what has been studied
 for LBA and Raksha.
@@
 -How are sources identified? How do you decide to drop the source?
-Sources are identified by instruction type or explicit SW instructions. 
-For example, for uninitialized
-memory check, all store instructions are configured to be marked as sources. In
-addition, custom instructions used for setting metadata (e.g., setting array
-bounds metadata) are marked as sources. Dropping sources can be done
-based on slack or using a probabilistic approach.
+Sources are identified by instruction type or explicit SW instructions. For
+example, for uninitialized memory check, all store instructions are configured
+to be marked as sources. In addition, custom instructions used for setting
+metadata (e.g., setting array bounds metadata) are marked as sources. Dropping
+sources can be done based on slack or using a probabilistic approach.
 
--Sample-based software tools.
-Sampling-based methods typically can reduce overheads but overheads can vary
+-Sampling-based software tools.
+Sampling-based methods can reduce overheads but overheads can vary
 across applications. Results typically show that average overheads are low but

### 2014-12-17 related.tex
Fix up related work formatting.

@@
 leaks~\cite{chilimbi-asplos04}, dataflow analysis~\cite{greathouse-cgo11}, and
-dynamic information flow tracking\cite{testudo-micro08}.
+dynamic information flow tracking~\cite{testudo-micro08}.
 These sampling-based techniques all limit the debugging capabilities in order

### 2014-12-18 related.tex
Update related work.

--- 2014-09-13 related.tex
@@
 
-% Sampling
-Statistical sampling is a method that has been applied to various debugging
-techniques in order to reduce their performance overheads. For example, Liblit
-et al. sampled program runs of thousands of users in order to find bugs
-\cite{liblit-pldi05}. This idea has also been applied to detecting memory
-leaks~\cite{chilimbi-asplos04}, dataflow analysis~\cite{greathouse-cgo11}, and
-dynamic information flow tracking\cite{testudo-micro08}.
-These sampling-based techniques all limit the debugging capabilities in order
-to provide lower performance overheads, which is one of the target applications of our architecture.
-% Limited Monitoring
-In terms of partial monitoring with adjustable overheads, the Quality Virtual Machine (QVM) is a
-modification of the Java Virtual Machine (JVM) that supports run-time
-monitoring with controllable overheads \cite{qvm-oopsla08}. Similarly, Huang et
-al. created a framework for controlling the overheads of software-based
-monitoring \cite{huang-sttt12}. 
-Lo et al. \cite{lo-rtas14} developed a hardware architecture to limit
-monitoring to meet deadlines for hard real-time systems.
+% Previous work on sampling for monitoring
+\begin{table}[t]
+  \begin{center}
+    \vspace{-0.0in}
+    \begin{footnotesize}
+    \input{tabs/previous_sampling}
+    \end{footnotesize}
+    \caption{Previous work on partial run-time monitoring.}
+    \vspace{-0.2in}
+    \label{tab:monitoring.previous_sampling}
+  \end{center}
+\end{table}
 
-In contrast, our work is the first that we know of to present a general hardware platform that enables
-partial monitoring. The architecture is designed to be applicable to a wide range of monitoring
-techniques as well as monitor implementations (core, FPGA, etc.). By studying various potential
-uses of partial monitoring, including sampling for cooperative debugging and
-partial protection for low overheads, we also identified important trade-offs in dropping
-policy decisions to consider in order to best meet the requirements of the
-application.
+There exists a number of previous projects that have looked into performing
+partial monitoring in order to reduce the performance overheads. These
+platforms differ from ours in a number of ways including how monitoring is
+implemented and the monitoring techniques targeted. However, we note three main
+properties that differentiates our work:
+\begin{enumerate}
+  \item \textbf{Generality:} Our architecture applies to a variety of monitoring techniques.
+  \item \textbf{Adjustable Overhead:} Our architecture allows an overhead to be
+  targeted. Other work performs sampling to reduce overheads but does not try
+  to bound overheads, which we have shown can vary greatly.
+  \item \textbf{Prevent False Positives:} We prevent a mechanism to prevent
+  false positives. Previous work either has false positives or targets monitoring
+  techniques which degrade gracefully with sampling rather than exhibit false
+  positives.
+\end{enumerate}
+Our work is the first that we know of to present a hardware platform for partial
+monitoring that is general, allows an overhead target to be specified, and
+explicitly prevents false positives. In contrast, almost all previous work on
+partial monitoring achieves only some, but not all, of these properties.
+Table~\ref{tab:monitoring.previous_sampling} summarizes these differences.
+
+For example, there exists previous work for using statistical sampling to
+reduce the performance overheads of various debugging techniques. These include
+sampling for bug isolation \cite{liblit-pldi05}, memory leak detection
+\cite{chilimbi-asplos04}, race detection \cite{literace-pldi09, pacer-pldi10},
+and information flow tracking \cite{testudo-micro08, greathouse-cgo11}. These
+techniques modify the monitoring to support statistical sampling and so are not
+generalizable. For those that prevent false positives, this is also done with
+monitor-specific modifications. Finally, with the exception of the work by
+Greathouse et al. \cite{greathouse-cgo11}, they do not allow an overhead target
+to be specified.
+
+There also exists several projects that have looked into more general partial
+monitoring platforms. Arnold and Ryder \cite{arnold-pldi01} presented a general
+platform for sampling of instrumented code, but do not allow an overhead target
+to be specified and do not prevent false positives. Huang et al.
+\cite{huang-sttt12} proposed a general framework for controlling the overheads
+of software-based monitoring. However, they also do not explicitly address
+false positives and only target monitors which degrade gracefully when
+performed partially. Lo et al. \cite{lo-rtas14} designed a hardware
+architecture for performing monitoring on hard real-time systems that also
+prevents false positives. Their architecture is designed to meet real-time
+deadlines rather than overhead targets. In addition, in order to give strong
+guarantees, the coverage achieved is lower than our system. Finally, QVM
+\cite{qvm-oopsla08} proposes a modification to the Java Virtual Machine (JVM) to
+support monitoring with adjustable overheads. QVM does prevents false positives
+but does so be enabling or disabling monitoring on a per-object basis. This is
+similar to the idea of performing source-only dropping which our results show
+can lead to overshooting the overhead target depending on the monitoring
+technique. In addition, QVM is limited monitoring for code run on a JVM while
+our framework works for any binary.
+
+% In contrast, our work is the first that we know of to present a general hardware platform that enables
+% partial monitoring. The architecture is designed to be applicable to a wide range of monitoring
+% techniques as well as monitor implementations (core, FPGA, etc.). By studying various potential
+% uses of partial monitoring, including sampling for cooperative debugging and
+% partial protection for low overheads, we also identified important trade-offs in dropping
+% policy decisions to consider in order to best meet the requirements of the
+% application.
+
+% % Sampling
+% Statistical sampling is a method that has been applied to various debugging
+% techniques in order to reduce their performance overheads. For example, Liblit
+% et al. sampled program runs of thousands of users in order to find bugs
+% \cite{liblit-pldi05}. This idea has also been applied to detecting memory
+% leaks~\cite{chilimbi-asplos04}, dataflow analysis~\cite{greathouse-cgo11}, and
+% dynamic information flow tracking~\cite{testudo-micro08}.
+% These sampling-based techniques all limit the debugging capabilities in order
+% to provide lower performance overheads, which is one of the target applications of our architecture.
+% % Limited Monitoring
+% In terms of partial monitoring with adjustable overheads, the Quality Virtual Machine (QVM) is a
+% modification of the Java Virtual Machine (JVM) that supports run-time
+% monitoring with controllable overheads \cite{qvm-oopsla08}. Similarly, Huang et
+% al. created a framework for controlling the overheads of software-based
+% monitoring \cite{huang-sttt12}. 
+% Lo et al. \cite{lo-rtas14} developed a hardware architecture to limit
+% monitoring to meet deadlines for hard real-time systems.
 

### 2014-12-18 dropping.tex
Update to camera-ready formatting.
Various edits throughout paper.

@@
 series of \emph{monitoring operations}. The monitor operates in parallel to the
-main core. These events are referred to as \emph{monitoring events}. Depending
+main core. Events that trigger monitoring are referred to as \emph{monitoring events}. Depending
 on the type of monitoring event, different monitoring operations are
@@
     \includegraphics[]{figs/example_full.pdf}
-    \vspace{-0.2in}
     \caption{Example of array bounds check monitor.}
@@
 Note that this monitoring is performed automatically and transparently with
-almost no modification needed to the main program. The only modifications to
-the main program that are needed are to initially set metadata, such as setting
+almost no modification needed to the main program. The only modification to
+the main program that is needed is to initially set metadata, such as setting
 the base and bounds addresses on a {\tt malloc} call for array bounds check.
-Later propagation and checks of the metadata occur automatically as
+The propagation and checks of the metadata occur automatically as
 instructions are forwarded.
@@
 out-of-bounds access and raise an error. However, if the monitoring operation
-for this is dropped, then the error is not detected. This is the trade-off that
-we make in order to reduce overheads. That is, instead of either 100\% coverage
+for this is dropped, then the error is not detected. This reflects the trade-off that
+we make in order to reduce overheads. Instead of either 100\% coverage
 with all the associated overheads or no coverage and no overheads, we 
@@
 for line 4 in Figure~\ref{fig:dropping.example_full}) can only cause
-false negatives but will never cause false positives. On the other hand,
+false negatives and will never cause false positives. On the other hand,
 skipping monitoring operations that update metadata can lead to false positives
@@
 shows an example of associating invalid flags with metadata. Suppose that the monitoring for
-line 1 is dropped. When a monitoring event is dropped, metadata that would have
+line 1 is dropped in order to meet the overhead target. When a monitoring event is dropped, metadata that would have
 been updated is marked as invalid. In this case, instead of the normal
@@
     \includegraphics[]{figs/dataflow_architecture.pdf}
-    \vspace{-0.3in}
+    \vspace{-0.1in}
     \caption{Hardware architecture of the dataflow engine.}
@@
 hardware module to perform this null metadata filtering for a variety of
-monitoring schemes.
+monitoring schemes. Our architecture is able to support this null metadata
+filtering with a small modification.
 
@@
 Here, the main core's code has been slightly modified from
-Figure~\ref{fig:dropping.example_full} and {\tt r2} on line 2 is no longer set as an array pointer.
+Figure~\ref{fig:dropping.example_full} and on line 2, {\tt r2} is no longer set as an array pointer.
 Without null metadata filtering, the instruction for line 4 would be

### 2014-12-18 evaluation.tex
Update to camera-ready formatting.
Various edits throughout paper.

@@
 running at 2.5
-GHz with 4-way set-associative private L1 I/D caches and a shared 8-way 2 MB L2
+GHz with 4-way set-associative 32 kB private L1 I/D caches and a shared 8-way 2 MB L2
 cache. This setup is similar to the Snapdragon 801 processor commonly found in
@@
 memory check (UMC), array bounds check (BC), dynamic information flow
-tracking (DIFT), instruction-mix profiling (IMP), and LockSet-based race detection (LS).  Uninitialized memory
+tracking (DIFT), instruction-mix profiling (IMP), and LockSet-based race
+detection (LS).  
+Since we implement the monitor using a
+processor core and our dataflow engine was designed to be generally applicable,
+the same hardware platform supports all of these monitoring techniques.
+Uninitialized memory
 check seeks to detect loading from
@@
 locations. If a shared memory location is accessed while unprotected
-then a race condition may exist. Since we implement the monitor using a
-processor core and our dataflow engine was designed to be generally applicable,
-the same hardware platform supports all of these monitoring techniques.
+then a race condition may exist.
 
@@
 the applications from the SPLASH-2 benchmark set \cite{splash-isca95}. We ran these benchmarks for
-200 million instructions eachi using 2 main cores to run the benchmark application.
+200 million instructions each using 2 main cores to run the benchmark application.
 Each main core was connected to a dataflow engine and a monitoring core.
@@
   \begin{center}
-    \includegraphics[width=\columnwidth, clip=true, trim=0 0.5in 0 0.1in]{figs/data_filtering.pdf}
+    %\includegraphics[width=\columnwidth, clip=true, trim=0 0.5in 0 0.1in]{figs/data_filtering.pdf}
+    \includegraphics[width=\columnwidth]{figs/data_filtering.pdf}
     \vspace{-0.2in}
@@
 DIFT sees an average of 1.1x slowdown with a maximum slowdown of 1.3x with null
-metadata filtering. This low overhead is due, in part, to the fact
+metadata filtering. This low overhead is due to the fact
 that for our implementation of DIFT on SPEC
@@
   \begin{center}
-    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_bc_sweep.pdf}
+    \includegraphics[width=\linewidth]{figs/data_hb_sweep.pdf}
     \vspace{-0.2in}
@@
   \begin{center}
-    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_umc_sweep.pdf}
+    \includegraphics[width=\linewidth]{figs/data_umc_sweep.pdf}
     \vspace{-0.2in}
@@
   \begin{center}
-    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_ls_sweep.pdf}
+    \includegraphics[width=\linewidth]{figs/data_ls_sweep.pdf}
     \vspace{-0.2in}
@@
   \begin{center}
-    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_imp_sweep.pdf}
+    \includegraphics[width=\linewidth, clip=true]{figs/data_imp_sweep.pdf}
     \vspace{-0.2in}
@@
 how likely an error/attack instance is to be detected on an individual system, 
-assuming that errors/attacks are uniformly distributed across checks. 
+assuming that errors/attacks are uniformly distributed across checks. While
+this is not necessarily true for actual errors/attacks, we believe this metric
+provides a good initial estimate of the detection capabilities of the system.
 %While we could not evaluate real errors/attacks due to difficulty in setting up
@@
 71\% coverage with a 1.2x overhead. 
-We do not show a graph for DIFT because the overhead target of 1.1x is enough for
-all benchmark tested to reach 100\% coverage.
+For DIFT (not pictured), an overhead target of 1.1x is enough for
+all benchmarks tested to reach 100\% coverage.
 %Since the overheads of DIFT are low to begin with, setting an overhead target
@@
   }
-  \vspace{-0.1in}
+  \\
   \subfloat[Coverage.]{
@@
   }
-  \vspace{-0.1in}
   \caption{Comparing dropping policies for UMC.}
@@
 dropping and source-only dropping due to the fact that UMC consists of a large
-number of short, independent monitoring chains.
+number of short, independent monitoring dependence chains.
 Figure~\ref{fig:evaluation.umc_coverage} shows the coverage of UMC for
@@
   }
-  \vspace{-0.1in}
   \caption{Comparing dropping policies for BC.}
@@
     \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_multirun_coverage.pdf}
-    \vspace{-0.2in}
+    \vspace{-0.1in}
     \caption{Coverage over multiple runs for BC with a 10\% probability to not drop events.}

### 2014-12-18 policies.tex
Update to camera-ready formatting.
Various edits throughout paper.

@@
     \includegraphics[width=\columnwidth]{figs/slack.pdf}
-    \vspace{-0.3in}
+    \vspace{-0.2in}
     \caption{Slack and its effect on monitoring over time.}
@@
 The overhead budget is specified as a percentage of the main program's execution
-cycles without monitoring. We define \emph{slack} as the number of cycles of
+cycles without monitoring. In order to estimate overheads at run-time, we define \emph{slack} as the number of cycles of
 monitoring overhead that can be incurred while staying within the budget
-target. Slack is essentially the difference between the actual overheads seen
+target. Slack is essentially the difference between the actual overhead seen
 and the budget specified. Slack is generated as the main program runs and consumed as monitoring overheads occur.  For
@@
   \end{center}
-  \vspace{-0.1in}
   \caption{Comparison of dropping policies using metadata dependence graphs.
@@
 An alternative dropping policy which eliminates this wasted work is to only make dropping decisions at
-the root of these metadata flows. That is, we will decide to either monitor or
-not monitor an entire metadata flow. We refer to this dropping decision policy
+the root or source of these metadata flows. That is, we will decide to either monitor or
+not monitor an entire metadata flow. An example of these source nodes is the
+monitoring done to initialize base and bounds information on {\tt malloc} for
+an array bounds check. We refer to this dropping decision policy
 as \emph{source-only dropping} and we refer to the previous policy of 
@@
 A} through {\tt E}) needs to be monitored in order for the monitoring check to
-be useful, randomly dropping any single event will cause the final check to be
+be useful, randomly dropping any event will cause the final check to be
 invalid.

### 2014-12-24 dropping.tex
Various minor edits. Figure clean-up.

@@
 testudo-micro08}, soft error detection \cite{argus-micro07}, data-race
-detection \cite{cord-hpca06, eraser-tocs97}, etc.  For example, an array bounds check (BC)
+detection \cite{cord-hpca06, eraser-tocs97, literace-pldi09, pacer-pldi10}, etc.  For example, an array bounds check (BC)
 \cite{hardbound-asplos08} can be implemented in order to detect

### 2014-12-31 evaluation.tex
Update data.

@@
     \vspace{-0.1in}
-    \includegraphics[width=\linewidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_fpga_umc_sweep.pdf}
+    \includegraphics[width=\linewidth]{figs/data_fpga_umc_sweep.pdf}
     \vspace{-0.2in}
@@
   \subfloat[Overhead budget error.]{
-    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_umc_exec_time.pdf}
+    \includegraphics[width=\columnwidth]{figs/data_umc_overhead.pdf}
     \label{fig:evaluation.umc_exec_time}
@@
   \subfloat[Coverage.]{
-    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_umc_coverage.pdf}
+    \includegraphics[width=\columnwidth]{figs/data_umc_coverage.pdf}
     \label{fig:evaluation.umc_coverage}
@@
   \subfloat[Overhead budget error.]{
-    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_bc_exec_time.pdf}
+    \includegraphics[width=\columnwidth]{figs/data_hb_overhead.pdf}
     \label{fig:evaluation.bc_exec_time}
@@
   \subfloat[Coverage.]{
-    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_bc_coverage.pdf}
+    \includegraphics[width=\columnwidth]{figs/data_hb_coverage.pdf}
     \label{fig:evaluation.bc_coverage}
@@
 mW, corresponding to 4-11\% of the main core's runtime power.
+
+% Runtime power of monitoring core
+\begin{table}[tb]
+  \begin{center}
+    \vspace{-0.0in}
+    \begin{footnotesize}
+    \input{tabs/monitor_power}
+    \end{footnotesize}
+    \caption{Average runtime power of the monitoring core.}
+    \vspace{-0.2in}
+    \label{tab:evaluation.monitor_power}
+  \end{center}
+\end{table}
+
+Table~\ref{tab:evaluation.monitor_power} shows the runtime power usage of the
+monitoring core averaged across all benchmarks. These results are shown for an
+overhead target of 1.5x as well as when full monitoring is performed.

### 2014-12-31 data/filter_mon.py

@@ -1,0 +1,57 @@
+#!/usr/bin/python
+
+import sys
+sys.path.extend(['../..'])
+import tsg_plot
+import math
+
+# Parse data
+f = open("filter_mon.csv", 'r')
+subcat = f.readline().strip().split(',')[1:]
+cat = []
+data = []
+for line in f:
+  if line.strip():
+    ls = line.strip().split(',')
+    cat.append(ls[0])
+    data.append([float(x) for x in ls[1:]])
+f.close()
+
+# Set up plotting options
+opts = tsg_plot.PlotOptions()
+opts.data = data
+opts.labels = [cat, subcat]
+
+attribute_dict = \
+    {
+        'show' : False,
+        'file_name' : '../figs/data_filtering.pdf',
+        'paper_mode' : True,
+        'figsize' : (3.5, 2.0),
+        'ylabel' : 'Normalized Execution Time',
+        'legend_ncol' : 5,
+        'legend_handlelength' : 1.0,
+        'legend_columnspacing' : 0.8,
+        'legend_handletextpad' : 0.5,
+        'legend_bbox' : [-0.05, 1.05, 1, 0.1],
+        'rotate_labels' : False,
+        'rotate_labels_angle' : -30,
+        'fontsize' : 8,
+
+        #'colors' : ["#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd"]
+
+        # 'colors' : ["#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba", "#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba"],
+        # 'hatch' : [' ']*5 + ['///']*5
+
+        #'colors' : ['#8dd3c7', '#ffffb3', '#bebada', '#8dd3c7', '#ffffb3', '#bebada', '#8dd3c7', '#ffffb3', '#bebada', '#8dd3c6'],
+        #'hatch' : ['']*3 + ['/']*3 + ['xx']*3 + ['//']
+
+        'colors' : ['#eff3ff', '#bdd7e7', '#6baed6', '#3182bd', '#08519c', '#eff3ff', '#bdd7e7', '#6baed6', '#3182bd', '#08519c'],
+        'hatch' : ['']*5 + ['//']*5
+
+    }
+for name, value in attribute_dict.iteritems():
[…]

### 2014-12-31 data/fpga_bc_sweep.py

@@ -1,0 +1,104 @@
+#!/usr/bin/python
+
+import sys
+sys.path.extend(['../..'])
+import tsg_plot
+import numpy
+
+def geomean(l):
+  """
+  Return the geometric mean of the passed list.
+  """
+  prod = 1.0
+  n = 0
+  for ll in l:
+    prod *= ll
+    n += 1
+  return prod**(1.0/n)
+
+
+def parse():
+  # Parse data
+  data = {}
+  f = open("fpga_bc_sweep.csv", 'r')
+  for line in f:
+    ls = line.strip().split(',')
+    # Find monitor names
+    if len(ls) == 1:
+      if ls[0] != '':
+        monitor = ls[0]
+        data[monitor] = {}
+        data[monitor]['overheads'] = []
+        data[monitor]['data'] = []
+    else:
+      # Benchmarks
+      if ls[0] == '':
+        data[monitor]['benchmarks'] = ls[1:] + ["geomean"]
+      # Data line
+      else:
+        overhead = float(ls[0].split('=')[1])
+        if overhead != 0 and overhead != 30:
+          data[monitor]['overheads'].append("%.2fx" % (overhead + 1))
+          d = [float(x)*100 for x in ls[1:]]
+          d.append(geomean(d))
+          data[monitor]['data'].append(d)
+  f.close()
+
+  # Reformat data
+  for monitor in data.keys():
+    d = data[monitor]['data']
+    # Get differences for stacked graph
+    diffs = [d[0]]
+    for i in range(1, len(d)):
+      diffs.append([x - y for (x, y) in zip(d[i], d[i-1])])
+    # Transpose
+    diffs = numpy.array(diffs).transpose()
+    data[monitor]['data'] = diffs
+
+  return data
+
+
+def plot(data, filename):
+  # Set up plotting options
+  opts = tsg_plot.PlotOptions()
+  # Benchmarks
+  cat = data['benchmarks']
+  # Overheads
+  subcat = data['overheads']
+  opts.data = data['data']
+  opts.labels = [cat, subcat]
+
+  attribute_dict = \
+      {
+          'plot_type' : 'stacked_bar',
+          'show' : False,
+          'file_name' : filename,
+          'paper_mode' : True,
+          'figsize' : (3.5, 2.0),
+          'ylabel' : 'Coverage [%]',
+          'legend_ncol' : 6,
+          'legend_handlelength' : 1.0,
[…]

### 2014-12-31 data/fpga_umc_sweep.py

@@ -1,0 +1,103 @@
+#!/usr/bin/python
+
+import sys
+sys.path.extend(['../..'])
+import tsg_plot
+import numpy
+
+def geomean(l):
+  """
+  Return the geometric mean of the passed list.
+  """
+  prod = 1.0
+  n = 0
+  for ll in l:
+    prod *= ll
+    n += 1
+  return prod**(1.0/n)
+
+
+def parse():
+  # Parse data
+  data = {}
+  f = open("fpga_umc_sweep.csv", 'r')
+  for line in f:
+    ls = line.strip().split(',')
+    # Find monitor names
+    if len(ls) == 1:
+      if ls[0] != '':
+        monitor = ls[0]
+        data[monitor] = {}
+        data[monitor]['overheads'] = []
+        data[monitor]['data'] = []
+    else:
+      # Benchmarks
+      if ls[0] == '':
+        data[monitor]['benchmarks'] = ls[1:] + ["geomean"]
+      # Data line
+      else:
+        overhead = float(ls[0].split('=')[1])
+        if overhead != 0 and overhead != 30:
+          data[monitor]['overheads'].append("%.2fx" % (overhead + 1))
+          d = [float(x)*100 for x in ls[1:]]
+          d.append(geomean(d))
+          data[monitor]['data'].append(d)
+  f.close()
+
+  # Reformat data
+  for monitor in data.keys():
+    d = data[monitor]['data']
+    # Get differences for stacked graph
+    diffs = [d[0]]
+    for i in range(1, len(d)):
+      diffs.append([x - y for (x, y) in zip(d[i], d[i-1])])
+    # Transpose
+    diffs = numpy.array(diffs).transpose()
+    data[monitor]['data'] = diffs
+
+  return data
+
+
+def plot(data, filename):
+  # Set up plotting options
+  opts = tsg_plot.PlotOptions()
+  # Benchmarks
+  cat = data['benchmarks']
+  # Overheads
+  subcat = data['overheads']
+  opts.data = data['data']
+  opts.labels = [cat, subcat]
+
+  attribute_dict = \
+      {
+          'plot_type' : 'stacked_bar',
+          'show' : False,
+          'file_name' : filename,
+          'paper_mode' : True,
+          'figsize' : (3.5, 2.0),
+          'ylabel' : 'Coverage [%]',
+          'legend_ncol' : 6,
+          'legend_handlelength' : 1.0,
[…]

### 2014-12-31 data/imp_sweep.py

@@ -1,0 +1,99 @@
+#!/usr/bin/python
+
+import sys
+sys.path.extend(['../..'])
+import tsg_plot
+import math
+
+# Parse data
+f = open("imp_sweep.csv", 'r')
+"""
+f.readline()
+subcat = f.readline().strip().split(',')[1:]
+
+cat = []
+average_errors = []
+for i in range(4):
+  ls = f.readline().split(',')
+  cat.append(ls[0])
+  average_errors.append([float(x) for x in ls[1:]])
+for i in range(5):
+  f.readline()
+max_errors = []
+for i in range(4):
+  ls = f.readline().split(',')
+  max_errors.append([float(x) for x in ls[1:]])
+for i in range(5):
+  f.readline()
+min_errors = []
+for i in range(4):
+  ls = f.readline().split(',')
+  min_errors.append([float(x) for x in ls[1:]])
+"""
+
+average_errors = []
+min_errors = []
+max_errors = []
+overheads = []
+mode = None
+line = f.readline()
+while line:
+  ls = line.strip()
+  # Skip empty lines
+  if not ls:
+    pass
+  # Switch to new data type
+  elif ls == "Average" or ls == "Min" or ls == "Max":
+    mode = ls
+    benchmarks = f.readline().strip().split(',')[1:]
+  elif mode == "Average":
+    ls = ls.split(',')
+    overheads.append(ls[0])
+    average_errors.append([float(x) for x in ls[1:]])
+  elif mode == "Min":
+    ls = ls.split(',')
+    min_errors.append([float(x) for x in ls[1:]])
+  elif mode == "Max":
+    ls = ls.split(',')
+    overheads.append(ls[0])
+    max_errors.append([float(x) for x in ls[1:]])
+  else:
+    pass
+
+  line = f.readline()
+
+f.close()
+
+import numpy
+average_errors = numpy.array(average_errors).transpose()
+min_errors = numpy.array(min_errors).transpose()
+max_errors = numpy.array(max_errors).transpose()
+subcat = benchmarks
+cat = overheads
+
+# Set up plotting options
+opts = tsg_plot.PlotOptions()
+opts.data = average_errors
+opts.labels = [subcat, cat]
+
+attribute_dict = \
+    {
+        'show' : False,
+        'file_name' : '../figs/data_imp_sweep.pdf',
+        'paper_mode' : True,
+        'figsize' : (7.0, 2.0),
[…]

### 2014-12-31 data/policies.py

@@ -1,0 +1,102 @@
+#!/usr/bin/python
+
+import sys
+sys.path.extend(['../..'])
+import tsg_plot
+import numpy
+
+monitors = ["BC", "UMC"]
+metrics = ["Coverage", "Overhead difference"]
+
+def geomean(l):
+  """
+  Return the geometric mean of the passed list.
+  """
+  prod = 1.0
+  n = 0
+  for ll in l:
+    prod *= ll
+    n += 1
+  return prod**(1.0/n)
+
+
+def parse():
+  # Parse data
+  data = {}
+  f = open("policies.csv", 'r')
+  for line in f:
+    ls = line.strip().split(',')
+    if ls == ['']:
+      continue
+    elif ls[0] in monitors:
+      monitor = ls[0]
+      data[monitor] = {}
+    else:
+      if ls[0] in metrics:
+        metric = ls[0]
+        data[monitor][metric] = {}
+        data[monitor][metric]['data'] = []
+        data[monitor][metric]['categories'] = []
+        data[monitor][metric]['benchmarks'] = ls[1:]
+      elif len(ls) > 0:
+        data[monitor][metric]['categories'].append(ls[0])
+        ls2 = [float(d[:-1]) for d in ls[1:]]
+        data[monitor][metric]['data'].append(ls2)
+  f.close()
+
+  # Transpose
+  for x in monitors:
+    for y in metrics:
+      d = data[x][y]['data']
+      d = numpy.array(d).transpose()
+      data[x][y]['data'] = d
+
+  return data
+
+def plot(data, filename):
+  # Set up plotting options
+  opts = tsg_plot.PlotOptions()
+  # Benchmarks
+  cat = data['benchmarks']
+  # Overheads
+  subcat = data['categories']
+  opts.data = data['data']
+  opts.labels = [cat, subcat]
+
+  attribute_dict = \
+      {
+          'plot_type' : 'bar',
+          'show' : False,
+          'file_name' : filename,
+          'paper_mode' : True,
+          'figsize' : (3.5, 2.0),
+          'ylabel' : 'Coverage [%]',
+          'legend_ncol' : 6,
+          'legend_handlelength' : 1.0,
+          'legend_columnspacing' : 0.8,
+          'legend_handletextpad' : 0.5,
+          'legend_bbox' : [-0.05, 1.05, 1, 0.1],
+          'rotate_labels' : True,
[…]

### 2014-12-31 data/policy_coverage.py

@@ -1,0 +1,98 @@
+#!/usr/bin/python
+
+import sys
+sys.path.extend(['../..'])
+import tsg_plot
+import numpy
+
+def geomean(l):
+  """
+  Return the geometric mean of the passed list.
+  """
+  prod = 1.0
+  n = 0
+  for ll in l:
+    prod *= ll
+    n += 1
+  return prod**(1.0/n)
+
+
+def parse():
+  # Parse data
+  data = {}
+  f = open("policy_coverage.csv", 'r')
+  for line in f:
+    ls = line.strip().split(',')
+    # Find monitor names
+    if len(ls) == 1:
+      if ls[0] != '':
+        monitor = ls[0]
+        data[monitor] = {}
+        data[monitor]['policies'] = []
+        data[monitor]['data'] = []
+    else:
+      # Benchmarks
+      if ls[0] == '':
+        data[monitor]['benchmarks'] = ls[1:] + ["geomean"]
+      # Data line
+      else:
+        policy = ls[0]
+        data[monitor]['policies'].append(policy)
+        d = [float(x)*100 for x in ls[1:]]
+        d.append(geomean(d))
+        data[monitor]['data'].append(d)
+  f.close()
+
+  # Reformat data
+  for monitor in data.keys():
+    # Transpose
+    d = data[monitor]['data']
+    d = numpy.array(d).transpose()
+    data[monitor]['data'] = d
+
+  return data
+
+
+def plot(data, filename):
+  # Set up plotting options
+  opts = tsg_plot.PlotOptions()
+  # Benchmarks
+  cat = data['benchmarks']
+  # Overheads
+  subcat = data['policies']
+  opts.data = data['data']
+  opts.labels = [cat, subcat]
+
+  attribute_dict = \
+      {
+          'plot_type' : 'bar',
+          'show' : False,
+          'file_name' : filename,
+          'paper_mode' : True,
+          'figsize' : (3.5, 2.0),
+          'ylabel' : 'Coverage [%]',
+          'legend_ncol' : 6,
+          'legend_handlelength' : 1.0,
+          'legend_columnspacing' : 0.8,
+          'legend_handletextpad' : 0.5,
+          'legend_bbox' : [-0.05, 1.05, 1, 0.1],
+          'rotate_labels' : True,
+          'rotate_labels_angle' : -45,
+          'fontsize' : 8,
+          'bar_width' : 0.5,
+
[…]

### 2014-12-31 data/policy_overhead.py

@@ -1,0 +1,100 @@
+#!/usr/bin/python
+
+import sys
+sys.path.extend(['../..'])
+import tsg_plot
+import numpy
+
+def geomean(l):
+  """
+  Return the geometric mean of the passed list.
+  """
+  prod = 1.0
+  n = 0
+  for ll in l:
+    prod *= ll
+    n += 1
+  return prod**(1.0/n)
+
+
+def parse():
+  # Parse data
+  data = {}
+  f = open("policy_overhead.csv", 'r')
+  for line in f:
+    ls = line.strip().split(',')
+    # Find monitor names
+    if len(ls) == 1:
+      if ls[0] != '':
+        monitor = ls[0]
+        data[monitor] = {}
+        data[monitor]['policies'] = []
+        data[monitor]['data'] = []
+    else:
+      # Benchmarks
+      if ls[0] == '':
+        #data[monitor]['benchmarks'] = ls[1:] + ["geomean"]
+        data[monitor]['benchmarks'] = ls[1:]
+      # Data line
+      else:
+        policy = ls[0]
+        data[monitor]['policies'].append(policy)
+        d = [float(x)*100 for x in ls[1:]]
+        #d.append(geomean(d))
+        data[monitor]['data'].append(d)
+  f.close()
+
+  # Reformat data
+  for monitor in data.keys():
+    # Transpose
+    d = data[monitor]['data']
+    d = numpy.array(d).transpose()
+    data[monitor]['data'] = d
+
+  return data
+
+
+def plot(data, filename):
+  # Set up plotting options
+  opts = tsg_plot.PlotOptions()
+  # Benchmarks
+  cat = data['benchmarks']
+  # Overheads
+  subcat = data['policies']
+  opts.data = data['data']
+  opts.labels = [cat, subcat]
+
+  attribute_dict = \
+      {
+          'plot_type' : 'bar',
+          'show' : False,
+          'file_name' : filename,
+          'paper_mode' : True,
+          'figsize' : (3.5, 2.0),
+          'ylabel' : 'Overhead Difference [%]',
+          'legend_ncol' : 6,
+          'legend_handlelength' : 1.0,
+          'legend_columnspacing' : 0.8,
+          'legend_handletextpad' : 0.5,
+          'legend_bbox' : [-0.05, 1.05, 1, 0.1],
+          'rotate_labels' : True,
+          'rotate_labels_angle' : -45,
[…]

### 2014-12-31 data/sweeps.py

@@ -1,0 +1,106 @@
+#!/usr/bin/python
+
+import sys
+sys.path.extend(['../..'])
+import tsg_plot
+import numpy
+
+def geomean(l):
+  """
+  Return the geometric mean of the passed list.
+  """
+  prod = 1.0
+  n = 0
+  for ll in l:
+    prod *= ll
+    n += 1
+  return prod**(1.0/n)
+
+
+def parse():
+  # Parse data
+  data = {}
+  f = open("sweeps.csv", 'r')
+  for line in f:
+    ls = line.strip().split(',')
+    # Find monitor names
+    if len(ls) == 1:
+      if ls[0] != '':
+        monitor = ls[0]
+        data[monitor] = {}
+        data[monitor]['overheads'] = []
+        data[monitor]['data'] = []
+    else:
+      # Benchmarks
+      if ls[0] == '':
+        data[monitor]['benchmarks'] = ls[1:] + ["geomean"]
+      # Data line
+      else:
+        overhead = float(ls[0].split('=')[1])
+        if overhead != 0 and overhead != 30:
+          if overhead < 0.1:
+            data[monitor]['overheads'].append("%.2fx" % (overhead + 1))
+          else:
+            data[monitor]['overheads'].append("%.1fx" % (overhead + 1))
+          d = [float(x)*100 for x in ls[1:]]
+          d.append(geomean(d))
+          data[monitor]['data'].append(d)
+  f.close()
+
+  # Reformat data
+  for monitor in data.keys():
+    d = data[monitor]['data']
+    # Get differences for stacked graph
+    diffs = [d[0]]
+    for i in range(1, len(d)):
+      diffs.append([x - y for (x, y) in zip(d[i], d[i-1])])
+    # Transpose
+    diffs = numpy.array(diffs).transpose()
+    data[monitor]['data'] = diffs
+
+  return data
+
+
+def plot(data, filename):
+  # Set up plotting options
+  opts = tsg_plot.PlotOptions()
+  # Benchmarks
+  cat = data['benchmarks']
+  # Overheads
+  subcat = data['overheads']
+  opts.data = data['data']
+  opts.labels = [cat, subcat]
+
+  attribute_dict = \
+      {
+          'plot_type' : 'stacked_bar',
+          'show' : False,
+          'file_name' : filename,
+          'paper_mode' : True,
+          'figsize' : (3.5, 2.0),
[…]

### 2014-12-31 data/tsg_plot.py

@@ -1,0 +1,460 @@
+#! /usr/bin/env python
+"""
+Library for plotting.
+
+Classes:
+  PlotOptions
+    Object passed to plotting function that defines options.
+
+Functions:
+  add_plot( opt )
+  add_scatter_plot( ax, opt )
+  add_bar_plot( ax, opt )
+  set_legend( ax, opt, legend, legend_labels )
+  set_common( ax, opt )
+  main()
+"""
+
+import matplotlib
+import numpy as np
+import matplotlib.pyplot as plt
+import random
+import sys
+
+
+class PlotOptions:
+
+  def __init__( self ):
+
+    # default values
+    self.title = "test bar plot"
+    self.ylabel = None
+    self.xlabel = None
+    self.bar_width = 0.70
+
+    self.valid_plot_types = ["bar", "stacked_bar", "scatter_bar", "scatter"]
+    self.plot_type = "bar"
+
+    # Error bars, lists of 2 lists. The first contains the minimums to draw and the second contains the maximums to draw.
+    self.errorbars = None
+
+    # if true, draws an arrow from the earlier to later
+    self.scatter_bar_arrow = False
+    # only draw arrow head if the length is this long
+    self.scatter_bar_arrow_minlen = 0.2
+
+    # Can provide multiple file names seperated by spaces
+    self.file_name = ""
+    # option to rotate labels when there are many of them
+    self.rotate_labels = False
+    self.rotate_labels_angle = -45
+
+    # set the font size for the rest of things
+    self.fontsize = 12
+
+    # set the font size for the labels
+    self.labels_fontsize = 12
+
+    # used to override default y range. you can provide the range in the
+    # form of [min, max]
+    self.yrange = None
+    # Override default x range.
+    self.xrange = None
+
+    # if not None, normalize line draws a horizontal line at indicated y
+    self.normalize_line = None
+
+    self.fig = None
+    # this is useful for drawing multiple plots in a page
+    self.plot_idx = 1
+    self.num_cols = 1
+    self.num_rows = 1
+
+    # if non-zero, shares axis with the previous plot. 1 indicates x axis
+    # sharing, 2 indicates y axis sharing
+    self.share_axis = 0
+
+    # if true, just show the plot, don't save
+    self.show = True
+
+    # paper mode squishes everything
[…]

### 2015-01-03 conclusion.tex
Updated data.

@@
 With partial monitoring, we see that with a 1.5x overhead target, BC can still
-achieve 82\% average coverage and UMC and LS can still achieve 43\%  and 94\%
-average coverage respectively.
-At this overhead target, IMP shows less than 3\% error on average.
+achieve 85\% average coverage and UMC can still achieve 14\% 
+average coverage.
+At this overhead target, IMP shows less than 2\% error on average.
+LS can be pushed down to 1.1x overhead and still achieve 61\% average coverage.
 

### 2015-01-03 evaluation.tex
Updated data.

@@
 for these benchmarks are similar to the other results shown.
-We simulated each benchmark for 2 billion instructions. 
+We fast-forwarded each benchmark for 1 billion instructions and then simulated for 2 billion instructions. 
 % An initial slack of 2
@@
 Since LockSet race detection requires multi-threaded programs, we tested it using
-the applications from the SPLASH-2 benchmark set \cite{splash-isca95}. We ran these benchmarks for
-200 million instructions each using 2 main cores to run the benchmark application.
+the applications from the SPLASH-2 benchmark set \cite{splash-isca95}. 
+Each benchmark was run using 2 main cores to run the benchmark application.
+{\tt fmm} and {\tt raytrace} were run without fast-forwarding because they ran to completion in under 2 billion instructions.
 Each main core was connected to a dataflow engine and a monitoring core.
@@
 to the execution time of the benchmarks without monitoring. UMC sees normalized
-execution times of 1.7-7x with an average of 4x.
+execution times of 2.6-9.8x with an average of 5.7x.
 For BC, normalized execution
-are 3x on average and range from 1.1x to 8.3x.
+are 2.7x on average and range from 1.02x to 13.6x.
 % This is not surprising since in the baseline
@@
 untrusted input data, we would observe higher overheads. IMP shows normalized
-execution times of 2-6x with an average of 4x. Overheads for LS (not pictured)
+execution times of 1.6-9.5x with an average of 5.8x. Overheads for LS (not pictured)
 applied to the SPLASH-2 benchmarks are
@@
 how likely an error/attack instance is to be detected on an individual system, 
-assuming that errors/attacks are uniformly distributed across checks. While
-this is not necessarily true for actual errors/attacks, we believe this metric
+assuming that errors/attacks are uniformly distributed across checks. This may
+not necessarily be true for actual errors/attacks and so the reported coverage
+may not be the same as the probability of detecting actual errors/attacks.
+However, we believe this metric
 provides a good initial estimate of the detection capabilities of the system.
@@
 With only a 1.1x overhead target (10\% slowdown), array bounds check still
-achieves 84\% monitoring coverage on average. The high coverage
+achieves over 80\% monitoring coverage on average. The high coverage
 achieved with such low overheads is due to two main effects.  The first is that
@@
 Figure~\ref{fig:evaluation.umc_sweep} shows the analogous graph for UMC. Again
-we see that varying overhead budgets enables partial monitoring. We see that
-with a 1.1x overhead target, UMC achieves 25\% monitoring coverage on average.
-By increasing this overhead budget to 1.5x, UMC achieves 43\% coverage on
-average. Even higher coverage can be achieved by allowing higher overheads.
+we see that varying overhead budgets enables partial monitoring.
+With a 2x overhead target, UMC achieves 22\% monitoring coverage on average and with a 4x overhead target, this increases to 49\%.
+%We see that
+%with a 1.1x overhead target, UMC achieves 25\% monitoring coverage on average.
+%By increasing this overhead budget to 1.5x, UMC achieves 43\% coverage on
+%average. 
+Even higher coverage can be achieved by allowing higher overheads.
 Similarly, Figure~\ref{fig:evaluation.ls_sweep} shows the coverage achieved by LS.
-With a 1.1x overhead, LS achieves 53\% coverage on average. This increases to
-71\% coverage with a 1.2x overhead. 
+With only a 1.01x overhead, LS achieves 30\% coverage on average. This increases to
+71\% coverage with a 1.1x overhead. 
 For DIFT (not pictured), an overhead target of 1.1x is enough for
@@
 with a 1.1x overhead, a max error of 9\% is observed across the benchmarks and the
-average error across instruction types and benchmarks is less than 2\%.
+average error across instruction types and benchmarks is 3\%.
 
@@
 monitoring range from 1.1-1.7x. Figure~\ref{fig:evaluation.fpga_umc_sweep} shows
-the coverage achieved as we sweep the overhead target from 1.01-1.7x. We see that partial
+the coverage achieved as we sweep the overhead target from 1.01-2.0x. We see that partial
 monitoring can allow the overheads of such a system to be pushed below 10\%
-while still providing over 50\% coverage.
+while still providing 45\% coverage.
 
 \subsection{Comparing Dropping Policies}
-
-% % UMC exec time
-% \begin{figure}
-%   \begin{center}
-%     \includegraphics[width=\columnwidth]{figs/data_umc_exec_time.pdf}
-%     \vspace{-0.2in}
-%     \caption{Error in meeting overhead budget for uninitialized memory check for different dropping policies.}
-%     \label{fig:evaluation.umc_exec_time}
-%     \vspace{-0.2in}
-%   \end{center}
-% \end{figure}
-% 
-% % UMC coverage across policies
-% \begin{figure}
-%   \begin{center}
-%     \includegraphics[width=\columnwidth]{figs/data_umc_coverage.pdf}
-%     \vspace{-0.2in}
-%     \caption{Coverage for uninitialized memory check for different dropping policies.}
-%     \label{fig:evaluation.umc_coverage}
-%     \vspace{-0.1in}
-%   \end{center}
-% \end{figure}
 
@@
 In this section we evaluate the trade-offs between unrestricted dropping and source-only dropping.
-The results are shown for an overhead target of 1.5x.
+The results are shown for an overhead target of 2.0x for UMC and 1.5x for BC.
 Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
@@
 means that the overhead target was overshot while a negative value indicates
-that the overhead budget was met. We see similar results for unrestricted
+that the overhead budget was met. For most benchmarks, we see similar results for unrestricted
 dropping and source-only dropping due to the fact that UMC consists of a large
-number of short, independent monitoring dependence chains.
+number of short, independent monitoring dependence chains. Source-only dropping
+causes overshoot of the overhead target for {\tt hmmer} and {\tt h264ref}.
 Figure~\ref{fig:evaluation.umc_coverage} shows the coverage of UMC for
 unrestricted dropping and source-only dropping. We see that, in several cases, source-only dropping
-achieves higher coverage than unrestricted dropping. On average, unrestricted
-dropping achieves 46\% coverage while source-only dropping achieves 49\%
-coverage. This is due
+achieves higher coverage than unrestricted dropping while still closely matching the overhead target.
+For example, {\tt perlbench} shows a 10\% increase in coverage and {\tt libquantum} shows an 18\% increase in coverage by using source-only dropping.
+This is due
 to the fact that some of the overheads of monitoring for unrestricted dropping
 are being spent on wasted work as discussed in Section~\ref{sec:policies.which}.
-
-% % BC exec time
-% \begin{figure}
-%   \begin{center}
-%     \includegraphics[width=\columnwidth]{figs/data_bc_exec_time.pdf}
-%     \vspace{-0.2in}
-%     \caption{Error in meeting overhead budget for array bounds check for different dropping policies.}
-%     \label{fig:evaluation.bc_exec_time}
-%     \vspace{-0.2in}
-%   \end{center}
-% \end{figure}
-% 
-% % BC coverage across policies
-% \begin{figure}
-%   \begin{center}
-%     \includegraphics[width=\columnwidth]{figs/data_bc_coverage.pdf}
-%     \vspace{-0.2in}
-%     \caption{Coverage for array bounds check for different dropping policies.}
-%     \label{fig:evaluation.bc_coverage}
-%     \vspace{-0.2in}
-%   \end{center}
-% \end{figure}
 
@@
 benchmarks, source-only dropping fails to meet the specified overhead target.
-The overshoot of the overhead target is quite high with overhead differences
-of 420\% for {\tt bzip2}, 120\% for {\tt gcc}, and 1500\% for {\tt hmmer}.
+The overshoot of the overhead target is quite high with overhead differences over 100\% for {\tt bzip2}, {\tt hmmer}, and {\tt h264ref}.
 Since only infrequently occurring array allocations are considered as source events for BC, it can be 
@@
 
-From these results, we see that for monitors with many short, independent
-chains, such as UMC, source-only dropping can provide better coverage than
-unrestricted dropping with the same overhead. For other monitors, such as BC,
-source-only dropping is prone to overshoot the overhead target.
+From these results, we see that depending on the monitor and the program
+behavior, source-only dropping can provide better coverage than unrestricted
+dropping with the same overhead. Overall, unrestricted dropping is better at
+meeting an overhead target.
 Thus, for applications where staying within an overhead target is especially
@@
   \begin{center}
-    \includegraphics[width=\columnwidth, clip=true, trim=0 0.3in 0 0.1in]{figs/data_multirun_coverage.pdf}
-    \vspace{-0.1in}
+    \includegraphics[width=\columnwidth]{figs/data_multirun_coverage.pdf}
+    \vspace{-0.3in}
     \caption{Coverage over multiple runs for BC with a 10\% probability to not drop events.}
     \label{fig:evaluation.multirun}
-    \vspace{-0.1in}
+    \vspace{-0.3in}
   \end{center}
@@
 least one of the runs. Each run
-was simulated for 200 million instructions. 
+was simulated for 500 million instructions. 
 We see that for the unrestricted dropping policy there is almost no increase in
@@
 dropping is much better suited for achieving high coverage over multiple runs.
-Source-only dropping shows a 5\% increase in coverage with only seven runs.
-Over many runs, source-only dropping should be able to
+Source-only dropping shows a 6\% increase in coverage with only eight runs.
+With more runs, source-only dropping should be able to
 reach 100\% code coverage.

### 2015-01-03 introduction.tex
Various edits throughout paper.
Updated data.

--- 2014-09-13 introduction.tex
@@
 In addition to enabling partial monitoring, we show how a simple
-extension to our dataflow engine enables this null metadata filtering.
+extension to our dataflow engine can enable null metadata filtering.
 For example, for an array bounds check, checking non-pointer accesses is
@@
 % overhead budget, this coverage rate can be increased. 
-These monitors show slowdowns of 1.1-4.3x with the null metadata filtering when
+These monitors show average slowdowns of 1.1-5.8x with the null metadata filtering when
 implemented on a multi-core platform.
 Our results show that the partial monitoring can still achieve significant
-coverage with reduced overheads. For example, for monitors which showed 3-4x
-average overheads, slowdown could be reduced to 1.5x while still achieving a
-coverage of 43-82\% on average.
+coverage with reduced overheads. For example, for the monitors which showed over 2x
+average overheads, slowdown could be cut to 1.5x while still achieving a
+coverage of 14-85\% on average.
 

### 2015-01-03 title.tex
Update to camera-ready formatting.
Updated data.

--- 2014-09-13 title.tex
@@
 \ifanonymized{\author{Anonymized}}
-{\author{Daniel Lo, Tao Chen, Mohamed Ismail, and G. Edward Suh}}
+{\author{\IEEEauthorblockN{Daniel Lo, Tao Chen, Mohamed Ismail, and G. Edward Suh}
+\IEEEauthorblockA{Cornell University\\
+Ithaca, NY 14850, USA\\
+\{dl575, tc466, mii5, gs272\}@cornell.edu}}
+}
 
-\date{}
 \maketitle
@@
 Our experimental results show that overheads can be reduced significantly 
-by trading off coverage. For example, with average overheads of 3-4x for full
-monitoring range, the proposed architecture can still
-achieve 43-82\% coverage when the slowdown is reduced to 1.5x.
+by trading off coverage. For example, for monitoring techniques with average
+overheads of 2-6x, the proposed architecture is able to reduce slowdowns to
+1.5x while still achieving 14-85\% average coverage.
 

### 2015-01-03 data/fpga_umc_sweep.py

@@ -22,6 +22,9 @@ def parse():
   data = {}
   f = open("fpga_umc_sweep.csv", 'r')
   for line in f:
+    # Treat line as comment, skip
+    if line[0] == '#':
+      continue
     ls = line.strip().split(',')
     # Find monitor names
     if len(ls) == 1:
@@ -41,6 +44,7 @@ def parse():
           data[monitor]['overheads'].append("%.2fx" % (overhead + 1))
           d = [float(x)*100 for x in ls[1:]]
           d.append(geomean(d))
+          print geomean(d)
           data[monitor]['data'].append(d)
   f.close()
 

### 2015-01-03 data/multi_run.py

@@ -1,0 +1,61 @@
+#!/usr/bin/python
+
+import sys
+sys.path.extend(['../..'])
+import tsg_plot
+import math
+
+# Parse data
+f = open("multi_run.csv", 'r')
+#subcat = f.readline().strip().split(',')[1:]
+cat = []
+data = []
+for line in f:
+  if line.strip():
+    ls = line.strip().split(',')
+    cat.append(ls[0])
+    data.append([float(x)*100 for x in ls[1:]])
+f.close()
+
+subcat = ["%d" % (x + 1) for x in range(len(data[0]))]
+subcat[-1] += " runs"
+
+# Set up plotting options
+opts = tsg_plot.PlotOptions()
+opts.data = data
+opts.labels = [cat, subcat]
+
+attribute_dict = \
+    {
+        'show' : False,
+        'file_name' : '../figs/data_multirun_coverage.pdf',
+        'paper_mode' : True,
+        'figsize' : (3.5, 2.0),
+        'ylabel' : 'Coverage [%]',
+        'legend_ncol' : 8,
+        'legend_handlelength' : 1.0,
+        'legend_columnspacing' : 0.8,
+        'legend_handletextpad' : 0.5,
+        'legend_bbox' : [-0.05, 1.05, 1, 0.1],
+        'rotate_labels' : False,
+        'rotate_labels_angle' : -30,
+        'fontsize' : 8,
+        'yrange' : [80, 100],
+
+        #'colors' : ["#8dd3c7", "#ffffb3", "#bebada", "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5", "#d9d9d9", "#bc80bd"]
+
+        # 'colors' : ["#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba", "#d7191c", "#fdae61", "#ffffbf", "#abdda4", "#2b83ba"],
+        # 'hatch' : [' ']*5 + ['///']*5
+
+        #'colors' : ['#8dd3c7', '#ffffb3', '#bebada', '#8dd3c7', '#ffffb3', '#bebada', '#8dd3c7', '#ffffb3', '#bebada', '#8dd3c6'],
+        #'hatch' : ['']*3 + ['/']*3 + ['xx']*3 + ['//']
+
+        'colors' : ['#eff3ff', '#bdd7e7', '#6baed6', '#2171b5', '#eff3ff', '#bdd7e7', '#6baed6', '#2171b5'],
[…]

### 2015-01-03 data/policy_coverage.py

@@ -23,6 +23,9 @@ def parse():
   f = open("policy_coverage.csv", 'r')
   for line in f:
     ls = line.strip().split(',')
+    # Ignore empty lines and comments (starts with #)
+    if not line.strip() or line[0] == '#':
+      continue
     # Find monitor names
     if len(ls) == 1:
       if ls[0] != '':
@@ -42,6 +45,9 @@ def parse():
         d.append(geomean(d))
         data[monitor]['data'].append(d)
   f.close()
+  for mon in data.keys():
+    print mon
+    print data[mon]
 
   # Reformat data
   for monitor in data.keys():

### 2015-01-03 data/policy_overhead.py

@@ -15,6 +15,8 @@ def geomean(l):
     prod *= ll
     n += 1
   return prod**(1.0/n)
+def average(l):
+  return sum([float(x) for x in l])/len(l)
 
 
 def parse():
@@ -22,6 +24,9 @@ def parse():
   data = {}
   f = open("policy_overhead.csv", 'r')
   for line in f:
+    # Skip empty lines and comments
+    if not line.strip() or line[0] == '#':
+      continue
     ls = line.strip().split(',')
     # Find monitor names
     if len(ls) == 1:
@@ -33,16 +38,16 @@ def parse():
     else:
       # Benchmarks
       if ls[0] == '':
-        #data[monitor]['benchmarks'] = ls[1:] + ["geomean"]
-        data[monitor]['benchmarks'] = ls[1:]
+        data[monitor]['benchmarks'] = ls[1:] + ["average"]
       # Data line
       else:
         policy = ls[0]
         data[monitor]['policies'].append(policy)
         d = [float(x)*100 for x in ls[1:]]
-        #d.append(geomean(d))
+        d.append(average(d))
         data[monitor]['data'].append(d)
   f.close()
+  print data
 
   # Reformat data
   for monitor in data.keys():
@@ -97,4 +102,5 @@ if __name__ == "__main__":
   data = parse()
   for mon in data.keys():
     if mon != "insttype":
-      plot(data[mon], "../figs/data_%s_overhead.pdf" % (mon))
+      #plot(data[mon], "../figs/data_%s_overhead.pdf" % (mon))
+      plot(data[mon], "temp.pdf" % (mon))

### 2015-01-03 data/sweeps.py

@@ -22,6 +22,9 @@ def parse():
   data = {}
   f = open("sweeps.csv", 'r')
   for line in f:
+    # Skip empty lines and comments
+    if not line.strip() or line[0] == '#':
+      continue
     ls = line.strip().split(',')
     # Find monitor names
     if len(ls) == 1:
@@ -46,6 +49,9 @@ def parse():
           d.append(geomean(d))
           data[monitor]['data'].append(d)
   f.close()
+  for mon in data.keys():
+    print mon
+    print data[mon]
 
   # Reformat data
   for monitor in data.keys():

### 2015-01-04 dropping.tex
Edits based on Ed's comments.

@@
 way that metadata is. Figure~\ref{fig:dropping.example_invalid}
-shows an example of associating invalid flags with metadata. Suppose that the monitoring for
+shows an example of associating invalidation flags with metadata. Suppose that the monitoring for
 line 1 is dropped in order to meet the overhead target. When a monitoring event is dropped, metadata that would have
@@
 since {\tt rf\_invalid[1]} is marked as true. Note that in this case, line 2
-also propagates this invalid flag to {\tt rf\_invalid[2]} and causes the check
+also propagates this invalidation flag to {\tt rf\_invalid[2]} and causes the check
 performed on line 4 to also be dropped. This is necessary because an error
-would have been raised even if the access on line 4 was in bounds.
+would have been raised even if the access on line 4 was within bounds.
 
@@
 \begin{enumerate}
-  \item Propagate invalid flags, following the dataflow of metadata.
-  \item Filter out monitoring operations based on invalid flags.
+  \item Propagate invalidation flags, following the dataflow of metadata.
+  \item Filter out monitoring operations based on invalidation flags.
 \end{enumerate}
@@
 Thus, the hardware acts effectively as a dataflow tracking engine in order to
-track a 1-bit invalid flag per metadata. Figure~\ref{fig:dropping.dataflow} shows a detailed block
+track a 1-bit invalidation flag per metadata. Figure~\ref{fig:dropping.dataflow} shows a detailed block
 diagram of this hardware module. The dataflow engine uses two address
-calculation units to read in up two invalid flags. These source invalid flags
+calculation units to read in up to two invalidation flags. These source invalidation flags
 are used to decide whether a monitoring event should filtered. A third address
 calculation unit is used to optionally specify a target to propagate the
-invalidation information. The module includes a register file to store invalid
+invalidation information. The module includes a register file to store invalidation
 flags corresponding to register file metadata. In addition, it uses a small
-memory-backed cache to handle invalid flags corresponding to memory metadata.
+memory-backed cache to handle invalidation flags corresponding to memory metadata.
 
@@
 monitor is to filter out monitoring events that correspond to operating on null
-metadata. Null metadata corresponds to events that
+metadata. Null metadata correspond to events that
 are not relevant to the monitor. For example, Hardbound
@@
 operations needed for invalidation shown in
-Figure~\ref{fig:dropping.example_invalid}. Instead of taking a logical OR of the
-source invalid flags to determine whether monitoring can be skipped, null
+Figure~\ref{fig:dropping.example_invalid} except for a small change in the propagation policy. Instead of taking a logical OR of the
+source invalidation flags to determine whether monitoring can be skipped, null
 metadata filtering takes a logical AND of the source null flags.

### 2015-01-04 evaluation.tex
Update full monitoring overheads for FPGA and add 14x for BC.
Edits based on Ed's comments.

@@
 cache. This setup is similar to the Snapdragon 801 processor commonly found in
-mobile systems. The dataflow engine uses a 1 kB cache for invalid and null flags.
+mobile systems. The dataflow engine uses a 1 kB cache for invalidation and null flags.
 
@@
 Figure~\ref{fig:evaluation.bc_sweep} shows the monitoring coverage achieved by
-array bounds check as we vary the overhead budget. 
+array bounds checking as we vary the overhead budget. 
 We define \emph{monitoring coverage} as the 
@@
 monitors show similar trends. In this case, the full overheads of
-monitoring range from 1.1-1.7x. Figure~\ref{fig:evaluation.fpga_umc_sweep} shows
+monitoring range from 1.1-1.9x. Figure~\ref{fig:evaluation.fpga_umc_sweep} shows
 the coverage achieved as we sweep the overhead target from 1.01-2.0x. We see that partial
@@
 behavior, source-only dropping can provide better coverage than unrestricted
-dropping with the same overhead. Overall, unrestricted dropping is better at
+dropping with the same overhead. Unrestricted dropping is better at
 meeting an overhead target.
@@
 Adding the dataflow engine in order to enable filtering and partial monitoring adds
-overheads in terms of area and power. We use McPat \cite{mcpat-micro09} to get
+overheads in terms of area and power consumption. We use McPat \cite{mcpat-micro09} to get
 a first-order estimate of these area and power overheads in a 40 nm technology

### 2015-01-04 introduction.tex
Edits based on Ed's comments.

@@
 example, Hardbound is a hardware-assisted technique to detect out-of-bound
-memory accesses, which can cause undesired behavior or create a security
-vulnerability if uncaught \cite{hardbound-asplos08}.  Intel has
+memory accesses \cite{hardbound-asplos08}.  Intel has
 recently announced plans to support buffer overflow protection similar to
@@
 In order to prevent these
-false positive, we need to track the dropped metadata flows.  We present a
+false positives, we need to track the dropped metadata flows.  We present a
 hardware-based dataflow tracking engine that keeps track of these dropped
@@
 which events to drop for partial monitoring. These different policies show a
-trade-off between closely matching the overhead budget and the monitoring
-coverage achieved.
+trade-off between closely matching the overhead budget and increasing the monitoring
+coverage.
 In addition to enabling partial monitoring, we show how a simple
@@
 coverage with reduced overheads. For example, for the monitors which showed over 2x
-average overheads, slowdown could be cut to 1.5x while still achieving a
+average overheads, slowdown could be cut down to 1.5x while still achieving a
 coverage of 14-85\% on average.

### 2015-01-04 policies.tex
Various minor edits. Figure clean-up.
Edits based on Ed's comments.

@@
 In this slack-based policy, if the slack falls below zero (i.e., the overhead budget is exceeded), then
-monitoring events should be dropped.
+monitoring events are dropped.
 
@@
   \begin{center}
-  \subfloat[unrestricted dropping]{
-    \includegraphics[width=\columnwidth, clip=true, trim=0 0.1in 0 0]{figs/all_drop.pdf}
+  \subfloat[Unrestricted Dropping]{
+    \includegraphics[width=\columnwidth]{figs/unrestricted_drop.pdf}
     \label{fig:policies.all_drop}
@@
   \\
-  \subfloat[source-only dropping]{
-    \includegraphics[width=\columnwidth, clip=true, trim=0 0.1in 0 0]{figs/source_drop.pdf}
+  \subfloat[Source-only Dropping]{
+    \includegraphics[width=\columnwidth]{figs/source_drop.pdf}
     \label{fig:policies.source_drop}
@@
 That is, by the time we decide to drop event {\tt E}, we have already 
-updated metadata due to events {\tt C} and {\tt D} even though they are no longer needed.
+updated metadata for events {\tt C} and {\tt D} even though they are no longer needed.
 
@@
 dropping will
-result in no wasted work and thus better coverage. However, because of the coarser-grained decision, it
-may be more difficult to closely match overheads. We expect source-only dropping 
+result in no wasted work and thus better coverage at a given overhead. However, because of the coarser-grained decision, it
+may be more difficult to closely match overhead targets. We expect source-only dropping 
 to work well when there are a large number of independent metadata flows.
 In addition, using a probabilistic dropping policy works poorly with
-unrestricted dropping. Since each event in a dependent chain (e.g., events {\tt
+unrestricted dropping. Since every event in a dependent chain (e.g., events {\tt
 A} through {\tt E}) needs to be monitored in order for the monitoring check to
-be useful, randomly dropping any event will cause the final check to be
-invalid.
+be useful, randomly dropping any event is likely to cause the final check to be
+invalid by dropping at least one event in the chain.
 

### 2015-01-04 related.tex
Edits based on Ed's comments.

@@
 monitoring that is general, allows an overhead target to be specified, and
-explicitly prevents false positives. In contrast, almost all previous work on
+explicitly prevents false positives. In contrast, previous work on
 partial monitoring achieves only some, but not all, of these properties.
@@
 \cite{qvm-oopsla08} proposes a modification to the Java Virtual Machine (JVM) to
-support monitoring with adjustable overheads. QVM does prevents false positives
-but does so be enabling or disabling monitoring on a per-object basis. This is
+support monitoring with adjustable overheads. QVM does prevent false positives
+but does so by enabling or disabling monitoring on a per-object basis. This is
 similar to the idea of performing source-only dropping which our results show

### 2015-01-04 title.tex
Edits based on Ed's comments.

@@
-\title{
-%Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring
-%Efficient and Adjustable Run-Time Monitoring Using Dataflow-Guided Filtering
-Run-Time Monitoring with Adjustable Overheads \\Using Dataflow-Guided Filtering
-}
+\title{Run-Time Monitoring with Adjustable Overhead \\Using Dataflow-Guided Filtering}
 
@@
 }
+% Reduce space between title and authors
+\IEEEspecialpapernotice{\vspace{-0.4in}}
 
@@
 In this paper, we introduce a hardware dataflow tracking engine that enables
-adjustable overheads through partial monitoring. This allows a trade-off to be
+adjustable overhead through partial monitoring. This allows a trade-off to be
 made between monitoring coverage and overhead. This dataflow engine
 can also be extended to filter out monitoring operations associated with null
-metadata in order to reduce overheads.
+metadata in order to reduce overhead.
 % In this paper, we introduce a hardware dataflow tracking engine that can be
@@
 decisions depending on the target application of partial monitoring.
-Our experimental results show that overheads can be reduced significantly 
+Our experimental results show that overhead can be reduced significantly 
 by trading off coverage. For example, for monitoring techniques with average
-overheads of 2-6x, the proposed architecture is able to reduce slowdowns to
+overheads of 2-6x, the proposed architecture is able to reduce overhead to
 1.5x while still achieving 14-85\% average coverage.

### 2015-01-04 evaluation.tex
Update full overhead of LS.

@@
 applied to the SPLASH-2 benchmarks are
-1.03-1.72x after null metadata filtering and the average
-overhead observed is 1.33x.
+1.03-1.29x after null metadata filtering and the average
+overhead observed is 1.13x.
 Our baseline system and overheads are similar to previous multi-core

### 2015-01-05 evaluation.tex
Updated full overhead of LS.

@@
 applied to the SPLASH-2 benchmarks are
-1.03-1.29x after null metadata filtering and the average
+1.04-1.29x after null metadata filtering and the average
 overhead observed is 1.13x.

### 2015-01-05 evaluation.tex
Add discussion of overhead target vs. overhead seen.

@@
 %1.1x is enough for all benchmarks tested to reach 100\% coverage.
+Note that the overheads shown are the target overhead. In some cases, a high
+overhead target is needed in order to achieve 100\% coverage. For example, for
+{\tt mcf} with UMC monitoring, a 12x overhead target is needed to achieve 100\%
+coverage while the overheads of performing monitoring in full were 2.6x (see
+Figure~\ref{fig:evaluation.filtering}. This is due to the fact that we
+accumulate slack gradually and so bursty monitoring events may require a higher
+overhead target in order for all monitoring to be performed. However, the
+actual overheads seen are in-line with the overhead of performing monitoring
+without dropping (e.g., 2.6x for {\tt mcf} with UMC monitoring at a 12x
+overhead target).
 

### 2015-01-05 policies.tex
Discussion on sub-flow dropping.

@@
-\section{Dropping Policies}
-\label{sec:policies}
+\section{Dropping Policies} \label{sec:policies}
 
@@
 section, we discuss some of the options and trade-offs for dropping policies.
-We split this decision into two components:
-\begin{enumerate}
-  \item When do we need to drop events in order to enable reduced overheads? (Section~\ref{sec:policies.when})
-  \item Which events should be dropped? (Section~\ref{sec:policies.which})
-\end{enumerate}
+We split this decision into two components: \begin{enumerate} \item When do we
+need to drop events in order to enable reduced overheads?
+(Section~\ref{sec:policies.when}) \item Which events should be dropped?
+(Section~\ref{sec:policies.which}) \end{enumerate}
 
-% Our goal with enabling partial monitoring is to allow for a specified overhead target or budget to be met while maximizing the amount of monitoring performed.
-% This is done by dynamically dropping a portion of the monitoring operations. In this section, we discuss in detail how this dropping decision is
-% made. Specifically, we first discuss how to decide when dropping should be performed 
-% Section~\ref{sec:policies.slack}. Next, we investigate trade-offs created by
-% selecting which events are dropped (Section~\ref{sec:policies.events}). Different policies in determining which
+% Our goal with enabling partial monitoring is to allow for a specified
+% overhead target or budget to be met while maximizing the amount of monitoring
+% performed.  This is done by dynamically dropping a portion of the monitoring
+% operations. In this section, we discuss in detail how this dropping decision
+% is made. Specifically, we first discuss how to decide when dropping should be
+% performed Section~\ref{sec:policies.slack}. Next, we investigate trade-offs
+% created by selecting which events are dropped
+% (Section~\ref{sec:policies.events}). Different policies in determining which
 % events can be dropped create a trade-off between how closely the overhead
@@
 
-\subsection{Deciding When to Drop}
-\label{sec:policies.when}
+\subsection{Deciding When to Drop} \label{sec:policies.when}
 
 % Slack
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\columnwidth]{figs/slack.pdf}
-    \vspace{-0.2in}
+\begin{figure} \begin{center}
+\includegraphics[width=\columnwidth]{figs/slack.pdf} \vspace{-0.2in}
 \caption{Slack and its effect on monitoring over time.}
-    \label{fig:policies.slack}
-    \vspace{-0.1in}
-  \end{center}
-\end{figure}
+\label{fig:policies.slack} \vspace{-0.1in} \end{center} \end{figure}
 
-In this section, we discuss two possible ways to determine when events should be dropped.
-The first possibility is to probabilistically drop events. By setting the probability of
-dropping events appropriately, overheads can be reduced. This works well for
-enabling partial monitoring for cooperative testing and debugging since the
-randomness allows different users and runs to monitor different portions of the
-program. However, using a probabilistic dropping policy can make it difficult
-to meet a target overhead without prior profiling.
+In this section, we discuss two possible ways to determine when events should
+be dropped.  The first possibility is to probabilistically drop events. By
+setting the probability of dropping events appropriately, overheads can be
+reduced. This works well for enabling partial monitoring for cooperative
+testing and debugging since the randomness allows different users and runs to
+monitor different portions of the program. However, using a probabilistic
+dropping policy can make it difficult to meet a target overhead without prior
+profiling.
 
 Alternatively, we can specify a target overhead and estimate, at run-time, the
-overheads of monitoring in order to decide whether dropping is needed.
-The overhead budget is specified as a percentage of the main program's execution
-cycles without monitoring. In order to estimate overheads at run-time, we define \emph{slack} as the number of cycles of
-monitoring overhead that can be incurred while staying within the budget
-target. Slack is essentially the difference between the actual overhead seen
-and the budget specified. Slack is generated as the main program runs and consumed as monitoring overheads occur.  For
-example, if no monitoring overheads occur during 1000 cycles of the main
+overheads of monitoring in order to decide whether dropping is needed.  The
+overhead budget is specified as a percentage of the main program's execution
+cycles without monitoring. In order to estimate overheads at run-time, we
+define \emph{slack} as the number of cycles of monitoring overhead that can be
+incurred while staying within the budget target. Slack is essentially the
+difference between the actual overhead seen and the budget specified. Slack is
+generated as the main program runs and consumed as monitoring overheads occur.
+For example, if no monitoring overheads occur during 1000 cycles of the main
 program's execution and the designer sets a 20\% overhead target, then the
@@
 then stalled for 50 cycles due to monitoring, then the remaining slack is 150
-cycles. 
-In addition to this accumulated slack, a small amount of initial slack
+cycles.  In addition to this accumulated slack, a small amount of initial slack
 can be given in order for monitoring to be performed at the start of a program.
-Figure~\ref{fig:policies.slack} shows an example of how slack can change over time.
-In this slack-based policy, if the slack falls below zero (i.e., the overhead budget is exceeded), then
-monitoring events are dropped.
+Figure~\ref{fig:policies.slack} shows an example of how slack can change over
+time.  In this slack-based policy, if the slack falls below zero (i.e., the
+overhead budget is exceeded), then monitoring events are dropped.
 
-% % Slack tracking module
-% \begin{figure}
-%   \begin{center}
-%     \includegraphics[width=\columnwidth]{figs/stm.pdf}
-%     \vspace{-0.2in}
-%     \caption{Slack tracking and drop decision hardware.}
-%     \label{fig:policies.stm}
-%     \vspace{-0.1in}
-%   \end{center}
-% \end{figure}
+% % Slack tracking module \begin{figure} \begin{center}
+% \includegraphics[width=\columnwidth]{figs/stm.pdf} \vspace{-0.2in}
+% \caption{Slack tracking and drop decision hardware.} \label{fig:policies.stm}
+% \vspace{-0.1in} \end{center} \end{figure}
 
@@
 % keeping track of slack. 
-Slack can be easily measured in hardware by using a counter that increments on every $k$-th
-cycle of the main core (e.g., every 5th cycle for a 20\% target budget). 
-% This $k$ can be calculated by taking the reciprocal of
-% the target budget. For example, if the target budget is 20\%, then the counter
-% increments on every 5th clock cycle. 
-The value of this counter is the
-accumulated slack. Whenever the main core is stalled due to the monitor, the
-measured slack is decremented. It is difficult to precisely determine the
-entire impact of monitoring on the main core due to the
-difficulty in measuring certain overheads such as contention for shared memory.
-However, we have found that using only the stalls due to FIFO back pressure
-works well in practice.
+Slack can be easily measured in hardware by using a counter that increments on
+every $k$-th cycle of the main core (e.g., every 5th cycle for a 20\% target
+budget). 
+% This $k$ can be calculated by taking the reciprocal of the target budget. For
+% example, if the target budget is 20\%, then the counter increments on every
+% 5th clock cycle. 
+The value of this counter is the accumulated slack. Whenever the main core is
+stalled due to the monitor, the measured slack is decremented. It is difficult
+to precisely determine the entire impact of monitoring on the main core due to
+the difficulty in measuring certain overheads such as contention for shared
+memory.  However, we have found that using only the stalls due to FIFO back
+pressure works well in practice.
 
-\subsection{Deciding Which Events to Drop} 
-\label{sec:policies.which}
+\subsection{Deciding Which Events to Drop} \label{sec:policies.which}
 
-% % Varying slack's impact on coverage for UMC
-% \begin{figure}
-%   \begin{center}
-%     \includegraphics[width=\columnwidth]{figs/dataflow_graph.pdf}
-%     \vspace{-0.3in}
-%     \caption{Example dependence graph for metadata. Square nodes represent
-%     events where checks are performed.} 
-%     \label{fig:policies.dataflow_graph}
-%     \vspace{-0.1in}
-%   \end{center}
-% \end{figure}
+% % Varying slack's impact on coverage for UMC \begin{figure} \begin{center}
+% \includegraphics[width=\columnwidth]{figs/dataflow_graph.pdf} \vspace{-0.3in}
+% \caption{Example dependence graph for metadata. Square nodes represent events
+% where checks are performed.} \label{fig:policies.dataflow_graph}
+% \vspace{-0.1in} \end{center} \end{figure}
 
-\begin{figure}
-  \begin{center}
-  \subfloat[Unrestricted Dropping]{
+\begin{figure} \begin{center} \subfloat[Unrestricted Dropping]{
 \includegraphics[width=\columnwidth]{figs/unrestricted_drop.pdf}
-    \label{fig:policies.all_drop}
-  }
-  \vspace{-0.1in}
-  \\
-  \subfloat[Source-only Dropping]{
-    \includegraphics[width=\columnwidth]{figs/source_drop.pdf}
-    \label{fig:policies.source_drop}
-  }
-  \vspace{-0.1in}
-%   \\
-%   \subfloat[sub-flow dropping]{
-%     \includegraphics[width=\columnwidth]{figs/min_drop.pdf}
-%     \label{fig:policies.min_drop}
-%   }
-  \end{center}
-  \caption{Comparison of dropping policies using metadata dependence graphs.
-  Square nodes represent events where check are performed. Blue (dark) nodes
-  indicate which nodes can be dropped.}
-  \label{fig:policies.policies}
-\end{figure}
+\label{fig:policies.all_drop} } \vspace{-0.1in} \\ \subfloat[Source-Only
+Dropping]{ \includegraphics[width=\columnwidth]{figs/source_drop.pdf}
+\label{fig:policies.source_drop} } \vspace{-0.1in} \\ \subfloat[Sub-Flow
+Dropping]{ \includegraphics[width=\columnwidth]{figs/subflow_drop.pdf}
+\label{fig:policies.subflow_drop} } \end{center} \caption{Comparison of
+dropping policies using metadata dependence graphs.  Square nodes represent
+events where check are performed. Blue (dark) nodes indicate which nodes can be
+dropped.} \label{fig:policies.policies} \end{figure}
 
 In addition to deciding when dropping is required, trade-offs also exist in
-deciding which events should be dropped.
-The simplest policy is to drop any monitoring events when slack is less than or equal to zero.  However, this can result in
-wasted work. For example, consider the metadata dependence graph shown in
-Figure~\ref{fig:policies.all_drop}. Here, an edge from node {\tt A} to node
-{\tt B} represents that if event {\tt A} is dropped, then due to its
-invalidated metadata, it will cause event {\tt B} to be dropped. Square
-nodes indicate events where monitoring checks are performed. In the
-example, suppose that event {\tt E} is meant to perform a check operation but is dropped.
-In this situation, the
-monitoring operations that were done for events {\tt C} and {\tt D} were wasted
-since their results were not used for any monitoring checks.
-That is, by the time we decide to drop event {\tt E}, we have already 
-updated metadata for events {\tt C} and {\tt D} even though they are no longer needed.
+deciding which events should be dropped.  The simplest policy is to drop any
+monitoring events when slack is less than or equal to zero.  However, this can
+result in wasted work. For example, consider the metadata dependence graph
+shown in Figure~\ref{fig:policies.all_drop}. Here, an edge from node {\tt A} to
+node {\tt B} represents that if event {\tt A} is dropped, then due to its
+invalidated metadata, it will cause event {\tt B} to be dropped. Square nodes
+indicate events where monitoring checks are performed. In the example, suppose
+that event {\tt E} is meant to perform a check operation but is dropped.  In
+this situation, the monitoring operations that were done for events {\tt C} and
+{\tt D} were wasted since their results were not used for any monitoring
+checks.  That is, by the time we decide to drop event {\tt E}, we have already
+updated metadata for events {\tt C} and {\tt D} even though they are no longer
+needed.
 
-An alternative dropping policy which eliminates this wasted work is to only make dropping decisions at
-the root or source of these metadata flows. That is, we will decide to either monitor or
-not monitor an entire metadata flow. An example of these source nodes is the
-monitoring done to initialize base and bounds information on {\tt malloc} for
-an array bounds check. We refer to this dropping decision policy
-as \emph{source-only dropping} and we refer to the previous policy of 
-dropping any event as \emph{unrestricted dropping}.
+An alternative dropping policy which eliminates this wasted work is to only
+make dropping decisions at the root or source of these metadata flows. That is,
+we will decide to either monitor or not monitor an entire metadata flow. An
+example of these source nodes is the monitoring done to initialize base and
+bounds information on {\tt malloc} for an array bounds check. These source
+nodes are typically easy to identify by the dropping hardware because they
+correspond to special instructions that are used to set up metadata
+information. Thus, it is not necessary to generate and analyze the monitoring
+dependence graph to identify source nodes. We refer to this dropping decision
+policy as \emph{source-only
+dropping} and we refer to the previous policy of dropping any event as
+\emph{unrestricted dropping}.
 
@@
 
+More complex dropping policies can be implemented given more detailed
+information about the monitoring dependence graph. Here, we describe one such
+policy, which we call \emph{sub-flow dropping}. Sub-flow dropping makes
+dropping decisions at a granularity in between unrestricted dropping and
+source-only dropping (see Figure~\ref{fig:policies.subflow_drop}). Sub-flow
+dropping allows source nodes to be dropped and nodes after branch points in the
+monitoring dependence graph to be dropped. From
+Figure~\ref{fig:policies.subflow_drop}, this corresponds to nodes {\tt A}, {\tt
+C}, and {\tt F}. The basic intuition is that dropping certain portions of the
+monitoring dependence graph still result in no wasted work. For example,
+dropping node {\tt C} causes nodes {\tt D} and {\tt E} to be skipped. However,
+performing monitoring on the remaining nodes allows the check at node {\tt G}
+to be performed with no wasted work. Similarly, dropping {\tt F} allows node
+{\tt E} to be checked with no wasted work.
+
 Depending on the target application of partial monitoring, different policies
@@
 multiple runs.
+Sub-flow dropping enables a design point in between unrestricted dropping and
+source-only dropping. Sub-flow dropping matches overheads better than
+source-only dropping and provides better coverage than unrestricted dropping.
+For a probabilistic policy, sub-flow dropping allows better overhead matching
+while still allowing good total coverage over multiple runs, though at a slower
+rate than source-only dropping.
 

### 2015-01-05 dropping.tex
Change "overheads" to "overhead".

@@
 In this section, we present our hardware architecture that enables partial
-monitoring for adjustable overheads. Section~\ref{sec:dropping.baseline} first
+monitoring for adjustable overhead. Section~\ref{sec:dropping.baseline} first
 describes our baseline model for run-time monitoring while the rest of the
@@
 
-Our goal is to drop some monitoring operations in order to reduce the overheads
+Our goal is to drop some monitoring operations in order to reduce the overhead
 of run-time monitoring. This dropping can affect the functionality of the
@@
 for this is dropped, then the error is not detected. This reflects the trade-off that
-we make in order to reduce overheads. Instead of either 100\% coverage
-with all the associated overheads or no coverage and no overheads, we 
+we make in order to reduce overhead. Instead of either 100\% coverage
+with all the associated overhead or no coverage and no overhead, we 
 enable middle points of partial coverage with some fraction of the full
-overheads.
+overhead.
 
@@
 incorrectly raised. Although false negatives are part of the trade-off we make
-to reduce overheads, we need to prevent false positives.
+to reduce overhead, we need to prevent false positives.
 

### 2015-01-05 evaluation.tex
Change "overheads" to "overhead".

@@
 
-\subsection{Baseline Monitoring Overheads}
+\subsection{Baseline Monitoring Overhead}
 
@@
     \vspace{-0.2in}
-    \caption{Monitoring overheads with null metadata filtering.}
+    \caption{Monitoring overhead with null metadata filtering.}
     \label{fig:evaluation.filtering}
@@
 targeted network or streaming applications, which have larger amounts of
-untrusted input data, we would observe higher overheads. IMP shows normalized
+untrusted input data, we would observe higher overhead. IMP shows normalized
 execution times of 1.6-9.5x with an average of 5.8x. Overheads for LS (not pictured)
@@
 In this section, we evaluate the effectiveness of
-using partial monitoring to trade-off coverage for reduced overheads.
+using partial monitoring to trade-off coverage for reduced overhead.
 For these results, we use a slack-based, unrestricted dropping policy.
@@
 achieves over 80\% monitoring coverage on average. The high coverage
-achieved with such low overheads is due to two main effects.  The first is that
+achieved with such low overhead is due to two main effects.  The first is that
 monitoring can be done in parallel, providing a certain level of monitoring coverage without
-introducing overheads. The second effect is that there may still exist a large number of
+introducing overhead. The second effect is that there may still exist a large number of
 monitoring events that do not lead to security or reliability checks. As a
-result, dropping these events can reduce overheads without a large impact on
+result, dropping these events can reduce overhead without a large impact on
 monitoring coverage.
@@
 %average. 
-Even higher coverage can be achieved by allowing higher overheads.
+Even higher coverage can be achieved by allowing higher overhead.
 Similarly, Figure~\ref{fig:evaluation.ls_sweep} shows the coverage achieved by LS.
@@
 shows the min, max, and average error across
-instruction types for each benchmark and for varying overheads. We see that
+instruction types for each benchmark and for varying overhead. We see that
 with a 1.1x overhead, a max error of 9\% is observed across the benchmarks and the
@@
 the coverage achieved as we sweep the overhead target from 1.01-2.0x. We see that partial
-monitoring can allow the overheads of such a system to be pushed below 10\%
+monitoring can allow the overhead of such a system to be pushed below 10\%
 while still providing 45\% coverage.
@@
 Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
-overhead budget and the run-time monitoring overheads for UMC. A positive value
+overhead budget and the run-time monitoring overhead for UMC. A positive value
 means that the overhead target was overshot while a negative value indicates
@@
 This is due
-to the fact that some of the overheads of monitoring for unrestricted dropping
-are being spent on wasted work as discussed in Section~\ref{sec:policies.which}.
+to the fact that some of the overhead of monitoring for unrestricted dropping
+is being spent on wasted work as discussed in Section~\ref{sec:policies.which}.
 
@@
 Figure~\ref{fig:evaluation.bc_coverage} shows higher coverage for
-source-only dropping, this is largely due to the fact that is running with
-higher overheads.
+source-only dropping, this is largely due to the fact that it is running with
+higher overhead.
 
@@
 
-One application of partial monitoring with low overheads is to enable
+One application of partial monitoring with low overhead is to enable
 cooperative debugging. 
 The idea with cooperative debugging is to use partial
-monitoring with very low overheads across a large number of users or runs. By
+monitoring with very low overhead across a large number of users or runs. By
 varying the pattern of partial monitoring done on each run, the goal is to

### 2015-01-05 introduction.tex
Change "overheads" to "overhead".

@@
 propose to enable low overhead monitors without sacrificing flexibility by
-using partial monitoring. In essence, this enables a trade-off between overheads and
+using partial monitoring. In essence, this enables a trade-off between overhead and
 monitoring coverage or accuracy.
@@
 monitoring can enable a level of protection even for systems where the full
-monitoring overheads are too high. This is especially true for energy- or
+monitoring overhead is too high. This is especially true for energy- or
 power-constrained systems or soft real-time systems where the monitoring
@@
 % Adjustable overhead
-Partial monitoring with adjustable overheads is achieved by dynamically
-dropping monitoring operations when overheads exceed a specified overhead
+Partial monitoring with adjustable overhead is achieved by dynamically
+dropping monitoring operations when the overhead exceeds a specified overhead
 budget. Although this results in reduced monitoring coverage (i.e., false
@@
   \item We extend the dataflow engine to enable filtering null metadata to
-  reduce overheads.
+  reduce overhead.
 \end{itemize}
@@
 Our results show that the partial monitoring can still achieve significant
-coverage with reduced overheads. For example, for the monitors which showed over 2x
-average overheads, slowdown could be cut down to 1.5x while still achieving a
+coverage with reduced overhead. For example, for the monitors which showed over 2x
+average overhead, slowdown could be cut down to 1.5x while still achieving a
 coverage of 14-85\% on average.

### 2015-01-05 monitoring.tex
Change "overheads" to "overhead".

@@
 slowdowns of several tens of percents in some cases.
-In today's monitoring systems, the overheads are also unpredictable because they
-depend heavily on the characteristics of applications and monitoring operations.
+In today's monitoring systems, the overhead is also unpredictable because it
+depends heavily on the characteristics of applications and monitoring operations.
 In summary, users currently need to either pay noticeable overhead or the cost of custom

### 2015-01-05 policies.tex
Change "overheads" to "overhead".

@@
 
-In order to use partial monitoring to enable adjustable overheads, we must also
+In order to use partial monitoring to enable adjustable overhead, we must also
 specify a policy for when and which monitoring events are dropped. In this
@@
 We split this decision into two components: \begin{enumerate} \item When do we
-need to drop events in order to enable reduced overheads?
+need to drop events in order to enable reduced overhead?
 (Section~\ref{sec:policies.when}) \item Which events should be dropped?
@@
 be dropped.  The first possibility is to probabilistically drop events. By
-setting the probability of dropping events appropriately, overheads can be
+setting the probability of dropping events appropriately, overhead can be
 reduced. This works well for enabling partial monitoring for cooperative
@@
 Alternatively, we can specify a target overhead and estimate, at run-time, the
-overheads of monitoring in order to decide whether dropping is needed.  The
+overhead of monitoring in order to decide whether dropping is needed.  The
 overhead budget is specified as a percentage of the main program's execution
-cycles without monitoring. In order to estimate overheads at run-time, we
+cycles without monitoring. In order to estimate the overhead at run-time, we
 define \emph{slack} as the number of cycles of monitoring overhead that can be
@@
 Finally, if the goal is to use partial monitoring to enable cooperative
-debugging and testing with very low overheads, then a probabilistic,
+debugging and testing with very low overhead, then a probabilistic,
 source-only dropping policy can be used to provide good total coverage over
@@
 Sub-flow dropping enables a design point in between unrestricted dropping and
-source-only dropping. Sub-flow dropping matches overheads better than
+source-only dropping. Sub-flow dropping matches overhead better than
 source-only dropping and provides better coverage than unrestricted dropping.

### 2015-01-05 related.tex
Change "overheads" to "overhead".

@@
 There exists a number of previous projects that have looked into performing
-partial monitoring in order to reduce the performance overheads. These
+partial monitoring in order to reduce the performance overhead. These
 platforms differ from ours in a number of ways including how monitoring is
@@
   \item \textbf{Adjustable Overhead:} Our architecture allows an overhead to be
-  targeted. Other work performs sampling to reduce overheads but does not try
-  to bound overheads, which we have shown can vary greatly.
+  targeted. Other work performs sampling to reduce overhead but does not try
+  to bound overhead, which we have shown can vary greatly.
   \item \textbf{Prevent False Positives:} We prevent a mechanism to prevent
@@
 For example, there exists previous work for using statistical sampling to
-reduce the performance overheads of various debugging techniques. These include
+reduce the performance overhead of various debugging techniques. These include
 sampling for bug isolation \cite{liblit-pldi05}, memory leak detection
@@
 to be specified and do not prevent false positives. Huang et al.
-\cite{huang-sttt12} proposed a general framework for controlling the overheads
+\cite{huang-sttt12} proposed a general framework for controlling the overhead
 of software-based monitoring. However, they also do not explicitly address
@@
 \cite{qvm-oopsla08} proposes a modification to the Java Virtual Machine (JVM) to
-support monitoring with adjustable overheads. QVM does prevent false positives
+support monitoring with adjustable overhead. QVM does prevent false positives
 but does so by enabling or disabling monitoring on a per-object basis. This is

### 2015-01-05 data/policy_overhead.py
Add subflow results to CSV files.

@@ -102,5 +102,4 @@ if __name__ == "__main__":
   data = parse()
   for mon in data.keys():
     if mon != "insttype":
-      #plot(data[mon], "../figs/data_%s_overhead.pdf" % (mon))
-      plot(data[mon], "temp.pdf" % (mon))
+      plot(data[mon], "../figs/data_%s_overhead.pdf" % (mon))

### 2015-01-05 evaluation.tex
Add discussion and eval on sub-flow dropping.

@@
   \caption{Comparing dropping policies for UMC.}
+  \label{fig:evaluation.umc_policies}
   \vspace{-0.2in}
   \end{center}
-  \label{fig:evaluation.umc_policies}
 \end{figure}
 
-In this section we evaluate the trade-offs between unrestricted dropping and source-only dropping.
+%In this section we evaluate the trade-offs between unrestricted dropping and source-only dropping.
+In this section we evaluate the trade-offs between different dropping policies.
+For sub-flow dropping, we identified sub-flow nodes through profiling. Because
+of the need to profile the applications to determine sub-flow nodes, these
+experiments were simulated for 1 billion instructions without fast-forwarding.
+These results (not shown) indicated that sub-flow dropping produced similar
+results in terms of coverage and overhead target matching to unrestricted
+dropping for the benchmarks and monitors that we tested. Monitors and
+benchmarks with difference monitoring dependence graphs may see a more
+pronounced trade-off between using unrestricted dropping and sub-flow dropping.
+
+Since, sub-flow dropping showed little difference from unrestricted dropping,
+we focus on comparing unrestricted dropping and source-only dropping. The
+results shown in Figures~\ref{fig:evaluation.umc_policies} and
+\ref{fig:evaluation.bc_policies} were run for 2 billion instructions after
+fast-forwarding for 1 billion instructions.
 The results are shown for an overhead target of 2.0x for UMC and 1.5x for BC.
@@
   \caption{Comparing dropping policies for BC.}
+  \label{fig:evaluation.bc_policies}
   \vspace{-0.2in}
   \end{center}
-  \label{fig:evaluation.bc_policies}
 \end{figure}
@@
 coverage achieved over multiple runs for array bounds check using
-unrestricted and source-only dropping policies with probabilistic dropping.
+unrestricted, sub-flow, and source-only dropping policies with probabilistic dropping.
 These numbers are averaged across all benchmarks. Here, we use a 10\%
 probability
-that events are not dropped. Since the goal is code coverage, coverage here is
+that events are not dropped.  Each run
+was simulated for 500 million instructions. 
+Since the goal is code coverage, coverage here is
 measured as the percentage of static instructions which are monitored in at
-least one of the runs. Each run
-was simulated for 500 million instructions. 
+least one of the runs.
 We see that for the unrestricted dropping policy there is almost no increase in
@@
 policy. 
-Instead, source-only
-dropping is much better suited for achieving high coverage over multiple runs.
-Source-only dropping shows a 6\% increase in coverage with only eight runs.
-With more runs, source-only dropping should be able to
+Instead, sub-flow dropping and source-only
+dropping are much better suited for achieving high coverage over multiple runs.
+Source-only dropping shows a 6\% increase in coverage with only eight runs
+while sub-flow dropping shows a 1\% increase in coverage over eight runs.
+While sub-flow dropping shows a slower increase in coverage than source-only
+dropping, sub-flow dropping is better able to meet overhead targets.
+With enough runs, both sub-flow dropping and source-only dropping should be able to
 reach 100\% code coverage.

### 2015-01-05 policies.tex
Add discussion and eval on sub-flow dropping.

@@
 
-\begin{figure} \begin{center} \subfloat[Unrestricted Dropping]{
+\begin{figure} 
+  \begin{center} 
+    \subfloat[Unrestricted Dropping] {
       \includegraphics[width=\columnwidth]{figs/unrestricted_drop.pdf}
-\label{fig:policies.all_drop} } \vspace{-0.1in} \\ \subfloat[Source-Only
-Dropping]{ \includegraphics[width=\columnwidth]{figs/source_drop.pdf}
-\label{fig:policies.source_drop} } \vspace{-0.1in} \\ \subfloat[Sub-Flow
-Dropping]{ \includegraphics[width=\columnwidth]{figs/subflow_drop.pdf}
-\label{fig:policies.subflow_drop} } \end{center} \caption{Comparison of
-dropping policies using metadata dependence graphs.  Square nodes represent
-events where check are performed. Blue (dark) nodes indicate which nodes can be
-dropped.} \label{fig:policies.policies} \end{figure}
+      \label{fig:policies.all_drop} 
+    }
+    \vspace{-0.1in} \\ 
+    \subfloat[Source-Only Dropping] { 
+      \includegraphics[width=\columnwidth]{figs/source_drop.pdf}
+      \label{fig:policies.source_drop} 
+    } 
+    \vspace{-0.1in} \\ 
+    \subfloat[Sub-Flow Dropping] { 
+      \includegraphics[width=\columnwidth]{figs/subflow_drop.pdf}
+      \label{fig:policies.subflow_drop} 
+    } 
+  \end{center} 
+  \caption{Comparison of dropping policies using metadata dependence graphs.
+  Square nodes represent events where check are performed. Blue (dark) nodes
+  indicate which nodes can be dropped.} 
+  \label{fig:policies.policies} 
+\end{figure}
 
 In addition to deciding when dropping is required, trade-offs also exist in
-deciding which events should be dropped.  The simplest policy is to drop any
+deciding which events should be dropped.  The simplest policy is to drop 
 monitoring events when slack is less than or equal to zero.  However, this can
@@
 An alternative dropping policy which eliminates this wasted work is to only
-make dropping decisions at the root or source of these metadata flows. That is,
+make dropping decisions at the root or source of these metadata flows (see Figure~\ref{fig:policies.source_drop}). That is,
 we will decide to either monitor or not monitor an entire metadata flow. An
@@
 nodes are typically easy to identify by the dropping hardware because they
-correspond to special instructions that are used to set up metadata
+correspond to the special instructions that are used to set up metadata
 information. Thus, it is not necessary to generate and analyze the monitoring
@@
 \emph{unrestricted dropping}.
-
-Figures~\ref{fig:policies.all_drop} and \ref{fig:policies.source_drop} show a
-comparison of where dropping decisions are made for these two policies. Source
-dropping will
-result in no wasted work and thus better coverage at a given overhead. However, because of the coarser-grained decision, it
-may be more difficult to closely match overhead targets. We expect source-only dropping 
-to work well when there are a large number of independent metadata flows.
-In addition, using a probabilistic dropping policy works poorly with
-unrestricted dropping. Since every event in a dependent chain (e.g., events {\tt
-A} through {\tt E}) needs to be monitored in order for the monitoring check to
-be useful, randomly dropping any event is likely to cause the final check to be
-invalid by dropping at least one event in the chain.
 
@@
 Figure~\ref{fig:policies.subflow_drop}, this corresponds to nodes {\tt A}, {\tt
-C}, and {\tt F}. The basic intuition is that dropping certain portions of the
-monitoring dependence graph still result in no wasted work. For example,
+C}, and {\tt F}. The basic idea of sub-flow dropping is to drop portions of the
+monitoring dependence graph at the smallest granularity such that no work is
+wasted.
+For example,
 dropping node {\tt C} causes nodes {\tt D} and {\tt E} to be skipped. However,
@@
 
+%Figures~\ref{fig:policies.all_drop} and \ref{fig:policies.source_drop} show a
+%comparison of where dropping decisions are made for these two policies. 
+Source-only dropping will
+result in no wasted work and thus better coverage at a given overhead than unrestricted dropping. However, because of the coarser-grained decision, it
+may be more difficult to closely match overhead targets. 
+Sub-flow dropping enables a design point in between unrestricted dropping and
+source-only dropping in terms of matching overhead and coverage achieved.
+Sub-flow dropping can still result in wasted work in there are merge points in
+the monitoring dependence graph, but should result in less wasted work than
+unrestricted dropping.
+Thus, its coverage falls in between that of source-only dropping and unrestricted dropping.
+Similarly, sub-flow dropping's ability to closely match overhead targets falls between that of source-only dropping and unrestricted dropping.
+We expect source-only dropping 
+to work well when there are a large number of independent metadata flows.
+In addition, using a probabilistic dropping policy works poorly with
+unrestricted dropping. Since every event in a dependent chain (e.g., events {\tt
+A} through {\tt E}) needs to be monitored in order for the monitoring check to
+be useful, randomly dropping any event is likely to cause the final check to be
+invalid by dropping at least one event in the chain.
+
 Depending on the target application of partial monitoring, different policies
@@
 slack-based, unrestricted dropping policy is appropriate. However, if matching
-the overhead target is not as important, then a slack-based, source-only dropping
+the overhead target is not as important, then a slack-based, sub-flow or source-only dropping
 policy could provide better coverage. 
@@
 multiple runs.
-Sub-flow dropping enables a design point in between unrestricted dropping and
-source-only dropping. Sub-flow dropping matches overhead better than
-source-only dropping and provides better coverage than unrestricted dropping.
-For a probabilistic policy, sub-flow dropping allows better overhead matching
+For a probabilistic policy, sub-flow dropping allows better overhead matching than source-only dropping
 while still allowing good total coverage over multiple runs, though at a slower

### 2015-01-05 dropping.tex
Revised throughout.

@@
-\section{Architecture for Adjustable Monitoring}
+\section{Architecture for Partial Monitoring}
 \label{sec:dropping}
@@
 pointer, then this would allow a future instruction to load the pointer and use
-it to access the corresponding array or data structure. In the second case,
+it to access its corresponding array or data structure. In the second case,
 {\tt y+12} (line 4) corresponds to {\tt x+20} which is not within the array's bounds and
@@
 The propagation and checks of the metadata occur automatically as
-instructions are forwarded.
+instructions are forwarded to the monitor.
 Here, we have shown the dynamic sequence of operations executed by the monitor but
@@
 of run-time monitoring. This dropping can affect the functionality of the
-monitoring schemes. There are three possible outcomes for dropping a
+monitoring scheme. There are three possible outcomes for dropping a
 monitoring operation. The first is that there is no difference in operation
@@
 corresponding metadata. Thus, our general approach for preventing false
-positives is to associate a 1-bit flag with each metadata in order to mark these
+positives is to associate a 1-bit invalidation flag with each metadata in order to mark these
 metadata as valid or
@@
 diagram of this hardware module. The dataflow engine uses two address
-calculation units to read in up to two invalidation flags. These source invalidation flags
+generation units to read in up to two invalidation flags. These source invalidation flags
 are used to decide whether a monitoring event should filtered. A third address
-calculation unit is used to optionally specify a target to propagate the
+generation unit is used to optionally specify a target to propagate the
 invalidation information. The module includes a register file to store invalidation
@@
 and operation of the address generation units are set based on instruction
-type. Note that the address calculation units also take information from the
+type. Note that the address generation units also take information from the
 monitored event as inputs. Thus, in the same way that the monitor selects

### 2015-01-05 evaluation.tex
Update multi-run results.
Revised throughout.

@@
 execution times of 2.6-9.8x with an average of 5.7x.
-For BC, normalized execution
+For BC, normalized execution times
 are 2.7x on average and range from 1.02x to 13.6x.
@@
 %uniformly distributed across checks.
-Note that this metric is different from the
-percentage of monitoring events that are not dropped, which includes
-non-check instructions.
+%Note that this metric is different from the
+%percentage of monitoring events that are not dropped, which includes
+%non-check instructions.
 In the figure, the bottom portion of each
@@
 We see that by varying the overhead budget, the coverage achieved also varies.
-With only a 1.1x overhead target (10\% slowdown), array bounds check still
+With only a 1.1x overhead target, array bounds check still
 achieves over 80\% monitoring coverage on average. The high coverage
@@
 introducing overhead. The second effect is that there may still exist a large number of
-monitoring events that do not lead to security or reliability checks. As a
+monitoring events that do not lead to a bounds checks. As a
 result, dropping these events can reduce overhead without a large impact on
@@
 %average. 
-Even higher coverage can be achieved by allowing higher overhead.
+Even higher coverage can be achieved by allowing higher overhead budgets.
 Similarly, Figure~\ref{fig:evaluation.ls_sweep} shows the coverage achieved by LS.
 With only a 1.01x overhead, LS achieves 30\% coverage on average. This increases to
-71\% coverage with a 1.1x overhead. 
-For DIFT (not pictured), an overhead target of 1.1x is enough for
+71\% coverage with a 1.1x overhead budget. 
+For DIFT (not pictured), an overhead target of 1.05x is enough for
 all benchmarks tested to reach 100\% coverage.
@@
 coverage while the overheads of performing monitoring in full were 2.6x (see
-Figure~\ref{fig:evaluation.filtering}. This is due to the fact that we
+Figure~\ref{fig:evaluation.filtering}). This is due to the fact that we
 accumulate slack gradually and so bursty monitoring events may require a higher
@@
 actual overheads seen are in-line with the overhead of performing monitoring
-without dropping (e.g., 2.6x for {\tt mcf} with UMC monitoring at a 12x
-overhead target).
+without dropping (e.g., {\tt mcf} with UMC monitoring at a 12x
+overhead target shows a 2.6x overhead).
 
@@
 fully-pipelined monitor running on an FPGA fabric at half the frequency of the
-main core. We only show results for UMC due to space constraints though other
+main core. We only show results for UMC due to space constraints, though other
 monitors show similar trends. In this case, the full overheads of
-monitoring range from 1.1-1.9x. Figure~\ref{fig:evaluation.fpga_umc_sweep} shows
+monitoring range from 1.1-1.9x with an average overhead of 1.4x. Figure~\ref{fig:evaluation.fpga_umc_sweep} shows
 the coverage achieved as we sweep the overhead target from 1.01-2.0x. We see that partial
-monitoring can allow the overhead of such a system to be pushed below 10\%
+monitoring can allow the overhead of such a system to be pushed to 1.1x
 while still providing 45\% coverage.
@@
 %In this section we evaluate the trade-offs between unrestricted dropping and source-only dropping.
-In this section we evaluate the trade-offs between different dropping policies.
+In this section, we evaluate the trade-offs between different dropping policies.
 For sub-flow dropping, we identified sub-flow nodes through profiling. Because
@@
 dropping for the benchmarks and monitors that we tested. Monitors and
-benchmarks with difference monitoring dependence graphs may see a more
+benchmarks with different monitoring dependence graphs may see a more
 pronounced trade-off between using unrestricted dropping and sub-flow dropping.
@@
 behavior, source-only dropping can provide better coverage than unrestricted
-dropping with the same overhead. Unrestricted dropping is better at
+dropping with the same overhead. However, unrestricted dropping is better at
 meeting an overhead target.

### 2015-01-05 introduction.tex
Revised throughout.

@@
 hardware designs that target a single or narrow-range of monitors have been
-shown to have low performance overheads. On the other hand, flexible systems
-that support a range of monitors have shown much higher overheads. In this
+shown to have low performance overhead. On the other hand, flexible systems
+that support a range of monitors have shown much higher overhead. In this
 paper, we explore a new trade-off dimension for run-time monitor designs. We
@@
 implemented on a multi-core platform.
-Our results show that the partial monitoring can still achieve significant
+Our results show that partial monitoring can still achieve significant
 coverage with reduced overhead. For example, for the monitors which showed over 2x
@@
 This paper is organized as follows. Section~\ref{sec:monitoring} introduces the notion
-of adjustable run-time monitoring with partial monitoring.
+of adjustable overhead through partial monitoring.
 Section~\ref{sec:dropping} discusses the hardware architecture that enables 

### 2015-01-05 monitoring.tex
Revised throughout.

@@
-\section{Adjustable Run-Time Monitoring}
+\section{Partial Run-Time Monitoring}
 \label{sec:monitoring}
@@
 Table~\ref{tab:monitoring.previous_overheads} summarizes some of the representative designs
-and their reported performance overhead. The previous studies clearly show that there
+and their reported performance overheads. The previous studies clearly show that there
 exist trade-offs between efficiency, flexibility, and hardware costs. 
@@
 {\bf Cooperative testing, debugging, and protection:}
-Recently, studies \cite{liblit-pldi05, chilimbi-asplos04, greathouse-cgo11, testudo-micro08} have shown that software testing, debugging,
+Recent studies have shown that software testing, debugging,
 or attack detection may be done in a cooperative fashion across
-a large number of systems. In this case, each system is only willing to pay very low
+a large number of systems \cite{liblit-pldi05, chilimbi-asplos04, greathouse-cgo11, testudo-micro08}. In this case, each system is only willing to pay very low
 overhead (e.g., a few percent) and only performs a small subset of checks. 
@@
 {\bf Monitoring of soft real-time systems:}
-Soft real-time systems or interactive systems need to meet deadlines or response
-time requirements. Unfortunately, the overhead of run-time monitoring is often 
+Soft real-time systems or interactive systems need to meet deadlines or response-time
+requirements. Unfortunately, the overhead of run-time monitoring is often 
 unpredictable and varies significantly depending on the application and monitor
@@
 monitoring allows partial protection on such systems. For example, array bounds
-may be checked for a subset of memory accesses. For DIFT, a subset of information flows
+may be checked for a subset of memory accesses. For dynamic information flow tracking (DIFT), a subset of information flows
 may be tracked for partial attack detection. In this scenario, the effectiveness of 
@@
 \begin{enumerate}
-  \item \textbf{General:} Because we mainly target flexible run-time monitoring systems,
+  \item \textbf{General:} Since we mainly target flexible run-time monitoring systems,
   which often have high overhead, the framework also needs to be general enough to be

### 2015-01-05 policies.tex
Revised throughout.

@@
 bounds information on {\tt malloc} for an array bounds check. These source
-nodes are typically easy to identify by the dropping hardware because they
+nodes are easily identifiable by the dropping hardware because they typically
 correspond to the special instructions that are used to set up metadata
@@
 More complex dropping policies can be implemented given more detailed
-information about the monitoring dependence graph. Here, we describe one such
-policy, which we call \emph{sub-flow dropping}. Sub-flow dropping makes
+information about the monitoring dependence graph. This information can be
+found through static analysis or profiling, though we do not explore how to
+find it in detail here. Given this information, we describe one possible
+policy to use it, which we call \emph{sub-flow dropping}. Sub-flow dropping makes
 dropping decisions at a granularity in between unrestricted dropping and
-source-only dropping (see Figure~\ref{fig:policies.subflow_drop}). Sub-flow
-dropping allows source nodes to be dropped and nodes after branch points in the
-monitoring dependence graph to be dropped. From
+source-only dropping (see Figure~\ref{fig:policies.subflow_drop}). 
+The basic idea of sub-flow dropping is to drop portions of the monitoring
+dependence graph at the smallest granularity such that no work is wasted.
+Sub-flow dropping allows source nodes to be dropped and nodes after branch
+points in the monitoring dependence graph to be dropped. From
 Figure~\ref{fig:policies.subflow_drop}, this corresponds to nodes {\tt A}, {\tt
-C}, and {\tt F}. The basic idea of sub-flow dropping is to drop portions of the
-monitoring dependence graph at the smallest granularity such that no work is
-wasted.
-For example,
+C}, and {\tt F}. For example,
 dropping node {\tt C} causes nodes {\tt D} and {\tt E} to be skipped. However,
@@
 source-only dropping in terms of matching overhead and coverage achieved.
-Sub-flow dropping can still result in wasted work in there are merge points in
-the monitoring dependence graph, but should result in less wasted work than
+Sub-flow dropping can still result in wasted work if there are merge points in
+the monitoring dependence graph and only one part of the incoming flows is
+dropped, but should result in less wasted work than
+unrestricted dropping. Thus, its coverage falls in between that of source-only
+dropping and unrestricted dropping. Similarly, sub-flow dropping's ability to
+closely match overhead targets falls between that of source-only dropping and
 unrestricted dropping.
-Thus, its coverage falls in between that of source-only dropping and unrestricted dropping.
-Similarly, sub-flow dropping's ability to closely match overhead targets falls between that of source-only dropping and unrestricted dropping.
 We expect source-only dropping 
@@
 debugging and testing with very low overhead, then a probabilistic,
-source-only dropping policy can be used to provide good total coverage over
+source-only or sub-flow dropping policy can be used to provide good total coverage over
 multiple runs.
-For a probabilistic policy, sub-flow dropping allows better overhead matching than source-only dropping
-while still allowing good total coverage over multiple runs, though at a slower
-rate than source-only dropping.
 

### 2015-01-05 related.tex
Revised throughout.

@@
   to bound overhead, which we have shown can vary greatly.
-  \item \textbf{Prevent False Positives:} We prevent a mechanism to prevent
+  \item \textbf{Prevent False Positives:} We present a mechanism to prevent
   false positives. Previous work either has false positives or targets monitoring
@@
 prevents false positives. Their architecture is designed to meet real-time
-deadlines rather than overhead targets. In addition, in order to give strong
+deadlines rather than enable adjustable overheads. In addition, in order to give strong
 guarantees, the coverage achieved is lower than our system. Finally, QVM
 \cite{qvm-oopsla08} proposes a modification to the Java Virtual Machine (JVM) to
-support monitoring with adjustable overhead. QVM does prevent false positives
-but does so by enabling or disabling monitoring on a per-object basis. This is
-similar to the idea of performing source-only dropping which our results show
-can lead to overshooting the overhead target depending on the monitoring
-technique. In addition, QVM is limited monitoring for code run on a JVM while
-our framework works for any binary.
+support monitoring with adjustable overhead. 
+QVM is limited to monitoring for code run on a JVM which limits its
+generalizability while our framework works for any binary. QVM prevents false
+positives by enabling or disabling monitoring on a per-object basis. This
+limits the monitoring schemes that can be implemented. Also, this method of
+preventing false positives is similar to the idea of performing source-only
+dropping which our results show can lead to overshooting the overhead target
+depending on the monitoring technique.
 

### 2015-01-05 dropping.tex
Spacing around tables/figures.
Edits from Andrew's comments.

@@
   \begin{center}
+    \vspace{0.1in}
     \includegraphics[width=\columnwidth]{figs/monitoring_architecture.pdf}
@@
     \label{fig:dropping.overview} 
-    \vspace{-0.1in}
+    \vspace{-0.3in}
   \end{center}
@@
     \label{fig:dropping.example_full}
-    \vspace{-0.1in}
+    \vspace{-0.2in}
   \end{center}
@@
 the main program that is needed is to initially set metadata, such as setting
-the base and bounds addresses on a {\tt malloc} call for array bounds check.
+the base and bounds addresses on a {\tt malloc} call for array bounds checking.
 The propagation and checks of the metadata occur automatically as
 instructions are forwarded to the monitor.
-Here, we have shown the dynamic sequence of operations executed by the monitor but
+Here, we have shown the dynamic sequence of operations executed by the monitor, but
 the actual static code consists of a set of instructions to be run for each
@@
     \label{fig:dropping.example_invalid}
-    \vspace{-0.1in}
+    \vspace{-0.2in}
   \end{center}
@@
   \begin{center}
-    \vspace{-0.1in}
     \includegraphics[width=\columnwidth]{figs/dataflow_overview.pdf}
@@
     \label{fig:dropping.dataflow_overview}
-    \vspace{-0.2in}
+    \vspace{-0.3in}
   \end{center}
@@
     \includegraphics[]{figs/dataflow_architecture.pdf}
-    \vspace{-0.1in}
+    \vspace{-0.0in}
     \caption{Hardware architecture of the dataflow engine.}
     \label{fig:dropping.dataflow} 
-    \vspace{-0.1in}
+    \vspace{-0.3in}
   \end{center}
@@
     \includegraphics[width=\columnwidth]{figs/example_null_filtering.pdf}
-    \vspace{-0.2in}
+    \vspace{-0.1in}
     \caption{Example of using information about null metadata to filter monitoring events.}
     \label{fig:dropping.example_null} 
-    \vspace{-0.1in}
+    \vspace{-0.2in}
   \end{center}

### 2015-01-05 introduction.tex
Edits from Andrew's comments.

@@
 overhead should not exceed energy/power limits or real-time deadlines.
-Additionally, partial monitoring can be used for debugging purposes
-especially in the context of sampling-based or cooperative debugging techniques
+Additionally, partial monitoring can be used for sampling-based or cooperative debugging techniques
 which expect low coverage per run but use a large number of runs and users to

### 2015-01-06 policies.tex
Spacing around tables/figures.
Edits based on Mohamed's comments.

@@
 
-\subsection{Deciding When to Drop} \label{sec:policies.when}
+\subsection{Deciding When to Drop} 
+\label{sec:policies.when}
 
 % Slack
-\begin{figure} \begin{center}
-\includegraphics[width=\columnwidth]{figs/slack.pdf} \vspace{-0.2in}
+\begin{figure} 
+  \begin{center}
+    \includegraphics[width=\columnwidth]{figs/slack.pdf} 
+    \vspace{-0.1in}
     \caption{Slack and its effect on monitoring over time.}
-\label{fig:policies.slack} \vspace{-0.1in} \end{center} \end{figure}
+    \label{fig:policies.slack} 
+    \vspace{-0.3in} 
+  \end{center} 
+\end{figure}
 
@@
 
-\subsection{Deciding Which Events to Drop} \label{sec:policies.which}
+\subsection{Deciding Which Events to Drop} 
+\label{sec:policies.which}
 
@@
     }
-    \vspace{-0.1in} \\ 
+    \\
     \subfloat[Source-Only Dropping] { 
@@
     } 
-    \vspace{-0.1in} \\ 
+    \\
     \subfloat[Sub-Flow Dropping] { 
@@
   \label{fig:policies.policies} 
+  \vspace{-0.2in}
 \end{figure}
@@
 {\tt E} to be checked with no wasted work.
+Note that sub-flow dropping can still result in wasted work if there are merge
+points in the monitoring dependence graph and only part of the incoming flows
+are dropped, but it should result in less wasted work than compared to an
+unrestricted dropping policy.
 
@@
 Source-only dropping will
-result in no wasted work and thus better coverage at a given overhead than unrestricted dropping. However, because of the coarser-grained decision, it
+result in no wasted work and thus better coverage at a given overhead than
+unrestricted dropping. However, because of the coarser-grained decision, it
 may be more difficult to closely match overhead targets. 
@@
 source-only dropping in terms of matching overhead and coverage achieved.
-Sub-flow dropping can still result in wasted work if there are merge points in
-the monitoring dependence graph and only one part of the incoming flows is
-dropped, but should result in less wasted work than
-unrestricted dropping. Thus, its coverage falls in between that of source-only
-dropping and unrestricted dropping. Similarly, sub-flow dropping's ability to
-closely match overhead targets falls between that of source-only dropping and
-unrestricted dropping.
-We expect source-only dropping 
+If the monitoring dependence graph is highly connected, then source-only
+dropping will perform poorly. On the other hand, we expect source-only dropping 
 to work well when there are a large number of independent metadata flows.
-In addition, using a probabilistic dropping policy works poorly with
-unrestricted dropping. Since every event in a dependent chain (e.g., events {\tt
+Monitoring dependence graphs with a large number of branches will favor a
+sub-flow dropping policy.
+
+The choice of dropping policy can also depend on whether probabilistic dropping
+is performed instead of slack-based dropping.
+Probabilistic dropping works poorly with an
+unrestricted dropping policy. Since every event in a dependent chain (e.g., events {\tt
 A} through {\tt E}) needs to be monitored in order for the monitoring check to
 be useful, randomly dropping any event is likely to cause the final check to be
-invalid by dropping at least one event in the chain.
+invalid by dropping at least one event in the chain. Instead, source-only
+dropping works well when performing probabilistic dropping.
 

### 2015-01-06 evaluation.tex
Spacing around tables/figures.
Edits based on Ed's comments.

@@
     \label{fig:evaluation.filtering}
-    \vspace{-0.1in}
+    \vspace{-0.3in}
   \end{center}
@@
     \label{fig:evaluation.bc_sweep}
-    \vspace{-0.2in}
+    \vspace{-0.3in}
   \end{center}
@@
     \label{fig:evaluation.umc_sweep}
-    \vspace{-0.1in}
+    \vspace{-0.2in}
   \end{center}
@@
     \label{fig:evaluation.ls_sweep}
-    \vspace{-0.1in}
+    \vspace{-0.3in}
   \end{center}
@@
     \label{fig:evaluation.imp_sweep}
-    \vspace{-0.1in}
+    \vspace{-0.3in}
   \end{center}
@@
 introducing overhead. The second effect is that there may still exist a large number of
-monitoring events that do not lead to a bounds checks. As a
+monitoring events that do not lead to bounds checks. As a
 result, dropping these events can reduce overhead without a large impact on
@@
     \label{fig:evaluation.fpga_umc_sweep}
-    \vspace{-0.1in}
+    \vspace{-0.3in}
   \end{center}
@@
 In this section, we evaluate the trade-offs between different dropping policies.
-For sub-flow dropping, we identified sub-flow nodes through profiling. Because
-of the need to profile the applications to determine sub-flow nodes, these
-experiments were simulated for 1 billion instructions without fast-forwarding.
-These results (not shown) indicated that sub-flow dropping produced similar
-results in terms of coverage and overhead target matching to unrestricted
-dropping for the benchmarks and monitors that we tested. Monitors and
-benchmarks with different monitoring dependence graphs may see a more
-pronounced trade-off between using unrestricted dropping and sub-flow dropping.
-
-Since, sub-flow dropping showed little difference from unrestricted dropping,
-we focus on comparing unrestricted dropping and source-only dropping. The
-results shown in Figures~\ref{fig:evaluation.umc_policies} and
-\ref{fig:evaluation.bc_policies} were run for 2 billion instructions after
-fast-forwarding for 1 billion instructions.
-The results are shown for an overhead target of 2.0x for UMC and 1.5x for BC.
 Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
-overhead budget and the run-time monitoring overhead for UMC. A positive value
+overhead budget and the run-time monitoring overhead for UMC when the overhead target is set to 2.0x. A positive value
 means that the overhead target was overshot while a negative value indicates
@@
 and Figure~\ref{fig:evaluation.bc_coverage} shows the coverage for BC.
+Here, the overhead target is 1.5x.
 From Figure~\ref{fig:evaluation.bc_exec_time}, we see that for several
@@
 
+The results for the sub-flow dropping policy are not included in these graphs
+because our profiling infrastructure to identify sub-flow nodes currently does
+not support fast-forwarding. Instead, we compared the sub-flow dropping with
+other policies by running simulations for 1 billion instructions without
+fast-forwarding. The results (not shown) showed that sub-flow dropping produced
+similar coverage and overhead target matching to unrestricted dropping for the
+benchmarks and monitors that we tested.
+
 \subsection{Multiple-Run Coverage}
@@
     \includegraphics[width=\columnwidth]{figs/data_multirun_coverage.pdf}
-    \vspace{-0.3in}
+    \vspace{-0.2in}
     \caption{Coverage over multiple runs for BC with a 10\% probability to not drop events.}
     \label{fig:evaluation.multirun}
-    \vspace{-0.3in}
+    \vspace{-0.2in}
   \end{center}
@@
 was simulated for 500 million instructions. 
-Since the goal is code coverage, coverage here is
+Since the effectiveness of cooperative debugging is often measured by code coverage, coverage here is
 measured as the percentage of static instructions which are monitored in at
@@
     are normalized to the main core power.}
-    \vspace{-0.2in}
+    \vspace{-0.1in}
     \label{tab:evaluation.area_power}
